### 2013-01-02 report.tex
Initial commit

%\documentclass[preprint]{sigplanconf}
\documentclass{article}
\usepackage[margin=1in]{geometry}

% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.
\usepackage{fontspec}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{color}
\usepackage{multirow}
\usepackage{sectsty}
\usepackage[titles]{tocloft}
\usepackage{sidecap}
\usepackage{multicol}

\title{Library-based fault tolerance for scientific applications}
\author{Sui Chen for EE7091 independent research, Fall 2012}
\begin{document}

\maketitle

%\conferenceinfo{WXYZ '05}{date, City.} 
%\copyrightyear{2005} 
%\copyrightdata{[to be supplied]}

%\titlebanner{banner above paper title}        % These are ignored unless
%\preprintfooter{short description of paper}   % 'preprint' option specified.

%\subtitle{Subtitle Text, if any}

%\authorinfo{Name1}
%           {Affiliation1}
%           {Email1}
%\authorinfo{Name2\and Name3}
%           {Affiliation2/3}
 %          {Email2/3}

%\maketitle

%\begin{abstract}
%This is the text of the abstract.
%\end{abstract}

%\category{CR-number}{subcategory}{third-level}

%\terms
%Accuracy metric and Running time metric

%\keywords
%Resilience, High-performance computing, Algorithmic fault tolerance, Scientific libraries

This report documents the work based on the DSN 2012 poster (attached) spanning from the summer break to the end of this semester.

\section{Introduction}

As the size of HPC system grows to millions of processor cores and chip features shrink to a fraction of nanometers, HPC applications are increasingly threatened by transient hardware faults. Such faults can cause application aborts and performance degradation. More importantly, they may corrupt application results. As such, applications must be
made resilient to soft faults.

Such faults can be detected by running the application multiple
times and comparing the results. While simple and effective,
this approach is very inefficient. This inefficiency is
particularly important for Exascale systems, which will be
highly vulnerable to soft faults but will also be severely power-
constrained. This makes it necessary to develop alternative
techniques that leverage application semantics to provide
resilience at a lower cost.

We present a library-based approach to algorithmic
resilience where only the key application libraries are made
resilient and no protection is provided to the rest of the
application’s logic. By ensuring that the bulk of algorithmic
modifications is localized to libraries, where most application
time is spent, this approach balances developer effort against
performance and resilience to achieve high overall
productivity. We also explore various trade-offs that could be made in the choice of configurations for the fault tolerance parameters.

\section{Manifestation of errors}

As from the view of a user of supercomputers, errors would manifest themselves by:
\begin{itemize}
\item{Corrupting pointers/data structures, ensued by segmentation faults that terminate the application of other unexpected control flow changes.}
\item{Corrupting results that are used as input into subsequent routines, rendering the output of the application useless.}
\end{itemize}

Over-frequent termination of applications results in longer expected time to get acceptable results if the user had no choice but re-start the application over and over. Specially, this could be illustrated with the following Markov Model \ref{markovModel}.

\begin{SCfigure}
\includegraphics[width=6cm]{routine_markov.png}
\caption{The Markov Model illustrating a numerical routine we're interested in. This routine, under the effects of soft errors, may abort and produce no results at all or it produce erroneous results or under some circumstances produce correct results. If the user keeps restart the application in case of error, the expected completion time of the routine could be computed as $T=(2-c)B+{{1-c} \over {c}} \cdot R$}
\label{markovModel}
\end{SCfigure}

This model reveals the efficneicy concerns of fault-tolerance routines. The user is facing the choice of running multiple times the original application and running the fault-tolerant version of the application which is slower. The overhead of fault-tolerant routines are acceptable only when the fault-tolerant routines are capable of bringing down the \emph{expected} completion time to some acceptable level in comparison to running the original application multiple times.

\section{Library-Based Algorithmic Fault Tolerance}

Library-based algorithmic fault tolerance leverages the special properties of the computation routines used in those applications to provide cost-effective resilience. Because soft faults can corrupt arbitrary parts of the application's execution state, algorithmic fault error checkers could be very hard to implement to cover all possible vulnerabilities in practice, and even if it's possible, the amount of manual labor required would be prohibitively high. As such, to achieve the best balance between the costs of incorrect algorithm results, power use and development time, the most cost-effective approach would be to focus the algorithmic approach on application libraries, whose execution take up most of the CPU time. Libraries such as numerical solvers take up most of applications' execution time, therefore high resilience could be achieved at a low developer cost by focusing on implementing resilience on just those software components.

\section{Fault Injection}

Since the cause of soft errors could be complicated and it's hard to simulate them in an physically accurate way, we use error injection to simulate soft faults. We use a tool based on LLVM originally developed to evaluate the Relax framework \cite{deKruijf:2010:RAF:1816038.1816026} . It compiles the application (together with the library) into LLVM code, chooses random LLVM instructions during the application's execution and corrupts the results. As has been illustrated in the previous section, the errors manifest themselves by corrupting the outputs, or causing unexpected aberration in the control flow of the applications. For errors, to error magnitude is measured multiplicatively: if $x$ is the correct value and $x'$ is the erroneous one, error is measured as $|{x' \over x - 1}|$. This ratio is afterwards used in later sections to compute an \emph{accuracy metric}. The change in control flow may bring the application to a complete stop, requiring a restart, prolonging the expected time needed to get meaningful results. The time is taken into account in the \emph{running time metric}.

Errors are injected into the routines and the fault-tolerance code as well.

\begin{SCfigure}
\includegraphics[width=6cm]{vuln-gsl.png}
\caption{Error vulnerability of GSL routines. The figure shows the fraction of entries in the output matrix (the vertical axis) that have an error large than some magnitude (the horizontal axis). The fault-tolerance mechanism we have used here have reduced the occurances of errors of all magnitudes by one magnitude for routines Rank-K update and Matrix-matrix multication, and has removed the worst errors in the TRSV linear solver. This proves the fault-tolerant versions of the routines are effective.}
\label{vulnerabilityGSL}
\end{SCfigure}

\subsection{List of fault injection applications}

We evaluate our library-based resilience approach by applying it to three applications.

\begin{itemize}
\item{Lasso: a parallel shrinkage and selection method for linear regression. It computes the equation $Ax=b$, where matrix $A$ and vector $b$ are known.}

\item{DRC: Digital Room Correction, a program used to generate correction filters for acoustic compensation of HiFi and audio system in general. It generates FIR correction filters, which can be used with a real time or offline convolver to provide correction.}

\item{Hattrick: An accurate N-Body integrator meant for high control over small systems, specifically meant for perturbing bodies.}

The three applications make intensive use of GSL (the GNU Scientific Library)'s linear algebra, FFT and differential equation solvers and therefore, they are ideal testbeds for our proposed fault-tolerance methods.

\end{itemize}

\section{Fault-Tolerant Routines}

As mentioned before, soft faults would cause corruption in outputs and change the application's control flow as well. Therefore countermeasures to each of these phenomenon is integrated into those fault-tolerance routines: algorithmic checkers and pointer/index replication. Construction of fault-tolerant routines consists of the following steps:

\begin{itemize}
\item{Identifying the most frequently used routines with GProf.}
\item{Constructing a wrapper function and set a signal handler to provide recovery ability upon segmentation faults.}
\item{Replicating critical data structures and pointers between library function calls.}
\item{In case of very high fault rate, replication of indices and pointers may provide further resilience to segmentation faults, albeit at a very high running time cost.}
\end{itemize}

\begin{SCfigure}
\includegraphics[width=8cm]{flow.png}
\caption{Block diagram of fault-tolerant routines showing various parts: input backup, the original routine, result checkers and input restoration. The entire fault-tolerant routine is wrapped in a function call that has the same argument format as the original one, making it easy to be seamlessly integrated into applications.}
\label{faultRoutinesBlockDiagram}
\end{SCfigure}

\section{Algorithmic Error Checkers}

For each routines used in above applications we develop corresponding fault checkers. As noted before, these checkers need be as fast as possible in order that the application could complete in a reasonable amount of time.

\begin{itemize}
\item{Checker for matrix-matrix multiplication:

Matrix-matrix multiplications are checked using a matrix vector multiplication, making use of this identity: $(A \cdot B) \cdot x = A \cdot (B \cdot x)$. Here x is a generated error-checking vector and $A$ and $B$ are inputs to the matrix-matrix multiplication. Time complexities of the original routine and the checker are $O(n^3)$ and $O(n^2)$ respectively.
}

\item{Checker for matrix-vector multiplication:

For matrix-vector multiplication $Ax=b$ , note that the sum of the elements in the result vector $b$ is equal to the dot product of the vector composed of column sums of $A$ and $x$. Time complexity of original routine is $O(n^2)$ multiplications and that of the checker is $O(n^2)$ additions.
}

\item{Checker for Symmetric Rank-K update:

Symmetric Rank-K update is a special case of matrix-matrix multiplication where only the upper/lower half of the output matrix is updated. By decomposing the upper/lower half of the output matrix as a series of sub-matrices (like the sub-cells in a quad tree), the checker takes $O(n^2 \cdot \log{n})$ time, while the original routine woudl take $O(n^3)$.
}

\item{Checker for Cholesky Decomposition:

By multiplying back the upper and lower halves, the checker takes as much time as a regular matrix-matrix multiplication would do. The original routine is an iterative one, but it generally runs longer than the checker.
}

\item{Checker for Fast Fourier Transform:

By using Parseval's Theorem, we can check the result of an FFT of width $n$ in $O(n)$ time. Time complexity of the original routine is $O(\log{n})$.
}

\end{itemize}

Those checkers can detect errors greater than a certain numeric threshold in those computations (determined by the user-defined error checker threshold). Once an error beyond the threshold has occurred, a re-calculation would be initiated.

In the following sections we would show that the error tolerance could not be arbitrarily small in order to be helpful in providing fault-tolerance when we take into account the checker itself may be unreliable since it's also vulnerable to soft errors.

\subsection{Picking an Arithmetic Error Detection Threshold}

Although checkers are faster than the routines they protect, the unreliability of those checkers and the ensuing ``false alarms'' might trigger redundant re-calculations that are unnecessary, causing the application to run for much longer. This is espqcially true with an overly small error detection threshold. On the other hand, an overly loose detection threshold may fail to detect data corruption in output files. The results would be discussed in the results section.

\section{Segmentation Fault Handlers}

Modern operating systems provide convenient means of protecting segmentation faults. A bombination of \texttt{siglongjmp} and \texttt{sigsetjmp} are used to recover from segmentation faults. The overhead of this API call is having to copy the processor state into memory, therefore installing them between computation-intensive method calls adds to neglible performance degradation.

\section{Input Data Protection}

Considering the fact that input data may be damaged, we need a fast yet simple mechanism with which we can easily identify and fix errors in the input. We have implemented a block-checksum-based data correction mechanism. By treating the input array as a matrix and comparing row sums and column sums of the submatrices of that matrix, we are able to detect the existence of errors and fix a fraction of the errors. The protection mechanism has the following parameters:

\begin{itemize}
\item{\texttt{N}, the size of the error correcting sum block.}
\item{\texttt{ECC\_ECC}, whether or not the error correcting code itself should be corrected.}
\end{itemize}

This mechanism is able to correct up to $1/{n^2}$ of the elements in the input array.

\section{Pointer and Index Replication}

The third application Hattrick uses a high-accuracy ODE solver that is capable of estimating errors and detecting and reducing errors to a certain level adaptively. Therefore we would only focus on the application's vulnerability to pointer and index corruptions. Replicating critical variables (pointers and indices) and leveraging Byzantine fault tolerance and correcting those pointers and indices during run-time would raise the applications' resistence to soft errors.

In order to determine a replication strategy we have to determine which variables to replicate and how many replicas to use for replicating them. The following two subsections would be focused on these problems.

\subsection{Determining which variables to replicate}

\begin{center}
\begin{verbatim}
1   for(i=0; i<M; i++) {
2     for(j=0; j<P; j++) {
3       double temp = 0;
4       for(k=0; k<N; k++) {
5         int idx_a = IX_A(i, k);
6         int idx_b = IX_B(k, j);
7         double elem_a = A[idx_a];
8         double elem_b = B[idx_b];
9         temp = temp + elem_a * elem_b;
10      {
11      int idx_c = IX_C(i, j);
12      C[idx_c] = temp;
13  }}
\end{verbatim}
\end{center}

And, suppose we have the following configurations for replication when M, N and P are 100.

\begin{figure}[h!]
\begin{center}
\begin{tabular}{|p{2.3cm}|p{8cm}|}
\hline
%\multicolumn{2}{|l|}{Configuration and pr}\\
Configuration & Protection applied \\
\hline
Baseline & without any protection for indices or pointers \\
Conf I-1 & protect indices j and k in innermost loop (line 4)\\
Conf I-2 & protect index j in second innermost loop (line 2); protect k in innermost loop(line 4) \\
Conf I-3 & protect index k in innermost loop (line 4)\\
Conf I-4 & protect indices i in outermost loop (line 1), j in second innermost loop, k in innermost loop \\
Conf P-1 & protect pointers A and B in innermost loop \\
Conf P-2 & protect pointers A and B in second innermost loop \\
Conf IP-1 & Conf I-4 plus Conf P-2 \\
\hline
\end{tabular}
\end{center}
\caption{Configurations applied upon the previous code snippet.}
\label{fig:confsAndProtections}
\end{figure}

\begin{figure}[h!]
\begin{center}
\begin{tabular}{|p{2.3cm}p{1cm}p{1cm}p{1cm}p{2.2cm}|}
\hline
Configuration & \multicolumn{3}{c}{Completion rate at errors per run} & Running time\\
\hline
Baseline & 0.86 & 0.60 & 0.00 & 1x \\
Conf I-1 & 0.92 & 0.60 & 0.00 & 1.87x \\
Conf I-2 & 0.92 & 0.67 & 0.00 & 1.47x \\
Conf I-3 & 0.92 & 0.65 & 0.00 & 1.14x \\
Conf I-4 & 0.91 & 0.61 & 0.01 & 1.47x \\
Conf P-1 & 0.97 & 0.65 & 0.03 & 2.04x \\
Conf P-2 & 0.95 & 0.69 & 0.05 & 1.28x \\
Conf IP-1 & 0.94 & 0.73 & 0.05 & 1.74x \\
\hline
\end{tabular}
\end{center}
\caption{Effectiveness of the pointer/indices replications configurations listed above}
\label{fig:effectivenesses}
\end{figure}

From the results (Figures \ref{fig:confsAndProtections}, \ref{fig:effectivenesses} on page \pageref{fig:confsAndProtections} and \pageref{fig:effectivenesses}) we see the best resilience is obtained when we protect both pointers and indices and that indices should be protected in the same loop level as where they were declared, so that their values are read and overwritten no more often than their values are incremented by the for loop.

\subsection{Finding the Best Replication Strategy}

In some circumstances we don't have much choice over where to protect those pointers or indices, for example:

\begin{center} \begin{verbatim}
1  double *in, *out;
2  for(i=0; i<M; i++) {
3     complicated_computation1(in, out);
4     complicated_computation2(in, out);
5     complicated_computation3(in, out);
6  }
\end{verbatim} \end{center}

In this scenario, the only place (in this source code) we can protect the values of pointers in and out would be between lines 2, 3, 4, 5 and 6. To make protection more effective, we would use more replicas instead of updating the replicas more often. As more replicas are used, one must be aware that when each replica is cold-loaded into the CPU's register file a load instruction is executed, and it is a possibly vulnerable instruction. Therefore, re-using replicas should reduce the vulnerabilities brought by load instructions. However the longer replicas are kept in memory the more likely the the replicas would also be corrupted. To find the number of replicas, we have tried different numbers at which the variables are replicated, as well as how many pointers we cover. The code experimented was the RK4 integrator used in Hattrick and the results are listed in Figure \ref{fig:greenMtx}.

\begin{figure}[h!]
\begin{center}
\includegraphics[width=10cm]{n_replica_correct_runs_cdb1.png}
\end{center}
\caption{Effectiveness of pointer and index replication in RK4 integrator. The X axis is how many times each pair of replica is used and the Y axis is how many occurrances of a loop variable to protect. Obviously, we have the best resilience when all occurrances are protected. The interesting discovery is that using each replica twice instead of once has the best resilience. The color of the cells indicate the relative resilience: the brighter, the more resilient.}
\label{fig:greenMtx}
\end{figure}

\section{The metrics}

We derive the metrics for expected running time and accuracy of outputs. These metrics make it possible to meaningfully compare results of the same application across different input files with different failure-free execution times. The combined metric is a product of both metrics and therefore would reflect the tradeoff between accuracy and expected running time.

The metrics are computed on each run of an application; the metrics require data collected from both faulty runs and fault-free runs.

\subsection{Accuracy metric}

The correctness metric is defined as:

\[
m_{accu} = \left \{
	\begin{array} {l l}
	10 & \: \text{if RMSD is less than $1e^{-10}$} \\
	-\log(RMSD) & \: \text{otherwise} \\
	\end{array} \right.
\]

The correctness metric makes it possible to make reasonable comparisons between
different configurations (i.e. fault-tolerance techniques used, input size, fault rate)
of the same application. The accuracy metric is bounded to avoid having to take the
logarithm of zero.

\subsection{Running time metric}

\[
m_{runt} = \left \{
	\begin{array} {l l}
	3-\log_2{\frac{RT_{faulty}}{RT_{faultfree}}} & \: \text{if $\frac{RT_{faulty}}{RT_{faultfree}}$ < 8} \\
	0 & \: \text{otherwise} \\
	\end{array} \right.
\]

The running time metric is capped at 0 to reflect our consideration that one run that takes too much time to complete would be much less preferrable for the user. An unreasonably long running time might be caused by an overly strict error detection threshold for the fault rate under which the experiment is executed. Therefore the user might wish to loosen the error detection threshold in exchange for faster execution.

\subsection{Combined metric}

The combined metric is simply a product of the two aforementioned metrics.

\[
metric = m_{accu} \cdot m_{runt}
\]

It would reflect the overall result of tradeoffs between running speed and accuracy: The user may choose to trade some running speed for accuracy and result in a better performance of the faulty application.

\section{Fault Tolerance Experiment Results}

\subsection{Results of LASSO}

We are generating input data for Lasso program. Data size varied from 10x to 100x, and we are injecting errors into Lasso at different fault rates. Figure \ref{fig:lassoTable} is the result of representative runs of Lasso, and figure \ref{fig:chickGraphLasso} has an overview of the influence of different algorithmic checker thresholds.

\begin{figure} \begin{center} \begin{tabular}{p{2.5cm}p{2cm}p{2cm}p{2cm}p{2.5cm}p{2.5cm}}
\hline
Configuration and fault rate & SegFault Rate & Errors handled per run & RMSD per run & Measured Running Time & Expected Running Time \\
\hline
\multicolumn{6}{l}{Input 1 (Size = 40x500)} \\
Baseline 1e-8 & 11\%  &         & 1.13e-06 & 1.339s & 1.504s \\
FT 1e-8       & 0\%   & 1.49    & 3.7e-07  & 1.617s & 1.617s \\
Baseline 1e-7 & 76\%  &         & 5.85e-4  & 1.343s & 5.595s \\
FT 1e-7       & 15\%  & 14.29   & 6.44e-5  & 1.630s & 1.630s \\
Baseline 1e-6 & 100\% &         &          &        & Infinity \\
FT 1e-6       & 95\%  & 224.2   & 3.48e-5  & 1.915s & 38.3s  \\
\hline
\multicolumn{6}{l}{Input 2 (Size = 400x500)} \\
Baseline 1e-9 & 23\%  &         & 9.06e-5  & 3.947s & 5.126s \\
FT 1e-9       & 2\%   & 3.21    & 2.01e-5  & 7.623s & 7.779s \\
Baseline 1e-8 & 96\%  &         & 6.5e-4   & 3.928s & 98.2s  \\
FT 1e-8       & 45\%  & 59.89   & 4.5e-3   & 14.67s & 26.67s \\
\hline
\multicolumn{6}{l}{Input 3 (Size = 4000x500)} \\
Baseline 1e-11 & 32\% &         & 3.36e-6  & 28.92s & 42.53s \\
FT 1e-11      & 0\%   & 3.4     & 3.61e-6  & 62.6s  & 62.6s  \\
Baseline 1e-10 & 62\% &         & 6.56e-7  & 28.92s & 76.10s \\
FT 1e-10      & 0\%   & 7.1     & 1.64e-6  & 71.56s & 71.56s \\
Baseline 1e-9 & 80\%  &         & 2.88e-5  & 28.92s & 144.6s \\
FT 1e-9       & 20\%  & 36.8    & 5.83e-6  & 159.07s& 198.8s \\
\hline

\end{tabular}
\caption{Raw results of epresentative runs of Lasso}
\label{fig:lassoTable}
\end{center}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=16cm]{OverallMetric_Lasso_Boxplot.png}
\end{center}
\caption{Visualization of overall metrics of 73718 runs of Lasso throuout 256 different configurations. The facets correspond to different input sizes. The X and Y axes inside one facet correspond to different error checker thresholds and the overall metrics.}
\end{figure}

According to figure \ref{fig:lassoTable} For small inputs the fault-tolerant version of Lasso redcued segmentation fault rate and also reduced the expected completion time. For larger inputs, the fault tolerant version has an estimated running time on par with the non-fault-tolerant counterparts, but has less erroneous results, especially when the error rate is large.

\subsection{Results of DRC}

\begin{figure}[h!]
 \begin{center} \begin{tabular}{p{2.5cm}p{2cm}p{2cm}p{2.5cm}p{2.5cm}p{2.5cm}}

\hline
Configuration & SegFault Rate & Errors handled per run & RMSD per run & Measured Running Time & Estimated Running Time\\
\hline
Baseline 1e-11 & 66\% &         & 1.98e-4 & 26.51s & 77.97s \\
FT 1e-11       & 4\%  & 3       & 3.61e-6 & 35.54s & 37.02s \\
Baseline 1e-10 & 68\% &         & 6.56e-7 & 26.51s & 82.84s \\
FT 1e-10       & 15\% & 30      & 1.64e-6 & 37.65s & 44.29s \\
\hline
\end{tabular} 
\end{center}

\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=16cm]{Overall_Metric_DRC.png}
\end{center}
\caption{Visualization of overall metrics of 14936 runs of DRC throughout 95 different configurations. The facets correspond to different input sizes. The X and Y axes inside one facet correspond to different error checker thresholds and the overall metrics.}
\end{figure}

From the results we could see our fault tolerance methods are both effective at reducing errors in output and reducing estimated running time. From the overall visualization we see the best choice of error checker threshold should be 1e-07. The optimal threshold is 1 magnitude smaller than that of Lasso due to the different characteristics of the routines used in both applications, the FFT and linear algebra.

\subsection{Results of Hattrick}

(Work still in progress)

\section{Estimation of running time under low error rates}

The results from previous sections have shown our resilience techniques are capable of protecting applications at a high error rate, when both the routines and the error checkers are vulnerable. As it's estimating runnning times for lower error rates may be inefficient and would require many runs, we propose using a Markov model to predict them.

Estimated completion time is computed in this way:

\[
T=(2-c) B + \frac{1-c}{c} \cdot R
\]

Where $c$ and $B$ are respectively the probability and the time it would take for the non-fault-tolerant version of this routine to complete under a certain fault rate and $R$ is the time spent in recovery from segmentation faults.

\begin{SCfigure}
\includegraphics[width=8cm]{model1.png}
\label{markovModel}
\caption{Markov Model used for modeling running time under low error rates.}
\end{SCfigure}

\section{Conclusion}

As far as the scope of this report is concerned, we have extended our experiments from the Lasso application in the DSN paper to another two applications, DRC and Hattrick (By the time the report is submitted, Hattrick is still in progress). We have also taken the step into exploring the configuration space of error checker thresholds and have found there should exist an optimal threshold choice that varies according to experiment configurations. We have devised two metrics, the accuracy metric and the running time metric, to quantify the effectiveness of our fault-tolerance mechanisms.

\nocite{5764677}
\nocite{Rinard:2006:PAB:1183401.1183447}
\nocite{deKruijf:2010:RAF:1816038.1816026}
\nocite{deKruijf:2010:RAF:1815961.1816026}
\nocite{Riesen:2011:SAR:2238436.2238466}
\nocite{1386657}
\nocite{journals/ijhpca/CappelloGGKKS09}
\nocite{Restrepo-Calle:2010:CIF:1811212.1811218}
\nocite{6264672}
\nocite{deKruijf:2011:IPA:2155620.2155637}
\nocite{Baek10green:a}
\nocite{Hoffmann:2011:DKR:1961295.1950390}
\nocite{Chan:2009:AMP:1654059.1654065}
\nocite{Cui:2012:EPO:2355585.2355587}
\nocite{1402092}
\nocite{deKruijf:2012:SAC:2345156.2254120}
\nocite{Abadi:2009:CIP:1609956.1609960}
\nocite{Reis:2005:SFT:1113841.1113843}
\nocite{lee2011sef}
\nocite{6012845}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\appendix
%\section{Appendix Title}

%This is the text of the appendix, if you need one.

%\acks

%Acknowledgments, if needed.

% We recommend abbrvnat bibliography style.

%\bibliographystyle{abbrvnat}
\bibliographystyle{plain}
% The bibliography should be embedded for final submission.

\bibliography{./bibs}

\end{document}

### 2013-01-07 paper.tex
Greg's initial text

\documentclass[prodmode,acmtecs]{acmsmall} % Aptara syntax
\usepackage{url}
%\usepackage{fontspec}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{color}
\usepackage{multirow}
\usepackage[ruled]{algorithm2e}

\newcommand{\sui}[1]{%
  \textcolor{green}{SC - #1}
}

\newcommand{\marc}[1]{%
  \textcolor{red}{[MC: #1]}
}

\newcommand{\greg}[1]{%
  \textcolor{blue}{GB: #1}
}

\title{Library-based fault tolerance for scientific applications}
\author{Sui Chen
\affil{Louisiana State University}
Greg Bronevetsky, Marc Casas-Guix
\affil{Lawrence Livermore National Laboratory}}

\begin{document}

\section{Introduction}
\label{sec:intro}

HPC systems are becoming larger and more vulnerable to soft faults.

While error correcting codes have been very effective at making memories and caches resilient to soft faults, they are more expensive for protecting core-internal state such as latches and are significantly less effective for checking the correctness of computations.
Traditional approaches like replication are expensive, making algorithmic resilience techniques a very promising alternative.

However, since such techniques depend on the semantics of the application, they require manual effort to incorporate into applications and may require deep insight into an application's invariants to deploy effectively.
Fortunately, the software components of most real-world applications are asymmetric in their processor utilization, with most of the time spent in a few components.
In scientific applications these components are usually various types of libraries, including numeric solvers domain-specific scientific packages.
This indicates that if by focusing efforts on making key library routines resilient to soft faults it would be possible to efficiently protect the most expensive portions of application code, allowing simpler but more expensive techniques such as replication for the remaining code regions.

This paper evaluates this approach in the context of three different types of applications: the Lasso linear solver~\cite{}, the Hattrick gravitational simulation~\cite{} and the DRC acoustic correction application~\cite{}.
We show how these three applications can be comprehensively protected from soft faults by adding three different resilience mechanisms to each of their routines: algorithmic error checks, replication of critical data structures and checkpoint-restart of individual routines.
We demonstrate via fault injection experiments that although none of these techniques can individually protect applications from soft faults, their combination is highly effective.
Further, we demonstrate how these techniques can be parameterized to offer the right level of protection at any given fault rate or input type, and further can be tuned to bound the probability that the application will produce erroneous results or executes for longer than some limit.

\section{Soft Faults}
\label{sec:soft_faults}

Causes of soft faults and how they manifest themselves in applications.

Discuss prior work on replication and algorithmic techniques, emphasize that in this paper we're focusing on software resilience since hardware resilience is extremely difficult to deploy in real processors since the costs have to be carried by non-HPC markets that care much less about it.

Can show plots of how routine results are affected by injected errors. Should show this for all the routines we'll be considering in this paper.

Describe fault injection framework and how it approximates real faults.

\section{Target Applications}
\label{sec:apps}

We evaluate our library-based resilience approach by applying it to the following three applications.
As they make intensive use library routines, they are ideal testbeds for our proposed fault-tolerance methods.

\greg{This section needs to include more detailed descriptions of the algorithms, the routines where they spend the bulk of their time and what fraction of time is spent there across the range of inputs. We also need to describe the types of inputs we'll be giving them in our experiments.}
\subsection{Lasso}
\label{sec:apps:lasso}
A parallel shrinkage and selection method for linear regression. It computes the equation $Ax=b$, where matrix $A$ and vector $b$ are known. \greg{is it a shrinkage and selection method or an actual solver?}. Utilizes GSL (the GNU Scientific Library)'s linear algebra routines.

\subsection{DRC: Digital Room Correction}
\label{sec:apps:drc}

A program used to generate correction filters for acoustic compensation of HiFi and audio system in general. It generates FIR correction filters, which can be used with a real time or offline convolver to provide correction. Utilizes the GSL FFT routine.

\subsection{Hattrick}
\label{sec:apps:hattrick}
An accurate N-Body integrator meant for high control over small systems, specifically meant for perturbing bodies. Utilizes a differential equation solver routine.

\section{Resilience Techniques}
\label{sec:res_tech}

\subsection{Error Detection}
\label{sec:res_tech:err_det}
\subsubsection{Algorithmic Detection}
\label{sec:res_tech:err_det:algo}

\greg{This section needs performance results. How much slower are the resilient routines and how much more resilient do they become across a range of detection thresholds.}

\paragraph{Matrix-matrix multiplication (MMM)}: 
Matrix-matrix multiplications are checked using a matrix vector multiplication, making use of this identity: $(A \cdot B) \cdot x = A \cdot (B \cdot x)$. Here x is a generated error-checking vector and $A$ and $B$ are inputs to the matrix-matrix multiplication. Time complexities of the original routine and the checker are $O(n^3)$ and $O(n^2)$ respectively.

\paragraph{Matrix-vector multiplication (MVM)}: 
For matrix-vector multiplication $Ax=b$ , note that the sum of the elements in the result vector $b$ is equal to the dot product of the vector composed of column sums of $A$ and $x$. Time complexity of original routine is $O(n^2)$ multiplications and that of the checker is $O(n^2)$ additions.

\paragraph{Symmetric Rank-K update (SYRK)}:
Symmetric Rank-K update is a special case of matrix-matrix multiplication where only the upper/lower half of the output matrix is updated. By decomposing the upper/lower half of the output matrix as a series of sub-matrices (like the sub-cells in a quad tree), the checker takes $O(n^2 \cdot \log{n})$ time, while the original routine woudl take $O(n^3)$.

\paragraph{Cholesky Decomposition}: \greg{Need to define this decomposition}
By multiplying back the upper and lower halves, the checker takes as much time as a regular matrix-matrix multiplication would do. The original routine is an iterative one, but it generally runs longer than the checker. \greg{How much longer? Need to quantify.}

\paragraph{Fast Fourier Transform (FFT)}:
By using Parseval's Theorem, we can check the result of an FFT of width $n$ in $O(n)$ time. Time complexity of the original routine is $O(\log{n})$.
\greg{What is this theorem, how is it used?}

\paragraph{Runge-Kutta PDE Solver (RK)}:
\greg{Describe the multiple step size checker in Hattrick}.

Those checkers can detect errors greater than a certain numeric threshold in those computations (determined by the user-defined error checker threshold). Once an error beyond the threshold has occurred, a re-calculation would be initiated.
\greg{You need to be more specific. Each routine produces a result, your checker produces a result. How are these results compared and how is the threshold used to determine if there is an error or not.}

In the following sections we would show that the error tolerance could not be arbitrarily small in order to be helpful in providing fault-tolerance when we take into account the checker itself may be unreliable since it's also vulnerable to soft errors.
\greg{How is this shown? We certainly do show tighter detection tolerances cause longer executions, which cause more vulnerability. However, this is not the same as what you've just described}

Although checkers are faster than the routines they protect, the unreliability of those checkers and the ensuing ``false alarms'' might trigger redundant re-calculations that are unnecessary, causing the application to run for much longer. This is specially true with an overly small error detection threshold. On the other hand, an overly loose detection threshold may fail to detect data corruption in output files. The results would be discussed in the results section.

\subsubsection{Memory Fault Detection}
\label{sec:res_tech:err_det:mem}

Modern operating systems provide convenient means of protecting segmentation faults. A combination of \texttt{siglongjmp} and \texttt{sigsetjmp} are used to recover from segmentation faults. The overhead of this API call is having to copy the processor state into memory, therefore installing them between computation-intensive method calls adds to negligible performance degradation.

\subsubsection{Targeted Replication}
\label{sec:res_tech:err_det:repl}

The third application Hattrick uses a high-accuracy ODE solver that is capable of estimating errors and detecting and reducing errors to a certain level adaptively. Therefore we would only focus on the application's vulnerability to pointer and index corruptions. Replicating critical variables (pointers and indices) and leveraging Byzantine fault tolerance and correcting those pointers and indices during run-time would raise the applications' resistance to soft errors.
\greg{Need to explain why we only do this for Hattrick.}

In order to determine a replication strategy we have to determine which variables to replicate and how many replicas to use for replicating them. The following two subsections would be focused on these problems.
\greg{The report spends a lot of space talking about this but in the end the conclusion is that we just a pick a few options and evaluate them experimentally. How much space should we devote to this discussion.}

\subsection{Recovery via Checkpoint-Restart}
\label{sec:res_tech:cr}

When errors are detected we roll the application back to the start of the current routine.
This is done by using \texttt{siglongjmp} and \texttt{sigsetjmp}. \greg{This is the place to describe these routines and how we use them.}

Further, since the fault may have corrupted the routine's initial state, it is necessary to protect it via a fast yet simple mechanism with which we can easily identify and fix errors in a routine's inputs. 
We have implemented a block-checksum-based data correction mechanism. 
By treating the input array as a matrix and comparing row sums and column sums of the submatrices of that matrix, we are able to detect the existence of errors and fix a fraction of the errors. 
The protection mechanism has the following parameters:
\begin{itemize}
\item{\texttt{N}, the size of the error correcting sum block.}
\item{\texttt{ECC\_ECC}, whether or not the error correcting code itself should be corrected.}
\end{itemize}
This mechanism is able to correct up to $1/{n^2}$ of the elements in the input array.
\greg{Which code is being used here? You just say that it is based on block-checksums}.

\section{Evaluation of Resilience Methodology}
\label{sec:eval}

\end{document}

### 2013-01-07 report.tex
Makefile and tex file for pdflatex

@@
 % authoryear    To obtain author/year citation style instead of numeric.
-\usepackage{fontspec}
+%\usepackage{fontspec} % Doesn't work with LaTeX
 \usepackage{graphicx}
@@
 \usepackage{color}
-\usepackage{multirow}
-\usepackage{sectsty}
-\usepackage[titles]{tocloft}
-\usepackage{sidecap}
+%\usepackage{multirow}
+%\usepackage{sectsty}
+%\usepackage[titles]{tocloft}
+%\usepackage{sidecap}
 \usepackage{multicol}
@@
 
-\begin{SCfigure}
-\includegraphics[width=6cm]{routine_markov.png}
+\begin{figure}
+\includegraphics[width=6cm,natwidth=398,natheight=162]{routine_markov.png}
 \caption{The Markov Model illustrating a numerical routine we're interested in.
@@
 \label{markovModel}
-\end{SCfigure}
+\end{figure}
 
@@
 
-\begin{SCfigure}
-\includegraphics[width=6cm]{vuln-gsl.png}
+\begin{figure}
+\includegraphics[width=6cm,natwidth=440,natheight=401]{vuln-gsl.png}
 \caption{Error vulnerability of GSL routines.
@@
 \label{vulnerabilityGSL}
-\end{SCfigure}
+\end{figure}
 
@@
 
-\begin{SCfigure}
-\includegraphics[width=8cm]{flow.png}
+\begin{figure}
+\includegraphics[width=8cm,natwidth=560,natheight=507]{flow.png}
 \caption{Block diagram of fault-tolerant routines showing various parts: input backup, the original routine, result checkers and input restoration.
@@
 \label{faultRoutinesBlockDiagram}
-\end{SCfigure}
+\end{figure}
 
@@
 \begin{center}
-\includegraphics[width=10cm]{n_replica_correct_runs_cdb1.png}
+\includegraphics[width=10cm,natwidth=1000,natheight=800]{n_replica_correct_runs_cdb1.png}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=16cm]{OverallMetric_Lasso_Boxplot.png}
+\includegraphics[width=16cm,natwidth=1156,natheight=725]{OverallMetric_Lasso_Boxplot.png}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=16cm]{Overall_Metric_DRC.png}
+\includegraphics[width=16cm,natwidth=1160,natheight=744]{Overall_Metric_DRC.png}
 \end{center}
@@
 
-\begin{SCfigure}
-\includegraphics[width=8cm]{model1.png}
+\begin{figure}
+\includegraphics[width=8cm,natwidth=488,natheight=243]{model1.png}
 \label{markovModel}
 \caption{Markov Model used for modeling running time under low error rates.}
-\end{SCfigure}
+\end{figure}
 

### 2013-01-07 paper.tex
More edits, good ACM format

@@
-\documentclass[prodmode,acmtecs]{acmsmall} % Aptara syntax
+\documentclass[prodmode,acmtecs]{acmconf} % Aptara syntax
 \usepackage{url}
-%\usepackage{fontspec}
 \usepackage{graphicx}
@@
 
+\ConferenceShortName{ICS}
+\ConferenceName{International Conference on Supercomputing}
+
 \title{Library-based fault tolerance for scientific applications}
 \author{Sui Chen
-\affil{Louisiana State University}
+%\affil{Louisiana State University}
 Greg Bronevetsky, Marc Casas-Guix
-\affil{Lawrence Livermore National Laboratory}}
+%\affil{Lawrence Livermore National Laboratory}
+}
 
 \begin{document}
+
+\maketitle
 
@@
 \label{sec:res_tech:err_det}
+
 \subsubsection{Algorithmic Detection}
@@
 
-Those checkers can detect errors greater than a certain numeric threshold in those computations (determined by the user-defined error checker threshold).
+We evaluate these algorithmic error checkers by measuring their costs and effectiveness when running at different error rates on different types of inputs.
+
+Experiment 1: Measure performance degradation of using error detectors as input sizes increase.
+In case of RK if we can run a version that doesn't do the multiple step sizes, do so but if this is difficult then don't worry about it.
+
+Experiment 2: Measure the error in the outputs of the algorithms with and without the algorithmic detectors. 
+First, show a probability distribution histogram for all the codes at one error rate, input size and algorithmic detection threshold.
+Then show a summary of the entire space using the same heatmaps that we're using for final application results.
+Put the input size and error rate on the y-axis and the detection threshold on the x-axis.
+The color in each box should be the fraction of entries in the output that have an error of a given magnitude.
+Show a separate plot for a few different magnitudes.
+
+Experiment 3: Measure the fraction fraction of routine runs that are interrupted by the detection of an error.
+Display this information via the same heatmaps as above.
+
+For each experiment we need to include a discussion of what it means for how these checkers behave within the full-application runs.
+
+%Those checkers can detect errors greater than a certain numeric threshold in those computations (determined by the user-defined error checker threshold).
 Once an error beyond the threshold has occurred, a re-calculation would be initiated.
-\greg{You need to be more specific.
+%\greg{You need to be more specific.
 Each routine produces a result, your checker produces a result.
@@
 
-In the following sections we would show that the error tolerance could not be arbitrarily small in order to be helpful in providing fault-tolerance when we take into account the checker itself may be unreliable since it's also vulnerable to soft errors.
-\greg{How is this shown? We certainly do show tighter detection tolerances cause longer executions, which cause more vulnerability. However, this is not the same as what you've just described}
+%In the following sections we would show that the error tolerance could not be arbitrarily small in order to be helpful in providing fault-tolerance when we take into account the checker itself may be unreliable since it's also vulnerable to soft errors.
+%\greg{How is this shown? We certainly do show tighter detection tolerances cause longer executions, which cause more vulnerability. However, this is not the same as what you've just described}
 
-Although checkers are faster than the routines they protect, the unreliability of those checkers and the ensuing ``false alarms'' might trigger redundant re-calculations that are unnecessary, causing the application to run for much longer.
+%Although checkers are faster than the routines they protect, the unreliability of those checkers and the ensuing ``false alarms'' might trigger redundant re-calculations that are unnecessary, causing the application to run for much longer.
 This is specially true with an overly small error detection threshold.
@@
 
-Modern operating systems provide convenient means of protecting segmentation faults.
-A combination of \texttt{siglongjmp} and \texttt{sigsetjmp} are used to recover from segmentation faults.
-The overhead of this API call is having to copy the processor state into memory, therefore installing them between computation-intensive method calls adds to negligible performance degradation.
+Modern operating systems and hardware efficiently detect accesses to memory regions that are not allocated to the application and make it easy for the application to detect such errors by catching Segmentation Fault and Bus Error signals.
+We have used this mechanism to protect applications from a subset of memory errors.
 
-\subsubsection{Targeted Replication}
-\label{sec:res_tech:err_det:repl}
+Experiment 1: Show the fraction of runs of each routine that is aborted as a result of a memory error
 
-The third application Hattrick uses a high-accuracy ODE solver that is capable of estimating errors and detecting and reducing errors to a certain level adaptively.
+Errors that do not cause the application to access an invalid memory location are detected by replicating key pointers and loop iteration variables.
+The Hattrick application uses a high-accuracy ODE solver that is capable of estimating errors and detecting and reducing errors to a certain level adaptively. 
 Therefore we would only focus on the application's vulnerability to pointer and index corruptions. 
@@
 
-In order to determine a replication strategy we have to determine which variables to replicate and how many replicas to use for replicating them.
+To determine a replication strategy we have to determine which variables to replicate and how many replicas to use for replicating them.
 The following two subsections would be focused on these problems.
@@
 
+Experiment 2: Measure the effectiveness and performance costs of replication. Need to explain why we don't use it for Lasso and DRC.
+
 \subsection{Recovery via Checkpoint-Restart}
@@
 
-When errors are detected we roll the application back to the start of the current routine.
-This is done by using \texttt{siglongjmp} and \texttt{sigsetjmp}. \greg{This is the place to describe these routines and how we use them.}
+When errors are detected we employ a recovery method based on checkpoint-restart.
+At the start of each routine the state of all inputs is recorded and an error correcting code is used to ensure that the contents of this backup copy are not corrupted.
+Further, we use the \texttt{sigsetjmp} routine to record the application's execution state at the routine's start.
+If an error is detected during the routine \texttt{siglongjmp} is used to return the application's execution back to the start of the routine (this function unrolls function calls as needed and reloads the process counter and register state).
+The recovery procedure then resets the routine's inputs from their backed-up copies, overwriting any changes or corruptions that may have occurred during the routine's execution, and restarts its execution.
+%A combination of \texttt{siglongjmp} and \texttt{sigsetjmp} are used to recover from segmentation faults.
+The overhead of this API call is having to copy the processor state into memory, therefore installing them between computation-intensive method calls adds to negligible performance degradation.
 
-Further, since the fault may have corrupted the routine's initial state, it is necessary to protect it via a fast yet simple mechanism with which we can easily identify and fix errors in a routine's inputs.
-We have implemented a block-checksum-based data correction mechanism. 
+The backup copy of routine inputs is protected via a block-checksum-based data correction mechanism.
 By treating the input array as a matrix and comparing row sums and column sums of the submatrices of that matrix, we are able to detect the existence of errors and fix a fraction of the errors.
@@
 
+Experiment 1: Measure the cost of setjmp and longjmp in our routines as the input sizes vary
+Experiment 2: Measure the cost of input backup and input recovery in our routines as the input sizes vary
+
 \section{Evaluation of Resilience Methodology}
@@
 
+\greg{This is the section where we put all the pieces together and present the experimental results on full applications.
+We should describe our main use-case here: user wants to achieve a given accuracy with a given probability and we configure the algorithm to provide this at the lowest cost.
+QUESTION: how do we pick the parameters of each resilience technique?
+Which metric is being optimized?}
+
+\subsection{Performance Impact}
+\label{sec:eval:perf}
+
+Show the Markov model that describes our rollback scheme.
+
+Experiment 1: Measure the slowdown of each application at the optimal configuration
+Experiment 2: Measure the rate of aborts due to algorithmic errors, segfaults and redundant copy mismatches (separate the causes)
+
+\subsection{Result Accuracy}
+\label{sec:eval:acc}
+
+Experiment: Measure the accuracy of the results produced in completed runs, including error bars on the variability of the errors
+
+\subsection{Overall Evaluation}
+\label{sec:eval:overall}
+
+Experiment: Final graph that shows slowdown for each desired probability of perfect results.
+
+\section{Summary}
+\label{sec:summary}
+
 \end{document}

### 2013-01-07 paper.tex
Added two experiments-to-do

@@
 \greg{This section needs performance results. How much slower are the resilient routines and how much more resilient do they become across a range of detection thresholds.}
+\sui{Performance experiments inserted here: running algorithmic checkers across different input sizes and fault rates: 62x62 to 500x500 matrixes for MM, RK, CD, MV checkers; FFT sizes of $2^8$ to $2^{24}$ for the FFT checker; Error bars should apply here, especially for the Cholesky Decomposition checker}
 
@@
 \greg{How much longer?
-Need to quantify.}
+Need to quantify.} \sui{Can we quantify it using the performance data?}
 
@@
 QUESTION: how do we pick the parameters of each resilience technique?
-Which metric is being optimized?}
+Which metric is being optimized?} \sui{Using the choices of those parameters derived from the experiments we have already performed (treat it as a ``training set'') and use those choices on some new experiment configurations (treat as ``test set''), and see what is the probability that those choices could satisfy the users' experiments.}
 

### 2013-01-11 paper.tex
test
Updated intro

@@
 
-HPC systems are becoming larger and more vulnerable to soft faults.
+The increasing size and complexity of (High-Performance Computing) HPC systems is making them increasingly vulnerable to soft faults.
+Soft faults are transient corruptions of circuit state caused by physical phenomena such as strikes by neutrons or alpha particles~\cite{baumann:2005, asciQSER:2005} or thermal electrical noise~\cite{therm_noise:2007}.
+They can affect the state of processor latches and registers and may cause the application to crash or worse, to silently return incorrect results~\cite{mpi_ser:reed:2004}.
+As the electronic feature sizes shrink technology scaling will make soft errors a larger problem~\cite{err_scaling:2012} due to the fact that each circuit element will hold less charge and can thus be disrupted by easily.
+These phenomena make it imperative to develop techniques to make HPC systems resilient to soft faults.
 
-While error correcting codes have been very effective at making memories and caches resilient to soft faults, they are more expensive for protecting core-internal state such as latches and are significantly less effective for checking the correctness of computations.
-Traditional approaches like replication are expensive, making algorithmic resilience techniques a very promising alternative.
+Resilience is a problem that must be addressed at all levels.
+While materials science and circuit design techniques can be used to improve resilience, they come at a cost in reduced power efficiency and performance that an become prohibitive if processors need to be sufficiently reliable to build a large HPC system.
+Techniques such as error correcting codes have been very effective at making memories and caches resilient to soft faults~\cite{mem_errors:2010}.
+However, they are more expensive for protecting core-internal state such as latches and are significantly less effective for checking the correctness of computations.
+Traditional approaches like cross-core or node replication~\cite{rmpi:2011, dyn_cmp_repl:2007} are expensive in their use of computing resources and power and their long error detection latencies require processor state to be frequently checkpointed.
+Processor designs that incorporate instruction replication~\cite{repl_smt:2000} offer fine-grained error detection and rollback but require more power as well as novel hardware features that are not likely to be included in the commodity processors must be used in HPC systems to limit their cost.
 
-However, since such techniques depend on the semantics of the application, they require manual effort to incorporate into applications and may require deep insight into an application's invariants to deploy effectively.
-Fortunately, the software components of most real-world applications are asymmetric in their processor utilization, with most of the time spent in a few components.
-In scientific applications these components are usually various types of libraries, including numeric solvers domain-specific scientific packages.
-This indicates that if by focusing efforts on making key library routines resilient to soft faults it would be possible to efficiently protect the most expensive portions of application code, allowing simpler but more expensive techniques such as replication for the remaining code regions.
+The limitations of automated resilience techniques has motivated the need for algorithmic techniques that can detect and tolerate errors by taking advantage of application-specific properties and invariants~\cite{amg_abft:2012, robustification:2010, 1bft:1984}.
+Although these techniques can be very efficient, they also require manual effort to incorporate into applications and may require deep insight into an application's invariants to deploy effectively.
+Fortunately, because real-world applications spend most of their execution is spent in a few application modules, most soft faults will occur in those modules rather than uniformly throughout the application.
+In scientific applications these modules are usually various types of libraries, including numeric solvers and domain-specific scientific packages.
+This indicates that if by focusing algorithmic resilience efforts on the key library routines, while leaving other modules unprotected or protected via more expensive mechanisms, it is possible to fully protect an application at a low cost.
 
-This paper evaluates this approach in the context of three different types of applications: the Lasso linear solver~\cite{}, the Hattrick gravitational simulation~\cite{} and the DRC acoustic correction application~\cite{}.
-We show how these three applications can be comprehensively protected from soft faults by adding three different resilience mechanisms to each of their routines: algorithmic error checks, replication of critical data structures and checkpoint-restart of individual routines.
+This paper evaluates this approach in the context of three different types of applications: the MPI implementation of using the Alternating Direction Method of Multipliers (ADDR) to solver Lasso problems~\cite{lasso:2011}, the Hattrick gravitational simulation~\cite{hattrick:2012} and the Digital Room Correction (DRC) acoustic correction application~\cite{drc:2012}.
+We show how these three applications can be comprehensively protected from soft faults by adding three different resilience mechanisms to the routines where they spend the most time: algorithmic error checks, replication of critical data structures and checkpoint-restart of individual routines.
 We demonstrate via fault injection experiments that although none of these techniques can individually protect applications from soft faults, their combination is highly effective.

### 2013-01-12 paper.tex
Wrote text through Sec 4.2

@@
-\documentclass[prodmode,acmtecs]{acmconf} % Aptara syntax
+%\documentclass[prodmode,acmtecs]{acmconf} % Aptara syntax
+\documentclass{sig-alternate}
 \usepackage{url}
@@
 
-\ConferenceShortName{ICS}
-\ConferenceName{International Conference on Supercomputing}
+%\ConferenceShortName{ICS}
+%\ConferenceName{International Conference on Supercomputing}
 
-\title{Library-based fault tolerance for scientific applications}
-\author{Sui Chen
+\title{Comprehensive Algorithmic Resilience for Scientific Applications}
+\author{Sui Chen, 
 %\affil{Louisiana State University}
@@
 
+\begin{abstract}
+As High-Performance Computing (HPC) systems become larger and more complex and the feature sizes of their electronic components grow smaller, the systems become increasingly vulnerable to soft faults.
+Since such faults may cause application crashes or may silently corrupt application results it is necessary to develop techniques to protect applications from such errors at a low cost in performance and power.
+Despite extensive work on general-purpose resilience techniques, they are cost effective for protecting only certain system components, such as memories.
+To provide comprehensive protection for the computations of HPC applications it is necessary to develop algorithmic resilience mechanisms that use the properties of algorithms to provide cost-effective protection.
+
+Algorithmic resilience can be very complex to deploy in applications because a comprehensive resilience strategy must address all the types of errors that an application may experience.
+This includes corruptions to numerical data as well as control information and data structures.
+Full protection requires the combination of multiple techniques including algorithmic error detection, crash detection, replication of key control information as well as support for restart when errors are detected.
+Further, it is necessary to configure the properties of the resilience mechanisms to make sure that they are cost-effectively meeting the user's accuracy and confidence targets.
+This paper shows how all these issues can be resolved comprehensively in the context of three numeric applications.
+
+\end{abstract}
+
 \section{Introduction}
@@
 
-The increasing size and complexity of (High-Performance Computing) HPC systems is making them increasingly vulnerable to soft faults.
+The increasing size and complexity of HPC systems is making them increasingly vulnerable to soft faults.
 Soft faults are transient corruptions of circuit state caused by physical phenomena such as strikes by neutrons or alpha particles~\cite{baumann:2005, asciQSER:2005} or thermal electrical noise~\cite{therm_noise:2007}.
@@
 
-The limitations of automated resilience techniques has motivated the need for algorithmic techniques that can detect and tolerate errors by taking advantage of application-specific properties and invariants~\cite{amg_abft:2012, robustification:2010, 1bft:1984}.
-Although these techniques can be very efficient, they also require manual effort to incorporate into applications and may require deep insight into an application's invariants to deploy effectively.
-Fortunately, because real-world applications spend most of their execution is spent in a few application modules, most soft faults will occur in those modules rather than uniformly throughout the application.
-In scientific applications these modules are usually various types of libraries, including numeric solvers and domain-specific scientific packages.
-This indicates that if by focusing algorithmic resilience efforts on the key library routines, while leaving other modules unprotected or protected via more expensive mechanisms, it is possible to fully protect an application at a low cost.
+The limitations of automated resilience techniques has motivated the need for algorithmic techniques that can detect and tolerate errors by taking advantage of application-specific properties and invariants~\cite{amg_abft:2012, robustification:2010, abft:1984}.
+Although these techniques can be very efficient most research work on such techniques has primarily focused on their algorithmic properties and how they can be leveraged to enable accurate error detection.
+While this goal is important, it ignores the full complexity of the problem since it focuses narrowly on errors in numeric data while ignoring errors in control information and data structures.
+In reality a comprehensive solution must incorporate a range of additional techniques such as key variable replication, memory violation detection and localized rollback on detected error, as well as techniques to configure these various resilience methods to achieve the best performance given the user's accuracy and confidence targets.
+%, they also require manual effort to incorporate into applications and may require deep insight into an application's invariants to deploy effectively.
+%Fortunately, because real-world applications spend most of their execution is spent in a few application modules, most soft faults will occur in those modules rather than uniformly throughout the application.
+%In scientific applications these modules are usually various types of libraries, including numeric solvers and domain-specific scientific packages.
+%This indicates that if by focusing algorithmic resilience efforts on the key library routines, while leaving other modules unprotected or protected via more expensive mechanisms, it is possible to fully protect an application at a low cost.
 
-This paper evaluates this approach in the context of three different types of applications: the MPI implementation of using the Alternating Direction Method of Multipliers (ADDR) to solver Lasso problems~\cite{lasso:2011}, the Hattrick gravitational simulation~\cite{hattrick:2012} and the Digital Room Correction (DRC) acoustic correction application~\cite{drc:2012}.
+This paper presents a comprehensive algorithmic resilience approach that considers all of the above issues.
+We demonstrate it on different types of numerical applications: the Alternating Direction Method of Multipliers (ADDR) for Lasso problems~\cite{lasso:2011}, the Hattrick gravitational simulation~\cite{hattrick:2012} and the Digital Room Correction (DRC) acoustic correction application~\cite{drc:2012}.
 We show how these three applications can be comprehensively protected from soft faults by adding three different resilience mechanisms to the routines where they spend the most time: algorithmic error checks, replication of critical data structures and checkpoint-restart of individual routines.
@@
 
-\greg{This section needs to include more detailed descriptions of the algorithms, the routines where they spend the bulk of their time and what fraction of time is spent there across the range of inputs.
-We also need to describe the types of inputs we'll be giving them in our experiments.}
-\subsection{Lasso}
+\greg{This section needs to include an experimental breakdown of where the routines spent their time. We also need to describe the types of inputs we'll be giving them in our experiments.}
+
+\subsection{Alternating Direction Method of Multipliers (ADDR)}
 \label{sec:apps:lasso}
-A parallel shrinkage and selection method for linear regression.
-It computes the equation $Ax=b$, where matrix $A$ and vector $b$ are known.
-\greg{is it a shrinkage and selection method or an actual solver?}.
-Utilizes GSL (the GNU Scientific Library)'s linear algebra routines.
+The ADDR method solves under-unconstrained linear problems $Ax=b$ for $x$ ($A$ has fewer rows than columns) while minimizing the function $f(x)$.
+This is an iterative algorithm that spends most of its time in the linear algebra operations, matrix-matrix multiplication, matrix-vector multiplication, rank-k update and Cholesky factorization, using the implementations of these routines from the GNU Scientific Library~\cite{gsl:2011}.
+We focused on a specific application that uses the ADDR method to solve Lasso problems, which are defined as minimizing the function $\frac{1}{2} \left|| Ax - b \right||_2^2 + \lambda*\left|| x \right||_1$.
+In our experiments we considered linear systems of size \textbf{???} that were generated \textbf{???}.
 
@@
 
-A program used to generate correction filters for acoustic compensation of HiFi and audio system in general.
-It generates FIR correction filters, which can be used with a real time or offline convolver to provide correction.
-Utilizes the GSL FFT routine.
+The DRC program generates filters for HiFi audio systems to compensate for the reflection of sounds in a room, using impulse response measurements of the audio equipment and considering the positions of the listeners.
+Although the algorithm is divided into many phases the most computationally intensive step consists of the GSL's implementation of the Fast Fourier Transform.
+\greg{What inputs?}
 
@@
 \label{sec:apps:hattrick}
-An accurate N-Body integrator meant for high control over small systems, specifically meant for perturbing bodies. Utilizes a differential equation solver routine.
+The Hattrick application simulates the motion of bodies under the effects of gravity.
+It is designed to help discover extra-solar planets by inferring their existence from Transit Timing Variations, where the difference between the times when a planet passes in front of its host star is used to infer the existence and properties of other planets in the system.
+This program spends most of its execution time in the GSL solver for Ordinary Differential Equations to solver the system's equations of motion.
 
@@
 
-\greg{This section needs performance results. How much slower are the resilient routines and how much more resilient do they become across a range of detection thresholds.}
-\sui{Performance experiments inserted here: running algorithmic checkers across different input sizes and fault rates: 62x62 to 500x500 matrixes for MM, RK, CD, MV checkers; FFT sizes of $2^8$ to $2^{24}$ for the FFT checker; Error bars should apply here, especially for the Cholesky Decomposition checker}
+%\sui{Performance experiments inserted here: running algorithmic checkers across different input sizes and fault rates: 62x62 to 500x500 matrixes for MM, RK, CD, MV checkers; FFT sizes of $2^8$ to $2^{24}$ for the FFT checker; Error bars should apply here, especially for the Cholesky Decomposition checker}
+
+Each algorithm used by the above application was enhanced with an algorithm-specific checker that validated its results by exploiting some algorithmic identity.
+Each checker computes some aggregate property of the algorithm's results using two different algorithm to produce vectors $v$ and $v'$ and their relative difference by more than some threshold $\tau$, the checker declares that an error has occurred.
+The relative difference is computed as $\frac{\left|| v-v' \right||}{max(\left||v\right||, \left||v'\right||)}$ ???.
 
 \paragraph{Matrix-matrix multiplication (MMM)}:
-Matrix-matrix multiplications are checked using a matrix vector multiplication, making use of this identity: $(A \cdot B) \cdot x = A \cdot (B \cdot x)$.
-Here x is a generated error-checking vector and $A$ and $B$ are inputs to the matrix-matrix multiplication.
-Time complexities of the original routine and the checker are $O(n^3)$ and $O(n^2)$ respectively.
+The MMM operation $A \cdot B$ is checked using a matrix vector multiplication (MVM), making use of this identity: $(A \cdot B) \cdot x = A \cdot (B \cdot x)$, where $x$ is an error-checking vector.
+This check is efficient because the checker is asymptotically faster than the MMM algorithm, with MVM taking $O(n^2)$ time and MMM $O(n^3)$ time.
+Since we make no assumptions about the distribution of data in the matrixes being multiplied, we set the check vector $x$ to contain all 1's.
 
 \paragraph{Matrix-vector multiplication (MVM)}:
-For matrix-vector multiplication $Ax=b$ , note that the sum of the elements in the result vector $b$ is equal to the dot product of the vector composed of column sums of $A$ and $x$.
-Time complexity of original routine is $O(n^2)$ multiplications and that of the checker is $O(n^2)$ additions.
+The MVM algorithm $Ax=b$ is checked using a similar identity $(x^TA)x = x^T(Ax) = x^Tb$.
+The product $x^TA$ is a checksum of the columns of matrix $A$ and when $x$ is the vector of 1's, the value used in our experiments, it is the sum of $A$'s columns.
+Although the computation of $x^TA$ involves additions rather than multiplications the complexity of computing $x^TA$ is $O(n^2)$, the same as the original MVM operation, making this check significantly less efficient.
+While this cost can be reduced significantly by computing $x^TA$ once and reusing it in subsequent multiplications by the same matrix, we did not consider it in this paper because MVM took up little application execution time.
+
+\greg{If MVM doesn't matter then why did we protect it? Also, why is the overhead >100\%? Where is the time spent?}
 
 \paragraph{Symmetric Rank-K update (SYRK)}:
-Symmetric Rank-K update is a special case of matrix-matrix multiplication where only the upper/lower half of the output matrix is updated.
-By decomposing the upper/lower half of the output matrix as a series of sub-matrices (like the sub-cells in a quad tree), the checker takes $O(n^2 \cdot \log{n})$ time, while the original routine woudl take $O(n^3)$.
+The SYRK update is a special case of matrix-matrix multiplication $C = \alpha A \cdot A^T + \beta*C$, where an $nxk$ matrix $A$ is multiplied by itself to produce an $nxn$ matrix with rank $k$, which is added to the $nxn$ matrix $C$.
+By decomposing the upper/lower half of the output matrix as a series of sub-matrices (like the sub-cells in a quad tree), the checker takes $O(n^2 \cdot \log{n})$ time, while the original routine takes $O(n^3)$ time.
 
-\paragraph{Cholesky Decomposition}: \greg{Need to define this decomposition}
-By multiplying back the upper and lower halves, the checker takes as much time as a regular matrix-matrix multiplication would do.
-The original routine is an iterative one, but it generally runs longer than the checker.
-\greg{How much longer?
-Need to quantify.} \sui{Can we quantify it using the performance data?}
+\paragraph{Cholesky Decomposition (CD)}:
+In this decomposition matrix $A$ is decomposed into $L \cdot L^T$ where $L$ is a lower-triangular matrix with a positive diagonal.
+This routine is checked just like MMM by checking the identity $Ac = L \cdot (L^T x)$ where $x$ is the vector of all 1's.
+This check runs in time $O(n^2)$ while the deterministic CD algorithm has $O(n^3)$ complexity.
+The GSL implements an iterative algorithm that runs faster than $O(n^3)$ time but as shown in the experiments below it is significantly slower than our checker.
 
 \paragraph{Fast Fourier Transform (FFT)}:
-By using Parseval's Theorem, we can check the result of an FFT of width $n$ in $O(n)$ time. Time complexity of the original routine is $O(\log{n})$.
-\greg{What is this theorem, how is it used?}
+The FFT algorithm decomposes a function into a sum of sine waves of different frequencies: $f(x) = \sum_{n=0}^{N-1} x_n e^{-i2\pi k n / N}$ for some constant $k$.
+We check the results of FFT by using Parseval's theorem, which states that $\sum_{n=0}^{N-1} \left| x[n] \right|^2 = \frac{1}{N} \sum_{k=0}^{N-1} \left| X[k] \right|^2$, where $x$ is the original function and $X$ is its transform.
+This means that the total energy of the original function is preserved by the transform and provides an $O(n)$ time check for the $O(n log(n))$ FFT algorithm.
 
 \paragraph{Runge-Kutta PDE Solver (RK)}:
-\greg{Describe the multiple step size checker in Hattrick}.
+This algorithm uses the ???-th order Runge-Kutta method for solving Ordinary Differential Equations of the form $\frac{dy}{dx} = f(y, x)$.
+It works by advancing the variable $x$ by steps of size $h$ and using estimates of the derivative $\frac{dy}{dx}$ at each point $x$ to find the value of $y$ at the next point $x+h$.
+The RK method in GSL uses adaptive step-size control where it simultaneously performs the RK method with two step sizes $h$ and $\frac{h}{2}$ (more precise).
+If the difference between the results obtained with the two step sizes are larger than a threshold, the algorithm switches to the smaller step size to maintain accuracy and resumes.
+Since the adaptive step control algorithm can also detect errors we utilize it to this end in our work.
 
-We evaluate these algorithmic error checkers by measuring their costs and effectiveness when running at different error rates on different types of inputs.
+\begin{figure}[ht!]
+%\vspace{-20pt}
+\centering
+\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp1_linalg}
+\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp1_fft}
+%\vspace{-5pt}
+\caption{Overhead of the algorithmic resilience techniques of key application routines}
+\label{fig:algo_ovhd}
+\end{figure}
 
-Experiment 1: Measure performance degradation of using error detectors as input sizes increase.
+%Experiment 1: Measure performance degradation of using error detectors as input sizes increase.
 In case of RK if we can run a version that doesn't do the multiple step sizes, do so but if this is difficult then don't worry about it.
+We evaluate the costs of the above algorithmic error checkers by measuring their costs and effectiveness when running at different error rates on different types of inputs.
+Figure~\ref{fig:algo_ovhd} measures the overheads of employing the checkers in each algorithms when no errors are injected.
+The linear algebra routines were executed on $nxn$ square matrixes where $n=(62, 250, 500, 750 and 1000)$.
+FFT ran on inputs upto ??? and RK ran on ???.
+Our experiments show that ???.
 
-Experiment 2: Measure the error in the outputs of the algorithms with and without the algorithmic detectors.
-First, show a probability distribution histogram for all the codes at one error rate, input size and algorithmic detection threshold.
-Then show a summary of the entire space using the same heatmaps that we're using for final application results.
-Put the input size and error rate on the y-axis and the detection threshold on the x-axis.
-The color in each box should be the fraction of entries in the output that have an error of a given magnitude.
-Show a separate plot for a few different magnitudes.
+%Experiment 2: Measure the error in the outputs of the algorithms with and without the algorithmic detectors.
+%First, show a probability distribution histogram for all the codes at one error rate, input size and algorithmic detection threshold.
+%Then show a summary of the entire space using the same heatmaps that we're using for final application results.
+%Put the input size and error rate on the y-axis and the detection threshold on the x-axis.
+%The color in each box should be the fraction of entries in the output that have an error of a given magnitude.
+%Show a separate plot for a few different magnitudes.
 
-Experiment 3: Measure the fraction fraction of routine runs that are interrupted by the detection of an error.
-Display this information via the same heatmaps as above.
+We now evaluate the effectiveness of our algorithmic detectors when errors are injected.
+Figure~\ref{fig:algo_err_dist} shows the fraction of entries in the outputs of each routine (y-axis) that have an error of a given magnitude (x-axis) as error magnitudes span from ???
+to ???
+and errors are injected at rate ???.
+In this experiment the liner algebra operations were run on matrixes of size ???, FFT on ??? and RK on ???.
+We show results for both the routines with no error checkers as well as routines where error checkers are used with detection thresholds ???.
+The data shows ???.
 
-For each experiment we need to include a discussion of what it means for how these checkers behave within the full-application runs.
+Figure~\ref{fig:algo_err_heatmap} shows the same information across the entire experimental space.
+The y-axis includes all algorithms, input sizes and error rates.
+The x-axis includes several segments for different degrees of error checking: none, as well as checking with thresholds ???. In each segment the x-axis spans over error magnitudes from ??? to ???.
+The box associated with each combination of the four parameters the figure indicates the fraction of outputs that have an error of this magnitude by the shade of gray it is colored with, as indicated in the legend.
+The data shows ???
+
+%Experiment 3: Measure the fraction fraction of routine runs that are interrupted by the detection of an error.
+%Display this information via the same heatmaps as above.
+
+Finally, Figure~\ref{fig:algor_term_heatmap} shows the fraction of algorithm runs during which the error checks detect errors, across all the algorithms, input sizes, error rates and detection types.
+This information is shown in the same format as Figure~\ref{fig:algo_err_heatmap} and is important for informing us about the performance impact of error detection at these error rates since each error detection results in a rollback of application execution.
+The experiments show ???.
+
+\greg{For each experiment we need to include a discussion of what it means for how these checkers behave within the full-application runs.}
 
@@
 
-Modern operating systems and hardware efficiently detect accesses to memory regions that are not allocated to the application and make it easy for the application to detect such errors by catching Segmentation Fault and Bus Error signals.
-We have used this mechanism to protect applications from a subset of memory errors.
+Detecting errors in application data structures requires checkers dedicated for this purpose.
+Depending on the algorithm's properties different error detectors are appropriate.
+In our work we considered the use of access protection hardware and replication of key data.
 
-Experiment 1: Show the fraction of runs of each routine that is aborted as a result of a memory error
+Since modern systems use 64-bit addressing the range of values that can be assumed by a given pointer is much larger than the amount of memory that can ever be assigned to an application.
+This means that there is a very high probability that the corruption of a given pointer will cause it to point to an unallocated address.
+Once this address is accessed this violation will be detected by the memory protection hardware and communicated to the application via a Segmentation Fault or Bus Error signal.
+This mechanism is very effective and cheap in practice but suffers from a long delay between an error and its detection.
+
+An alternative mechanism is to replicate key application variable and to compare the values of the replicas on each access to the variable.
+While more expensive this approach is effective if the number of key variables is small and the cost of rolling back computation is high.
+The next section discusses rollback and how its costs influence the choice of the memory fault detection strategy for each application.
+
+%Experiment 1: Show the fraction of runs of each routine that is aborted as a result of a memory error
+Figure~\ref{fig:mem_aborts} shows the fraction of runs of each routine that is aborted as a result of a memory error across all algorithm configurations, which indicates the performance impact of such errors on applications.
+The data shows that  ???.
 
@@
 
+\bibliographystyle{abbrv}
+\bibliography{bibs}
+
 \end{document}

### 2013-01-14 paper.tex
Added figures; explanation for using replication for rk4

@@
 %\affil{Lawrence Livermore National Laboratory}
+Lu Peng,
+%\affil{Louisiana State University}
 }
@@
 \paragraph{Runge-Kutta PDE Solver (RK)}:
-This algorithm uses the ???-th order Runge-Kutta method for solving Ordinary Differential Equations of the form $\frac{dy}{dx} = f(y, x)$.
+This algorithm uses the \sui{4-th order} Runge-Kutta method for solving Ordinary Differential Equations of the form $\frac{dy}{dx} = f(y, x)$.
 It works by advancing the variable $x$ by steps of size $h$ and using estimates of the derivative $\frac{dy}{dx}$ at each point $x$ to find the value of $y$ at the next point $x+h$.
@@
 
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp2_1_Example.png}
+\caption{Error magnitude histogram of one routine}
+\end{figure}
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp2_2_Heatmap_Error_ConfSpace.png}
+\caption{Error magnitude histogram of the entire configuration space}
+\end{figure}
+
 %Experiment 1: Measure performance degradation of using error detectors as input sizes increase.
@@
 We evaluate the costs of the above algorithmic error checkers by measuring their costs and effectiveness when running at different error rates on different types of inputs.
-Figure~\ref{fig:algo_ovhd} measures the overheads of employing the checkers in each algorithms when no errors are injected.
-The linear algebra routines were executed on $nxn$ square matrixes where $n=(62, 250, 500, 750 and 1000)$.
-FFT ran on inputs upto ??? and RK ran on ???.
-Our experiments show that ???.
+Figure \ref{fig:algo_ovhd} measures the overheads of employing the checkers in each algorithms when no errors are injected.
+The linear algebra routines were executed on $nxn$ square matrixes where $n=(62, 250, 500, 750\ and\ 1000)$.
+FFT ran on inputs upto 4M and RK ran on an integral with different intervals $[t_0, t_1]$ where $t_0=0$ and $t_1 \in \{10, 15, 20, 25, 30, 50, 75, 100\}$.
+Our experiments show ???\sui{the effectiveness of the error handlers and the error checkers. Note that the graph shows fault injection experiments at various rates (defined by errors per run).}
 
@@
 
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=1.00\columnwidth]{figs/4_1_2_Exp2_Overhead_Inputsize.png}
+\includegraphics[width=1.00\columnwidth]{figs/4_1_2_Exp2_Effectiveness.png}
+\includegraphics[width=1.00\columnwidth]{figs/4_1_2_Exp2_Expected_Running_Time_Needed.png}
+\caption{Overhead and effectiveness of the pointer protection techniques for the ODE solver}
+\label{fig:rk4_eff_ovhd}
+\end{figure}
+
 Errors that do not cause the application to access an invalid memory location are detected by replicating key pointers and loop iteration variables.
@@
 Therefore we would only focus on the application's vulnerability to pointer and index corruptions.
-Replicating critical variables (pointers and indices) and leveraging Byzantine fault tolerance and correcting those pointers and indices during run-time would raise the applications' resistance to soft errors.
-\greg{Need to explain why we only do this for Hattrick.}
+Replicating critical variables (pointers and indices) and leveraging Byzantine fault tolerance and correcting those pointers and indices during run-time would raise the applications' resistance to soft errors \sui{and increase the chance this application would output correct results.}
+
+\sui{Unlike the numerical and FFT routines used in the other two applications, the ODE solver has an stepping function called on the evaluation(s) of each integration step (GSL performs two evaluations per timestep in order to estimate integration error and perform adaptive step size control), which is called much more frequently, on the magnitude of millions of times during the lifetime of the application.
+Contrarily, the linear algebra routines are only called less than a hundred times during the execution of Lasso with a 10x input, and the FFT routines are called only several hundred times during the execution of DRC with a 30x input.
+Both are moderate-sized inputs in our experiments.}
+
+\greg{Need to explain why we only do this for Hattrick.} \sui{The checkpointing scheme used here has overhead of having to set signal handlers and perform up-to-date checksum computation of critical data structures.
+The more often the routines are called, the more frequent the critical data structures need be kept up-to-date, therefore this overhead is larger (non-trivial compated to that for FFT and linear algebra) for a frequently called routine like the integrator.
+This overhead makes the fault tolerance technique less efficient as it was in linear algebra and FFT routines, bringing it to the same ballpark as the expensive pointer replciation technique.
+Although it is possible to reduce the frequency of updating the checkpoints, it would also reduce the effectiveness of this method sharply.}
+
+\sui{
+The other reason is, pointer replication performs a better job at maintaining the chance the routine generates correct results (See figure \ref{fig:rk4_eff_ovhd}).
+This would lower the expected cost the user would have to pay to get the correct result despite the overhead.
+The method of computing this expected running time is described in greater detail in section \ref{sec:eval:perf}'s Markov Model.
+From \ref{fig:rk4_eff_ovhd} one could see the ratio of perfect runs out of all runs (that are completed or incomplete) and the number of faults per run forms an almost linear relationship.
+The linear regression model is used to evaluate the position of the guide points needed for the following experiments.
+}
+
+\sui{Refer to Figure \ref{fig:rk4_eff_ovhd} one could see the number of expected running time of different configurations if one wants to get a correct output from the application in a faulty environments.
+As the number of faults injected per run increases, the expected running time of the original, non-fault-tolerant application increases at a much faster rate than the other fault-tolerant counterparts.
+In real life scientific applications that run much longer, it's likely that a fault could hit the application several times before it completes.
+In such situations, the replication-only resilience technology produces the least expected number of runs needed to get a correct answer.}
 
@@
 
+\sui{When a user wants to run an application under a faulty environment and obtain correct results, s/he would have to expect to run a part of this application multiple times in case the application aborts due to memory access errors or return useless incorrect results.
+The user might want to have the application return to a certain checkpoint, or restart the entire application should any unrecoverable occurs.
+The cost of rollback, and the chance of each type of errors occurs at a different rate.
+The Markov model comes into play here with its capability to capture all these parameters of experiments we have already run and predict the running time of the applications at configurations not experimented with.}
+
 Experiment 1: Measure the slowdown of each application at the optimal configuration

### 2013-01-14 paper.tex
Some graphs and some text about single-routine experiments

@@
 \includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp1_fft}
+\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Overall_Breakdown.png}
 %\vspace{-5pt}
@@
 \includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp2_1_Example.png}
-\caption{Error magnitude histogram of one routine}
+\caption{Error magnitude histogram showing the effectiveness of the error detector}
 \end{figure}
@@
 \includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp2_2_Heatmap_Error_ConfSpace.png}
+\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp2_2_Heatmap_Error_ConfSpace_RK4.png}
 \caption{Error magnitude histogram of the entire configuration space}
@@
 Our experiments show ???\sui{the effectiveness of the error handlers and the error checkers. Note that the graph shows fault injection experiments at various rates (defined by errors per run).}
+
+\sui{There are roughly three types of fault rates: 1) very high fault rates (more than tens of errors per run), under which the application would barely complete running; 2) medium fault rates (1 to 10 errors per run), where the rate that the application completes running is very sensitive to the fault rate (because for a non-resilient application, just one error is enough to cause a segmentation fault and abort the application) and 3) low error rates, smaller than 1 errors per run.
+The effectiveness of the fault detector is most visible at fault rates from under 1 faults per run to several errors per run, which is visible from the histograms that the vertical bars indicating a run with algorithmic detectors is generally brighter, therefore less erroneous; when the error rates goes beyond 10 errors per run, the effectiveness of the signal handlers become more noticable, rising the chance the application would complete.}
+
+\sui{At higher error rates, the signal handlers enable more runs to complete as well as allowing erroneous results to go into the output.
+More erroneous results are allowed into the final output by the signal handler than the erroneous intermediate corrected by the algorithmic checker.
+That is why from the histograms, some vertical bars indicating fault-tolerant runs at a very high error rate would look darker.}
 
@@
 The data shows ???.
+
+\sui{The RK4 integrator does not have an algorithmic checker that works in a similar way and can precisely detect algorithmic error upon the completion of an integration, and we're relying on the integrator's built-in adaptive step size control capabilities that is able to control errors.
+This method is still not able to completely prevent algorithmic errors from occurring in the final output, as is shown in the histogram.}
 
@@
 The other reason is, pointer replication performs a better job at maintaining the chance the routine generates correct results (See figure \ref{fig:rk4_eff_ovhd}).
-This would lower the expected cost the user would have to pay to get the correct result despite the overhead.
+As mentioned before, the RK4 integrator does not have an effective checker that takes its output and tells the user whether the output is correct.
+As the user can't know whether a re-run is necessary and the user should be concerned by the expected time by which s/he can expect a correct output.
+Therefore, a higher chance of producing correct results would lower the expected cost the user would have to pay to get the correct result despite the overhead.
 The method of computing this expected running time is described in greater detail in section \ref{sec:eval:perf}'s Markov Model.

### 2013-01-16 paper.tex
Some more text

@@
 %\affil{Louisiana State University}
-Greg Bronevetsky, Marc Casas-Guix
+Greg Bronevetsky, Marc Casas-Guix,
 %\affil{Lawrence Livermore National Laboratory}
-Lu Peng,
+Lu Peng
 %\affil{Louisiana State University}
@@
 We focused on a specific application that uses the ADDR method to solve Lasso problems, which are defined as minimizing the function $\frac{1}{2} \left|| Ax - b \right||_2^2 + \lambda*\left|| x \right||_1$.
-In our experiments we considered linear systems of size \textbf{???} that were generated \textbf{???}.
+In our experiments we considered linear systems of size $\{40, 80, 120, ..., 4000\} \times 500$ that were generated by sampling a normal distribution (According to our experiments, the runing time of this solver is not dependent on the kind of distribution the inputs are sampled from.
 
@@
 The DRC program generates filters for HiFi audio systems to compensate for the reflection of sounds in a room, using impulse response measurements of the audio equipment and considering the positions of the listeners.
+The inputs are stored in PCM format, which is an array of 32-bit floating point numbers representing the samples at each sampling time step.
 Although the algorithm is divided into many phases the most computationally intensive step consists of the GSL's implementation of the Fast Fourier Transform.
+This program accepts user-controllable inputs to decide the width and number of iterations of FFT transforms it performs in these various phases.
+Different configurations would load the CPU for different lengths of time, virtually equivalent to inputs of different sizes.
 \greg{What inputs?}
@@
 Although the computation of $x^TA$ involves additions rather than multiplications the complexity of computing $x^TA$ is $O(n^2)$, the same as the original MVM operation, making this check significantly less efficient.
-While this cost can be reduced significantly by computing $x^TA$ once and reusing it in subsequent multiplications by the same matrix, we did not consider it in this paper because MVM took up little application execution time.
+%While this cost can be reduced significantly by computing $x^TA$ once and reusing it in subsequent multiplications by the same matrix, we did not consider it in this paper because MVM took up little application execution time.
+\sui{However, for applications where matrix re-use occurs frequently, for example, a transformation matrix is used to transform many vectors, re-using the column sums would reduce the overhead of this check substantially.}
 
@@
 We check the results of FFT by using Parseval's theorem, which states that $\sum_{n=0}^{N-1} \left| x[n] \right|^2 = \frac{1}{N} \sum_{k=0}^{N-1} \left| X[k] \right|^2$, where $x$ is the original function and $X$ is its transform.
+GSL implements optimized radix 2, 3, 4, 5, 6 and 7 subtransform routines as well as a general radix-N FFT routine.
+The subtransforms are optimized and consume $O(n log(n))$ time and the general one consumes $O(n^2)$ time.
+Therefore, GSL has a scheduling algorithm that factorizes the length of the input FFT, running the subtransforms whenever possible, using the fallback general radix-N routine only for the remaining factors.
+The theorem applies to FFT of any radix.
 This means that the total energy of the original function is preserved by the transform and provides an $O(n)$ time check for the $O(n log(n))$ FFT algorithm.
 
+\paragraph{FIR Filter Generation (FIR)}:
+The FIR filter generation algorithm used in DRC generates a series of samples over the sinc function $f(x)=\frac{sin(x)}{x} (f(x)=1\ when\ x\ is\ 0)$ and modulates it with a Blackman window.
+The sinc function has a nice property that $\int_{-\infty}^{\infty} sinc(x)dx = 1$. The Blackman window - modulated output of this routine
+preserves this same property for most of the time, and therefore we have the checker compute the sum of the output array and compare that with 1.
+Computing the sum requires merely $O(n)$ additions while this non-iterative FIR generation algorithm takes $O(n)$ trigonometric function evaluations, which much more complex than addition.
+Results have shown
+the overhead of this checker is very small (less than 5\%).
+Since this algorithm has only output and no input and therefore it does not need to have to be reinforced with input backup as other routines do.
+
 \paragraph{Runge-Kutta PDE Solver (RK)}:
-This algorithm uses the \sui{4-th order} Runge-Kutta method for solving Ordinary Differential Equations of the form $\frac{dy}{dx} = f(y, x)$.
+This algorithm uses the 4-th order Runge-Kutta method for solving Ordinary Differential Equations of the form $\frac{dy}{dx} = f(y, x)$.
 It works by advancing the variable $x$ by steps of size $h$ and using estimates of the derivative $\frac{dy}{dx}$ at each point $x$ to find the value of $y$ at the next point $x+h$.
-The RK method in GSL uses adaptive step-size control where it simultaneously performs the RK method with two step sizes $h$ and $\frac{h}{2}$ (more precise).
+The RK method in GSL uses adaptive step-size control (user-selectable) where it simultaneously performs the RK method with two step sizes $h$ and $\frac{h}{2}$ (more precise).
 If the difference between the results obtained with the two step sizes are larger than a threshold, the algorithm switches to the smaller step size to maintain accuracy and resumes.
@@
 We now evaluate the effectiveness of our algorithmic detectors when errors are injected.
-Figure~\ref{fig:algo_err_dist} shows the fraction of entries in the outputs of each routine (y-axis) that have an error of a given magnitude (x-axis) as error magnitudes span from ???
-to ???
-and errors are injected at rate ???.
-In this experiment the liner algebra operations were run on matrixes of size ???, FFT on ??? and RK on ???.
+Figure~\ref{fig:algo_err_dist} shows the fraction of entries in the outputs of each routine (y-axis) that have an error of a given magnitude (x-axis) as error magnitudes span from $10^{-14}$ to $10^{-1}$ and errors are injected at three rates for each routine.
+With a total of 5 input sizes and three fault rates, we have a combination of 15 error rates, from 0.1 error per run to 100 errors per run for each routine.
+In this experiment the liner algebra operations were run on matrixes of size 100, 150, 200, 250, 300, 500, FFT on input array sizes 512K, 1M, 2M, 4M and 8M and RK on ???.
 We show results for both the routines with no error checkers as well as routines where error checkers are used with detection thresholds ???.
@@
 \centering
-\includegraphics[width=1.00\columnwidth]{figs/4_1_2_Exp2_Overhead_Inputsize.png}
 \includegraphics[width=1.00\columnwidth]{figs/4_1_2_Exp2_Effectiveness.png}

### 2013-01-17 paper.tex
More edits

@@
 The relative difference is computed as $\frac{\left|| v-v' \right||}{max(\left||v\right||, \left||v'\right||)}$ ???.
+Each error detection causes the routine's execution to be restarted using the rollback mechanism described in Section~\ref{sec:res_tech:cr}.
 
-\paragraph{Matrix-matrix multiplication (MMM)}:
+\paragraph{Matrix-matrix multiplication (MMM)}
 The MMM operation $A \cdot B$ is checked using a matrix vector multiplication (MVM), making use of this identity: $(A \cdot B) \cdot x = A \cdot (B \cdot x)$, where $x$ is an error-checking vector.
@@
 
-\paragraph{Matrix-vector multiplication (MVM)}:
+\paragraph{Matrix-vector multiplication (MVM)}
 The MVM algorithm $Ax=b$ is checked using a similar identity $(x^TA)x = x^T(Ax) = x^Tb$.
@@
 
-\paragraph{Symmetric Rank-K update (SYRK)}:
-The SYRK update is a special case of matrix-matrix multiplication $C = \alpha A \cdot A^T + \beta*C$, where an $nxk$ matrix $A$ is multiplied by itself to produce an $nxn$ matrix with rank $k$, which is added to the $nxn$ matrix $C$.
+\paragraph{Symmetric Rank-K update (SYRK)}
+The SYRK update is a special case of matrix-matrix multiplication $C = \alpha A \cdot A^T + \beta*C$, where an $n \times k$ matrix $A$ is multiplied by itself to produce an $n \times n$ matrix with rank $k$, which is added to the $nxn$ matrix $C$.
 By decomposing the upper/lower half of the output matrix as a series of sub-matrices (like the sub-cells in a quad tree), the checker takes $O(n^2 \cdot \log{n})$ time, while the original routine takes $O(n^3)$ time.
 
-\paragraph{Cholesky Decomposition (CD)}:
+\paragraph{Cholesky Decomposition (CD)}
 In this decomposition matrix $A$ is decomposed into $L \cdot L^T$ where $L$ is a lower-triangular matrix with a positive diagonal.
@@
 
-\paragraph{Fast Fourier Transform (FFT)}:
+\paragraph{Fast Fourier Transform (FFT)}
 The FFT algorithm decomposes a function into a sum of sine waves of different frequencies: $f(x) = \sum_{n=0}^{N-1} x_n e^{-i2\pi k n / N}$ for some constant $k$.
@@
 
-\paragraph{Runge-Kutta PDE Solver (RK)}:
+\paragraph{Runge-Kutta PDE Solver (RK)}
 This algorithm uses the 4-th order Runge-Kutta method for solving Ordinary Differential Equations of the form $\frac{dy}{dx} = f(y, x)$.
 It works by advancing the variable $x$ by steps of size $h$ and using estimates of the derivative $\frac{dy}{dx}$ at each point $x$ to find the value of $y$ at the next point $x+h$.
-The RK method in GSL uses adaptive step-size control (user-selectable) where it simultaneously performs the RK method with two step sizes $h$ and $\frac{h}{2}$ (more precise).
-If the difference between the results obtained with the two step sizes are larger than a threshold, the algorithm switches to the smaller step size to maintain accuracy and resumes.
-Since the adaptive step control algorithm can also detect errors we utilize it to this end in our work.
-
-\begin{figure}[ht!]
-%\vspace{-20pt}
-\centering
-\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp1_linalg}
-\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp1_fft}
-\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Overall_Breakdown.png}
-%\vspace{-5pt}
-\caption{Overhead of the algorithmic resilience techniques of key application routines}
-\label{fig:algo_ovhd}
-\end{figure}
-
-\begin{figure}[ht!]
-\centering
-\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp2_1_Example.png}
-\caption{Error magnitude histogram showing the effectiveness of the error detector}
-\end{figure}
-
-\begin{figure}[ht!]
-\centering
-\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp2_2_Heatmap_Error_ConfSpace.png}
-\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp2_2_Heatmap_Error_ConfSpace_RK4.png}
-\caption{Error magnitude histogram of the entire configuration space}
-\end{figure}
-
-%Experiment 1: Measure performance degradation of using error detectors as input sizes increase.
-In case of RK if we can run a version that doesn't do the multiple step sizes, do so but if this is difficult then don't worry about it.
-We evaluate the costs of the above algorithmic error checkers by measuring their costs and effectiveness when running at different error rates on different types of inputs.
-Figure \ref{fig:algo_ovhd} measures the overheads of employing the checkers in each algorithms when no errors are injected.
-The linear algebra routines were executed on $nxn$ square matrixes where $n=(62, 250, 500, 750\ and\ 1000)$.
-FFT ran on inputs upto 4M and RK ran on an integral with different intervals $[t_0, t_1]$ where $t_0=0$ and $t_1 \in \{10, 15, 20, 25, 30, 50, 75, 100\}$.
-Our experiments show ???\sui{the effectiveness of the error handlers and the error checkers. Note that the graph shows fault injection experiments at various rates (defined by errors per run).}
-
-\sui{There are roughly three types of fault rates: 1) very high fault rates (more than tens of errors per run), under which the application would barely complete running; 2) medium fault rates (1 to 10 errors per run), where the rate that the application completes running is very sensitive to the fault rate (because for a non-resilient application, just one error is enough to cause a segmentation fault and abort the application) and 3) low error rates, smaller than 1 errors per run.
-The effectiveness of the fault detector is most visible at fault rates from under 1 faults per run to several errors per run, which is visible from the histograms that the vertical bars indicating a run with algorithmic detectors is generally brighter, therefore less erroneous; when the error rates goes beyond 10 errors per run, the effectiveness of the signal handlers become more noticable, rising the chance the application would complete.}
-
-\sui{At higher error rates, the signal handlers enable more runs to complete as well as allowing erroneous results to go into the output.
-More erroneous results are allowed into the final output by the signal handler than the erroneous intermediate corrected by the algorithmic checker.
-That is why from the histograms, some vertical bars indicating fault-tolerant runs at a very high error rate would look darker.}
+The RK method in GSL uses adaptive step-size control where it simultaneously performs the RK method with two step sizes $h$ and $\frac{h}{2}$ (more precise).
+If the difference between the results obtained with the two step sizes are larger than a high water-mark threshold, the algorithm switches to the smaller step size to maintain accuracy and resumes.
+If it is smaller than a low water-mark threshold, the algorithm switches to a larger step size.
+Because soft errors will usually cause significant differences between the two computations, the adaptive step size control algorithm is naturally fault tolerant.
+It will react to such errors by rolling back, temporarily increasing its step size and then reverting to the original step size when it observes that this is safe.
+As such, we utilize this algorithm to protect the computations within individual RK4 steps from soft errors.
+However, computations in user-provided function derivative functions are not checked since their invariants are unknown.
 
@@
 %Show a separate plot for a few different magnitudes.
+\subsubsection{Memory Fault Detection}
+\label{sec:res_tech:err_det:mem}
+
+Detecting errors in application data structures requires checkers dedicated for this purpose.
+Depending on the algorithm's properties different error detectors are appropriate.
+In our work we considered the use of access protection hardware and replication of key data.
+
+Since modern systems use 64-bit addressing the range of values that can be assumed by a given pointer is much larger than the amount of memory that can ever be assigned to an application.
+This means that there is a very high probability that the corruption of a given pointer will cause it to point to an unallocated address.
+Once this address is accessed this violation will be detected by the memory protection hardware and communicated to the application via a Segmentation Fault or Bus Error signal.
+This mechanism is cheap but suffers from a long delay between an error and its detection.
+
+An alternative mechanism is to replicate key application variable and to compare the values of the replicas on each access to the variable.
+While more expensive, this approach is effective if the number of key variables is small and the cost of rolling back computation is high.
+
+The low cost of hardware memory protection mechanisms means that they are deployed in protecting all our routines.
+In contrast, replication is only used to protect the RK routine because it is more sensitive to soft errors than the others and the low detection latency provided by replication ensures that any state corruptions are quickly corrected and have little impact on execution time.
+This is important for resilience as well as performance since slowdowns due to frequent rollbacks increase the total number of errors the application is exposed to.
+
+\subsection{Recovery via Checkpoint-Restart and Replication}
+\label{sec:res_tech:cr}
+
+When errors are detected we employ a recovery method based on checkpoint-restart, where routine state is recorded at key locations and whenever an algorithmic or memory error is detected we roll back to that point in its execution.
+Rollbacks are implemented via the \texttt{sigsetjmp}/\texttt{siglongjmp} routines.
+\texttt{sigsetjmp} record the application's execution state (registers and stack pointer) at the time of the checkpoint.
+On error detection \texttt{siglongjmp} returns the application to the same execution point, unrolling any function calls between the checkpoint and the error detection.
+This method is useful for rolling back to checkpoints recorded in the same function or in a caller function.
+When execution returns to the \texttt{sigsetjmp} call the checkpointed portions of the routine's state are recovered and execution repeats.
+% The overhead of this API call is having to copy the processor state into memory, therefore installing them between computation-intensive method calls adds to negligible performance degradation.
+
+Checkpoint data is protected using two different mechanisms.
+Pointers and scalars are replicated such that the checkpoint contains two copies of each.
+Arrays of floating point numbers are protected via a block-checksum-based error correcting code..
+By treating the input array as a matrix and comparing row sums and column sums of the submatrices of that matrix, we are able to detect the existence of errors and fix a fraction of the errors.
+The protection mechanism has the following parameters:
+\begin{itemize}
+\item{\texttt{N}, the size of the error correcting sum block.}
+\item{\texttt{ECC\_ECC}, whether or not the error correcting code itself should be corrected.}
+\end{itemize}
+This mechanism is able to correct up to $1/{n^2}$ of the elements in the input array.
+\greg{Which code is being used here? You just say that it is based on block-checksums}.
+
+The structure of each routine determines how checkpointing is deployed.
+All routines except RK are monolithic code blocks that take a relatively large amount of data as input, operate on it and then return.
+In contrast, RK takes in very little input and operates as a long sequence of steps, each of which involves a call to a user-provided derivative routine.
+As such, we checkpoint the state of non-RK routines immediately after the call to the routine and include in this checkpoint the routine's arguments and contents of the floating point input buffers \greg{is this correct?}.
+Because RK's checkpoints are very small and it is more sensitive to errors, it is checkpointed periodically during its execution with a fixed number of steps between checkpoints.
+As our experiments \greg{where?} show, the choice of the RK checkpoint period significantly affects its efficiency with longer periods suffering from more overhead due to rollbacks and short period spending more time checkpointing.
+
+Experiment 1: Measure the cost of setjmp and longjmp in our routines as the input sizes vary
+Experiment 2: Measure the cost of input backup and input recovery in our routines as the input sizes vary
+
+\subsection{Evaluation of Subroutine Resilience}
+\label{sec:res_tech:eval}
+
+\subsubsection{Linear Algebra and Integer Routines}
+\label{sec:res_tech:eval:la_int}
+
+\paragraph{Overheads}
+%\label{sec:res_tech:eval:ovhd}
+
+\begin{figure}[ht!]
+%\vspace{-20pt}
+\centering
+\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp1_linalg}
+\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp1_fft}
+%\vspace{-5pt}
+\caption{Overhead of resilience techniques when errors are not injected}
+\label{fig:routine_detect_ovhd}
+\end{figure}
+
+%Experiment 1: Measure performance degradation of using error detectors as input sizes increase.
+In case of RK if we can run a version that doesn't do the multiple step sizes, do so but if this is difficult then don't worry about it.
+We evaluate the costs of the above algorithmic error checkers by measuring their costs and effectiveness when running at different error rates on different types of inputs.
+Figure \ref{fig:routine_detect_ovhd} measures the overheads of employing the algorithmic error detectors when no errors are injected, reporting the average of \greg{???} runs.
+The linear algebra routines were executed on $nxn$ square matrixes where $n$ varies from 100 to 1000.
+FFT ran on inputs vectors with 128K to 4M entries and RK ran on an integral with intervals $[0, t_1]$ where $t_1 \in \{10, 15, 20, 25, 30, 50, 75, 100\}$.
+\greg{What function was being computed by RK?}
+As our experiments show, the cost of algorithmic error detection is low and drops with increasing input sizes.
+
+\begin{figure}[ht!]
+%\vspace{-20pt}
+\centering
+\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Overall_Breakdown.png}
+%\vspace{-5pt}
+\caption{Overhead of resilience techniques when errors are injected}
+\label{fig:routine_all_ovhd}
+\end{figure}
+
+Figure~\ref{fig:routine_all_ovhd} shows the overhead of all resilience techniques with no fault injection.
+The data shows that checkpointing consumes as much, if not more, time than error detection but that this cost grows more slowly than the routines' computational cost.
+In routines MV and RK the inputs are sufficiently small that checkpointing has negligible cost.
+Howveer, because MV runs quickly and because RK takes many checkpoints (\greg{what's the period in this experiment?}), the cost of saving processor state is small but not negligible.
+
+Finally, Figure~\ref{fig:routine_exp_exec} documents the overhead of resilience when errors are injected into application execution, accounting for the slowdowns due to the resilience mechanisms as well as rollbacks due to error detections.
+Each routine's graph shows $\frac{\texttt{Time with Errors}}{\texttt{Time No Errors}}$, where the \texttt{Time with Errors} is average increase execution time of a resilient or non-resilient version of the routine when running on the same range of inputs as above, and \texttt{Time No Errors} is the execution time of the non-resilient version on the same inputs when no errors are injected.
+The graph for each routine reports this metric for all of the above input sizes, error injection rates between 1e-8 and 1e-11 \greg{confirm!} and all resilience configurations (no resilience and resilience with detection thresholds between 1e-6 and 1e-8), as the average time of \greg{???} runs.
+The data points within each figure are organized horizontally according to the average number of errors injected into runs with the given input size and at the given injection rate.
+\greg{Do the error counts include errors after rollbacks?}
+This average error count is shown on the x-axis.
+All runs with the same resilience configuration are joined together into a line to highlight its properties.
+
+The data shows \greg{???}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\paragraph{Output Accuracy}
+%\label{sec:res_tech:eval:accu}
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp2_1_Example.png}
+\caption{Error magnitude histogram showing the effectiveness of the error detector}
+\label{fig:algo_err_dist}
+\end{figure}
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp2_2_Heatmap_Error_ConfSpace.png}
+\caption{Error magnitude histogram of the entire configuration space}
+\label{fig:algo_err_heatmap}
+\end{figure}
 
 We now evaluate the effectiveness of our algorithmic detectors when errors are injected.
-Figure~\ref{fig:algo_err_dist} shows the fraction of entries in the outputs of each routine (y-axis) that have an error of a given magnitude (x-axis) as error magnitudes span from $10^{-14}$ to $10^{-1}$ and errors are injected at three rates for each routine.
-With a total of 5 input sizes and three fault rates, we have a combination of 15 error rates, from 0.1 error per run to 100 errors per run for each routine.
-In this experiment the liner algebra operations were run on matrixes of size 100, 150, 200, 250, 300, 500, FFT on input array sizes 512K, 1M, 2M, 4M and 8M and RK on ???.
-We show results for both the routines with no error checkers as well as routines where error checkers are used with detection thresholds ???.
-The data shows ???.
+Figures~\ref{fig:algo_err_dist} and \ref{fig:algo_err_heatmap} show the fraction of entries in the outputs of each routine (y-axis) that have an error of a given magnitude (x-axis) as error magnitudes span from 1e-14 to 1e-4.
+Figure~\ref{fig:algo_err_dist} is a focused view on few representative cases (each routine run on a single input size and error injection rate) across different resilience configurations: no error detection as well as detection thresholds of 1e-6, 1e-7 and 1e-8.
+As the error magnitude increases the fraction of output entries with errors of that magnitude drops.
+Importantly, the fraction of erroneous entries is reduced at all magnitudes as a result of error detection and restart.
+Further, this reduction is more significant as the detection threshold shrinks since smaller magnitude errors are detected.
 
-\sui{The RK4 integrator does not have an algorithmic checker that works in a similar way and can precisely detect algorithmic error upon the completion of an integration, and we're relying on the integrator's built-in adaptive step size control capabilities that is able to control errors.
-This method is still not able to completely prevent algorithmic errors from occurring in the final output, as is shown in the histogram.}
+Figure~\ref{fig:algo_err_heatmap} shows the same information but for all experimental configurations: (i) input sizes of n=100 to 500 for linear algebra, 512K-8M size inputs for FFT and ???
+for RK; (ii) error injection rates of 1e-7 to 1e-10.
+The graph is a grid, with results for different routines listed vertically and different resilience configurations (no detection and different detection thresholds) listed horizontally.
+The graph for each routine and resilience configuration presents the curves from Figure~\ref{fig:algo_err_dist} as a heatmap.
+The y-axis shows different routine input sizes and error injection rates, sorted according to the average number of error injections per run, in increasing order from top to bottom.
+The x-axis shows the different error magnitudes and the shade of each point is the fraction of the routine's entries that have a given error magnitude, with darker shades indicating more such entries.
+Thus, each row of Figure~\ref{fig:algo_err_heatmap} represents the same information as each curve in Figure~\ref{fig:algo_err_dist}, with the shades of the squares representing the same information as the height of the line's points.
 
@@
 The box associated with each combination of the four parameters the figure indicates the fraction of outputs that have an error of this magnitude by the shade of gray it is colored with, as indicated in the legend.
-The data shows ???
+The data shows that the accurate of application results is directly related to the number of errors that impact a given application run, rather than their rate.
+Further, it shows that a reduction of detection threshold is consistently effective at reducing the fraction of errors in each algorithm's output.
+However, the effectiveness of algorithmic checks has a limit, with CD and FFT showing moderate error vulnerability even at threshold 1e-8.
+MM, MV and SYRK are significantly less vulnerable across the full range of error magnitudes.
+\greg{should reference vulnerability numbers more specifically}
 
@@
 %Display this information via the same heatmaps as above.
-
-Finally, Figure~\ref{fig:algor_term_heatmap} shows the fraction of algorithm runs during which the error checks detect errors, across all the algorithms, input sizes, error rates and detection types.
-This information is shown in the same format as Figure~\ref{fig:algo_err_heatmap} and is important for informing us about the performance impact of error detection at these error rates since each error detection results in a rollback of application execution.
-The experiments show ???.
 
@@
 
-\subsubsection{Memory Fault Detection}
-\label{sec:res_tech:err_det:mem}
+\subsubsection{Runge-Kutta}
+\label{sec:res_tech:eval:rk}
 
-Detecting errors in application data structures requires checkers dedicated for this purpose.
-Depending on the algorithm's properties different error detectors are appropriate.
-In our work we considered the use of access protection hardware and replication of key data.
+The difference between RK routine and others motivates special discussion of its resilience properties.
+Since RK's adaptive step size mechanism is already effective at tolerating many numerical errors we did not design any additional error detector for this routine.
+Instead, we focused on finding the optimal approach to memory error detection and an effective checkpointing strategy and considered different choices for these strategies as different resilience configurations.
+Configurations \texttt{Ckpt} and \texttt{NoChkpt} denote whether checkpointing was used and configurations \texttt{Re} and \texttt{NoRe} denote whether replication was used.
+\greg{What checkpointing period and replication properties?}
 
-Since modern systems use 64-bit addressing the range of values that can be assumed by a given pointer is much larger than the amount of memory that can ever be assigned to an application.
-This means that there is a very high probability that the corruption of a given pointer will cause it to point to an unallocated address.
-Once this address is accessed this violation will be detected by the memory protection hardware and communicated to the application via a Segmentation Fault or Bus Error signal.
-This mechanism is very effective and cheap in practice but suffers from a long delay between an error and its detection.
+\paragraph{Overheads}
+%\label{sec:res_tech:eval:ovhd}
 
-An alternative mechanism is to replicate key application variable and to compare the values of the replicas on each access to the variable.
-While more expensive this approach is effective if the number of key variables is small and the cost of rolling back computation is high.
-The next section discusses rollback and how its costs influence the choice of the memory fault detection strategy for each application.
+\begin{figure}[ht!]
+%\vspace{-20pt}
+\centering
+%\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp1_rk}
+%\vspace{-5pt}
+\caption{Overhead of resilience techniques for RK when errors are not injected}
+\label{fig:rk_routine_detect_ovhd}
+\end{figure}
 
-%Experiment 1: Show the fraction of runs of each routine that is aborted as a result of a memory error
-Figure~\ref{fig:mem_aborts} shows the fraction of runs of each routine that is aborted as a result of a memory error across all algorithm configurations, which indicates the performance impact of such errors on applications.
-The data shows that  ???.
+Figure~\ref{fig:rk_routine_detect_ovhd} shows the overhead of employing each of the four RK resilience techniques when no errors are injected.
+\greg{This should be the same format as Figure~\ref{fig:routine_all_ovhd}}
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=1.00\columnwidth]{figs/4_1_2_Exp2_Expected_Running_Time_Needed.png}
+\caption{Overhead of resilience techniques for RK when errors are injected}
+\label{fig:rk_routine_exp_exec}
+\end{figure}
+
+Figure~\ref{fig:rk_routine_exp_exec} shows the RK's overhead when errors are injected, using the same format as Figure~\ref{fig:routine_exp_exec}.
+Although the \texttt{NoRe\_NoCkpt} configuration has the lowest overhead when only a few errors are injected (low error rate or small execution times), as the number of errors per run increase this configuration quickly becomes extremely slow because it crashes on almost every run.
+In contrast, while the other configurations have a higher overhead for a few errors per run, their execution time scales much better with increasing error rates, making them significantly more practical.
+By varying the checkpointing interval and the degree to which various RK variables are replicated it is possible to achieve a wide range of design points that provide the best performance at various error rates.
+
+\paragraph{Output Accuracy}
+%\label{sec:res_tech:eval:accu}
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp2_2_Heatmap_Error_ConfSpace_RK4.png}
+\caption{Error magnitude histogram of the entire configuration space of RK}
+\label{fig:rt_algo_err_heatmap}
+\end{figure}
 
@@
 \includegraphics[width=1.00\columnwidth]{figs/4_1_2_Exp2_Effectiveness.png}
-\includegraphics[width=1.00\columnwidth]{figs/4_1_2_Exp2_Expected_Running_Time_Needed.png}
-\caption{Overhead and effectiveness of the pointer protection techniques for the ODE solver}
-\label{fig:rk4_eff_ovhd}
+\caption{Fraction of RK runs that achieve low error (smaller than 1e-7).}
+\label{fig:rk_effectiveness}
 \end{figure}
 
-Errors that do not cause the application to access an invalid memory location are detected by replicating key pointers and loop iteration variables.
-The Hattrick application uses a high-accuracy ODE solver that is capable of estimating errors and detecting and reducing errors to a certain level adaptively.
-Therefore we would only focus on the application's vulnerability to pointer and index corruptions.
-Replicating critical variables (pointers and indices) and leveraging Byzantine fault tolerance and correcting those pointers and indices during run-time would raise the applications' resistance to soft errors \sui{and increase the chance this application would output correct results.}
+Figure~\ref{fig:rt_algo_err_heatmap} shows the fraction of errors in RK's outputs for the four resilience configurations, in the same format as Figure~\ref{fig:algo_err_heatmap}.
+It shows that the use of either resilience approaches significantly reduces RK's vulnerability to soft errors and the use of both techniques is even more effective.
+Figure~\ref{fig:rk_effectiveness} shows this from a different perspective: the fraction of RK runs where the root-mean-squared error of the output vector is below 1e-7 (y-axis), relative to the number of error injections per run (x-axis).
+It shows that RK configurations with higher level of resilience are able to achieve this accuracy target more frequently.
+This is primarily because by avoiding frequent rollbacks due to crashes these techniques reduce the total number of errors that are injected into the application run, which also reduces the probability that one such error will be missed by RK's algorithmic error detector.
+As with overhead, the choice of checkpointing interval and replication degree make it possible to choose a wide range accuracy levels and probabilities that these targets will be satisfied.
 
-\sui{Unlike the numerical and FFT routines used in the other two applications, the ODE solver has an stepping function called on the evaluation(s) of each integration step (GSL performs two evaluations per timestep in order to estimate integration error and perform adaptive step size control), which is called much more frequently, on the magnitude of millions of times during the lifetime of the application.
+%
+%???\sui{the effectiveness of the error handlers and the error checkers. Note that the graph shows fault injection experiments at various rates (defined by errors per run).}
+%
+%\sui{There are roughly three types of fault rates: 1) very high fault rates (more than tens of errors per run), under which the application would barely complete running; 2) medium fault rates (1 to 10 errors per run), where the rate that the application completes running is very sensitive to the fault rate (because for a non-resilient application, just one error is enough to cause a segmentation fault and abort the application) and 3) low error rates, smaller than 1 errors per run.
+The effectiveness of the fault detector is most visible at fault rates from under 1 faults per run to several errors per run, which is visible from the histograms that the vertical bars indicating a run with algorithmic detectors is generally brighter, therefore less erroneous; when the error rates goes beyond 10 errors per run, the effectiveness of the signal handlers become more noticeable, rising the chance the application would complete.}
+%
+%\sui{At higher error rates, the signal handlers enable more runs to complete as well as allowing erroneous results to go into the output.
+More erroneous results are allowed into the final output by the signal handler than the erroneous intermediate corrected by the algorithmic checker.
+That is why from the histograms, some vertical bars indicating fault-tolerant runs at a very high error rate would look darker.}
+%
+%
+%
+%
+%
+%
+%Errors that do not cause the application to access an invalid memory location are detected by replicating key pointers and loop iteration variables.
+%The Hattrick application uses a high-accuracy ODE solver that is capable of estimating errors and detecting and reducing errors to a certain level adaptively.
+%Therefore we would only focus on the application's vulnerability to pointer and index corruptions.
+%Replicating critical variables (pointers and indices) and leveraging Byzantine fault tolerance and correcting those pointers and indices during run-time would raise the applications' resistance to soft errors \sui{and increase the chance this application would output correct results.}
+%
+%\sui{Unlike the numerical and FFT routines used in the other two applications, the ODE solver has an stepping function called on the evaluation(s) of each integration step (GSL performs two evaluations per timestep in order to estimate integration error and perform adaptive step size control), which is called much more frequently, on the magnitude of millions of times during the lifetime of the application.
 Contrarily, the linear algebra routines are only called less than a hundred times during the execution of Lasso with a 10x input, and the FFT routines are called only several hundred times during the execution of DRC with a 30x input.
 Both are moderate-sized inputs in our experiments.}
-
-\greg{Need to explain why we only do this for Hattrick.} \sui{The checkpointing scheme used here has overhead of having to set signal handlers and perform up-to-date checksum computation of critical data structures.
+%
+%\greg{Need to explain why we only do this for Hattrick.} \sui{The checkpointing scheme used here has overhead of having to set signal handlers and perform up-to-date checksum computation of critical data structures.
 The more often the routines are called, the more frequent the critical data structures need be kept up-to-date, therefore this overhead is larger (non-trivial compated to that for FFT and linear algebra) for a frequently called routine like the integrator.
@@
 Although it is possible to reduce the frequency of updating the checkpoints, it would also reduce the effectiveness of this method sharply.}
-
-\sui{
-The other reason is, pointer replication performs a better job at maintaining the chance the routine generates correct results (See figure \ref{fig:rk4_eff_ovhd}).
+%
+%\sui{
+%The other reason is, pointer replication performs a better job at maintaining the chance the routine generates correct results (See figure \ref{fig:rk4_eff_ovhd}).
 As mentioned before, the RK4 integrator does not have an effective checker that takes its output and tells the user whether the output is correct.
@@
 The linear regression model is used to evaluate the position of the guide points needed for the following experiments.
-}
-
-\sui{Refer to Figure \ref{fig:rk4_eff_ovhd} one could see the number of expected running time of different configurations if one wants to get a correct output from the application in a faulty environments.
+%}
+%
+%\sui{Refer to Figure \ref{fig:rk4_eff_ovhd} one could see the number of expected running time of different configurations if one wants to get a correct output from the application in a faulty environments.
 As the number of faults injected per run increases, the expected running time of the original, non-fault-tolerant application increases at a much faster rate than the other fault-tolerant counterparts.
@@
 In such situations, the replication-only resilience technology produces the least expected number of runs needed to get a correct answer.}
-
-To determine a replication strategy we have to determine which variables to replicate and how many replicas to use for replicating them.
+%
+%To determine a replication strategy we have to determine which variables to replicate and how many replicas to use for replicating them.
 The following two subsections would be focused on these problems.
-\greg{The report spends a lot of space talking about this but in the end the conclusion is that we just a pick a few options and evaluate them experimentally.
+%\greg{The report spends a lot of space talking about this but in the end the conclusion is that we just a pick a few options and evaluate them experimentally.
 How much space should we devote to this discussion.}
-
-Experiment 2: Measure the effectiveness and performance costs of replication. Need to explain why we don't use it for Lasso and DRC.
-
-\subsection{Recovery via Checkpoint-Restart}
-\label{sec:res_tech:cr}
-
-When errors are detected we employ a recovery method based on checkpoint-restart.
-At the start of each routine the state of all inputs is recorded and an error correcting code is used to ensure that the contents of this backup copy are not corrupted.
-Further, we use the \texttt{sigsetjmp} routine to record the application's execution state at the routine's start.
-If an error is detected during the routine \texttt{siglongjmp} is used to return the application's execution back to the start of the routine (this function unrolls function calls as needed and reloads the process counter and register state).
-The recovery procedure then resets the routine's inputs from their backed-up copies, overwriting any changes or corruptions that may have occurred during the routine's execution, and restarts its execution.
-%A combination of \texttt{siglongjmp} and \texttt{sigsetjmp} are used to recover from segmentation faults.
-The overhead of this API call is having to copy the processor state into memory, therefore installing them between computation-intensive method calls adds to negligible performance degradation.
-
-The backup copy of routine inputs is protected via a block-checksum-based data correction mechanism.
-By treating the input array as a matrix and comparing row sums and column sums of the submatrices of that matrix, we are able to detect the existence of errors and fix a fraction of the errors.
-The protection mechanism has the following parameters:
-\begin{itemize}
-\item{\texttt{N}, the size of the error correcting sum block.}
-\item{\texttt{ECC\_ECC}, whether or not the error correcting code itself should be corrected.}
-\end{itemize}
-This mechanism is able to correct up to $1/{n^2}$ of the elements in the input array.
-\greg{Which code is being used here? You just say that it is based on block-checksums}.
-
-Experiment 1: Measure the cost of setjmp and longjmp in our routines as the input sizes vary
-Experiment 2: Measure the cost of input backup and input recovery in our routines as the input sizes vary
-
+%
+%Experiment 2: Measure the effectiveness and performance costs of replication.
+Need to explain why we don't use it for Lasso and DRC.
 \section{Evaluation of Resilience Methodology}

### 2013-01-18 paper.tex
Updates on 01-18

@@
 %\affil{Louisiana State University}
-Greg Bronevetsky, Marc Casas-Guix,
+Greg Bronevetsky, Marc Casas,
 %\affil{Lawrence Livermore National Laboratory}
@@
 They can affect the state of processor latches and registers and may cause the application to crash or worse, to silently return incorrect results~\cite{mpi_ser:reed:2004}.
-As the electronic feature sizes shrink technology scaling will make soft errors a larger problem~\cite{err_scaling:2012} due to the fact that each circuit element will hold less charge and can thus be disrupted by easily.
+As the feature sizes of electronic circuits shrink, technology scaling will make soft errors a larger problem~\cite{err_scaling:2012} due to the fact that each circuit element will hold less charge and can thus be disrupted more easily.
 These phenomena make it imperative to develop techniques to make HPC systems resilient to soft faults.
@@
 Resilience is a problem that must be addressed at all levels.
-While materials science and circuit design techniques can be used to improve resilience, they come at a cost in reduced power efficiency and performance that an become prohibitive if processors need to be sufficiently reliable to build a large HPC system.
+While materials science and circuit design techniques can be used to improve resilience, they come at a cost in reduced power efficiency and performance that can become prohibitive if processors need to be sufficiently reliable to build a large HPC system.
 Techniques such as error correcting codes have been very effective at making memories and caches resilient to soft faults~\cite{mem_errors:2010}.
@@
 Traditional approaches like cross-core or node replication~\cite{rmpi:2011, dyn_cmp_repl:2007} are expensive in their use of computing resources and power and their long error detection latencies require processor state to be frequently checkpointed.
-Processor designs that incorporate instruction replication~\cite{repl_smt:2000} offer fine-grained error detection and rollback but require more power as well as novel hardware features that are not likely to be included in the commodity processors must be used in HPC systems to limit their cost.
+Processor designs that incorporate instruction replication~\cite{repl_smt:2000} offer fine-grained error detection and rollback but require more power as well as novel hardware features that are not likely to be included in the commodity processors used in HPC systems for cost reduction reasons.
 
 The limitations of automated resilience techniques has motivated the need for algorithmic techniques that can detect and tolerate errors by taking advantage of application-specific properties and invariants~\cite{amg_abft:2012, robustification:2010, abft:1984}.
-Although these techniques can be very efficient most research work on such techniques has primarily focused on their algorithmic properties and how they can be leveraged to enable accurate error detection.
-While this goal is important, it ignores the full complexity of the problem since it focuses narrowly on errors in numeric data while ignoring errors in control information and data structures.
-In reality a comprehensive solution must incorporate a range of additional techniques such as key variable replication, memory violation detection and localized rollback on detected error, as well as techniques to configure these various resilience methods to achieve the best performance given the user's accuracy and confidence targets.
+Although these techniques can be very efficient, most research work on such techniques has primarily focused on how their algorithmic invariants can be leveraged to enable accurate error detection.
+While this goal is important, it ignores the full complexity of the problem since it focuses on only a subset of the errors and does not consider combinations of different resilience techniques to efficiently and effectively address all the different types of faults.
+In reality a comprehensive solution must incorporate a range of additional techniques such the replication of key variables, detection of memory violations and localized rollback on detected errors, as well as techniques to configure these various resilience methods to achieve the best performance given the user's accuracy and confidence targets.
 %, they also require manual effort to incorporate into applications and may require deep insight into an application's invariants to deploy effectively.
@@
 We demonstrate via fault injection experiments that although none of these techniques can individually protect applications from soft faults, their combination is highly effective.
-Further, we demonstrate how these techniques can be parameterized to offer the right level of protection at any given fault rate or input type, and further can be tuned to bound the probability that the application will produce erroneous results or executes for longer than some limit.
+Further, we demonstrate how these techniques can be configured to offer the right level of protection at any given fault rate or input type.
+More specifically, these configurations can be used to produce a wide range of overheads and result accuracy levels and to tune the probabilistic confidence level that the given overhead or accuracy target will be met.
+
+The paper is organized as follows.
+Section~\ref{sec:soft_faults} discusses the problem of soft errors and presents the fault injection model we use to simulate them in our experiments.
+The applications to which we apply our resilience methodology are described in Section~\ref{sec:apps}.
+Section~\ref{sec:res_tech} summarizes the major routines of these applications and describes how these resilience techniques are applied to them.
+Further, it experimentally evaluates the resilience techniques, describing their overheads and the accuracy levels of the protected routines at a wide range of possible error rates.
+The effectiveness of our methodology for protecting the full applications is evaluated in Section~\ref{sec:eval}.
 
@@
 
-Causes of soft faults and how they manifest themselves in applications.
+Cosmic rays or alpha particles hitting the computer circuits and accumulating a sufficient amount of energy may invert the device logic states from "0" to "1" or from "1" to "0".
+These inversions are termed as soft errors or soft faults.
+The shrinking processor and memory feature size and the lower threshold voltage coming with relentless manufacturing technology scaling make modern high performance computers are highly vulnerable to soft errors which propagate as the program is being executed.
+Prior studies have demonstrated that soft errors occurring in all the important bits that are required for Architecturally Correct Execution (ACE)~\cite{avf:2003} eventually change the program outcome, i.e., manifest errors to the application.
+The ACE bits include critical information to keep the program logic such as the program counter register for a committed instruction.
+In the past decades, soft errors have brought significant losses on commercial servers and high performance computers.
 
-Discuss prior work on replication and algorithmic techniques, emphasize that in this paper we're focusing on software resilience since hardware resilience is extremely difficult to deploy in real processors since the costs have to be carried by non-HPC markets that care much less about it.
+%Discuss prior work on replication and algorithmic techniques, emphasize that in this paper we're focusing on software resilience since hardware resilience is extremely difficult to deploy in real processors since the costs have to be carried by non-HPC markets that care much less about it.
 
-Can show plots of how routine results are affected by injected errors. Should show this for all the routines we'll be considering in this paper.
-
-Describe fault injection framework and how it approximates real faults.
+We use a fault injection infrastructure~\cite{relax:2010} that simulates the architecture-level effects of soft faults.
+It works by compiling application into the Low Level Virtual Machine (LLVM) typed bytecode~\cite{LLVM}.
+This bytecode is transformed to allow a bit flip in each instruction's output with some probability $P$.
+The time between injections is thus exponentially distributed with period $\frac{1}{P}$.
+Our fault injector is a high-level approximation of the physical corruptions caused by soft errors that strikes a balance between model accuracy and cost, as do related approaches~\cite{fault_injection:iyer:1997, avf:2003}.
+Alternative models, such as injection into gate-level processor models~\cite{sesee:2004} or neutron beam irradiation~\cite{freq_dep:irom:2004} are possible but their significantly higher cost makes them impractical for conducting large-scale fault injection campaigns such as the ones reported in this paper.
 
@@
 
-We evaluate our library-based resilience approach by applying it to the following three applications.
-As they make intensive use library routines, they are ideal testbeds for our proposed fault-tolerance methods.
+We evaluate our resilience methodology by applying it to the following three applications.
+Since they spend most of their execution time in library routines, we focus our resilience efforts on protecting these routines, which are described in Section~\ref{sec:res_tech:err_det:algo}.
 
-\greg{This section needs to include an experimental breakdown of where the routines spent their time. We also need to describe the types of inputs we'll be giving them in our experiments.}
+%\greg{This section needs to include an experimental breakdown of where the routines spent their time. We also need to describe the types of inputs we'll be giving them in our experiments.}
 
@@
 
+In this paper we consider a comprehensive resilience strategy that combines error detection and recovery.
+Error detection techniques are discussed in Section~\ref{sec:res_tech:err_det}, with Section~\ref{sec:res_tech:err_det:algo} focusing on algorithmic techniques to detect errors in the numeric portions of application state, while Section~\ref{sec:res_tech:err_det:mem} focuses on protection of data structures.
+Section~\ref{sec:res_tech:cr} then considers recovery techniques based on both checkpoint-restart, which offers cheaper but higher-latency recovery, as well as replication of key variables, which is more expensive but enables fast recovery.
+Finally, Section~\ref{sec:res_tech:eval} presents an experimental evaluation of the performance and accuracy properties of the resilient versions of the major routines used by our target applications, under a wide range of inputs and error injection rates.
+
 \subsection{Error Detection}
@@
 The product $x^TA$ is a checksum of the columns of matrix $A$ and when $x$ is the vector of 1's, the value used in our experiments, it is the sum of $A$'s columns.
-Although the computation of $x^TA$ involves additions rather than multiplications the complexity of computing $x^TA$ is $O(n^2)$, the same as the original MVM operation, making this check significantly less efficient.
+Although the computation of $x^TA$ involves additions rather than multiplications, the complexity of computing $x^TA$ is $O(n^2)$, the same as the original MVM operation.
 %While this cost can be reduced significantly by computing $x^TA$ once and reusing it in subsequent multiplications by the same matrix, we did not consider it in this paper because MVM took up little application execution time.
-\sui{However, for applications where matrix re-use occurs frequently, for example, a transformation matrix is used to transform many vectors, re-using the column sums would reduce the overhead of this check substantially.}
-
-\greg{If MVM doesn't matter then why did we protect it? Also, why is the overhead >100\%? Where is the time spent?}
+However, for applications such as ADDR where the MVM is applied to same matrix many times with different vectors, the column sums can be reused, amortizing the cost of this computation.
+Our experimental evaluation below thus focuses on the overhead of the $x^T(Ax)$ computation, which has $O(n)$ complexity.
 
 \paragraph{Symmetric Rank-K update (SYRK)}
-The SYRK update is a special case of matrix-matrix multiplication $C = \alpha A \cdot A^T + \beta*C$, where an $n \times k$ matrix $A$ is multiplied by itself to produce an $n \times n$ matrix with rank $k$, which is added to the $nxn$ matrix $C$.
-By decomposing the upper/lower half of the output matrix as a series of sub-matrices (like the sub-cells in a quad tree), the checker takes $O(n^2 \cdot \log{n})$ time, while the original routine takes $O(n^3)$ time.
+The SYRK update is a special case of matrix-matrix multiplication $B = \alpha A \cdot A^T + \beta B$, where an $n \times k$ matrix $A$ is multiplied by itself to produce an $n \times n$ matrix with rank $k$, which is added to the $nxn$ matrix $C$.
+Its results are checked using the same algorithm as MMM by checking the identity $B x = \alpha A \cdot (A^Tx) + \beta B x$, where $x$ is the vector of all 1's.
+This check runs in time $O(n^2)$, as compared to the $O(n^3)$ cost of SYRK.
 
@@
 In this decomposition matrix $A$ is decomposed into $L \cdot L^T$ where $L$ is a lower-triangular matrix with a positive diagonal.
-This routine is checked just like MMM by checking the identity $Ac = L \cdot (L^T x)$ where $x$ is the vector of all 1's.
-This check runs in time $O(n^2)$ while the deterministic CD algorithm has $O(n^3)$ complexity.
-The GSL implements an iterative algorithm that runs faster than $O(n^3)$ time but as shown in the experiments below it is significantly slower than our checker.
+This routine is checked just like MMM by checking the identity $Ac = L \cdot (L^T x)$, which is cheaper than the $O(n^3)$ complexity of the deterministic CD algorithm.
+The GSL implements an iterative algorithm that runs faster than $O(n^3)$ time but as shown in the experiments below, it is significantly slower than our checker.
 
@@
 As our experiments \greg{where?} show, the choice of the RK checkpoint period significantly affects its efficiency with longer periods suffering from more overhead due to rollbacks and short period spending more time checkpointing.
-
-Experiment 1: Measure the cost of setjmp and longjmp in our routines as the input sizes vary
-Experiment 2: Measure the cost of input backup and input recovery in our routines as the input sizes vary
 
@@
 
+\marc{The y-axis should show just the number of errors injected. We can say that we have used several fault injection rates and sizes to get those numbers. Right now the picture is confusing.}
+
 Figure~\ref{fig:algo_err_heatmap} shows the same information across the entire experimental space.
@@
 Figure~\ref{fig:rk_routine_exp_exec} shows the RK's overhead when errors are injected, using the same format as Figure~\ref{fig:routine_exp_exec}.
-Although the \texttt{NoRe\_NoCkpt} configuration has the lowest overhead when only a few errors are injected (low error rate or small execution times), as the number of errors per run increase this configuration quickly becomes extremely slow because it crashes on almost every run.
+Although the \texttt{NoRe\_NoCkpt} configuration has the lowest overhead when only a few errors are injected (low error rate or small execution times), as the number of errors per run increase this configuration quickly becomes extremely slow because an error is detected in almost every run.
+Since each such detection requires a restart of the routine, at higher error rates this causes RK to restart many times before it completes.
 In contrast, while the other configurations have a higher overhead for a few errors per run, their execution time scales much better with increasing error rates, making them significantly more practical.
@@
 %
-%Experiment 2: Measure the effectiveness and performance costs of replication.
-Need to explain why we don't use it for Lasso and DRC.
+%Experiment 2: Measure the effectiveness and performance costs of replication. Need to explain why we don't use it for Lasso and DRC.
 \section{Evaluation of Resilience Methodology}

### 2013-01-19 paper.tex
Contains a list of problems to be tackled

@@
 \usepackage{color}
-\usepackage{multirow}
 \usepackage[ruled]{algorithm2e}
@@
 
-\title{Comprehensive Algorithmic Resilience for Scientific Applications}
+\title{Comprehensive Algorithmic Resilience for Scientific Applications }
 \author{Sui Chen,
@@
 More specifically, these configurations can be used to produce a wide range of overheads and result accuracy levels and to tune the probabilistic confidence level that the given overhead or accuracy target will be met.
+\greg{Summarize real results}
 
@@
 
-Cosmic rays or alpha particles hitting the computer circuits and accumulating a sufficient amount of energy may invert the device logic states from "0" to "1" or from "1" to "0".
+Cosmic rays or alpha particles hitting the computer circuits and accumulating a sufficient amount of energy may invert the device logic states from "0" to "1" or from "1" to "0"~\cite{ibm_softerr:1996,baumann:2005}.
 These inversions are termed as soft errors or soft faults.
@@
 The ACE bits include critical information to keep the program logic such as the program counter register for a committed instruction.
-In the past decades, soft errors have brought significant losses on commercial servers and high performance computers.
+In the past decades, soft errors have brought significant losses on commercial servers and high performance computers such as the recall of Sun servers due to memory errors~\cite{SoftErr:Baumann:2002} or the cache corruptions in the ASCI Q system~\cite{asciQSER:2005}.
 
@@
 The ADDR method solves under-unconstrained linear problems $Ax=b$ for $x$ ($A$ has fewer rows than columns) while minimizing the function $f(x)$.
-This is an iterative algorithm that spends most of its time in the linear algebra operations, matrix-matrix multiplication, matrix-vector multiplication, rank-k update and Cholesky factorization, using the implementations of these routines from the GNU Scientific Library~\cite{gsl:2011}.
-We focused on a specific application that uses the ADDR method to solve Lasso problems, which are defined as minimizing the function $\frac{1}{2} \left|| Ax - b \right||_2^2 + \lambda*\left|| x \right||_1$.
-In our experiments we considered linear systems of size $\{40, 80, 120, ..., 4000\} \times 500$ that were generated by sampling a normal distribution (According to our experiments, the runing time of this solver is not dependent on the kind of distribution the inputs are sampled from.
+This is an iterative algorithm uses 64-bit precision and that spends most of its time in the following linear algebra operations, matrix-matrix multiplication, matrix-vector multiplication, rank-k update and Cholesky factorization, using the implementations of these routines from the GNU Scientific Library~\cite{gsl:2011}.
+We focus on a specific application that uses the ADDR method to solve Lasso problems, which are defined as minimizing the function $\frac{1}{2} \left|| Ax - b \right||_2^2 + \lambda*\left|| x \right||_1$.
+In our experiments we consider linear systems of size $\{40, 80, 120, ..., 4000\} \times 500$ that are generated by sampling a normal distribution.
+According to our experiments, the running time of this solver is not dependent on the kind of distribution the inputs are sampled from.
 
@@
 The DRC program generates filters for HiFi audio systems to compensate for the reflection of sounds in a room, using impulse response measurements of the audio equipment and considering the positions of the listeners.
-The inputs are stored in PCM format, which is an array of 32-bit floating point numbers representing the samples at each sampling time step.
-Although the algorithm is divided into many phases the most computationally intensive step consists of the GSL's implementation of the Fast Fourier Transform.
+The inputs are stored in Pulse Code Modulation (PCM) format, which is an array of 32-bit floating point numbers representing the samples at each sampling time step and the computations are performed using 32-bit precision.
+Although the algorithm is divided into many phases the most computationally intensive step consists of the GSL's implementation of the Fast Fourier Transform and an DRC-internal implementation of Finite Impulse Response filter generation.
 This program accepts user-controllable inputs to decide the width and number of iterations of FFT transforms it performs in these various phases.
@@
 \label{sec:apps:hattrick}
-The Hattrick application simulates the motion of bodies under the effects of gravity.
+The Hattrick application simulates the motion of bodies under the effects of gravity using 64-bit precision.
 It is designed to help discover extra-solar planets by inferring their existence from Transit Timing Variations, where the difference between the times when a planet passes in front of its host star is used to infer the existence and properties of other planets in the system.
 This program spends most of its execution time in the GSL solver for Ordinary Differential Equations to solver the system's equations of motion.
+\greg{What inputs?}
 
@@
 Error detection techniques are discussed in Section~\ref{sec:res_tech:err_det}, with Section~\ref{sec:res_tech:err_det:algo} focusing on algorithmic techniques to detect errors in the numeric portions of application state, while Section~\ref{sec:res_tech:err_det:mem} focuses on protection of data structures.
-Section~\ref{sec:res_tech:cr} then considers recovery techniques based on both checkpoint-restart, which offers cheaper but higher-latency recovery, as well as replication of key variables, which is more expensive but enables fast recovery.
+Section~\ref{sec:res_tech:cr} then considers recovery techniques based on both checkpoint-restart, which offers lower-overhead but higher-latency recovery, as well as replication of key variables, which reduces performance but enables fast recovery.
 Finally, Section~\ref{sec:res_tech:eval} presents an experimental evaluation of the performance and accuracy properties of the resilient versions of the major routines used by our target applications, under a wide range of inputs and error injection rates.
@@
 
-%\sui{Performance experiments inserted here: running algorithmic checkers across different input sizes and fault rates: 62x62 to 500x500 matrixes for MM, RK, CD, MV checkers; FFT sizes of $2^8$ to $2^{24}$ for the FFT checker; Error bars should apply here, especially for the Cholesky Decomposition checker}
+%\sui{Performance experiments inserted here: running algorithmic checkers across different input sizes and fault rates: 62x62 to 500x500 matrices for MM, RK, CD, MV checkers; FFT sizes of $2^8$ to $2^{24}$ for the FFT checker; Error bars should apply here, especially for the Cholesky Decomposition checker}
 
@@
 Each checker computes some aggregate property of the algorithm's results using two different algorithm to produce vectors $v$ and $v'$ and their relative difference by more than some threshold $\tau$, the checker declares that an error has occurred.
-The relative difference is computed as $\frac{\left|| v-v' \right||}{max(\left||v\right||, \left||v'\right||)}$ ???.
+The relative difference is computed as $\frac{\left|| v-v' \right||}{max(\left||v\right||, \left||v'\right||)}$ \greg{is this the right formula???}.
 Each error detection causes the routine's execution to be restarted using the rollback mechanism described in Section~\ref{sec:res_tech:cr}.
@@
 The MMM operation $A \cdot B$ is checked using a matrix vector multiplication (MVM), making use of this identity: $(A \cdot B) \cdot x = A \cdot (B \cdot x)$, where $x$ is an error-checking vector.
+Both sides compute the sums of the columns of matrix $A \cdot B$ using a single matrix-vector multiplication on the left and two such operations on the right.
+Most errors in the computation of $A \cdot B$ will affect the sums of its columns and will thus be detected.
 This check is efficient because the checker is asymptotically faster than the MMM algorithm, with MVM taking $O(n^2)$ time and MMM $O(n^3)$ time.
-Since we make no assumptions about the distribution of data in the matrixes being multiplied, we set the check vector $x$ to contain all 1's.
+Since we make no assumptions about the distribution of data in the matrices being multiplied, we set the check vector $x$ to contain all 1's.
 
@@
 \paragraph{Symmetric Rank-K update (SYRK)}
-The SYRK update is a special case of matrix-matrix multiplication $B = \alpha A \cdot A^T + \beta B$, where an $n \times k$ matrix $A$ is multiplied by itself to produce an $n \times n$ matrix with rank $k$, which is added to the $nxn$ matrix $C$.
+The SYRK update is a special case of matrix-matrix multiplication $B = \alpha A \cdot A^T + \beta B$, where an $n \times k$ matrix $A$ is multiplied by itself to produce an $n \times n$ matrix with rank $k$, which is incremented into the $nxn$ matrix $B$.
 Its results are checked using the same algorithm as MMM by checking the identity $B x = \alpha A \cdot (A^Tx) + \beta B x$, where $x$ is the vector of all 1's.
@@
 
-\paragraph{FIR Filter Generation (FIR)}:
-The FIR filter generation algorithm used in DRC generates a series of samples over the sinc function $f(x)=\frac{sin(x)}{x} (f(x)=1\ when\ x\ is\ 0)$ and modulates it with a Blackman window.
-The sinc function has a nice property that $\int_{-\infty}^{\infty} sinc(x)dx = 1$. The Blackman window - modulated output of this routine
-preserves this same property for most of the time, and therefore we have the checker compute the sum of the output array and compare that with 1.
-Computing the sum requires merely $O(n)$ additions while this non-iterative FIR generation algorithm takes $O(n)$ trigonometric function evaluations, which much more complex than addition.
-Results have shown
-the overhead of this checker is very small (less than 5\%).
-Since this algorithm has only output and no input and therefore it does not need to have to be reinforced with input backup as other routines do.
+\paragraph{Finite Impulse Response Filter Generation (FIR)}
+The FIR filter generation algorithm used in DRC generates a series of samples over the $sinc$ function $sinc(x)=\frac{sin(x)}{x} (sinc(x)=1\ when\ x\ is\ 0)$ and modulates it with a Blackman window.
+This function has the property that $\int_{-\infty}^{\infty} sinc(x)dx = 1$. 
+The Blackman window - modulated output of this routine preserves this same property for most of the time, making it possible to check the output by computing the sum and comparing it to 1.
+Computing the sum requires $O(n)$ additions, as compared to the $O(n)$ trigonometric function evaluations of the non-iterative FIR generation algorithm.
+%As shown below, the overhead of this checker is very small (less than 5\%).
+%Since this algorithm has only output and no input and therefore it does not need to have to be reinforced with input backup as other routines do.
 
@@
 As such, we checkpoint the state of non-RK routines immediately after the call to the routine and include in this checkpoint the routine's arguments and contents of the floating point input buffers \greg{is this correct?}.
+The one exception is the FIR routine, which takes no input and thus requires only the processor's state to be checkpointed..
 Because RK's checkpoints are very small and it is more sensitive to errors, it is checkpointed periodically during its execution with a fixed number of steps between checkpoints.
@@
 \label{sec:res_tech:eval:la_int}
+
+This section evaluates the routines MMM, MVM, SYRK, CD, FFT and FIR.
 
@@
 %\vspace{-5pt}
-\caption{Overhead of resilience techniques when errors are not injected}
+\caption{Overhead of resilience techniques when errors are not injected \greg{Change y-axis name to say overhead rather than ratio} \sui{Done}}
 \label{fig:routine_detect_ovhd}
@@
 Figure \ref{fig:routine_detect_ovhd} measures the overheads of employing the algorithmic error detectors when no errors are injected, reporting the average of \greg{???} runs.
-The linear algebra routines were executed on $nxn$ square matrixes where $n$ varies from 100 to 1000.
+The linear algebra routines were executed on $nxn$ square matrices where $n$ varies from 100 to 1000.
 FFT ran on inputs vectors with 128K to 4M entries and RK ran on an integral with intervals $[0, t_1]$ where $t_1 \in \{10, 15, 20, 25, 30, 50, 75, 100\}$.
-\greg{What function was being computed by RK?}
+\greg{What function was being computed by RK?} \sui{The ODE system consists of a second-order nonlinear Van der Pol oscillator equation $x''(t)+\mu x'(t)(x(t)^2-1)+x(t)=0$}
 As our experiments show, the cost of algorithmic error detection is low and drops with increasing input sizes.
@@
 %\vspace{-5pt}
-\caption{Overhead of resilience techniques when errors are injected}
+\caption{Overhead of resilience techniques when no errors are injected}
 \label{fig:routine_all_ovhd}
@@
 In routines MV and RK the inputs are sufficiently small that checkpointing has negligible cost.
-Howveer, because MV runs quickly and because RK takes many checkpoints (\greg{what's the period in this experiment?}), the cost of saving processor state is small but not negligible.
+Howveer, because MV runs quickly and because RK takes many checkpoints (\greg{what's the period in this experiment?} \sui{if we take a checkpoint each time RK is called, we can end up taking millions of checkpoints, several magnitudes more than we do for MV and for other linear algebra routines and FFT)}), the cost of saving processor state is small but not negligible.
 
@@
 \includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp2_1_Example.png}
-\caption{Error magnitude histogram showing the effectiveness of the error detector}
+\caption{Error magnitude histogram showing the effectiveness of the error detector \greg{Why is FFT not improving with tighter threshold?
+Why is MM differently vulnerable than SYRK?} \sui{Reason for FFT: 1.
+Errors smaller than 10e-5 cannot be captured by the checker, even at low error rates, so this should be the fundamental limitation of the checker.
+2.
+We actually are seeing reduction in large errors, but those large errors account for such a small fraction that they are not obviously visible on the graph.}}
 \label{fig:algo_err_dist}
@@
 
-Figure~\ref{fig:algo_err_heatmap} shows the same information but for all experimental configurations: (i) input sizes of n=100 to 500 for linear algebra, 512K-8M size inputs for FFT and ???
-for RK; (ii) error injection rates of 1e-7 to 1e-10.
+Figure~\ref{fig:algo_err_heatmap} shows the same information but for all experimental configurations: (i) input sizes of n=100 to 500 for linear algebra, 512K-8M size inputs for FFT and \greg{???} \sui{10 to 100} for RK; (ii) error injection rates of 1e-7 to 1e-10.
 The graph is a grid, with results for different routines listed vertically and different resilience configurations (no detection and different detection thresholds) listed horizontally.
-The graph for each routine and resilience configuration presents the curves from Figure~\ref{fig:algo_err_dist} as a heatmap.
-The y-axis shows different routine input sizes and error injection rates, sorted according to the average number of error injections per run, in increasing order from top to bottom.
+The graph for each routine and resilience configuration presents the curves from Figure \ref{fig:algo_err_dist} as a heatmap.
+The y-axis shows different routine input sizes and error injection rates, sorted according to the average number of error injections per run, in decreasing order from top to bottom.
 The x-axis shows the different error magnitudes and the shade of each point is the fraction of the routine's entries that have a given error magnitude, with darker shades indicating more such entries.
@@
 The y-axis includes all algorithms, input sizes and error rates.
-The x-axis includes several segments for different degrees of error checking: none, as well as checking with thresholds ???. In each segment the x-axis spans over error magnitudes from ??? to ???.
+The x-axis includes several segments for different degrees of error checking: none, as well as checking with thresholds \greg{???} \sui{1e-06, 1e-07 and 1e-08}.
+In each segment the x-axis spans over error magnitudes from \greg{???} \sui{1e-14} to \greg{???} \sui{1e-01}.
 The box associated with each combination of the four parameters the figure indicates the fraction of outputs that have an error of this magnitude by the shade of gray it is colored with, as indicated in the legend.
@@
 \greg{For each experiment we need to include a discussion of what it means for how these checkers behave within the full-application runs.}
+
+\sui{The checkers have performed well in reducing erroneous results from the application outputs.
+Checkers for linear algebra are useful in removing errors from Lasso, as one could see from the results.
 
@@
 The difference between RK routine and others motivates special discussion of its resilience properties.
-Since RK's adaptive step size mechanism is already effective at tolerating many numerical errors we did not design any additional error detector for this routine.
-Instead, we focused on finding the optimal approach to memory error detection and an effective checkpointing strategy and considered different choices for these strategies as different resilience configurations.
+Since RK's adaptive step size mechanism is already effective at tolerating many numerical errors we used it as the error detector for this routine without adjusting its tolerance parameters.
+The configurable portions of the RK resilience method are thus the choice of checkpointing and/or replication strategy.
 Configurations \texttt{Ckpt} and \texttt{NoChkpt} denote whether checkpointing was used and configurations \texttt{Re} and \texttt{NoRe} denote whether replication was used.
@@
 %Experiment 2: Measure the effectiveness and performance costs of replication. Need to explain why we don't use it for Lasso and DRC.
-\section{Evaluation of Resilience Methodology}
+\section{Evaluation of Resilience Methodology in Applications}
 \label{sec:eval}
 
-\greg{This is the section where we put all the pieces together and present the experimental results on full applications.
+Having evaluated the resilience properties of the individual application routines and resilience techniques we now consider our three target applications ADDR, Hattrick and DRC.
+Section~\ref{sec:eval:perf} reports on the impact of errors and resilience techniques on application performance.
+Section~\ref{sec:eval:acc} describes the distribution of errors in the results of these applications.
+These results are based on a fault injection campaign that includes a range of fault injection rates, input sizes and configurations of resilience mechanisms, with 100 application runs for each combination of parameters.
+This corresponds to a total of \greg{???} error injection experiments.
+
+A given application run may terminate successfully or crash due to an unrecoverable error.
+Unrecoverable errors occur when there is a segmentation fault in code outside our protected routines or if 5 rollbacks of a routine have been observed in a given run since such repeated rollbacks almost always indicate that the corruption in application state cannot be repaired.
+Since we assume that the user needs application results regardless of the error rate, if the application crashes it is restarted from the beginning as many times as needed for it to complete.
+
+To evaluate the behavior of these applications in a broad range of error environments we considered error injection rates that range from 1e-16 \greg{???} errors per cycle of the LLVM virtual machine to 1e-6 errors per cycle.
+For reference, assuming a typical FIT of 100-100,000 for individual processing cores (FIT = failure in one billion hours of operation and this range has been experimentally observed for commodity electronics~\cite{mem_errors:2010, DRAMerr:2009}), an HPC system with one million cores will experience .1-100 errors per hour, which is equivalent to 3e-11 - 3e-14 errors per processor cycle on a 1Ghz processor.
+The range of error injection rates used used in our experiments were chosen such that (i) on average one or more errors were injected in the executions of our target applications and (ii) at least some application runs completed without crashing.
+Results for error rates that induce fewer than one error per run were extrapolated from our experiments with higher error rates.
+This was done by considering only the runs where exactly one error was injected and trivially scaling the results to correspond to 1 in every $n$ runs being injected with an error while remaining $n-1$ runs complete with no injection.
+
+%\greg{This is the section where we put all the pieces together and present the experimental results on full applications.
 We should describe our main use-case here: user wants to achieve a given accuracy with a given probability and we configure the algorithm to provide this at the lowest cost.
@@
 
-Show the Markov model that describes our rollback scheme.
+Our fault injection experiments ran each application configuration to completion, recording whether it completed and if so, the magnitude of the error.
+We calculate the execution time of each configuration, accounting for application rollbacks via the formula $\frac{p_c*t_c}{1-p_f}$, where $p_c$ is the probability of a run completing, $t_c$ is the average completion time of such runs and $p_f$ is the probability of a run failing to complete\greg{is this right}.
+Figures~\ref{fig:Lasso_EstdCost}, \ref{fig:DRC_EstdCost} and \ref{fig:Hattrick_EstdCost} show the expected execution time, relative to the execution time of the non-fault tolerant version of the application when no errors are injected.
+The graphs from left to right correspond to different application input sizes.
+Within each graph the x-axis corresponds to error injection rates, while the y-axis shows the relative execution time, both logarithmic.
+Each graph shows data for the non-fault tolerant configuration of the application, connecting all the data points with a line.
+They also show the relative execution times for all the configurations of the fault tolerant version, with the line connecting the results for the configuration that is optimal for each input size and fault rate.
 
-\sui{When a user wants to run an application under a faulty environment and obtain correct results, s/he would have to expect to run a part of this application multiple times in case the application aborts due to memory access errors or return useless incorrect results.
+For each configuration we plot a 95\% confidence interval.
+This is the range of relative execution times at the given error rate and application configuration in that will occur in 95\% of all trials.
+The confidence interval is computed by modeling the number of times application results will meet the error bound using a Binomial distribution.
+Given a binary random variable that is equal to 1 (i.e.
+error bound achieved) with probability $p$ and 0 (i.e.
+bound missed) with probability $1-p$ the Binomial distribution is the number of trials where it is 1.
+
+\greg{Sui, please confirm that the following is true}
+The data shows that the non-fault tolerant ADDR and Hattrick shows little slowdown as a result of injected faults.
+This is because these applications spend most of its time in linear algebra and differential equation routines which are numerically-intensive but perform relatively few memory manipulations or pointer accesses.
+As such, until it reaches error rates of 1e-9 and above ADDR does not see a significant number of aborts due to memory errors.
+The result is that the addition of resilience techniques does little to improve ADDR's performance.
+
+In contrast, since DRC performs many more integer operations and memory copies, it is more vulnerable to errors that corrupt the values of pointers or cause loads and stores to target invalid memory locations.
+Such errors cause memory segmentation faults, which force DRC to restart its execution and thus incur a very high overhead.
+In contrast, the localized rollback and correction provided by our checkpointing mechanisms ensures that each restart has a much lower cost and that a single error does not propagate past the end of a single call to FFT or FIR.
+
+\begin{figure*}[ht!]
+\centering
+\includegraphics[width=2.00\columnwidth]{figs/Lasso_EstdCost.png}
+\vspace{-10pt}
+\caption{Execution time of ADDR relative to non-fault tolerant fault-free execution time}
+\vspace{-10pt}
+\label{fig:Lasso_EstdCost}
+\end{figure*}
+
+\begin{figure*}[ht!]
+\centering
+\includegraphics[width=2.00\columnwidth]{figs/DRC_EstdCost.png}
+\vspace{-10pt}
+\caption{Execution time of DRC relative to non-fault tolerant fault-free execution time \greg{Need confidence intervals}}
+\vspace{-10pt}
+\label{fig:DRC_EstdCost}
+\end{figure*}
+
+\begin{figure*}[ht!]
+\centering
+\includegraphics[width=2.00\columnwidth]{figs/Hattrick_EstdCost.png}
+\vspace{-10pt}
+\caption{Execution time of Hattrick relative to non-fault tolerant fault-free execution time \greg{Need confidence intervals}}
+\vspace{-10pt}
+\label{fig:Hattrick_EstdCost}
+\end{figure*}
+
+%\sui{When a user wants to run an application under a faulty environment and obtain correct results, s/he would have to expect to run a part of this application multiple times in case the application aborts due to memory access errors or return useless incorrect results.
 The user might want to have the application return to a certain checkpoint, or restart the entire application should any unrecoverable occurs.
@@
 
-Experiment 1: Measure the slowdown of each application at the optimal configuration
-Experiment 2: Measure the rate of aborts due to algorithmic errors, segfaults and redundant copy mismatches (separate the causes)
+%Experiment 1: Measure the slowdown of each application at the optimal configuration
+%Experiment 2: Measure the rate of aborts due to algorithmic errors, segfaults and redundant copy mismatches (separate the causes)
 
@@
 
-Experiment: Measure the accuracy of the results produced in completed runs, including error bars on the variability of the errors
+%Experiment: Measure the accuracy of the results produced in completed runs, including error bars on the variability of the errors
+
+If the probability that a given run completes without an unrecoverable error is greater than 0 then the application will eventually complete with some output after some number of restarts.
+This section evaluates the distribution of errors in the resulting outputs.
+The outputs of the different applications are as follows:
+\begin{itemize}
+    \item ADDR: vector that holds the solution to the linear system.
+    \item Hattrick: vector that holds the state of the n-body system at the final application time-step, with 6 numbers per body.
+    \item DRC: a Pulse Code Modulation-formatted file, which is a sequence of numbers.
+\end{itemize}
+Since the outputs of each routine can be represented as vectors we calculate the error in the outputs by comparing the vector $v_e$ returned by each routine when errors are injected to the correct vector $v_c$ computed during a fault-free run and computing the L2 norm of the difference: $\sqrt{\frac{\sum_{i} (v_c-v_e)^2}{\left| v_c \right|}}$.
+
+Figures~\ref{fig:Lasso_ImperfectRate}, \ref{fig:DRC_ImperfectRate} and \ref{fig:Hattrick_ImperfectRate} show the errors in application results from a user-centric perspective.
+Given a user-specified error magnitude that must be satisfied by application results these figures plot the fraction of application runs for which this target was satisfied
+From left to right each figure shows graphs for different input sizes and from top to bottom it shows results for different error magnitude bounds.
+The x-axis in each graph shows the fault injection rate (logarithmic) while the y-axis shows the fraction of runs where the error target was satisfied (linear).
+Each graph shows data for the non-fault tolerant version of the application (all points connected by the line) as well as all the the fault tolerant versions (optimal configurations connected by a line).
+The 95\% confidence interval is plotted and denoted range of probabilities that the error bound will be achieved at the given error rate and application configuration in 95\% of all trials.
+
+\greg{Sui, please confirm that the following is true}
+The data shows that the probability that the non-fault tolerant version of ADDR satisfies the user-provided error bound is low, for bounds randing from 1e-10 to 1e-4.
+In contrast, the fault tolerant version of the application is significantly more successful, especially for error rates of 1e-12 and lower (\greg{need log plot to show the low accuracy rate more precisely}).
+\greg{Can mention the fact that the fault tolerant versions are most successful for loose error bounds and small inputs but the success rate is poor for the 10x input. Why?}
+
+\greg{I won't comment on the DRC errors because Sui is currently trying to improve them.
+Right now we're not being very successful.
+However, since DRC is a single precision code, we should really be considering much larger error bounds, like 1e-6 or 1e-4}.
+
+\greg{Its hard to comment on Hattrick until we have a version of the plots with a logscale y-axis since it'll tell us exactly how successful the fault tolerant versions of the code are.
+Also, what are the relative execution times of these input files?
+This will give us an idea of how vulnerable they are.}
+
+\begin{figure*}[ht!]
+\centering
+\includegraphics[height=5in]{figs/Lasso_ImperfectRate.png}
+\caption{Probability of not satisfying the user-provided error bound for ADDR \greg{why do the probabilities of the no-FT version not change as the error threshold changes?
+Also, it may be good to show fewer columns but add graphs where the y-scale is a log-plot}}}
+\label{fig:Lasso_ImperfectRate}
+\end{figure*}
+
+\begin{figure*}[ht!]
+\centering
+\includegraphics[height=3in]{figs/DRC_ImperfectRate.png}
+\caption{Probability of not satisfying the user-provided error bound for DRC \greg{why do the probabilities of the no-FT version not change as the error threshold changes?
+Also, it may be good to show fewer columns but add graphs where the y-scale is a log-plot}}
+\label{fig:DRC_ImperfectRate}
+\end{figure*}
+
+\begin{figure*}[ht!]
+\centering
+\includegraphics[height=3in]{figs/Hattrick_ImperfectRate.png}
+\caption{Probability of not satisfying the user-provided error bound for Hattrick \greg{why do the probabilities of the no-FT version not change as the error threshold changes?
+Also, it may be good to show fewer columns but add graphs where the y-scale is a log-plot}}
+\label{fig:Hattrick_ImperfectRate}
+\end{figure*}
 
@@
 
-Experiment: Final graph that shows slowdown for each desired probability of perfect results.
+%Experiment: Final graph that shows slowdown for each desired probability of perfect results.
 

### 2013-01-20 paper.tex
Application results; fixing some problems

@@
 More specifically, these configurations can be used to produce a wide range of overheads and result accuracy levels and to tune the probabilistic confidence level that the given overhead or accuracy target will be met.
-\greg{Summarize real results}
+\greg{Summarize real results} \sui{A tighter algorithmic checker bounds can be enabled at low error rates to completely eliminate erroneous results at a lower overhead, while a more lenient algorithmic checker would allow the user to produce more acceptable results in face of a higher error rate (such as 1e-09).}
 
@@
 Although the algorithm is divided into many phases the most computationally intensive step consists of the GSL's implementation of the Fast Fourier Transform and an DRC-internal implementation of Finite Impulse Response filter generation.
-This program accepts user-controllable inputs to decide the width and number of iterations of FFT transforms it performs in these various phases.
-Different configurations would load the CPU for different lengths of time, virtually equivalent to inputs of different sizes.
+This program accepts user-controllable configurations to decide the width and number of iterations of FFT transforms it uses to perform various passes on the input file, such as windowing, deconvolution, pre- and post-filtering.
+Different configurations would load the CPU for different lengths of time, resulting in the program generating inputs of different sizes \sui{to the FFT routines.}.
 \greg{What inputs?}
@@
 This program spends most of its execution time in the GSL solver for Ordinary Differential Equations to solver the system's equations of motion.
-\greg{What inputs?}
+\greg{What inputs?} \sui{Input files are four different planetary systems.
+The first, third and fourth input files simulate one star and one planet and the second simulates one star and two planets.
+The third and fourth inputs are the same planetary system with different lengths of simulation time.}
 
@@
 Each checker computes some aggregate property of the algorithm's results using two different algorithm to produce vectors $v$ and $v'$ and their relative difference by more than some threshold $\tau$, the checker declares that an error has occurred.
-The relative difference is computed as $\frac{\left|| v-v' \right||}{max(\left||v\right||, \left||v'\right||)}$ \greg{is this the right formula???}.
+The relative difference is computed as $\frac{\left|| v-v' \right||}{max(\left||v\right||, \left||v'\right||)}$ \greg{is this the right formula???} \sui{yes}.
 Each error detection causes the routine's execution to be restarted using the rollback mechanism described in Section~\ref{sec:res_tech:cr}.
@@
 By treating the input array as a matrix and comparing row sums and column sums of the submatrices of that matrix, we are able to detect the existence of errors and fix a fraction of the errors.
+\sui{This algorithm takes in an array of numbers, divide it into blocks of $N \cdot N$ numbers, computes and stores the row sums and column sums that are used for recovery.
+It is capable of correcting one corrupt element that does not share the same row or column with another element.}
 The protection mechanism has the following parameters:
@@
 In contrast, RK takes in very little input and operates as a long sequence of steps, each of which involves a call to a user-provided derivative routine.
-As such, we checkpoint the state of non-RK routines immediately after the call to the routine and include in this checkpoint the routine's arguments and contents of the floating point input buffers \greg{is this correct?}.
+As such, we checkpoint the state of non-RK routines immediately after the call to the routine and include in this checkpoint the routine's arguments and contents of the floating point input buffers \greg{is this correct?} \sui{Yes}.
 The one exception is the FIR routine, which takes no input and thus requires only the processor's state to be checkpointed..
 Because RK's checkpoints are very small and it is more sensitive to errors, it is checkpointed periodically during its execution with a fixed number of steps between checkpoints.
-As our experiments \greg{where?} show, the choice of the RK checkpoint period significantly affects its efficiency with longer periods suffering from more overhead due to rollbacks and short period spending more time checkpointing.
+As our experiments \greg{where?} \sui{This should be ready tomorrow} show, the choice of the RK checkpoint period significantly affects its efficiency with longer periods suffering from more overhead due to rollbacks and short period spending more time checkpointing.
 
@@
 We evaluate the costs of the above algorithmic error checkers by measuring their costs and effectiveness when running at different error rates on different types of inputs.
-Figure \ref{fig:routine_detect_ovhd} measures the overheads of employing the algorithmic error detectors when no errors are injected, reporting the average of \greg{???} runs.
+Figure \ref{fig:routine_detect_ovhd} measures the overheads of employing the algorithmic error detectors when no errors are injected, reporting the average of \greg{???} \sui{100} runs.
 The linear algebra routines were executed on $nxn$ square matrices where $n$ varies from 100 to 1000.
@@
 %\vspace{-5pt}
-\caption{Overhead of resilience techniques when no errors are injected}
+\caption{Overhead of resilience techniques when no errors are injected \sui{(Bars) and when errors are injected (points).
+(The SYRK overhead is from the old algorithm (the O($n^2 logn$) version; will update today!
+}}
 \label{fig:routine_all_ovhd}
@@
 Each routine's graph shows $\frac{\texttt{Time with Errors}}{\texttt{Time No Errors}}$, where the \texttt{Time with Errors} is average increase execution time of a resilient or non-resilient version of the routine when running on the same range of inputs as above, and \texttt{Time No Errors} is the execution time of the non-resilient version on the same inputs when no errors are injected.
-The graph for each routine reports this metric for all of the above input sizes, error injection rates between 1e-8 and 1e-11 \greg{confirm!} and all resilience configurations (no resilience and resilience with detection thresholds between 1e-6 and 1e-8), as the average time of \greg{???} runs.
+The graph for each routine reports this metric for all of the above input sizes, error injection rates between 1e-8 and \sui{(1e-8 is the highest error rate that we can have for most runs to complete) and 1e-10 (is the lowest error rate we would observe errors from mosts of the runs) } \greg{confirm!} and all resilience configurations (no resilience and resilience with detection thresholds between 1e-6 and 1e-8), as the average time of \greg{???} runs \sui{100 runs per configuration}.
 The data points within each figure are organized horizontally according to the average number of errors injected into runs with the given input size and at the given injection rate.
-\greg{Do the error counts include errors after rollbacks?}
+\greg{Do the error counts include errors after rollbacks?} \sui{No, because the error count is the number of errors for Non-FT runs with errors.
+Since there is no FT and no checkpointing in these configurations, the number doesn't include errors after rollbacks.
+}
 This average error count is shown on the x-axis.
@@
 
-The data shows \greg{???}
+The data shows \greg{???} \sui{Soft errors may cause the routines to run longer due to rollback in attempt to correct the errors.
+The higher the error rate, the more likely rollback is to occur.
+For example, the Cholesky Decomposition routine with input size 100, when exposed to an error rate of 1e-06, would run on average twice as long as when no fault is injected.
+That means in this case, every one of the runs tend to have 100\% overhead.
+In some cases the algorithm would give up retrying after the fifth time because it is unable to recover from the corrupt inputs.
+For example, for FFT exposed at a fault rate of 1e-08, the average overhead, as computed by dividing the time spent in the fault-tolerance codes by the baseline running time, decreases, because more runs tend to early-terminate when too many elements in the input have been corrupted.}
+
+\sui{I will add a graph showing how many of the inputs have been corrupted and failed to recover. This is an important issue for FFT.}
 
@@
 Why is MM differently vulnerable than SYRK?} \sui{Reason for FFT: 1.
-Errors smaller than 10e-5 cannot be captured by the checker, even at low error rates, so this should be the fundamental limitation of the checker.
+Errors smaller than a certain value cannot be captured by the checker, even at low error rates, so this should be the fundamental limitation of the checker.
 2.
-We actually are seeing reduction in large errors, but those large errors account for such a small fraction that they are not obviously visible on the graph.}}
+We actually are seeing reduction in large errors, but those large errors account for such a small fraction that they are not obviously visible on the graph.
+For the difference between MM and SYRK: MM is more vulnerable therefore the difference between ``corrected'' and ``not corrected'' would look more noticable.}}
 \label{fig:algo_err_dist}
@@
 \marc{The y-axis should show just the number of errors injected. We can say that we have used several fault injection rates and sizes to get those numbers. Right now the picture is confusing.}
+\sui{How does this look?}
 
@@
 Further, it shows that a reduction of detection threshold is consistently effective at reducing the fraction of errors in each algorithm's output.
-However, the effectiveness of algorithmic checks has a limit, with CD and FFT showing moderate error vulnerability even at threshold 1e-8.
-MM, MV and SYRK are significantly less vulnerable across the full range of error magnitudes.
+%However, the effectiveness of algorithmic checks has a limit, with CD and FFT showing moderate error vulnerability even at threshold 1e-8.
+\sui{(Rephrase) However, the algorithmic checkers cannot cover 100\% of the errors for CD, FFT and FIR, mainly fixing errors of large magnitudes which turned out to be effective in real applications (refer to the imperfect run rate plot for DRC.)}
+%MM, MV and SYRK are significantly less vulnerable across the full range of error magnitudes.
+\sui{(Rephrase) The errors shown in MM, MV and SYRK are much more noticably fixed.}
 \greg{should reference vulnerability numbers more specifically}
@@
 Configurations \texttt{Ckpt} and \texttt{NoChkpt} denote whether checkpointing was used and configurations \texttt{Re} and \texttt{NoRe} denote whether replication was used.
-\greg{What checkpointing period and replication properties?}
+\greg{What checkpointing period and replication properties?} \sui{Checkpointing: every 1/100/1000/10000/100000 calls to RK4 stepping function.
+Replication: No replication, using 1 pair of replicas and using maximum number of replicas.}
 
@@
 These results are based on a fault injection campaign that includes a range of fault injection rates, input sizes and configurations of resilience mechanisms, with 100 application runs for each combination of parameters.
-This corresponds to a total of \greg{???} error injection experiments.
+This corresponds to a total of \greg{???} \sui{66100 for lasso, 43200 for DRC, 102000 for Hattrick, as counted by my scripts.} error injection experiments.
 
@@
 
-To evaluate the behavior of these applications in a broad range of error environments we considered error injection rates that range from 1e-16 \greg{???} errors per cycle of the LLVM virtual machine to 1e-6 errors per cycle.
+To evaluate the behavior of these applications in a broad range of error environments we considered error injection rates that range from 1e-16 \greg{???}\sui{Yes.
+At least this is the argument I sent to the injector.} errors per cycle of the LLVM virtual machine to 1e-6 errors per cycle.
 For reference, assuming a typical FIT of 100-100,000 for individual processing cores (FIT = failure in one billion hours of operation and this range has been experimentally observed for commodity electronics~\cite{mem_errors:2010, DRAMerr:2009}), an HPC system with one million cores will experience .1-100 errors per hour, which is equivalent to 3e-11 - 3e-14 errors per processor cycle on a 1Ghz processor.
@@
 We calculate the execution time of each configuration, accounting for application rollbacks via the formula $\frac{p_c*t_c}{1-p_f}$, where $p_c$ is the probability of a run completing, $t_c$ is the average completion time of such runs and $p_f$ is the probability of a run failing to complete\greg{is this right}.
+\sui{I don't see anything wrong}
 Figures~\ref{fig:Lasso_EstdCost}, \ref{fig:DRC_EstdCost} and \ref{fig:Hattrick_EstdCost} show the expected execution time, relative to the execution time of the non-fault tolerant version of the application when no errors are injected.
@@
 
-\greg{Sui, please confirm that the following is true}
+\greg{Sui, please confirm that the following is true}}
 The data shows that the non-fault tolerant ADDR and Hattrick shows little slowdown as a result of injected faults.
@@
 As such, until it reaches error rates of 1e-9 and above ADDR does not see a significant number of aborts due to memory errors.
-The result is that the addition of resilience techniques does little to improve ADDR's performance.
+The result is that the addition of resilience techniques does little to improve ADDR's performance. \sui{The addition of resilience techniques greatly improve its accuracy}}
 
@@
 \vspace{-10pt}
-\caption{Execution time of DRC relative to non-fault tolerant fault-free execution time \greg{Need confidence intervals}}
+\caption{Execution time of DRC relative to non-fault tolerant fault-free execution time \greg{Need confidence intervals} \sui{Added.}}
 \vspace{-10pt}
@@
 \vspace{-10pt}
-\caption{Execution time of Hattrick relative to non-fault tolerant fault-free execution time \greg{Need confidence intervals}}
+\caption{Execution time of Hattrick relative to non-fault tolerant fault-free execution time \greg{Need confidence intervals} \sui{Added.}}
 \vspace{-10pt}
@@
 
-\greg{Sui, please confirm that the following is true}
+\greg{Sui, please confirm that the following is true} \sui{Yes.
+ADDR is sensitive to errors, and even a small number of error would cause noticable errors in the results.
+Comparing the leftmost data points on the ``probability of not satisfying'' graphs of the 3 applications.}
 The data shows that the probability that the non-fault tolerant version of ADDR satisfies the user-provided error bound is low, for bounds randing from 1e-10 to 1e-4.
-In contrast, the fault tolerant version of the application is significantly more successful, especially for error rates of 1e-12 and lower (\greg{need log plot to show the low accuracy rate more precisely}).
-\greg{Can mention the fact that the fault tolerant versions are most successful for loose error bounds and small inputs but the success rate is poor for the 10x input. Why?}
+In contrast, the fault tolerant version of the application is significantly more successful, especially for error rates of 1e-12 and lower (\greg{need log plot to show the low accuracy rate more precisely} \sui{Changed to log plot.}).
+\greg{Can mention the fact that the fault tolerant versions are most successful for loose error bounds and small inputs but the success rate is poor for the 10x input.
+Why?} \sui{I know why.
+After inspecting Lasso's code, it turns out Lasso contains TWO branches of code: one on ``fat'' matrices and the other on ``skinny'' matrices.
+Only 10x input triggers the ``fat matrix'' code and inputs 20 through 100 would trigger the ``skinny'' code.
+The two versions of code are not the same therefore they would be vulnerable to soft errors to different degrees.}
 
@@
 However, since DRC is a single precision code, we should really be considering much larger error bounds, like 1e-6 or 1e-4}.
+\sui{Expanding the ranges makes sense.
+Plus, I think the range for Hattrick can be as tight as 1e-20.
+I have used 1e-17, 1e-14 and 1e-10 in the updated visualization.}
 
@@
 Also, what are the relative execution times of these input files?
-This will give us an idea of how vulnerable they are.}
+This will give us an idea of how vulnerable they are.} \sui{Input 1=55s; Input 2=155s; Input 3=82s; Input 4=120s }
 
@@
 \includegraphics[height=5in]{figs/Lasso_ImperfectRate.png}
+\includegraphics[height=5in]{figs/Lasso_ImperfectRate_log.png}
 \caption{Probability of not satisfying the user-provided error bound for ADDR \greg{why do the probabilities of the no-FT version not change as the error threshold changes?
-Also, it may be good to show fewer columns but add graphs where the y-scale is a log-plot}}}
+Also, it may be good to show fewer columns but add graphs where the y-scale is a log-plot}} \sui{Added a log plot.}}
 \label{fig:Lasso_ImperfectRate}
@@
 \includegraphics[height=3in]{figs/DRC_ImperfectRate.png}
+\includegraphics[height=3in]{figs/DRC_ImperfectRate_log.png}
 \caption{Probability of not satisfying the user-provided error bound for DRC \greg{why do the probabilities of the no-FT version not change as the error threshold changes?
-Also, it may be good to show fewer columns but add graphs where the y-scale is a log-plot}}
+Also, it may be good to show fewer columns but add graphs where the y-scale is a log-plot} \sui{Added a log plot (on the right).}}
 \label{fig:DRC_ImperfectRate}
@@
 \includegraphics[height=3in]{figs/Hattrick_ImperfectRate.png}
+\includegraphics[height=3in]{figs/Hattrick_ImperfectRate_log.png}
 \caption{Probability of not satisfying the user-provided error bound for Hattrick \greg{why do the probabilities of the no-FT version not change as the error threshold changes?
-Also, it may be good to show fewer columns but add graphs where the y-scale is a log-plot}}
+Also, it may be good to show fewer columns but add graphs where the y-scale is a log-plot} \sui{The change is small therefore not visible on a linear scale.
+Switching to log plot revealed those tiny differences.}}
 \label{fig:Hattrick_ImperfectRate}

### 2013-01-22 paper.tex
Fixed 4 figs
Marc's changes

@@
 \sui{The checkers have performed well in reducing erroneous results from the application outputs.
-Checkers for linear algebra are useful in removing errors from Lasso, as one could see from the results.
+Checkers for linear algebra are useful in removing errors from Lasso, as one could see from the results.}
 
@@
 \centering
-%\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp1_rk}
+\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Overall_Breakdown_RK4}
 %\vspace{-5pt}
@@
 
-\greg{Sui, please confirm that the following is true}}
+\greg{Sui, please confirm that the following is true}
 The data shows that the non-fault tolerant ADDR and Hattrick shows little slowdown as a result of injected faults.
@@
 As such, until it reaches error rates of 1e-9 and above ADDR does not see a significant number of aborts due to memory errors.
-The result is that the addition of resilience techniques does little to improve ADDR's performance. \sui{The addition of resilience techniques greatly improve its accuracy}}
+The result is that the addition of resilience techniques does little to improve ADDR's performance. \sui{The addition of resilience techniques greatly improve its accuracy}
 
@@
 \caption{Probability of not satisfying the user-provided error bound for ADDR \greg{why do the probabilities of the no-FT version not change as the error threshold changes?
-Also, it may be good to show fewer columns but add graphs where the y-scale is a log-plot}} \sui{Added a log plot.}}
+Also, it may be good to show fewer columns but add graphs where the y-scale is a log-plot} \sui{Added a log plot.}}
 \label{fig:Lasso_ImperfectRate}

### 2013-01-22 paper.tex
Fixed Fig 4

@@
 More specifically, these configurations can be used to produce a wide range of overheads and result accuracy levels and to tune the probabilistic confidence level that the given overhead or accuracy target will be met.
-\greg{Summarize real results} \sui{A tighter algorithmic checker bounds can be enabled at low error rates to completely eliminate erroneous results at a lower overhead, while a more lenient algorithmic checker would allow the user to produce more acceptable results in face of a higher error rate (such as 1e-09).}
+\greg{Summarize real results} \sui{A tighter algorithmic checker bounds can be enabled at low error rates to completely eliminate erroneous results at a lower overhead, while a more lenient algorithmic checker would allow the user to produce more acceptable results in face of a higher error rate (such as 1e-09).
+Through high and low error rates, we correct more than 50\% of segmentation faults and reduce the error in the outputs by magnitudes by paying lower than 20\% overheads for linear algebra, FFT and FIR filter generation and a reasonable overhead for the integrator.}
 
@@
 
-Cosmic rays or alpha particles hitting the computer circuits and accumulating a sufficient amount of energy may invert the device logic states from "0" to "1" or from "1" to "0"~\cite{ibm_softerr:1996,baumann:2005}.
+Cosmic rays or alpha particles hitting the computer circuits and accumulating a sufficient amount of energy may invert the device logic states from "0" to "1" or from "1" to "0"~\cite{Ziegler:1996:TCR:226354.226356,baumann:2005}.
 These inversions are termed as soft errors or soft faults.
@@
 The ACE bits include critical information to keep the program logic such as the program counter register for a committed instruction.
-In the past decades, soft errors have brought significant losses on commercial servers and high performance computers such as the recall of Sun servers due to memory errors~\cite{SoftErr:Baumann:2002} or the cache corruptions in the ASCI Q system~\cite{asciQSER:2005}.
+In the past decades, soft errors have brought significant losses on commercial servers and high performance computers such as the recall of Sun servers due to memory errors~\cite{baumann:2005} or the cache corruptions in the ASCI Q system~\cite{asciQSER:2005}.
 
@@
 This program accepts user-controllable configurations to decide the width and number of iterations of FFT transforms it uses to perform various passes on the input file, such as windowing, deconvolution, pre- and post-filtering.
-Different configurations would load the CPU for different lengths of time, resulting in the program generating inputs of different sizes \sui{to the FFT routines.}.
-\greg{What inputs?}
+The input used in this paper is an audio file of size 768 kilobytes.
 
@@
 This program spends most of its execution time in the GSL solver for Ordinary Differential Equations to solver the system's equations of motion.
-\greg{What inputs?} \sui{Input files are four different planetary systems.
+Input files are four different planetary systems.
 The first, third and fourth input files simulate one star and one planet and the second simulates one star and two planets.
-The third and fourth inputs are the same planetary system with different lengths of simulation time.}
+The third and fourth inputs are the same planetary system with different lengths of simulation time.
 
@@
 Each checker computes some aggregate property of the algorithm's results using two different algorithm to produce vectors $v$ and $v'$ and their relative difference by more than some threshold $\tau$, the checker declares that an error has occurred.
-The relative difference is computed as $\frac{\left|| v-v' \right||}{max(\left||v\right||, \left||v'\right||)}$ \greg{is this the right formula???} \sui{yes}.
+The relative difference is computed as $\frac{\left|| v-v' \right||}{max(\left||v\right||, \left||v'\right||)}$.
 Each error detection causes the routine's execution to be restarted using the rollback mechanism described in Section~\ref{sec:res_tech:cr}.
@@
 This check is efficient because the checker is asymptotically faster than the MMM algorithm, with MVM taking $O(n^2)$ time and MMM $O(n^3)$ time.
-Since we make no assumptions about the distribution of data in the matrices being multiplied, we set the check vector $x$ to contain all 1's.
+Since this is dense martix matrix multiplication, we set the check vector $x$ to contain all 1's.
 
@@
 When errors are detected we employ a recovery method based on checkpoint-restart, where routine state is recorded at key locations and whenever an algorithmic or memory error is detected we roll back to that point in its execution.
-Rollbacks are implemented via the \texttt{sigsetjmp}/\texttt{siglongjmp} routines.
+Rollbacks are implemented via the \texttt{sigsetjmp}/\texttt{siglongjmp} \cite{sigsetjmp:1997} routines.
 \texttt{sigsetjmp} record the application's execution state (registers and stack pointer) at the time of the checkpoint.
@@
 By treating the input array as a matrix and comparing row sums and column sums of the submatrices of that matrix, we are able to detect the existence of errors and fix a fraction of the errors.
-\sui{This algorithm takes in an array of numbers, divide it into blocks of $N \cdot N$ numbers, computes and stores the row sums and column sums that are used for recovery.
-It is capable of correcting one corrupt element that does not share the same row or column with another element.}
-The protection mechanism has the following parameters:
-\begin{itemize}
-\item{\texttt{N}, the size of the error correcting sum block.}
-\item{\texttt{ECC\_ECC}, whether or not the error correcting code itself should be corrected.}
-\end{itemize}
-This mechanism is able to correct up to $1/{n^2}$ of the elements in the input array.
-\greg{Which code is being used here? You just say that it is based on block-checksums}.
+This algorithm takes in an array of numbers, divide it into blocks of $N \cdot N$ numbers and for each block it computes and stores the row sums and column sums that are used for recovery.
+The protection mechanism has one parameters,\texttt{N}, the size of the error correcting sum block.
 
@@
 In contrast, RK takes in very little input and operates as a long sequence of steps, each of which involves a call to a user-provided derivative routine.
-As such, we checkpoint the state of non-RK routines immediately after the call to the routine and include in this checkpoint the routine's arguments and contents of the floating point input buffers \greg{is this correct?} \sui{Yes}.
+As such, we checkpoint the state of non-RK routines immediately after the call to the routine and include in this checkpoint the routine's arguments and contents of the floating point input buffers.
 The one exception is the FIR routine, which takes no input and thus requires only the processor's state to be checkpointed..
 Because RK's checkpoints are very small and it is more sensitive to errors, it is checkpointed periodically during its execution with a fixed number of steps between checkpoints.
-As our experiments \greg{where?} \sui{This should be ready tomorrow} show, the choice of the RK checkpoint period significantly affects its efficiency with longer periods suffering from more overhead due to rollbacks and short period spending more time checkpointing.
+As our experiments in \ref{sec:res_tech:eval:rk} show, the choice of the RK checkpoint period significantly affects its efficiency with longer periods suffering from more overhead due to rollbacks and short period spending more time checkpointing.
 
@@
 %\vspace{-5pt}
-\caption{Overhead of resilience techniques when errors are not injected \greg{Change y-axis name to say overhead rather than ratio} \sui{Done}}
+\caption{Overhead of resilience techniques when errors are not injected \greg{Change y-axis name to say overhead rather than ratio}}
 \label{fig:routine_detect_ovhd}
@@
 We evaluate the costs of the above algorithmic error checkers by measuring their costs and effectiveness when running at different error rates on different types of inputs.
-Figure \ref{fig:routine_detect_ovhd} measures the overheads of employing the algorithmic error detectors when no errors are injected, reporting the average of \greg{???} \sui{100} runs.
+Figure \ref{fig:routine_detect_ovhd} measures the overheads of employing the algorithmic error detectors when no errors are injected, reporting the average of 100 runs.
 The linear algebra routines were executed on $nxn$ square matrices where $n$ varies from 100 to 1000.
 FFT ran on inputs vectors with 128K to 4M entries and RK ran on an integral with intervals $[0, t_1]$ where $t_1 \in \{10, 15, 20, 25, 30, 50, 75, 100\}$.
-\greg{What function was being computed by RK?} \sui{The ODE system consists of a second-order nonlinear Van der Pol oscillator equation $x''(t)+\mu x'(t)(x(t)^2-1)+x(t)=0$}
+The ODE system consists of a second-order nonlinear Van der Pol oscillator equation $x''(t)+\mu x'(t)(x(t)^2-1)+x(t)=0$.
 As our experiments show, the cost of algorithmic error detection is low and drops with increasing input sizes.
@@
 %\vspace{-5pt}
-\caption{Overhead of resilience techniques when no errors are injected \sui{(Bars) and when errors are injected (points).
-(The SYRK overhead is from the old algorithm (the O($n^2 logn$) version; will update today!
-}}
+\caption{Overhead of resilience techniques when no errors are injected and when errors are injected (points).}
 \label{fig:routine_all_ovhd}
@@
 In routines MV and RK the inputs are sufficiently small that checkpointing has negligible cost.
-Howveer, because MV runs quickly and because RK takes many checkpoints (\greg{what's the period in this experiment?} \sui{if we take a checkpoint each time RK is called, we can end up taking millions of checkpoints, several magnitudes more than we do for MV and for other linear algebra routines and FFT)}), the cost of saving processor state is small but not negligible.
+Howveer, because MV runs quickly and because RK takes many checkpoints ($9000 \cdot input size)$), the cost of saving processor state is small but not negligible.
 
-Finally, Figure~\ref{fig:routine_exp_exec} documents the overhead of resilience when errors are injected into application execution, accounting for the slowdowns due to the resilience mechanisms as well as rollbacks due to error detections.
-Each routine's graph shows $\frac{\texttt{Time with Errors}}{\texttt{Time No Errors}}$, where the \texttt{Time with Errors} is average increase execution time of a resilient or non-resilient version of the routine when running on the same range of inputs as above, and \texttt{Time No Errors} is the execution time of the non-resilient version on the same inputs when no errors are injected.
-The graph for each routine reports this metric for all of the above input sizes, error injection rates between 1e-8 and \sui{(1e-8 is the highest error rate that we can have for most runs to complete) and 1e-10 (is the lowest error rate we would observe errors from mosts of the runs) } \greg{confirm!} and all resilience configurations (no resilience and resilience with detection thresholds between 1e-6 and 1e-8), as the average time of \greg{???} runs \sui{100 runs per configuration}.
+%Finally, Figure~\ref{fig:routine_exp_exec} documents the overhead of resilience when errors are injected into application execution, accounting for the slowdowns due to the resilience mechanisms as well as rollbacks due to error detections.
+
+\sui{(Describe Fig.2 when faults are injected)}
+
+%Each routine's graph shows $\frac{\texttt{Time with Errors}}{\texttt{Time No Errors}}$, where the \texttt{Time with Errors} is average increase execution time of a resilient or non-resilient version of the routine when running on the same range of inputs as above, and \texttt{Time No Errors} is the execution time of the non-resilient version on the same inputs when no errors are injected.
+%The graph for each routine reports this metric for all of the above input sizes, error injection rates between 1e-8 and \sui{(1e-8 is the highest error rate that we can have for most runs to complete) and 1e-10 (is the lowest error rate we would observe errors from mosts of the runs) } \greg{confirm!} and all resilience configurations (no resilience and resilience with detection thresholds between 1e-6 and 1e-8), as the average time of \greg{???} runs \sui{100 runs per configuration}.
 The data points within each figure are organized horizontally according to the average number of errors injected into runs with the given input size and at the given injection rate.
@@
 
-The data shows \greg{???} \sui{Soft errors may cause the routines to run longer due to rollback in attempt to correct the errors.
+\sui{These are about Figure 2}The data shows \greg{???} \sui{Soft errors may cause the routines to run longer due to rollback in attempt to correct the errors.
 The higher the error rate, the more likely rollback is to occur.
@@
 \includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp2_1_Example.png}
-\caption{Error magnitude histogram showing the effectiveness of the error detector \greg{Why is FFT not improving with tighter threshold?
-Why is MM differently vulnerable than SYRK?} \sui{Reason for FFT: 1.
-Errors smaller than a certain value cannot be captured by the checker, even at low error rates, so this should be the fundamental limitation of the checker.
-2.
-We actually are seeing reduction in large errors, but those large errors account for such a small fraction that they are not obviously visible on the graph.
-For the difference between MM and SYRK: MM is more vulnerable therefore the difference between ``corrected'' and ``not corrected'' would look more noticable.}}
+\caption{Error magnitude histogram showing the effectiveness of the error detector}
 \label{fig:algo_err_dist}

### 2013-01-22 paper.tex
Fig 7, Ref to sigsetjmp

@@
 Configurations \texttt{Ckpt} and \texttt{NoChkpt} denote whether checkpointing was used and configurations \texttt{Re} and \texttt{NoRe} denote whether replication was used.
-\greg{What checkpointing period and replication properties?} \sui{Checkpointing: every 1/100/1000/10000/100000 calls to RK4 stepping function.
-Replication: No replication, using 1 pair of replicas and using maximum number of replicas.}
+We use the following checkpointing intervals: every 1/100/1000/10000/100000 calls to RK4 stepping function and the following replication configurations: No replication, 1 pair of replicas for each variable inthe source code and 1 pair of replicas for each occurrance of each variable in the source code.
 
@@
 
-Figure~\ref{fig:rk_routine_detect_ovhd} shows the overhead of employing each of the four RK resilience techniques when no errors are injected.
-\greg{This should be the same format as Figure~\ref{fig:routine_all_ovhd}}
+Figure~\ref{fig:rk_routine_detect_ovhd} shows the overhead of employing each of the four RK resilience techniques when no errors are injected. \sui{Marc: please explain Fig 5}
 
@@
 
-Figure~\ref{fig:rk_routine_exp_exec} shows the RK's overhead when errors are injected, using the same format as Figure~\ref{fig:routine_exp_exec}.
+Figure~\ref{fig:rk_routine_exp_exec} shows the RK's overhead when errors are injected.%, using the same format as Figure~\ref{fig:routine_exp_exec}.
 Although the \texttt{NoRe\_NoCkpt} configuration has the lowest overhead when only a few errors are injected (low error rate or small execution times), as the number of errors per run increase this configuration quickly becomes extremely slow because an error is detected in almost every run.

### 2013-01-22 paper.tex
Marc's changes

@@
 More specifically, these configurations can be used to produce a wide range of overheads and result accuracy levels and to tune the probabilistic confidence level that the given overhead or accuracy target will be met.
-\greg{Summarize real results} \sui{A tighter algorithmic checker bounds can be enabled at low error rates to completely eliminate erroneous results at a lower overhead, while a more lenient algorithmic checker would allow the user to produce more acceptable results in face of a higher error rate (such as 1e-09).
-Through high and low error rates, we correct more than 50\% of segmentation faults and reduce the error in the outputs by magnitudes by paying lower than 20\% overheads for linear algebra, FFT and FIR filter generation and a reasonable overhead for the integrator.}
+%\greg{Summarize real results} 
+%\sui{A tighter algorithmic checker bounds can be enabled at low error rates to completely eliminate erroneous results at a lower overhead, while a more lenient algorithmic checker would allow the user to produce more acceptable results in face of a higher error rate (such as 1e-09)}
+For the three considered scientific codes, we correct more than 50\% of segmentation faults and we reduce the error in the outputs by one order of magnitude. The overhead is always lower than 20\%.
 
@@
 
-Checkpoint data is protected using two different mechanisms.
+Checkpoint data is protected using two different mechanisms:
 Pointers and scalars are replicated such that the checkpoint contains two copies of each.
-Arrays of floating point numbers are protected via a block-checksum-based error correcting code..
-By treating the input array as a matrix and comparing row sums and column sums of the submatrices of that matrix, we are able to detect the existence of errors and fix a fraction of the errors.
-This algorithm takes in an array of numbers, divide it into blocks of $N \cdot N$ numbers and for each block it computes and stores the row sums and column sums that are used for recovery.
-The protection mechanism has one parameters,\texttt{N}, the size of the error correcting sum block.
+Arrays of floating point numbers are protected via a block-checksum-based error correcting approach:
+By treating the input array as a matrix and comparing row sums and column sums of the submatrices of that matrix, it detects the existence of errors and fixes them. 
+This algorithm takes an array of numbers as input, divides it into blocks of $N \cdot N$ numbers and for each block it computes and stores the row sums and column sums, which are used for recovery.
 
@@
 %\vspace{-5pt}
-\caption{Overhead of resilience techniques when errors are not injected \greg{Change y-axis name to say overhead rather than ratio}}
+\caption{Overhead of resilience techniques when errors are not injected}
 \label{fig:routine_detect_ovhd}
@@
 FFT ran on inputs vectors with 128K to 4M entries and RK ran on an integral with intervals $[0, t_1]$ where $t_1 \in \{10, 15, 20, 25, 30, 50, 75, 100\}$.
-The ODE system consists of a second-order nonlinear Van der Pol oscillator equation $x''(t)+\mu x'(t)(x(t)^2-1)+x(t)=0$.
+The ODE system consists of a second-order nonlinear Van der Pol oscillator equation.
 As our experiments show, the cost of algorithmic error detection is low and drops with increasing input sizes.

### 2013-01-22 paper.tex
Changed Figs 9~11; Marc's changes

@@
 %\vspace{-5pt}
-\caption{Overhead of resilience techniques when no errors are injected and when errors are injected (points).}
+\caption{Overhead of resilience techniques. Results when no errors are injected are shown in the bars while overheads when errors are injected are expressed by using points.}
 \label{fig:routine_all_ovhd}
@@
 
-Figure~\ref{fig:routine_all_ovhd} shows the overhead of all resilience techniques with no fault injection.
-The data shows that checkpointing consumes as much, if not more, time than error detection but that this cost grows more slowly than the routines' computational cost.
-In routines MV and RK the inputs are sufficiently small that checkpointing has negligible cost.
-Howveer, because MV runs quickly and because RK takes many checkpoints ($9000 \cdot input size)$), the cost of saving processor state is small but not negligible.
-
-%Finally, Figure~\ref{fig:routine_exp_exec} documents the overhead of resilience when errors are injected into application execution, accounting for the slowdowns due to the resilience mechanisms as well as rollbacks due to error detections.
-
-\sui{(Describe Fig.2 when faults are injected)}
-
-%Each routine's graph shows $\frac{\texttt{Time with Errors}}{\texttt{Time No Errors}}$, where the \texttt{Time with Errors} is average increase execution time of a resilient or non-resilient version of the routine when running on the same range of inputs as above, and \texttt{Time No Errors} is the execution time of the non-resilient version on the same inputs when no errors are injected.
-%The graph for each routine reports this metric for all of the above input sizes, error injection rates between 1e-8 and \sui{(1e-8 is the highest error rate that we can have for most runs to complete) and 1e-10 (is the lowest error rate we would observe errors from mosts of the runs) } \greg{confirm!} and all resilience configurations (no resilience and resilience with detection thresholds between 1e-6 and 1e-8), as the average time of \greg{???} runs \sui{100 runs per configuration}.
+Figure~\ref{fig:routine_all_ovhd} shows in the bars the overhead of all resilience techniques with no fault injection.
+The data shows that checkpointing routine's input (input backup bar) consumes more time than algorithm specific data corruption mechanisms (checker bar).
+In all cases, the overheads decrease as the input set size increses.
+The same figure shows in the points the computational overhead when errors are injected. 
 The data points within each figure are organized horizontally according to the average number of errors injected into runs with the given input size and at the given injection rate.
-\greg{Do the error counts include errors after rollbacks?} \sui{No, because the error count is the number of errors for Non-FT runs with errors.
-Since there is no FT and no checkpointing in these configurations, the number doesn't include errors after rollbacks.
-}
-This average error count is shown on the x-axis.
-All runs with the same resilience configuration are joined together into a line to highlight its properties.
-
-\sui{These are about Figure 2}The data shows \greg{???} \sui{Soft errors may cause the routines to run longer due to rollback in attempt to correct the errors.
-The higher the error rate, the more likely rollback is to occur.
-For example, the Cholesky Decomposition routine with input size 100, when exposed to an error rate of 1e-06, would run on average twice as long as when no fault is injected.
-That means in this case, every one of the runs tend to have 100\% overhead.
-In some cases the algorithm would give up retrying after the fifth time because it is unable to recover from the corrupt inputs.
-For example, for FFT exposed at a fault rate of 1e-08, the average overhead, as computed by dividing the time spent in the fault-tolerance codes by the baseline running time, decreases, because more runs tend to early-terminate when too many elements in the input have been corrupted.}
-
-\sui{I will add a graph showing how many of the inputs have been corrupted and failed to recover. This is an important issue for FFT.}
+In general, those overheads also decrease when the input set size increases.
 
@@
 We now evaluate the effectiveness of our algorithmic detectors when errors are injected.
-Figures~\ref{fig:algo_err_dist} and \ref{fig:algo_err_heatmap} show the fraction of entries in the outputs of each routine (y-axis) that have an error of a given magnitude (x-axis) as error magnitudes span from 1e-14 to 1e-4.
-Figure~\ref{fig:algo_err_dist} is a focused view on few representative cases (each routine run on a single input size and error injection rate) across different resilience configurations: no error detection as well as detection thresholds of 1e-6, 1e-7 and 1e-8.
+Figures~\ref{fig:algo_err_dist} show the fraction of entries in the outputs of each routine (y-axis) that have an error of a given magnitude (x-axis) as error magnitudes span from 1e-14 to 1e-4.
+This figure is a focused view on few representative cases (each routine run on a single input size and error injection rate) across different resilience configurations: no error detection as well as detection thresholds of 1e-6, 1e-7 and 1e-8.
 As the error magnitude increases the fraction of output entries with errors of that magnitude drops.
@@
 
-Figure~\ref{fig:algo_err_heatmap} shows the same information but for all experimental configurations: (i) input sizes of n=100 to 500 for linear algebra, 512K-8M size inputs for FFT and \greg{???} \sui{10 to 100} for RK; (ii) error injection rates of 1e-7 to 1e-10.
+Figure~\ref{fig:algo_err_heatmap} shows the same information but for all experimental configurations: (i) input sizes of n=100 to 500 for linear algebra, 512K-8M size inputs for FFT and 10 100 for RK; (ii) error injection rates of 1e-7 to 1e-10.
 The graph is a grid, with results for different routines listed vertically and different resilience configurations (no detection and different detection thresholds) listed horizontally.
@@
 
-\marc{The y-axis should show just the number of errors injected. We can say that we have used several fault injection rates and sizes to get those numbers. Right now the picture is confusing.}
-\sui{How does this look?}
-
-Figure~\ref{fig:algo_err_heatmap} shows the same information across the entire experimental space.
-The y-axis includes all algorithms, input sizes and error rates.
-The x-axis includes several segments for different degrees of error checking: none, as well as checking with thresholds \greg{???} \sui{1e-06, 1e-07 and 1e-08}.
+%We dont have the following figure and there is no time to generate it. 
+%Figure~\ref{fig:algo_err_heatmap} shows the same information across the entire experimental space.
+%The y-axis includes all algorithms, input sizes and error rates.
+%The x-axis includes several segments for different degrees of error checking: none, as well as checking with thresholds \greg{???} \sui{1e-06, 1e-07 and 1e-08}.
 In each segment the x-axis spans over error magnitudes from \greg{???} \sui{1e-14} to \greg{???} \sui{1e-01}.
-The box associated with each combination of the four parameters the figure indicates the fraction of outputs that have an error of this magnitude by the shade of gray it is colored with, as indicated in the legend.
-The data shows that the accurate of application results is directly related to the number of errors that impact a given application run, rather than their rate.
-Further, it shows that a reduction of detection threshold is consistently effective at reducing the fraction of errors in each algorithm's output.
+%The box associated with each combination of the four parameters the figure indicates the fraction of outputs that have an error of this magnitude by the shade of gray it is colored with, as indicated in the legend.
+%The data shows that the accurate of application results is directly related to the number of errors that impact a given application run, rather than their rate.
+%Further, it shows that a reduction of detection threshold is consistently effective at reducing the fraction of errors in each algorithm's output.
 %However, the effectiveness of algorithmic checks has a limit, with CD and FFT showing moderate error vulnerability even at threshold 1e-8.
-\sui{(Rephrase) However, the algorithmic checkers cannot cover 100\% of the errors for CD, FFT and FIR, mainly fixing errors of large magnitudes which turned out to be effective in real applications (refer to the imperfect run rate plot for DRC.)}
+%\sui{(Rephrase) However, the algorithmic checkers cannot cover 100\% of the errors for CD, FFT and FIR, mainly fixing errors of large magnitudes which turned out to be effective in real applications (refer to the imperfect run rate plot for DRC.)}
 %MM, MV and SYRK are significantly less vulnerable across the full range of error magnitudes.
-\sui{(Rephrase) The errors shown in MM, MV and SYRK are much more noticably fixed.}
-\greg{should reference vulnerability numbers more specifically}
+%\sui{(Rephrase) The errors shown in MM, MV and SYRK are much more noticably fixed.}
+%\greg{should reference vulnerability numbers more specifically}
 
@@
 
-\greg{For each experiment we need to include a discussion of what it means for how these checkers behave within the full-application runs.}
+%\greg{For each experiment we need to include a discussion of what it means for how these checkers behave within the full-application runs.}
 
-\sui{The checkers have performed well in reducing erroneous results from the application outputs.
+%\sui{The checkers have performed well in reducing erroneous results from the application outputs.
 Checkers for linear algebra are useful in removing errors from Lasso, as one could see from the results.}
@@
 Configurations \texttt{Ckpt} and \texttt{NoChkpt} denote whether checkpointing was used and configurations \texttt{Re} and \texttt{NoRe} denote whether replication was used.
-We use the following checkpointing intervals: every 1/100/1000/10000/100000 calls to RK4 stepping function and the following replication configurations: No replication, 1 pair of replicas for each variable inthe source code and 1 pair of replicas for each occurrance of each variable in the source code.
+We consider checkpointing periods equal to 1, $10^2$, $10^3$, $10^4$ and $10^5$ calls to the RK4 stepping function and the following replication configurations: No replication, 1 pair of replicas for each variable in the source code and 1 pair of replicas for each occurrance of each variable in the source code.
 
@@
 
-\begin{figure}[ht!]
-%\vspace{-20pt}
-\centering
-\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Overall_Breakdown_RK4}
-%\vspace{-5pt}
-\caption{Overhead of resilience techniques for RK when errors are not injected}
-\label{fig:rk_routine_detect_ovhd}
-\end{figure}
+%\begin{figure}[ht!]
+%%\vspace{-20pt}
+%\centering
+%\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Overall_Breakdown_RK4}
+%%\vspace{-5pt}
+%\caption{Overhead of resilience techniques for RK when errors are not injected}
+%\label{fig:rk_routine_detect_ovhd}
+%\end{figure}
 
-Figure~\ref{fig:rk_routine_detect_ovhd} shows the overhead of employing each of the four RK resilience techniques when no errors are injected. \sui{Marc: please explain Fig 5}
+%Figure~\ref{fig:rk_routine_detect_ovhd} shows the overhead of employing each of the four RK resilience techniques when no errors are injected.
+We use here the same format as in figure\ref{fig:routine_all_ovhd}.
+%%MARC: I decided to remove the above figure because it doesnt make sense to me and the same information is displayed in the following figure with more readibility.
 
@@
 
-Figure~\ref{fig:rk_routine_exp_exec} shows the RK's overhead when errors are injected.%, using the same format as Figure~\ref{fig:routine_exp_exec}.
+Figure~\ref{fig:rk_routine_exp_exec} shows the RK's overhead when 0 up to 10 errors per run are injected.%, using the same format as Figure~\ref{fig:routine_exp_exec}.
 Although the \texttt{NoRe\_NoCkpt} configuration has the lowest overhead when only a few errors are injected (low error rate or small execution times), as the number of errors per run increase this configuration quickly becomes extremely slow because an error is detected in almost every run.
@@
 These results are based on a fault injection campaign that includes a range of fault injection rates, input sizes and configurations of resilience mechanisms, with 100 application runs for each combination of parameters.
-This corresponds to a total of \greg{???} \sui{66100 for lasso, 43200 for DRC, 102000 for Hattrick, as counted by my scripts.} error injection experiments.
+This corresponds to a total of 66100 for lasso, 43200 for DRC, 102000 for Hattrick error injection experiments.
 
@@
 
-To evaluate the behavior of these applications in a broad range of error environments we considered error injection rates that range from 1e-16 \greg{???}\sui{Yes.
-At least this is the argument I sent to the injector.} errors per cycle of the LLVM virtual machine to 1e-6 errors per cycle.
+To evaluate the behavior of these applications in a broad range of error environments we considered several error injection rates.
 For reference, assuming a typical FIT of 100-100,000 for individual processing cores (FIT = failure in one billion hours of operation and this range has been experimentally observed for commodity electronics~\cite{mem_errors:2010, DRAMerr:2009}), an HPC system with one million cores will experience .1-100 errors per hour, which is equivalent to 3e-11 - 3e-14 errors per processor cycle on a 1Ghz processor.

### 2013-01-22 paper.tex
Redrawn figures 12~15
Marc's changes

@@
 
-To evaluate the behavior of these applications in a broad range of error environments we considered several error injection rates.
-For reference, assuming a typical FIT of 100-100,000 for individual processing cores (FIT = failure in one billion hours of operation and this range has been experimentally observed for commodity electronics~\cite{mem_errors:2010, DRAMerr:2009}), an HPC system with one million cores will experience .1-100 errors per hour, which is equivalent to 3e-11 - 3e-14 errors per processor cycle on a 1Ghz processor.
+To evaluate the behavior of these applications in a broad range of error environments we consider several error injection rates.
+For reference, assuming a typical FIT of 100-100,000 for individual processing cores (FIT = failure in one billion hours of operation and this range has been experimentally observed for commodity electronics~\cite{mem_errors:2010}), an HPC system with one million cores will experience .1-100 errors per hour, which is equivalent to 3e-11 - 3e-14 errors per processor cycle on a 1Ghz processor.
 The range of error injection rates used used in our experiments were chosen such that (i) on average one or more errors were injected in the executions of our target applications and (ii) at least some application runs completed without crashing.
@@
 Our fault injection experiments ran each application configuration to completion, recording whether it completed and if so, the magnitude of the error.
-We calculate the execution time of each configuration, accounting for application rollbacks via the formula $\frac{p_c*t_c}{1-p_f}$, where $p_c$ is the probability of a run completing, $t_c$ is the average completion time of such runs and $p_f$ is the probability of a run failing to complete\greg{is this right}.
-\sui{I don't see anything wrong}
-Figures~\ref{fig:Lasso_EstdCost}, \ref{fig:DRC_EstdCost} and \ref{fig:Hattrick_EstdCost} show the expected execution time, relative to the execution time of the non-fault tolerant version of the application when no errors are injected.
+We calculate the execution time of each configuration, accounting for application rollbacks via the formula $\frac{t_c}{1-p_f}$, $t_c$ is the average completion time of such runs and $p_f$ is the probability of a run failing to complete.
+Figures~\ref{fig:Lasso_EstdCost}, \ref{fig:DRC_EstdCost} and \ref{fig:Hattrick_EstdCost} show the expected execution time, relative to the execution time of the non-fault tolerant version of the application when no errors are injected, of the non-fault tolerant version (red) and the optimal fault tolerant version (blue).
+The optimal fault tolerant version is the one that suffers from smaller overhead among the ones that provide more resilience enhancement. 
+To select the optimal one we consider error checker tresholds from $10^{-4}$ to $10^{-7}$ for ADDR, from $10^{-6}$ to $3 \cdot 10^{-8}$ for DRC and from using one pair of copies for each variable occurence to using a different pair of replicas for each variable occurrence for Hattrick.
 The graphs from left to right correspond to different application input sizes.
 Within each graph the x-axis corresponds to error injection rates, while the y-axis shows the relative execution time, both logarithmic.
-Each graph shows data for the non-fault tolerant configuration of the application, connecting all the data points with a line.
-They also show the relative execution times for all the configurations of the fault tolerant version, with the line connecting the results for the configuration that is optimal for each input size and fault rate.
 
@@
 
-\greg{Sui, please confirm that the following is true}
-The data shows that the non-fault tolerant ADDR and Hattrick shows little slowdown as a result of injected faults.
-This is because these applications spend most of its time in linear algebra and differential equation routines which are numerically-intensive but perform relatively few memory manipulations or pointer accesses.
-As such, until it reaches error rates of 1e-9 and above ADDR does not see a significant number of aborts due to memory errors.
-The result is that the addition of resilience techniques does little to improve ADDR's performance. \sui{The addition of resilience techniques greatly improve its accuracy}
+Figure~\ref{fig:Lasso_EstdCost} shows that the non-fault tolerant ADDR slows down from 2x to 9x, depending on the input set size, as a result of the injected faults.
+The optimal fault tolerance version the algorithm shows degradation that are around one half of the non-faul tolerant one, which implies that our technique improve ADDR's resilience by 50\%.
 
-In contrast, since DRC performs many more integer operations and memory copies, it is more vulnerable to errors that corrupt the values of pointers or cause loads and stores to target invalid memory locations.
+Figure~\ref{fig:DRC_EstdCost} shows results regarding the DRC application. 
+Since it performs many more integer operations and memory copies than ADDR, it is more vulnerable to errors that corrupt the values of pointers or cause loads and stores to target invalid memory locations.
 Such errors cause memory segmentation faults, which force DRC to restart its execution and thus incur a very high overhead. 
-In contrast, the localized rollback and correction provided by our checkpointing mechanisms ensures that each restart has a much lower cost and that a single error does not propagate past the end of a single call to FFT or FIR.
+That is the reason why the performance slowdown of its non-fault tolerant version is much larger than the ADDR one. 
+In contrast, the optimal fault tolerant version reduces almost completely the slowdown due to fault injection.
+
+Figure~\ref{fig:Hattrick_EstdCost} shows how the performance slowdown of the fault tolerance version it is very unstable but it can reach values close to 100x. 
+The performance of the optimal fault tolerant version is consistently much better.
 
@@
 \vspace{-10pt}
-\caption{Execution time of ADDR relative to non-fault tolerant fault-free execution time}
+\caption{Execution time of ADDR relative to non-fault tolerant fault-free execution time. Results for the non-fault tolerant version (red) and the optimal fault tolerant version (blue) are shown.}
 \vspace{-10pt}
@@
 \vspace{-10pt}
-\caption{Execution time of DRC relative to non-fault tolerant fault-free execution time \greg{Need confidence intervals} \sui{Added.}}
+\caption{Execution time of DRC relative to non-fault tolerant fault-free execution time. Results for the non-fault tolerant version (red) and the optimal fault tolerant version (blue) are shown.}
 \vspace{-10pt}
@@
 \vspace{-10pt}
-\caption{Execution time of Hattrick relative to non-fault tolerant fault-free execution time \greg{Need confidence intervals} \sui{Added.}}
+\caption{Execution time of Hattrick relative to non-fault tolerant fault-free execution time. Results for the non-fault tolerant version (red) and the optimal fault tolerant version (blue) are shown.}
 \vspace{-10pt}

### 2013-01-22 paper.tex
Removed 4 figures from Sec 5

@@
 
-Figure~\ref{fig:algo_err_heatmap} shows the same information but for all experimental configurations: (i) input sizes of n=100 to 500 for linear algebra, 512K-8M size inputs for FFT and 10 100 for RK; (ii) error injection rates of 1e-7 to 1e-10.
+Figure~\ref{fig:algo_err_heatmap} shows the same information but for all experimental configurations: (i) input sizes of n=100 to 500 for linear algebra, 512K-8M size inputs for FFT and 10, 20, 30, 50 and 100 for RK; (ii) error injection rates of 1e-7 to 1e-10.
 The graph is a grid, with results for different routines listed vertically and different resilience configurations (no detection and different detection thresholds) listed horizontally.
@@
 \centering
-\includegraphics[height=5in]{figs/Lasso_ImperfectRate.png}
-\includegraphics[height=5in]{figs/Lasso_ImperfectRate_log.png}
+%\includegraphics[height=5in]{figs/Lasso_ImperfectRate.png}
+\includegraphics[height=3in]{figs/Lasso_ImperfectRate_log.png}
 \caption{Probability of not satisfying the user-provided error bound for ADDR \greg{why do the probabilities of the no-FT version not change as the error threshold changes?
@@
 \centering
-\includegraphics[height=3in]{figs/DRC_ImperfectRate.png}
+%\includegraphics[height=3in]{figs/DRC_ImperfectRate.png}
 \includegraphics[height=3in]{figs/DRC_ImperfectRate_log.png}
@@
 \centering
-\includegraphics[height=3in]{figs/Hattrick_ImperfectRate.png}
+%\includegraphics[height=3in]{figs/Hattrick_ImperfectRate.png}
 \includegraphics[height=3in]{figs/Hattrick_ImperfectRate_log.png}

### 2013-01-22 paper.tex
Marc's additions

@@
 %\vspace{-5pt}
-\caption{Overhead of resilience techniques. Results when no errors are injected are shown in the bars while overheads when errors are injected are expressed by using points.}
+\caption{Overhead of resilience techniques. Results when no errors are injected are shown.}
 \label{fig:routine_all_ovhd}
@@
 In all cases, the overheads decrease as the input set size increses.
-The same figure shows in the points the computational overhead when errors are injected. 
-The data points within each figure are organized horizontally according to the average number of errors injected into runs with the given input size and at the given injection rate.
-In general, those overheads also decrease when the input set size increases.
 
@@
 \centering
-\includegraphics[width=2.00\columnwidth]{figs/Lasso_EstdCost.png}
+\includegraphics[height=2.75in]{figs/Lasso_EstdCost.png}
 \vspace{-10pt}
@@
 \centering
-\includegraphics[width=2.00\columnwidth]{figs/DRC_EstdCost.png}
+\includegraphics[height=2.75in]{figs/DRC_EstdCost.png}
 \vspace{-10pt}
@@
 \centering
-\includegraphics[width=2.00\columnwidth]{figs/Hattrick_EstdCost.png}
+\includegraphics[height=2.75in]{figs/Hattrick_EstdCost.png}
 \vspace{-10pt}
@@
 This section evaluates the distribution of errors in the resulting outputs.
-The outputs of the different applications are as follows:
-\begin{itemize}
-    \item ADDR: vector that holds the solution to the linear system.
-    \item Hattrick: vector that holds the state of the n-body system at the final application time-step, with 6 numbers per body.
-    \item DRC: a Pulse Code Modulation-formatted file, which is a sequence of numbers.
-\end{itemize}
+The outputs of are, for ADDR, vectors that holds the solution to the linear system.
+For Hattrick, they are vectoris that hold the state of the n-body system at the final application time-step, with 6 numbers per body
+and and for DRC they are a Pulse Code Modulation-formatted file, which is a sequence of numbers.
 Since the outputs of each routine can be represented as vectors we calculate the error in the outputs by comparing the vector $v_e$ returned by each routine when errors are injected to the correct vector $v_c$ computed during a fault-free run and computing the L2 norm of the difference: $\sqrt{\frac{\sum_{i} (v_c-v_e)^2}{\left| v_c \right|}}$.
 
-Figures~\ref{fig:Lasso_ImperfectRate}, \ref{fig:DRC_ImperfectRate} and \ref{fig:Hattrick_ImperfectRate} show the errors in application results from a user-centric perspective.
-Given a user-specified error magnitude that must be satisfied by application results these figures plot the fraction of application runs for which this target was satisfied
+Figures~\ref{fig:Lasso_ImperfectRate}, \ref{fig:DRC_ImperfectRate} and \ref{fig:Hattrick_ImperfectRate} show the errors in ADDR, DRC and Hattrick results from a user-centric perspective.
+They plot the fraction of application runs for which a numerical error equal or smaller than $10^{-8}$ is achieved.
 From left to right each figure shows graphs for different input sizes and from top to bottom it shows results for different error magnitude bounds.
-The x-axis in each graph shows the fault injection rate (logarithmic) while the y-axis shows the fraction of runs where the error target was satisfied (linear).
+The x-axis in each graph shows the fault injection rate while the y-axis shows the fraction of runs where the error target was satisfied.
+Both axis are expressed in logarithmic scale.
 Each graph shows data for the non-fault tolerant version of the application (all points connected by the line) as well as all the the fault tolerant versions (optimal configurations connected by a line).
@@
 
-\greg{Sui, please confirm that the following is true} \sui{Yes.
-ADDR is sensitive to errors, and even a small number of error would cause noticable errors in the results.
-Comparing the leftmost data points on the ``probability of not satisfying'' graphs of the 3 applications.}
-The data shows that the probability that the non-fault tolerant version of ADDR satisfies the user-provided error bound is low, for bounds randing from 1e-10 to 1e-4.
-In contrast, the fault tolerant version of the application is significantly more successful, especially for error rates of 1e-12 and lower (\greg{need log plot to show the low accuracy rate more precisely} \sui{Changed to log plot.}).
-\greg{Can mention the fact that the fault tolerant versions are most successful for loose error bounds and small inputs but the success rate is poor for the 10x input.
-Why?} \sui{I know why.
-After inspecting Lasso's code, it turns out Lasso contains TWO branches of code: one on ``fat'' matrices and the other on ``skinny'' matrices.
-Only 10x input triggers the ``fat matrix'' code and inputs 20 through 100 would trigger the ``skinny'' code.
-The two versions of code are not the same therefore they would be vulnerable to soft errors to different degrees.}
-
-\greg{I won't comment on the DRC errors because Sui is currently trying to improve them.
-Right now we're not being very successful.
-However, since DRC is a single precision code, we should really be considering much larger error bounds, like 1e-6 or 1e-4}.
-\sui{Expanding the ranges makes sense.
-Plus, I think the range for Hattrick can be as tight as 1e-20.
-I have used 1e-17, 1e-14 and 1e-10 in the updated visualization.}
-
-\greg{Its hard to comment on Hattrick until we have a version of the plots with a logscale y-axis since it'll tell us exactly how successful the fault tolerant versions of the code are.
-Also, what are the relative execution times of these input files?
-This will give us an idea of how vulnerable they are.} \sui{Input 1=55s; Input 2=155s; Input 3=82s; Input 4=120s }
+Figures~\ref{fig:Lasso_ImperfectRate}, \ref{fig:DRC_ImperfectRate} and \ref{fig:Hattrick_ImperfectRate} show that the probability that the non-fault tolerant version of each one the three application satisfies the user-provided error bound is low especially for fault injection rates $10^{-9}$ to $10^{-11}$.
+In contrast, the fault tolerant version of the application is significantly more successful, especially for error rates of $10^{-9}$ and lower.
 
@@
 \includegraphics[height=3in]{figs/Lasso_ImperfectRate_log.png}
-\caption{Probability of not satisfying the user-provided error bound for ADDR \greg{why do the probabilities of the no-FT version not change as the error threshold changes?
-Also, it may be good to show fewer columns but add graphs where the y-scale is a log-plot} \sui{Added a log plot.}}
+\caption{Probability of not satisfying the user-provided error bound for ADDR. Results for the non-fault tolerant version (red) and the optimal fault tolerant version (blue) are shown.}
 \label{fig:Lasso_ImperfectRate}
@@
 \includegraphics[height=3in]{figs/DRC_ImperfectRate_log.png}
-\caption{Probability of not satisfying the user-provided error bound for DRC \greg{why do the probabilities of the no-FT version not change as the error threshold changes?
-Also, it may be good to show fewer columns but add graphs where the y-scale is a log-plot} \sui{Added a log plot (on the right).}}
+\caption{Probability of not satisfying the user-provided error bound for DRC. Results for the non-fault tolerant version (red) and the optimal fault tolerant version (blue) are shown.}
 \label{fig:DRC_ImperfectRate}
@@
 \includegraphics[height=3in]{figs/Hattrick_ImperfectRate_log.png}
-\caption{Probability of not satisfying the user-provided error bound for Hattrick \greg{why do the probabilities of the no-FT version not change as the error threshold changes?
-Also, it may be good to show fewer columns but add graphs where the y-scale is a log-plot} \sui{The change is small therefore not visible on a linear scale.
-Switching to log plot revealed those tiny differences.}}
+\caption{Probability of not satisfying the user-provided error bound for Hattrick. Results for the non-fault tolerant version (red) and the optimal fault tolerant version (blue) are shown.}
 \label{fig:Hattrick_ImperfectRate}
@@
 
-\subsection{Overall Evaluation}
-\label{sec:eval:overall}
+%\subsection{Overall Evaluation}
+%\label{sec:eval:overall}
 
-%Experiment: Final graph that shows slowdown for each desired probability of perfect results.
+%%Experiment: Final graph that shows slowdown for each desired probability of perfect results.
 
@@
 \label{sec:summary}
+
+In this paper we show that it is possible to effectively protect three real scientific applications from a wide range of faults by properly combining three different fault tolerant mechanisms:       
+Algorithmic error checks, replication of critical data structures and checkpoint-restart of individual routines.
+We demonstrate that very significant improvements can be achieved in terms of reduction of performance slowdown and output numerical accuracy when scientific problems face error injection rates from $3 \cdot 10^{-11}$ and $3 \cdot 10^{-14}$ per processor cycle in a 1Ghz, which are the expected rates in future high performance computing systems.
+This work demonstrates that current scientific application can be safely executed on such systems if the programmers combine resilience techniques in the way it is described.
 

### 2013-01-23 paper.tex
Fix figures; add shade under figure 9~14

@@
 \includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp2_2_Heatmap_Error_ConfSpace.png}
+\begin{tabular}{|p{0.5in}|p{0.7in}|p{0.7in}|p{0.7in}|}
+\hline
+Routine & Threshold 1 & Threshold 2 & Threshold 3 \\ 
+\hline
+\hline
+CD & 1e-06 & 1e-07 & 1e-08 \\
+\hline
+FFT & 1e-06 & 1e-09 & 1e-14 \\
+\hline
+FIR & 1e-06 & 1e-09 & 1e-14 \\
+\hline
+MM & 1e-06 & 1e-07 & 1e-08 \\
+\hline
+MV & 1e-06 & 1e-07 & 1e-08 \\
+\hline
+SYRK & 1e-06 & 1e-07 & 1e-08 \\
+\hline
+\end{tabular}
 \caption{Error magnitude histogram of the entire configuration space}
@@
 Figures~\ref{fig:algo_err_dist} show the fraction of entries in the outputs of each routine (y-axis) that have an error of a given magnitude (x-axis) as error magnitudes span from 1e-14 to 1e-4.
-This figure is a focused view on few representative cases (each routine run on a single input size and error injection rate) across different resilience configurations: no error detection as well as detection thresholds of 1e-6, 1e-7 and 1e-8.
+This figure is a focused view on few representative cases (each routine run on a single input size and error injection rate) across different resilience configurations: no error detection as well as three error detection thresholds.
 As the error magnitude increases the fraction of output entries with errors of that magnitude drops.
@@
 Configurations \texttt{Ckpt} and \texttt{NoChkpt} denote whether checkpointing was used and configurations \texttt{Re} and \texttt{NoRe} denote whether replication was used.
-We consider checkpointing periods equal to 1, $10^2$, $10^3$, $10^4$ and $10^5$ calls to the RK4 stepping function and the following replication configurations: No replication, 1 pair of replicas for each variable in the source code and 1 pair of replicas for each occurrance of each variable in the source code.
+We consider checkpointing periods equal to 1, $10^2$, $10^3$, $10^4$ and $10^5$ calls to the RK4 stepping function and the following replication configurations: No replication, 1 pair of replicas for each variable in the source code and 1 pair of replicas for each occurrance of each variable in the source code (these configurations correspond to ``No repilcation'', ``Medium replication'' and ``Maximum replication'' in figure \ref{fig:rt_algo_err_heatmap}).
 
@@
 Figure~\ref{fig:rk_routine_exp_exec} shows the RK's overhead when 0 up to 10 errors per run are injected.%, using the same format as Figure~\ref{fig:routine_exp_exec}.
-Although the \texttt{NoRe\_NoCkpt} configuration has the lowest overhead when only a few errors are injected (low error rate or small execution times), as the number of errors per run increase this configuration quickly becomes extremely slow because an error is detected in almost every run.
+Although the \texttt{NoRe\_NoCkpt} configuration has the lowest overhead when only a few errors are injected (low error rate or small execution times), as the number of errors per run increases this configuration quickly becomes extremely slow because an error is detected in almost every run.
 Since each such detection requires a restart of the routine, at higher error rates this causes RK to restart many times before it completes.
@@
 
-Figure~\ref{fig:rt_algo_err_heatmap} shows the fraction of errors in RK's outputs for the four resilience configurations, in the same format as Figure~\ref{fig:algo_err_heatmap}.
-It shows that the use of either resilience approaches significantly reduces RK's vulnerability to soft errors and the use of both techniques is even more effective.
+Figure~\ref{fig:rt_algo_err_heatmap} shows the fraction of errors in RK's outputs for various checkpointing and replication configurations, in the same format as Figure~\ref{fig:algo_err_heatmap}.
+It shows that the use of either resilience approaches reduces RK's vulnerability to soft errors, preserving the correctness of its output, and the use of both techniques is even more effective.
 Figure~\ref{fig:rk_effectiveness} shows this from a different perspective: the fraction of RK runs where the root-mean-squared error of the output vector is below 1e-7 (y-axis), relative to the number of error injections per run (x-axis).
@@
 Our fault injection experiments ran each application configuration to completion, recording whether it completed and if so, the magnitude of the error.
-We calculate the execution time of each configuration, accounting for application rollbacks via the formula $\frac{t_c}{1-p_f}$, $t_c$ is the average completion time of such runs and $p_f$ is the probability of a run failing to complete.
+We calculate the execution time of each configuration, accounting for application rollbacks via the formula $\frac{t_c}{1-p_f}$, where $t_c$ is the average completion time of such runs and $p_f$ is the probability of a run failing to complete.
 Figures~\ref{fig:Lasso_EstdCost}, \ref{fig:DRC_EstdCost} and \ref{fig:Hattrick_EstdCost} show the expected execution time, relative to the execution time of the non-fault tolerant version of the application when no errors are injected, of the non-fault tolerant version (red) and the optimal fault tolerant version (blue).
@@
 %\includegraphics[height=5in]{figs/Lasso_ImperfectRate.png}
-\includegraphics[height=3in]{figs/Lasso_ImperfectRate_log.png}
+\includegraphics[height=2.75in]{figs/Lasso_ImperfectRate_log.png}
 \caption{Probability of not satisfying the user-provided error bound for ADDR. Results for the non-fault tolerant version (red) and the optimal fault tolerant version (blue) are shown.}
@@
 %\includegraphics[height=3in]{figs/DRC_ImperfectRate.png}
-\includegraphics[height=3in]{figs/DRC_ImperfectRate_log.png}
+\includegraphics[height=2.75in]{figs/DRC_ImperfectRate_log.png}
 \caption{Probability of not satisfying the user-provided error bound for DRC. Results for the non-fault tolerant version (red) and the optimal fault tolerant version (blue) are shown.}
@@
 %\includegraphics[height=3in]{figs/Hattrick_ImperfectRate.png}
-\includegraphics[height=3in]{figs/Hattrick_ImperfectRate_log.png}
+\includegraphics[height=2.75in]{figs/Hattrick_ImperfectRate_log.png}
 \caption{Probability of not satisfying the user-provided error bound for Hattrick. Results for the non-fault tolerant version (red) and the optimal fault tolerant version (blue) are shown.}

### 2013-01-23 paper.tex
Changes in text; change the facet names on figs 9~14

@@
 We focus on a specific application that uses the ADDR method to solve Lasso problems, which are defined as minimizing the function $\frac{1}{2} \left|| Ax - b \right||_2^2 + \lambda*\left|| x \right||_1$.
-In our experiments we consider linear systems of size $\{40, 80, 120, ..., 4000\} \times 500$ that are generated by sampling a normal distribution.
+It is parallelized using MPI and our experiments focus on 4-processors runs of ADDR which computes linear systems of size $\{40, 80, 120, ..., 4000\} \times 500$ (40 to 4000 measurements with 500 variables each) and in the experiments below are denoted $S40$, $S80$, ..., $S4000$.
+The values in $A$ and $b$ are generated by sampling a normal distribution with a mean of 0.
 According to our experiments, the running time of this solver is not dependent on the kind of distribution the inputs are sampled from.
@@
 
-The DRC program generates filters for HiFi audio systems to compensate for the reflection of sounds in a room, using impulse response measurements of the audio equipment and considering the positions of the listeners.
+DRC is a sequential program generates filters for HiFi audio systems to compensate for the reflection of sounds in a room, using impulse response measurements of the audio equipment and considering the positions of the listeners.
 The inputs are stored in Pulse Code Modulation (PCM) format, which is an array of 32-bit floating point numbers representing the samples at each sampling time step and the computations are performed using 32-bit precision.
-Although the algorithm is divided into many phases the most computationally intensive step consists of the GSL's implementation of the Fast Fourier Transform and an DRC-internal implementation of Finite Impulse Response filter generation.
+Although the algorithm is divided into many phases the most computationally intensive step consists of the GSL's implementation of the Fast Fourier Transform and a DRC-internal implementation of Finite Impulse Response filter generation.
 This program accepts user-controllable configurations to decide the width and number of iterations of FFT transforms it uses to perform various passes on the input file, such as windowing, deconvolution, pre- and post-filtering.
-The input used in this paper is an audio file of size 768 kilobytes.
+The input used in this paper is an audio file of size 768 kilobytes, which is sampled at the rate of 30Khz, 40Khz, ..., 90Khz.
 
@@
 \label{sec:apps:hattrick}
-The Hattrick application simulates the motion of bodies under the effects of gravity using 64-bit precision.
+Hattrick is a sequential application simulates the motion of bodies under the effects of gravity using 64-bit precision.
 It is designed to help discover extra-solar planets by inferring their existence from Transit Timing Variations, where the difference between the times when a planet passes in front of its host star is used to infer the existence and properties of other planets in the system.
-This program spends most of its execution time in the GSL solver for Ordinary Differential Equations to solver the system's equations of motion.
-Input files are four different planetary systems.
-The first, third and fourth input files simulate one star and one planet and the second simulates one star and two planets.
-The third and fourth inputs are the same planetary system with different lengths of simulation time.
+This program spends most of its execution time in the GSL solver for Ordinary Differential Equations to solve the system's equations of motion.
+A given input is described using three parameters: $P$=number of planets, $T$=amount of time to simulate, and $A$=the accuracy target.
+In our experiments we considered the following four inputs: $P2T2090A15$, $P2T3090A15$, $P2T4090A15$ and $P3T2090A11$, where an $A15$ and $A11$ denote accuracy targets of $1e-15$ and $1e-11$, respectively.
 
@@
 
-In this paper we consider a comprehensive resilience strategy that combines error detection and recovery.
+In this paper we consider a comprehensive resilience strategy that combines various types of error detection and recovery.
 Error detection techniques are discussed in Section~\ref{sec:res_tech:err_det}, with Section~\ref{sec:res_tech:err_det:algo} focusing on algorithmic techniques to detect errors in the numeric portions of application state, while Section~\ref{sec:res_tech:err_det:mem} focuses on protection of data structures.
@@
 
+\vspace{-5pt}
 \subsubsection{Algorithmic Detection}
@@
 Each algorithm used by the above application was enhanced with an algorithm-specific checker that validated its results by exploiting some algorithmic identity.
-Each checker computes some aggregate property of the algorithm's results using two different algorithm to produce vectors $v$ and $v'$ and their relative difference by more than some threshold $\tau$, the checker declares that an error has occurred.
+Each checker computes some aggregate property of the algorithm's results using two different algorithms to produce vectors $v$ and $v'$.
+If their relative difference by more than some threshold $\tau$, the checker declares that an error has occurred.
 The relative difference is computed as $\frac{\left|| v-v' \right||}{max(\left||v\right||, \left||v'\right||)}$.
@@
 
+\vspace{-5pt}
 \paragraph{Matrix-matrix multiplication (MMM)}
-The MMM operation $A \cdot B$ is checked using a matrix vector multiplication (MVM), making use of this identity: $(A \cdot B) \cdot x = A \cdot (B \cdot x)$, where $x$ is an error-checking vector.
-Both sides compute the sums of the columns of matrix $A \cdot B$ using a single matrix-vector multiplication on the left and two such operations on the right.
+The operation $A \cdot B$ is checked using a matrix vector multiplication (MVM), via the identity: $(A \cdot B) \cdot x = A \cdot (B \cdot x)$, where $x$ is an error-checking vector.
+Since we set the check vector $x$ to be all 1s this is equivalent to both sides computing the sums of the columns of matrix $A \cdot B$ using a single MVM on the left and two MVMs on the right.
 Most errors in the computation of $A \cdot B$ will affect the sums of its columns and will thus be detected.
 This check is efficient because the checker is asymptotically faster than the MMM algorithm, with MVM taking $O(n^2)$ time and MMM $O(n^3)$ time.
-Since this is dense martix matrix multiplication, we set the check vector $x$ to contain all 1's.
 
+\vspace{-5pt}
 \paragraph{Matrix-vector multiplication (MVM)}
 The MVM algorithm $Ax=b$ is checked using a similar identity $(x^TA)x = x^T(Ax) = x^Tb$.
-The product $x^TA$ is a checksum of the columns of matrix $A$ and when $x$ is the vector of 1's, the value used in our experiments, it is the sum of $A$'s columns.
+We again set $x$ to be the vector of 1's, making the product $x^TA$ equivalent to sum of the columns of matrix $A$.
 Although the computation of $x^TA$ involves additions rather than multiplications, the complexity of computing $x^TA$ is $O(n^2)$, the same as the original MVM operation.
@@
 
+\vspace{-5pt}
 \paragraph{Symmetric Rank-K update (SYRK)}
-The SYRK update is a special case of matrix-matrix multiplication $B = \alpha A \cdot A^T + \beta B$, where an $n \times k$ matrix $A$ is multiplied by itself to produce an $n \times n$ matrix with rank $k$, which is incremented into the $nxn$ matrix $B$.
+The SYRK update is a special case of matrix-matrix multiplication $B = \alpha A \cdot A^T + \beta B$, where an $n \times k$ matrix $A$ is multiplied by itself to produce an $n \times n$ matrix with rank $k$, which is incremented into the $n \times n$ matrix $B$.
 Its results are checked using the same algorithm as MMM by checking the identity $B x = \alpha A \cdot (A^Tx) + \beta B x$, where $x$ is the vector of all 1's.
@@
 
+\vspace{-5pt}
 \paragraph{Cholesky Decomposition (CD)}
 In this decomposition matrix $A$ is decomposed into $L \cdot L^T$ where $L$ is a lower-triangular matrix with a positive diagonal.
-This routine is checked just like MMM by checking the identity $Ac = L \cdot (L^T x)$, which is cheaper than the $O(n^3)$ complexity of the deterministic CD algorithm.
+This routine is checked just like MMM by checking the identity $Ax = L \cdot (L^T x)$, which is cheaper than the $O(n^3)$ complexity of the deterministic CD algorithm.
 The GSL implements an iterative algorithm that runs faster than $O(n^3)$ time but as shown in the experiments below, it is significantly slower than our checker.
 
+\vspace{-5pt}
 \paragraph{Fast Fourier Transform (FFT)}
@@
 We check the results of FFT by using Parseval's theorem, which states that $\sum_{n=0}^{N-1} \left| x[n] \right|^2 = \frac{1}{N} \sum_{k=0}^{N-1} \left| X[k] \right|^2$, where $x$ is the original function and $X$ is its transform.
+Intuitively it means that the energy of the original function is preserved by the transform.
 GSL implements optimized radix 2, 3, 4, 5, 6 and 7 subtransform routines as well as a general radix-N FFT routine.
@@
 Therefore, GSL has a scheduling algorithm that factorizes the length of the input FFT, running the subtransforms whenever possible, using the fallback general radix-N routine only for the remaining factors.
-The theorem applies to FFT of any radix.
-This means that the total energy of the original function is preserved by the transform and provides an $O(n)$ time check for the $O(n log(n))$ FFT algorithm.
+The theorem applies to FFT of any radix and provides an $O(n)$ time check for the $O(n log(n))$ or $O(n^2)$ FFT algorithms.
 
+\vspace{-5pt}
 \paragraph{Finite Impulse Response Filter Generation (FIR)}
-The FIR filter generation algorithm used in DRC generates a series of samples over the $sinc$ function $sinc(x)=\frac{sin(x)}{x} (sinc(x)=1\ when\ x\ is\ 0)$ and modulates it with a Blackman window.
+The FIR filter generation algorithm used in DRC generates a series of samples over the function $sinc(x)=\frac{sin(x)}{x}$
+%(\text{if}~x \ne 0)$ and $sinc(x)=1 (\text{if}~x=0)$ 
+and modulates it with a Blackman window.
 This function has the property that $\int_{-\infty}^{\infty} sinc(x)dx = 1$.
@@
 
+\vspace{-5pt}
 \paragraph{Runge-Kutta PDE Solver (RK)}
-This algorithm uses the 4-th order Runge-Kutta method for solving Ordinary Differential Equations of the form $\frac{dy}{dx} = f(y, x)$.
-It works by advancing the variable $x$ by steps of size $h$ and using estimates of the derivative $\frac{dy}{dx}$ at each point $x$ to find the value of $y$ at the next point $x+h$.
-The RK method in GSL uses adaptive step-size control where it simultaneously performs the RK method with two step sizes $h$ and $\frac{h}{2}$ (more precise).
-If the difference between the results obtained with the two step sizes are larger than a high water-mark threshold, the algorithm switches to the smaller step size to maintain accuracy and resumes.
-If it is smaller than a low water-mark threshold, the algorithm switches to a larger step size.
+RK is a 4-th order Runge-Kutta method for solving Ordinary Differential Equations of the form $\frac{dy}{dx} = f(y, x)$.
+It advances the variable $x$ by steps of size $h$ and uses estimates of the derivative $\frac{dy}{dx}$ at each point $x$ to find the value of $y$ at the next point $x+h$.
+The implementation of RK in GSL uses adaptive step-size control where it simultaneously performs the RK method with two step sizes $h$ and $\frac{h}{2}$ (more precise).
+If the difference between the results obtained with the two step sizes are larger than a threshold $\tau$, the algorithm switches to the smaller step size to maintain accuracy and resumes.
+If it is smaller than $\frac{\tau}{2}$, the algorithm switches to a larger step size.
 Because soft errors will usually cause significant differences between the two computations, the adaptive step size control algorithm is naturally fault tolerant.
 It will react to such errors by rolling back, temporarily increasing its step size and then reverting to the original step size when it observes that this is safe.
-As such, we utilize this algorithm to protect the computations within individual RK4 steps from soft errors.
-However, computations in user-provided function derivative functions are not checked since their invariants are unknown.
+As such, we utilize this algorithm to protect the computations within individual RK steps from soft errors.
+However, computations in user-provided function derivative functions are not checked since their invariants are notknown.
 
@@
 This means that there is a very high probability that the corruption of a given pointer will cause it to point to an unallocated address.
-Once this address is accessed this violation will be detected by the memory protection hardware and communicated to the application via a Segmentation Fault or Bus Error signal.
+Once this address is accessed the resulting violation will be detected by the memory protection hardware and communicated to the application via a Segmentation Fault or Bus Error signal.
 This mechanism is cheap but suffers from a long delay between an error and its detection.
 
-An alternative mechanism is to replicate key application variable and to compare the values of the replicas on each access to the variable.
+An alternative mechanism is to replicate key application variables and to compare the values of the replicas on each access to the variable.
 While more expensive, this approach is effective if the number of key variables is small and the cost of rolling back computation is high.
@@
 
-\subsection{Recovery via Checkpoint-Restart and Replication}
+\subsection{Recovery via Checkpoint-Restart and \\Replication}
 \label{sec:res_tech:cr}
@@
 Rollbacks are implemented via the \texttt{sigsetjmp}/\texttt{siglongjmp} \cite{sigsetjmp:1997} routines.
-\texttt{sigsetjmp} record the application's execution state (registers and stack pointer) at the time of the checkpoint.
-On error detection \texttt{siglongjmp} returns the application to the same execution point, unrolling any function calls between the checkpoint and the error detection.
+\texttt{sigsetjmp} records the application's execution state (registers and stack pointer) at the time of the checkpoint.
+On the detection of an error \texttt{siglongjmp} returns the application to the same execution point, unrolling any function calls between the checkpoint and the error detection.
 This method is useful for rolling back to checkpoints recorded in the same function or in a caller function.
@@
 Pointers and scalars are replicated such that the checkpoint contains two copies of each.
-Arrays of floating point numbers are protected via a block-checksum-based error correcting approach:
+Arrays of floating point numbers are protected via a block-checksum-based error correcting approach.
 By treating the input array as a matrix and comparing row sums and column sums of the submatrices of that matrix, it detects the existence of errors and fixes them.
-This algorithm takes an array of numbers as input, divides it into blocks of $N \cdot N$ numbers and for each block it computes and stores the row sums and column sums, which are used for recovery.
+This algorithm takes an array of numbers as input, divides it into blocks of $N \times N$ numbers and for each block it computes and stores the row and column sums, which are used for recovery.
 
@@
 As such, we checkpoint the state of non-RK routines immediately after the call to the routine and include in this checkpoint the routine's arguments and contents of the floating point input buffers.
-The one exception is the FIR routine, which takes no input and thus requires only the processor's state to be checkpointed..
+The one exception is the FIR routine, which takes no input and thus requires only the processor's state to be checkpointed.
 Because RK's checkpoints are very small and it is more sensitive to errors, it is checkpointed periodically during its execution with a fixed number of steps between checkpoints.
@@
 This corresponds to a total of 66100 for lasso, 43200 for DRC, 102000 for Hattrick error injection experiments.
+It was performed on compute nodes with two 2.33 GHz dual-core Xeon E5345 processors with 4GB Ram and RHEL 4 OS.
 
@@
 
-\begin{figure*}[ht!]
+\begin{figure}[ht!]
 \centering
@@
 \label{fig:Hattrick_EstdCost}
-\end{figure*}
+\end{figure}
 
@@
 \label{fig:Hattrick_ImperfectRate}
-\end{figure*}
+\end{figure}
 

### 2013-01-23 paper.tex
Describing replication levels 4.3.2

@@
 Configurations \texttt{Ckpt} and \texttt{NoChkpt} denote whether checkpointing was used and configurations \texttt{Re} and \texttt{NoRe} denote whether replication was used.
-We consider checkpointing periods equal to 1, $10^2$, $10^3$, $10^4$ and $10^5$ calls to the RK4 stepping function and the following replication configurations: No replication, 1 pair of replicas for each variable in the source code and 1 pair of replicas for each occurrance of each variable in the source code (these configurations correspond to ``No repilcation'', ``Medium replication'' and ``Maximum replication'' in figure \ref{fig:rt_algo_err_heatmap}).
+We consider checkpointing periods equal to 1, $10^2$, $10^3$, $10^4$ and $10^5$ calls to the RK4 stepping function and the following replication configurations: No replication, 1 pair of replicas for each pointer to the vector elements of the ODE system in the source code and 1 pair of replicas for each occurrance of each pointer in the source code (these configurations correspond to ``No repilcation'', ``Medium replication'' and ``Maximum replication'' in figure \ref{fig:rt_algo_err_heatmap}).
+In the latter two options, the replicas are used to correct the pointers before each access to it in the source code of the integrator.
 

### 2013-01-23 paper.tex
Fix typo error: occurrance -> occurrence

@@
 Configurations \texttt{Ckpt} and \texttt{NoChkpt} denote whether checkpointing was used and configurations \texttt{Re} and \texttt{NoRe} denote whether replication was used.
-We consider checkpointing periods equal to 1, $10^2$, $10^3$, $10^4$ and $10^5$ calls to the RK4 stepping function and the following replication configurations: No replication, 1 pair of replicas for each pointer to the vector elements of the ODE system in the source code and 1 pair of replicas for each occurrance of each pointer in the source code (these configurations correspond to ``No repilcation'', ``Medium replication'' and ``Maximum replication'' in figure \ref{fig:rt_algo_err_heatmap}).
+We consider checkpointing periods equal to 1, $10^2$, $10^3$, $10^4$ and $10^5$ calls to the RK4 stepping function and the following replication configurations: No replication, 1 pair of replicas for each pointer to the vector elements of the ODE system in the source code and 1 pair of replicas for each occurrence of each pointer in the source code (these configurations correspond to ``No repilcation'', ``Medium replication'' and ``Maximum replication'' in figure \ref{fig:rt_algo_err_heatmap}).
 In the latter two options, the replicas are used to correct the pointers before each access to it in the source code of the integrator.

### 2013-10-14 accept.tex
2013-10-13

%%
%% This is file `accept.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% acmconf.dtx  (with options: `accept,head,body')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from accept.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file acmconf.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% to test the checksum, uncomment    \OnlyDescription
%% in the driver
%%\fi
%%
%% \CharacterTable
%%  {Upper-case    \A\B\C\D\E\F\G\H\I\J\K\L\M\N\O\P\Q\R\S\T\U\V\W\X\Y\Z
%%   Lower-case    \a\b\c\d\e\f\g\h\i\j\k\l\m\n\o\p\q\r\s\t\u\v\w\x\y\z
%%   Digits        \0\1\2\3\4\5\6\7\8\9
%%   Exclamation   \!     Double quote  \"     Hash (number) \#
%%   Dollar        \$     Percent       \%     Ampersand     \&
%%   Acute accent  \'     Left paren    \(     Right paren   \)
%%   Asterisk      \*     Plus          \+     Comma         \,
%%   Minus         \-     Point         \.     Solidus       \/
%%   Colon         \:     Semicolon     \;     Less than     \<
%%   Equals        \=     Greater than  \>     Question mark \?
%%   Commercial at \@     Left bracket  \[     Backslash     \\
%%   Right bracket \]     Circumflex    \^     Underscore    \_
%%   Grave accent  \`     Left brace    \{     Vertical bar  \|
%%   Right brace   \}     Tilde         \~}

\documentclass[box,accept]{acmconf}
\CopyrightText{\copyright ACM 2000, ....., used with the \texttt{box} option.}
\IfFileExists{graphicx.sty}{\usepackage{graphicx}}{}
\ConferenceName{1. Conference on Designing a \LaTeX2e Class for
  Typesetting ACM Papers, Hawaii 2000}
\ConferenceShortName{CONF-2000}

\def\XX{More text should follow, but keep in mind that a limit of 6
  pages has been set, including figures and references.  More text
  should follow, but keep in mind that a limit of 6 pages has been
  set, including figures and references.  More text should follow, but
  keep in mind that a limit of 6 pages has been set, including figures
  and references.  More text should follow, but keep in mind that a
  limit of 6 pages has been set, including figures and references.
  \par
}

\begin{document}
\date{31. December 1999}
\title{A New Intuitionistic Proof of Usability\\
       of the Recommended Style File for the ACM Conference Papers}
\author{\Author{J\"urgen Vollmer\thanks{Happy \LaTeX{}ing}}\\
         \Address{Karlsruhe}\\
         \Email{Juergen.Vollmer@acm.org}\\
         \and
         \Author{Mickey Mouse}\\
         \Address{Enthausen University}\\
         \Email{Mickey.Mouse@entenhausen.org}
       }
\maketitle

\begin{abstract}
  This document demonstrates how to use the \LaTeX2e \verb|acmconf|
  class by exhibiting itself as an example.  You are expected to be
  familiar with~\cite{Lam94}.  The best way to use this file is to use
  it as a template, i.e., replace the prose in it by your
  own\footnote{And may use footnotes.}.
\end{abstract}

\begin{keywords}
\LaTeX2e-class, ACM proceedings
\end{keywords}

\section{Introduction}
To understand this file read the \emph{source} and not the typeset
version.  If you are reading this in the typeset version you might as
well stop --- it is not supposed to make sense.

\section{The Story Begins\ldots}
A real article is supposed to have some deep results and good
explanations.  That, however, is your job and not mine so you should
replace this text with something more appropriate\footnote{Another a
  footnote}..

\section{Some often used \LaTeX\ commands}

\subsection{\texttt{emph}, etc.}
Text may be set as \emph{emph}.\\
Text may be set as \texttt{texttt}.\\
Text may be set as \underline{unterline}.\\
Text may be set as \textbf{textbf}.\\
Text may be set as \textrm{textrm}.\\
Text may be set as {\tiny tiny}.\\
Text may be set as {\scriptsize scriptsize}.\\
Text may be set as {\footnotesize footnotesize}.\\
Text may be set as {\normalfont normalsize}.\\
Text may be set as {\large large}.\\
Text may be set as {\Large Large}.\\
Text may be set as {\LARGE LARGE}.\\
Text may be set as {\huge huge}.\\
Text may be set as {\Huge Huge}.\\
Text may have$^{\textrm{super}}$ and$_{\textrm{sub}}$scripts.

\subsection{\texttt{itemize}}
\begin{itemize}
\item More text should follow, but keep in mind that a limit of 6
     pages has been set, including figures and references.  More text
     should follow, but keep in mind that a limit of 6 pages has been
     set, including figures and references.
\item More text should follow, but keep in mind that a limit of 6
     pages has been set, including figures and references.  More text
     should follow, but keep in mind that a limit of 6 pages has been
     set, including figures and references.
\end{itemize}

\subsection{\texttt{enumerate}}
\begin{enumerate}
\item More text should follow, but keep in mind that a limit of 6
     pages has been set, including figures and references.  More text
     should follow, but keep in mind that a limit of 6 pages has been
     set, including figures and references.
\item More text should follow, but keep in mind that a limit of 6
     pages has been set, including figures and references.  More text
     should follow, but keep in mind that a limit of 6 pages has been
     set, including figures and references.
\end{enumerate}

\subsection{\texttt{description}}
\begin{description}
\item[Foo] More text should follow, but keep in mind that a limit of 6
     pages has been set, including figures and references.  More text
     should follow, but keep in mind that a limit of 6 pages has been
     set, including figures and references.
\item[Bar] More text should follow, but keep in mind that a limit of 6
     pages has been set, including figures and references.  More text
     should follow, but keep in mind that a limit of 6 pages has been
     set, including figures and references.
\end{description}

\subsection{\texttt{center} and \texttt{tabular}}
\begin{center}
\begin{tabular}{|l|c|r|}\hline
left     & center   & right    \\\hline\hline
AAAAAAAA & BBBBBBBB & CCCCCCCC \\
AAAAAAAA & BBBBBBBB & CCCCCCCC \\\cline{3-3}
AAAAAAAA & BBBBBBBB & CCCCCCCC \\\cline{2-2}
AAAAAAAA & BBBBBBBB & CCCCCCCC \\\cline{1-2}
AAAAAAAA & BBBBBBBB & CCCCCCCC \\\hline
AAAAAAAA & BBBBBBBB & CCCCCCCC \\\hline
1          & \multicolumn{2}{|c|}{2} \\\hline
\end{tabular}
\end{center}

\subsection{\texttt{figure} and Postscript pictures}
Have a look to to figure~\ref{fig-1} and~\ref{fig-2}.

\begin{figure}
\hrule
Nice Postscript, isn't it?
\begin{center}
\IfFileExists{graphicx.sty}{
  \includegraphics{body.eps}
}{
  Sorry, package \texttt{graphicx} not present.
}
\end{center}

Same, a little bit smaller:
\begin{center}
\IfFileExists{graphicx.sty}{
  \includegraphics[scale=.5]{body.eps}
  }{
  Sorry, package \texttt{graphicx} not present.
}
\end{center}
\caption{\label{fig-1}This is a nice floating figure}
\hrule
\end{figure}

\begin{figure*}
\hrule
This figure uses both columns, using \texttt{figure*}
\begin{center}
\IfFileExists{graphicx.sty}{
  \includegraphics[scale=.5]{body.eps}
  \hspace{1cm}
  \includegraphics[scale=.5]{body.eps}
}{
  Sorry, package \texttt{graphicx} not present.
}
\end{center}
\caption{\label{fig-2}This is a nice floating figure}
\hrule
\end{figure*}

\section{The Story Continues 1}

This is a \verb+\section+.

\XX\XX

\subsection{The Story Continues 2}

This is a \verb+\subsection+.

\XX\XX

\subsubsection{The Story Continues 3}

This is a \verb+\subsubsection+.

\XX\XX

\subsubsubsection{The Story Continues 4}

This is a \verb+\subsubsubsection+.

\XX\XX

\subsubsubsubsection{The Story Continues 5}

This is a \verb+\subsubsubsubsection+.

\XX\XX

\paragraph{The Story Continues 6}

This is a \verb+\paragraph+.
\XX\XX

\subparagraph{The Story Continues 7}
This is a \verb+\subparagraph+.
\XX\XX\XX

\section{Conclusion}
The end, at last!  In this example there really are no results or
points to summarize but I trust your article has more food for though
and thus will need a conclusion.

\appendix
\section{Appendices}
If you have any, appendices might go here.  Note that appendices
should not be used to circumvent the word count limit.

This is "doing it by hand" --- you might be better off using BibTeX.

\begin{thebibliography}{X}
\bibitem[1]{Lam94} Leslie Lamport: {\em \LaTeX, A Document
    Preparation System,} Addison Wesley~1994.
\end{thebibliography}
\IfPrepare{
  \tableofcontents
  \listoffigures
  \listoftables
}{}
\end{document}
\endinput
%%
%% End of file `accept.tex'.

### 2013-10-14 paper.tex
2013-10-13

--- 2013-01-23 paper.tex
@@
 \usepackage{url}
+\usepackage{ulem}
 \usepackage{graphicx}
@@
 \newcommand{\sui}[1]{%
-  \textcolor{green}{SC - #1}
+  \textcolor{green}{SC-#1}
 }
@@
 
-\title{Comprehensive Algorithmic Resilience for Scientific Applications }
-\author{Sui Chen,
+\title{Comprehensive Algorithmic Resilience for \\Numeric Applications}
+%\author{Sui Chen,
 %\affil{Louisiana State University}
-Greg Bronevetsky, Marc Casas,
+%Greg Bronevetsky, Marc Casas,
 %\affil{Lawrence Livermore National Laboratory}
-Lu Peng
+%Lu Peng
 %\affil{Louisiana State University}
-}
+%}
 
@@
 
-Algorithmic resilience can be very complex to deploy in applications because a comprehensive resilience strategy must address all the types of errors that an application may experience.
-This includes corruptions to numerical data as well as control information and data structures.
-Full protection requires the combination of multiple techniques including algorithmic error detection, crash detection, replication of key control information as well as support for restart when errors are detected.
-Further, it is necessary to configure the properties of the resilience mechanisms to make sure that they are cost-effectively meeting the user's accuracy and confidence targets.
-This paper shows how all these issues can be resolved comprehensively in the context of three numeric applications.
+\sui{
+Algorithmic resilience exploiting the invariants of certain algorithms is an old technique that has been around for many years. The invariants can serve as
+correctness unit tests during development and turned off during run-time in favor of speed, but under the context of the upcoming exascale reliability challenges,
+they can usually serve as useful fault checkers. However, unless other issues including input protection, control information
+and critical data structure protection are resolved will algorithmic invariants help detecting and correcting errors.
+% can be very complex to deploy in applications because a comprehensive resilience strategy must address all the types of errors that an application may experience.
+%This includes corruptions to numerical data as well as control information and data structures.
+%Full protection requires the combination of multiple techniques including algorithmic error detection, crash detection, replication of key control information as well as support for restart when errors are detected.
+%Further, it is necessary to configure the properties of the resilience mechanisms to make sure that they are cost-effectively meeting the user's accuracy and confidence targets.
+Further, certain design choices in exploiting algorithmic resilience imply trade-offs for cost-effectiveness of the final resilience strategy.
+This paper presents our case study of exploiting algorithmic invariants for constructing comprehensive fault tolerance mechanisms and balancing performance and result accuracy in the context of large-scale computing.
+}
+%This paper shows how all these issues can be resolved comprehensively in the context of three numeric applications.
 
@@
 Although these techniques can be very efficient, most research work on such techniques has primarily focused on how their algorithmic invariants can be leveraged to enable accurate error detection.
-While this goal is important, it ignores the full complexity of the problem since it focuses on only a subset of the errors and does not consider combinations of different resilience techniques to efficiently and effectively address all the different types of faults.
+While this goal is important, it ignores the full complexity of the problem since it focuses on only a subset of the errors \sui{(algorithmic correctness)} and 
+does not consider combinations of different resilience techniques to efficiently and effectively address all the different types of faults.
 In reality a comprehensive solution must incorporate a range of additional techniques such the replication of key variables, detection of memory violations and localized rollback on detected errors, as well as techniques to configure these various resilience methods to achieve the best performance given the user's accuracy and confidence targets.
@@
 
-This paper presents a comprehensive algorithmic resilience approach that considers all of the above issues.
+This paper presents a comprehensive algorithmic resilience approach \sui{that combines existing checkpointing and algorithmic checker techniques that considers all of the above issues and demonstrate {\bf with fault injection experiments and statistical analyses} that such approach is able to significantly reduce the side effects of soft errors on parallel programs at a  slight performance cost.}
 We demonstrate it on different types of numerical applications: the Alternating Direction Method of Multipliers (ADDR) for Lasso problems~\cite{lasso:2011}, the Hattrick gravitational simulation~\cite{hattrick:2012} and the Digital Room Correction (DRC) acoustic correction application~\cite{drc:2012}.
@@
 %\sui{A tighter algorithmic checker bounds can be enabled at low error rates to completely eliminate erroneous results at a lower overhead, while a more lenient algorithmic checker would allow the user to produce more acceptable results in face of a higher error rate (such as 1e-09)}
-For the three considered scientific codes, we correct more than 50\% of segmentation faults and we reduce the error in the outputs by one order of magnitude. The overhead is always lower than 20\%.
+%For the three considered scientific codes, we correct more than 50\% of segmentation faults and we reduce the error in the outputs by one order of magnitude. The overhead is always lower than 20\%.
 
@@
 
+\sui{As a metric for quantifying errors in output caused by soft faults, we perturb inputs to numeric applications ...}
+
 \section{Target Applications}
@@
 We focus on a specific application that uses the ADDR method to solve Lasso problems, which are defined as minimizing the function $\frac{1}{2} \left|| Ax - b \right||_2^2 + \lambda*\left|| x \right||_1$.
-It is parallelized using MPI and our experiments focus on 4-processors runs of ADDR which computes linear systems of size $\{40, 80, 120, ..., 4000\} \times 500$ (40 to 4000 measurements with 500 variables each) and in the experiments below are denoted $S40$, $S80$, ..., $S4000$.
+It is parallelized using MPI and our experiments focus on 4-processors runs of ADDR which computes linear systems of size $\{800, 1200, ..., 3600\} \times 500$ (800 to 3600 measurements with 500 variables each) and in the experiments below are denoted $S800$, $S1200$, ..., $S3600$.
 The values in $A$ and $b$ are generated by sampling a normal distribution with a mean of 0.
@@
 
+%\subsection{Barnes Hut}
+%\label{sec:apps:barneshut}
+%\sui{
+%Barnes Hut is a parallel application for fast simulation of the motion of large amounts of bodies under the effects of gravity.
+%It is designed to divide the cubic simulation volume using a quad-tree and approximating particles in distant cells
+%by treating them as a large particle on the center of the cell's center of mass. 
+%This dramatically reduces the number of interactions that needed to compute and reduces time complexity to $O(nlogn)$.
+%When evaluating contribution of gravitational force on a certain point, a parameter $\theta$ dictates whether a cell is far enough from
+%this point.
+Only when $\theta$ is larger than the quotient of the size of the cell and the distance from the cell's center of mass to the point ${\lvert \vec{s} \rvert} / {%\lvert \vec{d} \rvert}$ is the approximation used.
+Consequently, the algorithm is approximate and $\theta$ determines its accuracy.
+We use input ......
+%}
+
 \section{Resilience Techniques}
@@
 Error detection techniques are discussed in Section~\ref{sec:res_tech:err_det}, with Section~\ref{sec:res_tech:err_det:algo} focusing on algorithmic techniques to detect errors in the numeric portions of application state, while Section~\ref{sec:res_tech:err_det:mem} focuses on protection of data structures.
-Section~\ref{sec:res_tech:cr} then considers recovery techniques based on both checkpoint-restart, which offers lower-overhead but higher-latency recovery, as well as replication of key variables, which reduces performance but enables fast recovery.
+Section~\ref{sec:res_tech:cr} then considers recovery techniques based on both checkpoint-restart, which offers lower-overhead but higher-latency recovery, as well as replication of key pointers, which reduces performance but enables fast recovery.
 Finally, Section~\ref{sec:res_tech:eval} presents an experimental evaluation of the performance and accuracy properties of the resilient versions of the major routines used by our target applications, under a wide range of inputs and error injection rates.
@@
 Most errors in the computation of $A \cdot B$ will affect the sums of its columns and will thus be detected.
-This check is efficient because the checker is asymptotically faster than the MMM algorithm, with MVM taking $O(n^2)$ time and MMM $O(n^3)$ time.
+This check is efficient because the checker is asymptotically faster than MMM, with MVM taking $O(n^2)$ time and MMM $O(n^3)$ time.
 
@@
 %Show a separate plot for a few different magnitudes.
-\subsubsection{Memory Fault Detection}
-\label{sec:res_tech:err_det:mem}
 
-Detecting errors in application data structures requires checkers dedicated for this purpose.
-Depending on the algorithm's properties different error detectors are appropriate.
-In our work we considered the use of access protection hardware and replication of key data.
-
-Since modern systems use 64-bit addressing the range of values that can be assumed by a given pointer is much larger than the amount of memory that can ever be assigned to an application.
-This means that there is a very high probability that the corruption of a given pointer will cause it to point to an unallocated address.
-Once this address is accessed the resulting violation will be detected by the memory protection hardware and communicated to the application via a Segmentation Fault or Bus Error signal.
-This mechanism is cheap but suffers from a long delay between an error and its detection.
-
-An alternative mechanism is to replicate key application variables and to compare the values of the replicas on each access to the variable.
-While more expensive, this approach is effective if the number of key variables is small and the cost of rolling back computation is high.
-
-The low cost of hardware memory protection mechanisms means that they are deployed in protecting all our routines.
-In contrast, replication is only used to protect the RK routine because it is more sensitive to soft errors than the others and the low detection latency provided by replication ensures that any state corruptions are quickly corrected and have little impact on execution time.
-This is important for resilience as well as performance since slowdowns due to frequent rollbacks increase the total number of errors the application is exposed to.
+% 4.1.2 is way too verbose, vague and not understandable
+%\subsubsection{Memory Fault Detection}
+%\label{sec:res_tech:err_det:mem}
+%
+%Detecting errors in application data structures requires checkers dedicated for this purpose.
+%Depending on the algorithm's properties different error detectors are appropriate.
+%In our work we considered the use of access protection hardware and replication of key data.
+%
+%Since modern systems use 64-bit addressing the range of values that can be assumed by a given pointer is much larger than the amount of memory that can ever be assigned to %an application.
+%This means that there is a very high probability that the corruption of a given pointer will cause it to point to an unallocated address.
+%Once this address is accessed the resulting violation will be detected by the memory protection hardware and communicated to the application via a Segmentation Fault or %Bus Error signal.
+%This mechanism is cheap but suffers from a long delay between an error and its detection.
+%
+%An alternative mechanism is to replicate key application variables and to compare the values of the replicas on each access to the variable.
+%While more expensive, this approach is effective if the number of key variables is small and the cost of rolling back computation is high.
+%
+%The low cost of hardware memory protection mechanisms means that they are deployed in protecting all our routines.
+%In contrast, replication is only used to protect the RK routine because it is more sensitive to soft errors than the others and the low detection latency provided by %replication ensures that any state corruptions are quickly corrected and have little impact on execution time.
+%This is important for resilience as well as performance since slowdowns due to frequent rollbacks increase the total number of errors the application is exposed to.
 
@@
 \label{sec:res_tech:cr}
+Since bit-flip errors may corrupt the address of a load or store instruction and a bit-flip in a 64-bit address can easily become an invalid value pointing to
+unallocated memory, a Segmentation Fault or Bus Error can be easily detected by the operating system.\\
 
@@
 
-\begin{figure}[ht!]
-%\vspace{-20pt}
-\centering
-\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp1_linalg}
-\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp1_fft}
-%\vspace{-5pt}
-\caption{Overhead of resilience techniques when errors are not injected}
-\label{fig:routine_detect_ovhd}
-\end{figure}
+%\begin{figure}[ht!]
+%\centering
+%\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp1_linalg}
+%\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp1_fft}
+%\caption{Overhead of resilience techniques when errors are not injected}
+%\label{fig:routine_detect_ovhd}
+%\end{figure}
 
@@
 We evaluate the costs of the above algorithmic error checkers by measuring their costs and effectiveness when running at different error rates on different types of inputs.
-Figure \ref{fig:routine_detect_ovhd} measures the overheads of employing the algorithmic error detectors when no errors are injected, reporting the average of 100 runs.
-The linear algebra routines were executed on $nxn$ square matrices where $n$ varies from 100 to 1000.
-FFT ran on inputs vectors with 128K to 4M entries and RK ran on an integral with intervals $[0, t_1]$ where $t_1 \in \{10, 15, 20, 25, 30, 50, 75, 100\}$.
-The ODE system consists of a second-order nonlinear Van der Pol oscillator equation.
-As our experiments show, the cost of algorithmic error detection is low and drops with increasing input sizes.
+%Figure \ref{fig:routine_detect_ovhd} measures the overheads of employing the algorithmic error detectors when no errors are injected, 
+We report the averages of 100 runs.
+The linear algebra routines were executed on $n \times n$ square matrices where $n$ varies from 100 to 1000.
+FFT ran on inputs vectors with 128K to 16M entries and FIR on window widths of 1M to 8M.
+RK ran on an integral with intervals $[0, t_1]$ where $t_1 \in \{10, 15, 20, 25, 30, 50, 75, 100\}$. 
+Its ODE system is a second-order nonlinear Van der Pol oscillator equation.
+Our experiments show that the cost of algorithmic error detection is low (< 10\%) for small and drops with increasing input sizes.
 
@@
 
-Figure~\ref{fig:routine_all_ovhd} shows in the bars the overhead of all resilience techniques with no fault injection.
-The data shows that checkpointing routine's input (input backup bar) consumes more time than algorithm specific data corruption mechanisms (checker bar).
-In all cases, the overheads decrease as the input set size increses.
+Figure~\ref{fig:routine_all_ovhd} shows in the bars the overhead of all resilience techniques with no fault injection, separating into overheads due to algorithmic error detection, the checkpointing of routine inputs as well as checkpointing processor state.
+The data shows that checkpointing the input of routines consumes more time than algorithmic error detection and that the cost of checkpointing processor state is negligible. 
+In all cases, these overheads decrease as the input set size increases.
 
@@
 We now evaluate the effectiveness of our algorithmic detectors when errors are injected.
+In each case when an error is detected by our algorithmic detectors the routine's execution would restart from the beginning upto five restarts.
 Figures~\ref{fig:algo_err_dist} show the fraction of entries in the outputs of each routine (y-axis) that have an error of a given magnitude (x-axis) as error magnitudes span from 1e-14 to 1e-4.
@@
 
-Figure~\ref{fig:algo_err_heatmap} shows the same information but for all experimental configurations: (i) input sizes of n=100 to 500 for linear algebra, 512K-8M size inputs for FFT and 10, 20, 30, 50 and 100 for RK; (ii) error injection rates of 1e-7 to 1e-10.
-The graph is a grid, with results for different routines listed vertically and different resilience configurations (no detection and different detection thresholds) listed horizontally.
-The graph for each routine and resilience configuration presents the curves from Figure \ref{fig:algo_err_dist} as a heatmap.
-The y-axis shows different routine input sizes and error injection rates, sorted according to the average number of error injections per run, in decreasing order from top to bottom.
-The x-axis shows the different error magnitudes and the shade of each point is the fraction of the routine's entries that have a given error magnitude, with darker shades indicating more such entries.
+Figure~\ref{fig:algo_err_heatmap} shows the same information but for all experimental configurations: (i) input sizes as above (ii) error injection rates of 1e-7 to 1e-10.
+The graph is a grid, with results for different routines listed vertically and different resilience configurations (no detection and the different detection thresholds shown in the table at the bottom) listed horizontally.
+The graph for each routine and resilience configuration presents the curves from Figure~\ref{fig:algo_err_dist} as a heatmap.
+The y-axis shows different routine input sizes and error injection rates, sorted according to the average number of error injections per non-fault injected run, in decreasing order from top to bottom.
+The x-axis shows the different error magnitudes and the shade of each point is the fraction of the routine's entries that have a given error magnitude, with darker shades indicating more such entries (each shade corresponds to a fraction that is some power of 10).
 Thus, each row of Figure~\ref{fig:algo_err_heatmap} represents the same information as each curve in Figure~\ref{fig:algo_err_dist}, with the shades of the squares representing the same information as the height of the line's points.
+
+The data shows that in general as the error detection threshold drops there are fewer entries that have an error of a given magnitude.
+However, there are exceptions where the fraction of erroneous entries rises as the detection threshold shrinks (e.g. MMM and FT between NoFT and Threshold 1).
+This is caused by the fact that we restart each routine on an error detection, which can causes the routine's execution to be exposed to additional error injections.
+As the threshold is further tightened the amount of error in the output shrinks, indicating that error detection is ultimately effective if it is configured to be sufficiently accurate.
+Another phenomenon that is observed is that as the number of errors injected in a given run decreases the degree of error in routine outputs also generally decreases.
+However, this is not a fully consistent phenomenon, with some configurations that have more error injections being less vulnerable than others with a higher number of injections because the former had a higher injection rate but the latter a larger input size.
+This indicates that while the number of injected errors is strongly correlated with overall application vulnerability to errors, other properties such as input size also play a role.
 
@@
 Configurations \texttt{Ckpt} and \texttt{NoChkpt} denote whether checkpointing was used and configurations \texttt{Re} and \texttt{NoRe} denote whether replication was used.
-We consider checkpointing periods equal to 1, $10^2$, $10^3$, $10^4$ and $10^5$ calls to the RK4 stepping function and the following replication configurations: No replication, 1 pair of replicas for each variable in the source code and 1 pair of replicas for each occurrance of each variable in the source code (these configurations correspond to ``No repilcation'', ``Medium replication'' and ``Maximum replication'' in figure \ref{fig:rt_algo_err_heatmap}).
+We consider checkpointing periods equal to 1, $10^2$, $10^3$, $10^4$ and $10^5$ calls to the RK stepping function.
+Further, we consider the following replication configurations: (i) ``No replication'', (ii) `Medium replication'', where routine pointers are replicated but the replicas of each pointer are compared at a single code location where it is used, and (iii) ``Maximum replication'', where the comparison is performed on every pointer use.
 
@@
 
-Figure~\ref{fig:rk_routine_exp_exec} shows the RK's overhead when 0 up to 10 errors per run are injected.%, using the same format as Figure~\ref{fig:routine_exp_exec}.
+Figure~\ref{fig:rk_routine_exp_exec} shows the RK's overhead when between 0 and 10 errors per run are injected.%, using the same format as Figure~\ref{fig:routine_exp_exec}.
 Although the \texttt{NoRe\_NoCkpt} configuration has the lowest overhead when only a few errors are injected (low error rate or small execution times), as the number of errors per run increases this configuration quickly becomes extremely slow because an error is detected in almost every run.
@@
 In contrast, while the other configurations have a higher overhead for a few errors per run, their execution time scales much better with increasing error rates, making them significantly more practical.
-By varying the checkpointing interval and the degree to which various RK variables are replicated it is possible to achieve a wide range of design points that provide the best performance at various error rates.
+In particular, the overheads of the configurations with more replication and checkpointing scale better as the number of errors per run increase.
+By varying the checkpointing interval and the degree to which various RK pointers are replicated it is possible to achieve a wide range of design points that provide the best performance at various error rates.
 
@@
 
-Figure~\ref{fig:rt_algo_err_heatmap} shows the fraction of errors in RK's outputs for various checkpointing and replication configurations, in the same format as Figure~\ref{fig:algo_err_heatmap}.
-It shows that the use of either resilience approaches reduces RK's vulnerability to soft errors, preserving the correctness of its output, and the use of both techniques is even more effective.
-Figure~\ref{fig:rk_effectiveness} shows this from a different perspective: the fraction of RK runs where the root-mean-squared error of the output vector is below 1e-7 (y-axis), relative to the number of error injections per run (x-axis).
-It shows that RK configurations with higher level of resilience are able to achieve this accuracy target more frequently.
+Figure~\ref{fig:rt_algo_err_heatmap} shows the fraction of errors in RK's outputs for various checkpointing and replication configurations, in the same format and using the same experimental setup as in Figure~\ref{fig:algo_err_heatmap}.
+It shows that while the use replication reduces the amount of error in RK's outputs, increased checkpointing doesn't have a significant effect on this accuracy metric.
+Figure~\ref{fig:rk_effectiveness} evaluates the effect of checkpointing on accuracy from another perspective: the fraction of RK runs where the root-mean-squared error of the output vector is below 1e-7 (y-axis), relative to the number of error injections per run (x-axis).
+It shows that RK configurations with higher level of resilience and in particular, more frequent checkpoints, are able to achieve this accuracy target more frequently.
 This is primarily because by avoiding frequent rollbacks due to crashes these techniques reduce the total number of errors that are injected into the application run, which also reduces the probability that one such error will be missed by RK's algorithmic error detector.
@@
 These results are based on a fault injection campaign that includes a range of fault injection rates, input sizes and configurations of resilience mechanisms, with 100 application runs for each combination of parameters.
-This corresponds to a total of 66100 for lasso, 43200 for DRC, 102000 for Hattrick error injection experiments.
-It was performed on compute nodes with two 2.33 GHz dual-core Xeon E5345 processors with 4GB Ram and RHEL 4 OS.
+This corresponds to a total of 66,100 fault injection experiments for ADDR (4-process parallel), 43,200 for DRC (sequential), 102,000 for Hattrick (sequential).
+Our experiments were performed on compute nodes with two 2.33 GHz dual-core Xeon E5345 processors with 4GB Ram and RHEL 4 OS.
 
 A given application run may terminate successfully or crash due to an unrecoverable error.
-Unrecoverable errors occur when there is a segmentation fault in code outside our protected routines or if 5 rollbacks of a routine have been observed in a given run since such repeated rollbacks almost always indicate that the corruption in application state cannot be repaired.
+Unrecoverable errors occur when there is a segmentation fault in code outside our protected routines or if five rollbacks of a routine have been observed in a given run since such repeated rollbacks almost always indicate that the corruption in application state cannot be repaired.
 Since we assume that the user needs application results regardless of the error rate, if the application crashes it is restarted from the beginning as many times as needed for it to complete.
@@
 To evaluate the behavior of these applications in a broad range of error environments we consider several error injection rates.
-For reference, assuming a typical FIT of 100-100,000 for individual processing cores (FIT = failure in one billion hours of operation and this range has been experimentally observed for commodity electronics~\cite{mem_errors:2010}), an HPC system with one million cores will experience .1-100 errors per hour, which is equivalent to 3e-11 - 3e-14 errors per processor cycle on a 1Ghz processor.
+For reference, assuming a typical FIT of 100-100,000 for individual processing cores (FIT = failure in one billion hours of operation and this range has been experimentally observed for commodity electronics~\cite{mem_errors:2010, dram_error:2009}), an HPC system with one million cores will experience .1-100 errors per hour, which is equivalent to 3e-11 - 3e-14 errors per processor cycle on a 1Ghz processor.
+\sui{Hence, the experiments are referred to as ``accelerated experiments.''}
 The range of error injection rates used used in our experiments were chosen such that (i) on average one or more errors were injected in the executions of our target applications and (ii) at least some application runs completed without crashing.
-Results for error rates that induce fewer than one error per run were extrapolated from our experiments with higher error rates.
-This was done by considering only the runs where exactly one error was injected and trivially scaling the results to correspond to 1 in every $n$ runs being injected with an error while remaining $n-1$ runs complete with no injection.
+Results for error rates that induce fewer than one error per run were extrapolated from our experiments with higher error rates \sui{as described in \ref{sec:eval:extrapolation}}.
+This was done by considering only the runs where exactly one error was injected and trivially scaling the results to correspond to 1 in every $n$ runs being injected with an error while remaining $n-1$ runs complete with no injection.\\
 
-%\greg{This is the section where we put all the pieces together and present the experimental results on full applications.
-We should describe our main use-case here: user wants to achieve a given accuracy with a given probability and we configure the algorithm to provide this at the lowest cost.
-QUESTION: how do we pick the parameters of each resilience technique?
-Which metric is being optimized?} \sui{Using the choices of those parameters derived from the experiments we have already performed (treat it as a ``training set'') and use those choices on some new experiment configurations (treat as ``test set''), and see what is the probability that those choices could satisfy the users' experiments.}
+\subsection{Extrapolation Model}
+\label{sec:eval:extrapolation}
+\sui{
+This section explains our modelling of massively parallel and long-running applications with the experiments mentioned in page \pageref{sec:eval}, which in nature
+are accelerated experiments. We model larger scale experiments by extrapolating our experiments in the following ways:\\
+}
 
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=1.00\columnwidth]{figs/extrapolates.png}
+\caption{Extrapolation models}
+\label{fig:extrapolations}
+\end{figure}
+
+\begin{figure}[ht!]
+\centering
+\begin{tabular}{|p{1.0in}|p{2.0in}|}
+\hline
+ Model Type & Description \\
+\hline
+Accelerated Experiment & FIT > 1e+7, containing > 1e10 dynamic instructions, running on 1 node, wall time < 10 min \\
+\hline
+Serial Extrapolation by factor N & FIT > $\frac{1e+7}{N}$, containing > $N \cdot 1e10$ dynamic instructions, running on < 1 node \\
+\hline
+Parallel Extrapolation by factor N & FIT > $\frac{1e+7}{N}$, containing > $N \cdot 1e10$ dynamic instructions, running on N nodes \\
+\hline
+\end{tabular}
+\end{figure}
+
+\sui{
+There is 1 instance drawn from the accelerated tests in both serial and parallel extrapolation models. Each time one instance is sampled
+from the accelerated tests, it is combined with N-1 fault-free instances to become a large serial or parallel execution. The value of N is determined
+by each fault rate that needs be extrapolated to. For example, a value N=10000 is equivalent to extrapolating from 1e+7 FIT to 1e+3 FIT.
+In order to determine the magnitude of output errors from the combined executions, we make the following assumptions to the extrapolation models
+in order to simplify our analysis:
+}
+
+\sui{
+\begin{itemize}
+\item{
+The magnitude of output errors in a combined serial execution is that of the accelerated instance with the largest output error magnitude.
+Analysis of the output error magnitude is in \ref{sec:eval:errorsensitivity}.\\
+$RMSD_{serial} = max_{i=1}^{N} (RMSD_n)$}
+\item{The magnitude of output errors in a combined parallel execution is the error magnitude of the average of all outputs. 
+This reflects a classic reducing access pattern frequently seen in parallel programs. Since all instances except the first one
+are fault-less, the error magnitude, and the RMSD, of the combined instance would effectively be shrunk by a factor of N.\\
+$RMSD_{parallel} = \sum_{i=1}^{N}{RMSD_n}/N$
+}
+\item{The running time of a combined serial execution is the sum of every individual instance.}
+\item{The running time of a combined parallel execution is the running of the longest individual instance. }
+\item{At least 1 of the individual instances is injected with one error.}
+\end{itemize}
+}
+
+\subsection{Error sensitivity}
+\label{sec:eval:errorsensitivity}
+\sui{
+	I intend to show with the figure that 1) error propagation at an early point of time in a long serial run will result in an error that keeps being oscillated and magnified, resulting in numerical instability or simulation explosion,
+	and be terminated or 2) the speed the magnitude of errors grows keeps dropping until the magnitude of errors stabilize and converge. \\
+	In the first case, the magnitude of errors in the output should be small enough not to trigger numerical unstability or simulation explosion. In the second case,
+	the magnitude of error would peak the instant after the errors occur. Either way, we can assume the magnitude of errors in a serial run to be the largest among
+	individual instances.
+}
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=1.00\columnwidth]{figs/InputVarianceOutputRMSD.png}
+\caption{Extrapolation models}
+\label{fig:inputvarianceoutputrmsd}
+\end{figure}
+
+\subsection{Experiment setup}
+\label{sec:eval:confs}
+
+\sui{
+As we have seen in section \ref{sec:res_tech:eval}, varying the error checker thresholds in linear algebra routines and the checkpointing intervals and number of 
+replicas in the RK4 integrator results in different levels of accuracy. As they are deployed in applications, choice of these parameters may influence how often
+the applications need to rollback in order to obtain correct results. An intuition is we should strike a balance between the tradeoff between correctness
+and performance.
+To explore the possible tradeoff in applications, we consider a range of parameters for our resilience techniques as described in section \ref{sec:res_tech:eval} when deploying the algorithmic checkers in these applications.
+The variable parameters and fixed parameters are listed as below.
+}
+
+\begin{figure}[ht!]
+\centering
+\begin{tabular}{|p{0.7in}|p{2.0in}|}
+\hline
+{\bf Application} & {\bf Resilience Techniques} \\
+\hline
+\hline
+Lasso & Checkers for MM, MV, CD, SYRK; input backup; protection for critical data structures \\ 
+\hline
+DRC & Checkers for FFT, FIR; input backup; protection for critical data structures \\
+\hline
+Hattrick & Checkpointing; protection for critical data structures; replication of pointers \\
+\hline
+\hline
+{\bf Application} & {\bf Variable Parameters} \\
+\hline
+\hline
+Lasso & Checker thresholds 1e-4 to 1e-7 \\
+\hline
+DRC   & Checker thresholds 1e-6 to 3e-8 \\
+\hline
+Hattrick & Checkpointing intervals 0, 1e3, 1e5; 0, 1, 2 replicas for each pointer \\
+\hline
+\end{tabular}
+\caption{Experiment Configurations}
+\label{fig:exp_confs}
+\end{figure}
+
+\sui{
+	In each accelerated fault injection experiment, we ran each application configuration to completion, recording whether it completed and if so, the magnitude of the
+	error.
+	We calculate the execution time of each configuration, accounting for restarting the application in cases of unrecoverable failures via the formula $\frac{t_c}{1-p_f}$, where $t_c$ is the average completion time of such runs and $p_f$ is the probability of an unrecoverable failure.
+}
+
+% Configurations are described twice.
 \subsection{Performance Impact}
 \label{sec:eval:perf}
+Figures~\ref{fig:Lasso_EstdCost}, \ref{fig:DRC_EstdCost} and \ref{fig:Hattrick_EstdCost} show the expected execution time \sui{of parallel (in dashed lines) and sequential (in solid lines) extrapolated instances}, relative to the execution time of the non-fault tolerant version of the extrapolation of the application when no errors are injected.
 
-Our fault injection experiments ran each application configuration to completion, recording whether it completed and if so, the magnitude of the error.
-We calculate the execution time of each configuration, accounting for application rollbacks via the formula $\frac{t_c}{1-p_f}$, where $t_c$ is the average completion time of such runs and $p_f$ is the probability of a run failing to complete.
-Figures~\ref{fig:Lasso_EstdCost}, \ref{fig:DRC_EstdCost} and \ref{fig:Hattrick_EstdCost} show the expected execution time, relative to the execution time of the non-fault tolerant version of the application when no errors are injected, of the non-fault tolerant version (red) and the optimal fault tolerant version (blue).
-The optimal fault tolerant version is the one that suffers from smaller overhead among the ones that provide more resilience enhancement.
-To select the optimal one we consider error checker tresholds from $10^{-4}$ to $10^{-7}$ for ADDR, from $10^{-6}$ to $3 \cdot 10^{-8}$ for DRC and from using one pair of copies for each variable occurence to using a different pair of replicas for each variable occurrence for Hattrick.
+%Each figure shows a separate graph for each input size and along with the ID of the input shows the application's execution time on that input to quantify its vulnerability to errors on the input.
+Data for the non-fault tolerant version is shown in red and the fault tolerant version with the optimal configuration for each input size is shown in blue.
+The optimal configuration was selected to minimize overhead from across the following configuration spaces.
+
 The graphs from left to right correspond to different application input sizes.
-Within each graph the x-axis corresponds to error injection rates, while the y-axis shows the relative execution time, both logarithmic.
+Within each graph the x-axis corresponds to error injection rates (the real-system region of 3e-11 to 3e-14 highlighted in grey), while the y-axis shows the relative execution time.
+Both axes use logarithmic scales.
 
-For each configuration we plot a 95\% confidence interval.
-This is the range of relative execution times at the given error rate and application configuration in that will occur in 95\% of all trials.
-The confidence interval is computed by modeling the number of times application results will meet the error bound using a Binomial distribution.
-Given a binary random variable that is equal to 1 (i.e.
-error bound achieved) with probability $p$ and 0 (i.e.
-bound missed) with probability $1-p$ the Binomial distribution is the number of trials where it is 1.
+%For each configuration we plot a 95\% confidence interval.
+%This is the range of relative execution times at the given error rate and application configuration in that will occur in 95\% of all trials.
+%The confidence interval is computed by modeling the application's execution time as a Normal distribution.
 
-Figure~\ref{fig:Lasso_EstdCost} shows that the non-fault tolerant ADDR slows down from 2x to 9x, depending on the input set size, as a result of the injected faults.
-The optimal fault tolerance version the algorithm shows degradation that are around one half of the non-faul tolerant one, which implies that our technique improve ADDR's resilience by 50\%.
+Figure~\ref{fig:Lasso_EstdCost} shows that the non-fault tolerant ADDR slows down from 1.2x to 6x, depending on the input set size, as a result of the injected faults.
+\sui{Although the fault tolerant version incurs overheads from the algorithmic checker and backing up input, the benefits from reducing the expected running time begins to outweigh the cost when the fault rate approaches 1e-11.} The optimal fault tolerance version the algorithm shows degradation that are around one half of the non-fault tolerant one \sui{when the fault rate exceeds 1e-10}, which implies that our technique improve ADDR's resilience by 50\%.
+\sui{With the parallel extrapolation model, ADDR sees a higher running time overhead than with the serial extrapolation model.
+However it is much worth noting that as input size grows the overhead of the checkers become lower, as we have seen in section \ref{sec:res_tech:eval} with the individual linear algebra routines.}
 
-Figure~\ref{fig:DRC_EstdCost} shows results regarding the DRC application.
-Since it performs many more integer operations and memory copies than ADDR, it is more vulnerable to errors that corrupt the values of pointers or cause loads and stores to target invalid memory locations.
-Such errors cause memory segmentation faults, which force DRC to restart its execution and thus incur a very high overhead.
-That is the reason why the performance slowdown of its non-fault tolerant version is much larger than the ADDR one.
-In contrast, the optimal fault tolerant version reduces almost completely the slowdown due to fault injection.
+Figure~\ref{fig:DRC_EstdCost} 
+%shows that since DRC performs many more integer operations and memory copies than ADDR, it is more vulnerable to errors that corrupt the values of pointers or cause loads and stores to target invalid memory locations.
+\sui{shows that in DRC, the crossing point where the benefit fault-tolerant version outweights the cost of the algorithmic checker comes at a lower error rate than
+that of Lasso. Just as we have seen in section \ref{sec:res_tech:eval}, the overhead from the algorithmic checkers become lower as the sampling rate increases.
+Contrary to ADDR, DRC sees a larger overhead with the serial extrapolation model. This is because DRC is much more sensitive to input errors and is more likely
+not to satisfy the user-specified error bound under the sequential extrapolation model than under the parallel extrapolation model.
+}
+%Such errors cause memory segmentation faults, which force DRC to restart its execution and thus incur a very high overhead.
+%That is the reason why the performance slowdown of its non-fault tolerant version is much larger than for ADDR at high error rates.
+%In contrast, the optimal fault tolerant version almost completely eliminates the performance impact of faults.
+%However, at low error rates the cost of resilience is sufficiently high that at low error rates the non-fault tolerant version of DRC performs slightly better.
 
-Figure~\ref{fig:Hattrick_EstdCost} shows how the performance slowdown of the fault tolerance version it is very unstable but it can reach values close to 100x.
-The performance of the optimal fault tolerant version is consistently much better.
+Figure~\ref{fig:Hattrick_EstdCost} 
+\sui{shows that the cost of pointer replication and checkpointing are almost negligible.
+Just as we see in DRC, this application sees a higher overhead with the serial extrapolation model than with the parallel extrapolation model, because the overhead induced by the fault tolerance mechanisms are
+even smaller than those of DRC.
+So, from these figures, we learn that
+\begin{itemize}
+\item{Under the sequential extrapolation model, the overhead of the application would mainly come from not being able to meet the error bound and having to roll back.}
+\item{Under the parallel extrapolation model, the overhead of the application would mainly come from the overhead of the checkers themselves.}
+\item{Whether the checker overhead or the roll-back overhead is more dominant is application-dependent.}
+\end{itemize}
+}
 
@@
 \centering
-\includegraphics[height=2.75in]{figs/Lasso_EstdCost.png}
+\includegraphics[width=7in]{figs/Lasso_Par_Seq_EstdCost_log.png}
 \vspace{-10pt}
@@
 \centering
-\includegraphics[height=2.75in]{figs/DRC_EstdCost.png}
+\includegraphics[width=7in]{figs/DRC_Par_Seq_EstdCost_log.png}
 \vspace{-10pt}
@@
 \centering
-\includegraphics[height=2.75in]{figs/Hattrick_EstdCost.png}
+\includegraphics[width=3.4in]{figs/Hattrick_Par_Seq_EstdCost_log.png}
 \vspace{-10pt}
@@
 
-If the probability that a given run completes without an unrecoverable error is greater than 0 then the application will eventually complete with some output after some number of restarts.
 This section evaluates the distribution of errors in the resulting outputs.
-The outputs of are, for ADDR, vectors that holds the solution to the linear system.
-For Hattrick, they are vectoris that hold the state of the n-body system at the final application time-step, with 6 numbers per body
-and and for DRC they are a Pulse Code Modulation-formatted file, which is a sequence of numbers.
-Since the outputs of each routine can be represented as vectors we calculate the error in the outputs by comparing the vector $v_e$ returned by each routine when errors are injected to the correct vector $v_c$ computed during a fault-free run and computing the L2 norm of the difference: $\sqrt{\frac{\sum_{i} (v_c-v_e)^2}{\left| v_c \right|}}$.
+First of all, if the probability that a given run completes without an unrecoverable error is greater than 0 then the application will eventually complete with some output after some number of restarts.
+\sui{The restarts translate into the roll-back head reflected in section \ref{sec:eval:perf}.}
 
-Figures~\ref{fig:Lasso_ImperfectRate}, \ref{fig:DRC_ImperfectRate} and \ref{fig:Hattrick_ImperfectRate} show the errors in ADDR, DRC and Hattrick results from a user-centric perspective.
+The accuracy is quantified in the following way:
+The output of ADDR is the vector that holds the solution to the linear system.
+For Hattrick it is the vector that holds the state of the n-body system at the final application time-step, with 6 numbers per body
+Finally, DRC's output is a Pulse Code Modulation-formatted file, which is a sequence of numbers.
+Since the outputs of each routine can be represented as a vector we calculate the error in the outputs by comparing the vector $v_e$ returned by each application when errors are injected to the correct vector $v_c$ computed during a fault-free run and computing the L2 norm of the difference: $\sqrt{\frac{\sum_{i} (v_c-v_e)^2}{\left| v_c \right|}}$.
+
+Figures~\ref{fig:Lasso_ImperfectRate}, \ref{fig:DRC_ImperfectRate} and \ref{fig:Hattrick_ImperfectRate} show the errors in ADDR, DRC and Hattrick outputs from a user-centric perspective.
 They plot the fraction of application runs for which a numerical error equal or smaller than $10^{-8}$ is achieved.
@@
 The x-axis in each graph shows the fault injection rate while the y-axis shows the fraction of runs where the error target was satisfied.
-Both axis are expressed in logarithmic scale.
-Each graph shows data for the non-fault tolerant version of the application (all points connected by the line) as well as all the the fault tolerant versions (optimal configurations connected by a line).
-The 95\% confidence interval is plotted and denoted range of probabilities that the error bound will be achieved at the given error rate and application configuration in 95\% of all trials.
+Both axes are expressed in logarithmic scale.
+Each graph shows data for the non-fault tolerant version of the application (all points connected by the line) as well as all the fault tolerant versions (optimal configurations connected by a line).
+%The 95\% confidence interval is plotted and denotes the range of probabilities that the error bound will be achieved at the given error rate and application configuration in 95\% of all trials.
+%It is computed by modeling the number of times application results will meet the error bound using a Binomial distribution.
+%Given a binary random variable that is equal to 1 (i.e.
+error bound achieved) with probability $p$ and 0 (i.e.
+bound missed) with probability $1-p$ the Binomial distribution is the number of trials where it is 1.
 
-Figures~\ref{fig:Lasso_ImperfectRate}, \ref{fig:DRC_ImperfectRate} and \ref{fig:Hattrick_ImperfectRate} show that the probability that the non-fault tolerant version of each one the three application satisfies the user-provided error bound is low especially for fault injection rates $10^{-9}$ to $10^{-11}$.
-In contrast, the fault tolerant version of the application is significantly more successful, especially for error rates of $10^{-9}$ and lower.
+The data shows that the probability that the non-fault tolerant version of each one the three application violates the user-provided error bound is above 10\% for Hattrick and DRC in the region of realistic error rates (3e-11 to 3e-14) and above 1\% for ADDR across all inputs.
+When resilience is applied to these applications the probability of violating the error bound drops to below 1\% for ADDR and Hattrick and below 10\% for DRC in the same region.
+\sui{And, not surprisingly, an instance under the parallel extrapolation model is more likely to meet the error bound than one under the sequential extrapolation model.}
+This shows that the use of all the resilience techniques is critical for ensuring that application outputs are valid with a high probability even on very large HPC systems.
 
@@
 %\includegraphics[height=5in]{figs/Lasso_ImperfectRate.png}
-\includegraphics[height=2.75in]{figs/Lasso_ImperfectRate_log.png}
+\includegraphics[width=7in]{figs/Lasso_Par_Seq_ImperfectRate_log.png}
+\vspace{-10pt}
 \caption{Probability of not satisfying the user-provided error bound for ADDR. Results for the non-fault tolerant version (red) and the optimal fault tolerant version (blue) are shown.}
+\vspace{-10pt}
 \label{fig:Lasso_ImperfectRate}
@@
 %\includegraphics[height=3in]{figs/DRC_ImperfectRate.png}
-\includegraphics[height=2.75in]{figs/DRC_ImperfectRate_log.png}
+\includegraphics[width=7in]{figs/DRC_Par_Seq_ImperfectRate_log.png}
+\vspace{-10pt}
 \caption{Probability of not satisfying the user-provided error bound for DRC. Results for the non-fault tolerant version (red) and the optimal fault tolerant version (blue) are shown.}
+\vspace{-10pt}
 \label{fig:DRC_ImperfectRate}
@@
 %\includegraphics[height=3in]{figs/Hattrick_ImperfectRate.png}
-\includegraphics[height=2.75in]{figs/Hattrick_ImperfectRate_log.png}
+\includegraphics[width=3.4in]{figs/Hattrick_Par_Seq_ImperfectRate_log.png}
+\vspace{-10pt}
 \caption{Probability of not satisfying the user-provided error bound for Hattrick. Results for the non-fault tolerant version (red) and the optimal fault tolerant version (blue) are shown.}
+\vspace{-10pt}
 \label{fig:Hattrick_ImperfectRate}
+\end{figure}
+
+\subsection{Choice of Optimal Fault Tolerance Techniques}
+\label{sec:eval:optchoice}
+
+\sui{Figures XXX show the choice of optimal fault tolerance techniques in the parallel and sequential extrapolated runs of all three applications.}
+
+\sui{(All Suggested Fault Tolerance Configuration figures)
+There are cells in the figures that are not shaded.
+These cells mean there is not any possible fault tolerance configuration available to satisfy the user-specified probability of not satisfying the error bound because the specified probability is too low and error rate is to high.}
+
+\sui{(All Probability of not satisfying user's error bound figures) 
+Serial extrapolated runs have a slightly higher than parallel extrapolated ones.}
+
+\sui{(Comparing Lasso's Parallel and Serial Estimated Optimal Cost)
+When error rates are low, the overhead of both serial and parallel fault-tolerant runs are dominated by the algorithm checker.
+By the nature of the error checkers, their overheads decrease as input size goes up.
+As fault rate increases, the parallel run sees a more rapid gain in overhead than a serial run.
+This is because the overhead of rolling back the faulty runs becomes more pronounced in parallel runs than serial runs.}
+
+\sui{(Comparing Lasso's Parallel and Serial Suggested Fault Tolerance Configurations)
+From the cells that are shaded, as can be expected from the extrapolation model, a serial run requires a tighter error bound. }
+
+\sui{(Comparing DRC's Parallel and Serial Estimated Optimal Cost)
+Due to a much smaller proportion of the cost of rolling back in the samples of DRC, the difference between serial and parallel runs of DRC is much smaller than that of the serial and parallel runs of Lasso.}
+
+\sui{(Comparing DRC's Parallel and Serial Suggested Fault Tolerance Configurations)
+From the cells that are shaded, a serial run requires a tighter error bound.}
+
+\sui{(Comparing Hattrick's Parallel and Sequential Estimated Optimal Cost)
+Since the cost of Hattrick's checkpointing code is almost negligible, the difference between the performance overheads of the parallel and serial extrapolation models is also almost negligible.}
+
+\sui{(Comparing Hattrick's Parallel and Serial Suggested Fault Tolerance Configurations)
+From the cells that are shaded, a serial run requires a tighter error bound.}
+
+\begin{figure*}[ht!]
+\centering
+%\includegraphics[height=3in]{figs/DRC_ImperfectRate.png}
+\includegraphics[width=7in]{figs/Lasso_Parallel_SuggestedConf.png}
+\vspace{-10pt}
+\caption{Suggested Resilience Technique for ADDR Under the Parallel Extrapolation Model Under Different Fault Rates}
+\vspace{-10pt}
+\label{fig:Lasso_Parallel_SuggestedConf}
+\end{figure*}
+
+\begin{figure*}[ht!]
+\centering
+%\includegraphics[height=3in]{figs/DRC_ImperfectRate.png}
+\includegraphics[width=7in]{figs/Lasso_Serial_SuggestedConf.png}
+\vspace{-10pt}
+\caption{Suggested Resilience Technique for ADDR Under the Serial Extrapolation Model Under Different Fault Rates}
+\vspace{-10pt}
+\label{fig:Lasso_Serial_SuggestedConf}
+\end{figure*}
+
+\begin{figure*}[ht!]
+\centering
+%\includegraphics[height=3in]{figs/DRC_ImperfectRate.png}
+\includegraphics[width=7in]{figs/DRC_Parallel_SuggestedConf.png}
+\vspace{-10pt}
+\caption{Suggested Resilience Technique for DRC Under the Parallel Extrapolation Model Under Different Fault Rates}
+\vspace{-10pt}
+\label{fig:DRC_Parallel_SuggestedConf}
+\end{figure*}
+
+\begin{figure*}[ht!]
+\centering
+%\includegraphics[height=3in]{figs/DRC_ImperfectRate.png}
+\includegraphics[width=7in]{figs/DRC_Serial_SuggestedConf.png}
+\vspace{-10pt}
+\caption{Suggested Resilience Technique for DRC Under the Serial Extrapolation Model Under Different Fault Rates}
+\vspace{-10pt}
+\label{fig:DRC_Serial_SuggestedConf}
+\end{figure*}
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=3.4in]{figs/Hattrick_Parallel_SuggestedConf.png}
+\vspace{-10pt}
+\caption{Suggested Resilience Technique for Hattrick Under the Parallel Extrapolation Model Under Different Fault Rates}
+\vspace{-10pt}
+\label{fig:Hattrick_Parallel_SuggestedConf}
+\end{figure}
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=3.4in]{figs/Hattrick_Serial_SuggestedConf.png}
+\vspace{-10pt}
+\caption{Suggested Resilience Technique for Hattrick Under the Serial Extrapolation Model Under Different Fault Rates}
+\vspace{-10pt}
+\label{fig:Hattrick_Serial_SuggestedConf}
 \end{figure}
@@
 Algorithmic error checks, replication of critical data structures and checkpoint-restart of individual routines.
-We demonstrate that very significant improvements can be achieved in terms of reduction of performance slowdown and output numerical accuracy when scientific problems face error injection rates from $3 \cdot 10^{-11}$ and $3 \cdot 10^{-14}$ per processor cycle in a 1Ghz, which are the expected rates in future high performance computing systems.
-This work demonstrates that current scientific application can be safely executed on such systems if the programmers combine resilience techniques in the way it is described.
+We demonstrate that very significant improvements can be achieved in terms of reduction of performance slowdown and output numerical accuracy when scientific problems face error injection rates from $3 \cdot 10^{-11}$ and $3 \cdot 10^{-14}$ per processor cycle at 1Ghz, which are the expected rates in future high performance computing systems.
+\sui{We have also demonstrated that the choice of the error checker threshold is important in achieveing a balance between result accuracy and overall performance.
+The choice is dependent on both the fault rate and the characteristics of the application.}
+This work demonstrates that current scientific application can be safely executed on such systems if the programmers comprehensively combine a range of resilience techniques.
 

### 2013-10-14 pubform.tex
2013-10-13

%%
%% This is file `pubform.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% acmconf.dtx  (with options: `pubform')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from pubform.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file acmconf.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% to test the checksum, uncomment    \OnlyDescription
%% in the driver
%%\fi
%%
%% \CharacterTable
%%  {Upper-case    \A\B\C\D\E\F\G\H\I\J\K\L\M\N\O\P\Q\R\S\T\U\V\W\X\Y\Z
%%   Lower-case    \a\b\c\d\e\f\g\h\i\j\k\l\m\n\o\p\q\r\s\t\u\v\w\x\y\z
%%   Digits        \0\1\2\3\4\5\6\7\8\9
%%   Exclamation   \!     Double quote  \"     Hash (number) \#
%%   Dollar        \$     Percent       \%     Ampersand     \&
%%   Acute accent  \'     Left paren    \(     Right paren   \)
%%   Asterisk      \*     Plus          \+     Comma         \,
%%   Minus         \-     Point         \.     Solidus       \/
%%   Colon         \:     Semicolon     \;     Less than     \<
%%   Equals        \=     Greater than  \>     Question mark \?
%%   Commercial at \@     Left bracket  \[     Backslash     \\
%%   Right bracket \]     Circumflex    \^     Underscore    \_
%%   Grave accent  \`     Left brace    \{     Vertical bar  \|
%%   Right brace   \}     Tilde         \~}

\documentclass[submit]{acmconf}
\usepackage{flushend}

\CopyrightText{LEAVE BLANK THE LAST 3.81 cm
               (1.5'') OF THE LEFT COLUMN ON THE FIRST PAGE
               FOR THE COPYRIGHT NOTICE }
\ConferenceName{1. Conference on Designing a \LaTeX2e Class for
                Typesetting ACM Papers, Hawaii 2000}
\ConferenceShortName{CONF-2000}

\begin{document}
\date{31. December 1999}
\title{ACM SIG Proceedings Format}
\author{\Author{1st Author}                               \\
        \Address{First author's affiliation}              \\
        \Address{1st line of address}                     \\
        \Address{2st line of address}                     \\
        \Address{Telephone number, incl.\ country code}   \\
        \Email{1st author's email address}                \\
        \and
        \Author{2nd Author}                               \\
        \Address{First author's affiliation}              \\
        \Address{1st line of address}                     \\
        \Address{2st line of address}                     \\
        \Address{Telephone number, incl.\ country code}   \\
        \Email{2nd author's email address}                \\
        \and
        \Author{3rd Author}                               \\
        \Address{First author's affiliation}              \\
        \Address{1st line of address}                     \\
        \Address{2st line of address}                     \\
        \Address{Telephone number, incl.\ country code}   \\
        \Email{3rd author's email address}                \\
       }
\maketitle

\begin{abstract}
In this paper, we describe the formatting guidelines for ACM SIG
Proceedings.
\end{abstract}

\begin{keywords}
Guides, instructions, authors kit, conference publications.
\end{keywords}

\section{Introduction}

The proceedings are the records of the conference. ACM hopes to
give these conference by-products a single, high-quality appearance.
To do this, we ask that authors follow some simple guidelines. In
essence, we ask you to make your paper look exactly like this
document. The easiest way to do this is simply to down-load a
template from \cite{template.1999}, and replace the content with
your own material.

\section{Page Size}

All material on each page should fit within a rectangle of
18 x 23.5 cm (6" x 9.25"), centered on the page, beginning
2.54 cm (1") from the top of the page and ending with
2.54 cm (1") from the bottom.  The right and left margins
should be 1.9 cm (.75'').  The text should be in two
8.45 cm (3.33") columns with a .83 cm (.33") gutter.

\section{Typeset Text}

Prepare your submissions on a typesetter or word processor.

\subsection{Normal or Body Text}

Please use a 9-point Times Roman font, or other Roman font with
serifs, as close as possible in appearance to Times Roman in
which these guidelines have been set. The goal is to have a
9-point text, as you see here. Please use sans-serif or
non-proportional fonts only for special purposes, such as
distinguishing source code text. The Press 9-point font available
to users of Script is a good substitute for Times Roman. If
Times Roman is not available, try the font named Computer
Modern Roman. On a Macintosh, use the font named Times.
Right margins should be justified, not ragged.

\subsection{Title and Authors}

The title (Helvetica 18-point bold), authors' names (Helvetica
12-point) and affiliations (Helvetica 10-point) run across the full
width of the page - one column wide. We also recommend phone number
(Helvetica 10-point) and e-mail address (Helvetica 12-point). See the
top of this page for three addresses. If only one address is needed,
center all address text. For two addresses, use two centered tabs, and
so on. For more than three authors, you may have to
improvise\footnote{If necessary, you may place some address
information in a foortone, or in a named section at the end of your
paper.}

\subsection{First Page Copyright Notice}

Leave 3.81 cm (1.5") of blank space at the bottom of the left
column of the first page for the copyright notice.

\subsection{Subsequent Pages}

For pages other than the first page, start at the top of the page, and
continue in double-column format.  The two columns on the last page
should be of equal length.

\subsection{References and Citations}

Footnotes should be Times New Roman 9-point.

Use the standard Communications of the ACM format for references -
that is, a numbered list at the end of the article, ordered
alphabetically by first author, and referenced by numbers in
brackets \cite{anderson.1992}. See the examples of citations at
the end of this document
\cite{conger.1995,mackay.1995,schwartz.1995}. Within this template
file, use the style named references for the text of your citation.

References should be published materials accessible to the public.
Internal technical reports may be cited only if they are easily
accessible (i.e. you can give the address to obtain the report within
your citation) and may be obtained by any reader. Proprietary
information may not be cited. Private communications should be
acknowledged, not referenced  (e.g., ``[Robertson, personal
communication]'').

\subsection{Page Numbering, Headers and Footers}

Do not include headers, footers or page numbers in your submission.
These will be added when the publications are assembled.

\section{Sections}

The heading of a section should be in Times New Roman 12-point bold in
all-capitals flush left.  Sections and subsequent subsections should
be numbered and flush left.

\subsection{Subsections}

The heading of subsections should be in Times New Roman 12-point bold
with only the initial letters capitalized. (Note: For subsections and
subsubsections, a word like the or a is not capitalized unless it is
the first word of the header.)

\subsubsection{Subsubsections}

The heading for subsubsections should be in Times New Roman 11-point
italic with initial letters capitalized.

\subsubsubsection{Subsubsubsection}

The heading for subsubsubsections should be in
Times New Roman 11-point italic with initial letters capitalized.

\subsubsubsubsection{Subsubsubsubsection}

The heading for subsubsubsubsections should be in
Times New Roman 11-point italic with initial letters capitalized.

\section{Figures/Captions}

Place Tables/Figures/Images in text as close to the reference as
possible.  It may extend across both columns to a maximum width of
17.78 cm (7'').

Captions should be Times New Roman 9-point bold.  They should be
numbered (e.g., ``Table 1'' or ``Figure 2'') and be centered beneath
each table, figure or image.

\section{Acknowledgments}

Our thanks to ACM SIGCHI for allowing us to modify templates they had
developed.

\medskip

\bibliographystyle{plain}
\bibliography{pubform}

Columns on Last Page Should Be Made Equal Length
\end{document}
\endinput
%%
%% End of file `pubform.tex'.

### 2013-10-14 paper.tex
Updating abstract and intro for current paper angle.

@@
 Since such faults may cause application crashes or may silently corrupt application results it is necessary to develop techniques to protect applications from such errors at a low cost in performance and power.
-Despite extensive work on general-purpose resilience techniques, they are cost effective for protecting only certain system components, such as memories.
-To provide comprehensive protection for the computations of HPC applications it is necessary to develop algorithmic resilience mechanisms that use the properties of algorithms to provide cost-effective protection.
+\greg{Since general-purpose resilience mechanisms, such as replication, can be very expensive for general applications it is imperative to develop ways to tailor resilience techniques to the needs and properties of each application.}
 
-\sui{
-Algorithmic resilience exploiting the invariants of certain algorithms is an old technique that has been around for many years. The invariants can serve as
-correctness unit tests during development and turned off during run-time in favor of speed, but under the context of the upcoming exascale reliability challenges,
-they can usually serve as useful fault checkers. However, unless other issues including input protection, control information
-and critical data structure protection are resolved will algorithmic invariants help detecting and correcting errors.
-% can be very complex to deploy in applications because a comprehensive resilience strategy must address all the types of errors that an application may experience.
-%This includes corruptions to numerical data as well as control information and data structures.
-%Full protection requires the combination of multiple techniques including algorithmic error detection, crash detection, replication of key control information as well as support for restart when errors are detected.
-%Further, it is necessary to configure the properties of the resilience mechanisms to make sure that they are cost-effectively meeting the user's accuracy and confidence targets.
-Further, certain design choices in exploiting algorithmic resilience imply trade-offs for cost-effectiveness of the final resilience strategy.
-This paper presents our case study of exploiting algorithmic invariants for constructing comprehensive fault tolerance mechanisms and balancing performance and result accuracy in the context of large-scale computing.
+\greg{Extensive ongoing work on algorithmic techniques to detect and tolerate errors in common scientific sub-routines and kernels has shown that such techniques can be very effective at protecting the numeric properties of these code regions.
+However, the practical utility of these techniques has been limited by the fact that real scientific applications (i) are significantly more complex than these simple kernels and (i) are vulnerable to a wider range of errors than most algorithmic checkers focus on.
+To make scientific applications resilient to soft errors it is necessary to comprehensively combine resilience techniques that detect all the major types of errors that can affect an application and provide techniques that can tolerate them.
+Further, the techniques applied to a given application must be tailored to an application's needs and configured to provide the performance for each input, hardware error rate and application user's accuracy tolerance.
+This paper presents a case study of the design of comprehensive resilience mechanisms for four(\textbf{???}) different applications.
+In addition to exploring the space of mechanism design and configuration we project our results to massively parallel applications and show how application performance and resilience configuration choices vary with error frequency, application scale and accuracy goals.
 }
-%This paper shows how all these issues can be resolved comprehensively in the context of three numeric applications.
 
@@
 
-The limitations of automated resilience techniques has motivated the need for algorithmic techniques that can detect and tolerate errors by taking advantage of application-specific properties and invariants~\cite{amg_abft:2012, robustification:2010, abft:1984}.
+\greg{The limitations of automated resilience techniques has motivated the need for algorithmic techniques that can detect and tolerate errors by taking advantage of application-specific properties and invariants~\cite{amg_abft:2012, robustification:2010, abft:1984}.
 Although these techniques can be very efficient, most research work on such techniques has primarily focused on how their algorithmic invariants can be leveraged to enable accurate error detection.
-While this goal is important, it ignores the full complexity of the problem since it focuses on only a subset of the errors \sui{(algorithmic correctness)} and 
-does not consider combinations of different resilience techniques to efficiently and effectively address all the different types of faults.
-In reality a comprehensive solution must incorporate a range of additional techniques such the replication of key variables, detection of memory violations and localized rollback on detected errors, as well as techniques to configure these various resilience methods to achieve the best performance given the user's accuracy and confidence targets.
-%, they also require manual effort to incorporate into applications and may require deep insight into an application's invariants to deploy effectively.
-%Fortunately, because real-world applications spend most of their execution is spent in a few application modules, most soft faults will occur in those modules rather than uniformly throughout the application.
-%In scientific applications these modules are usually various types of libraries, including numeric solvers and domain-specific scientific packages.
-%This indicates that if by focusing algorithmic resilience efforts on the key library routines, while leaving other modules unprotected or protected via more expensive mechanisms, it is possible to fully protect an application at a low cost.
+While this work is crucial, it addresses only a portion of the full challenge of making scientific applications resilient to soft faults.
+A comprehensive resilience strategy must address three areas.
+First, it must deal all the possible application-level manifestations of soft faults including numeric errors (the focus of traditional algorithmic checkers), as well as corruptions in data structures and other application state and computations.
+Second, it must address tolerance to soft faults in addition to their detection.
+Finally, given the many ways in which resilience techniques can be configured (e.g.
+different detection tolerances for an algorithmic checker) it must provide ways to choose the best configuration for each system and for each user's performance and accuracy needs.}
 
-This paper presents a comprehensive algorithmic resilience approach \sui{that combines existing checkpointing and algorithmic checker techniques that considers all of the above issues and demonstrate {\bf with fault injection experiments and statistical analyses} that such approach is able to significantly reduce the side effects of soft errors on parallel programs at a  slight performance cost.}
-We demonstrate it on different types of numerical applications: the Alternating Direction Method of Multipliers (ADDR) for Lasso problems~\cite{lasso:2011}, the Hattrick gravitational simulation~\cite{hattrick:2012} and the Digital Room Correction (DRC) acoustic correction application~\cite{drc:2012}.
-We show how these three applications can be comprehensively protected from soft faults by adding three different resilience mechanisms to the routines where they spend the most time: algorithmic error checks, replication of critical data structures and checkpoint-restart of individual routines.
-We demonstrate via fault injection experiments that although none of these techniques can individually protect applications from soft faults, their combination is highly effective.
-Further, we demonstrate how these techniques can be configured to offer the right level of protection at any given fault rate or input type.
-More specifically, these configurations can be used to produce a wide range of overheads and result accuracy levels and to tune the probabilistic confidence level that the given overhead or accuracy target will be met.
-%\greg{Summarize real results}
-%\sui{A tighter algorithmic checker bounds can be enabled at low error rates to completely eliminate erroneous results at a lower overhead, while a more lenient algorithmic checker would allow the user to produce more acceptable results in face of a higher error rate (such as 1e-09)}
-%For the three considered scientific codes, we correct more than 50\% of segmentation faults and we reduce the error in the outputs by one order of magnitude. The overhead is always lower than 20\%.
+\greg{
+This paper presents a case study of how existing resilience techniques can be combined to make a usable resilience solution for scientific applications.
+Within the context of three real applications we consider various combinations of algorithmic error detection, light-weight checkpoint restart to tolerate detected errors, as well as replication of pointers within data structures.
+We focus on the Alternating Direction Method of Multipliers (ADDR) for Lasso problems~\cite{lasso:2011}, the Hattrick gravitational simulation~\cite{hattrick:2012} and the Digital Room Correction (DRC) acoustic correction application~\cite{drc:2012}.
+Each application requires different resilience techniques, each of which can be configured in different ways.
+We use extensive fault injection experiments to measure the performance/resilience properties of a large space of resilience configuration within each application.
+For each configuration we measure the slowdown induced by the execution of resilience logic and roll-backs due to error detections, as well as the distribution of magnitudes of error in the application's output.
+We then demonstrate how these measurements can be used to configure the resilience mechanisms used with each application to minimize overhead while meeting the user's accuracy requirements.
+Since in practice users do not have a clear idea of the level of accuracy needed from an algorithm's results, we demonstrate a novel alternative method to specify accuracy requirements as the level noise in the application's input required to deviate the output by a given magnitude (i.e.
+similar to backward error analysis).
+Finally, we extrapolate from the small-scale application runs considered in our fault injection campaign (the many runs required to cover the large space of resilience configurations limited the length of each run) to application runs with 1 million CPU hours via three different models of error propagation.
+Our results demonstrate that when applications use all the relevant resilience methods, configured for maximum efficiency, the overhead of resilient execution on expected Exascale designs is low.
+
+\greg{Need a few highlight with concrete numbers}
 

### 2013-10-16 paper.tex
First major pass through paper by Greg.

@@
 %\documentclass[prodmode,acmtecs]{acmconf} % Aptara syntax
-\documentclass{sig-alternate}
+\documentclass[10pt, conference, compsocconf]{IEEEtran}
 \usepackage{url}
@@
 \usepackage[ruled]{algorithm2e}
+\usepackage{multirow}
 
@@
 %\affil{Louisiana State University}
-%Greg Bronevetsky, Marc Casas,
+%Greg Bronevetsky,
 %\affil{Lawrence Livermore National Laboratory}
+%Marc Casas,
+%\affil{Barcelona Supercomputing Center}
 %Lu Peng
-%\affil{Louisiana State University}
-%}
+%\affil{Louisiana State University}}
+
+\author{\IEEEauthorblockN{Sui Chen}
+\IEEEauthorblockA{Louisiana State \\ University}
+\and
+\IEEEauthorblockN{Greg Bronevetsky}
+\IEEEauthorblockA{Lawrence Livermore \\ National Laboratory}
+\and
+\IEEEauthorblockN{Marc Casas-Guix}
+\IEEEauthorblockA{Barcelona Supercomputing \\ Center}
+\and
+\IEEEauthorblockN{Lu Peng}
+\IEEEauthorblockA{Louisiana State \\ University}
+}
 
@@
 \maketitle
+\begin{abstract}
 
-\begin{abstract}
-As High-Performance Computing (HPC) systems become larger and more complex and the feature sizes of their electronic components grow smaller, the systems become increasingly vulnerable to soft faults.
+As High-Performance Computing (HPC) systems become larger and more complex and their circuit feature sizes shrink, the systems become increasingly vulnerable to soft faults.
 Since such faults may cause application crashes or may silently corrupt application results it is necessary to develop techniques to protect applications from such errors at a low cost in performance and power.
-\greg{Since general-purpose resilience mechanisms, such as replication, can be very expensive for general applications it is imperative to develop ways to tailor resilience techniques to the needs and properties of each application.}
+Since general-purpose resilience mechanisms, such as replication, can be very expensive it is imperative to develop ways to tailor resilience techniques to each application's properties and the needs of its users.
 
-\greg{Extensive ongoing work on algorithmic techniques to detect and tolerate errors in common scientific sub-routines and kernels has shown that such techniques can be very effective at protecting the numeric properties of these code regions.
-However, the practical utility of these techniques has been limited by the fact that real scientific applications (i) are significantly more complex than these simple kernels and (i) are vulnerable to a wider range of errors than most algorithmic checkers focus on.
-To make scientific applications resilient to soft errors it is necessary to comprehensively combine resilience techniques that detect all the major types of errors that can affect an application and provide techniques that can tolerate them.
-Further, the techniques applied to a given application must be tailored to an application's needs and configured to provide the performance for each input, hardware error rate and application user's accuracy tolerance.
-This paper presents a case study of the design of comprehensive resilience mechanisms for four(\textbf{???}) different applications.
-In addition to exploring the space of mechanism design and configuration we project our results to massively parallel applications and show how application performance and resilience configuration choices vary with error frequency, application scale and accuracy goals.
-}
+Extensive ongoing work on algorithmic techniques to detect and tolerate errors in common scientific sub-routines and kernels has proven their effectiveness.
+However, their practical utility has been limited by the fact that real scientific applications (i) are significantly more complex than these simple kernels and (i) are vulnerable to a wider range of errors than most algorithmic checkers focus on.
+To make scientific applications resilient to soft errors it is necessary to comprehensively combine resilience techniques that tolerate \emph{all} the major types of application errors.
+Further, each technique must be tailored to an application's needs and configured to provide the performance for each input, hardware error rate and application user's accuracy tolerance.
+This paper presents a case study of the design of comprehensive resilience mechanisms for three application domains.
+We describe the combinations of techniques required to make a single application from each domain resilient and quantify their performance and resilience on small-scale inputs.
+Further, we use these experiments to predict the behavior of large-scale instances (1 million CPU hours) of applications in each domain when using the most efficient application configuration for each fault rate and user-specified accuracy requirement.
 
@@
 Soft faults are transient corruptions of circuit state caused by physical phenomena such as strikes by neutrons or alpha particles~\cite{baumann:2005, asciQSER:2005} or thermal electrical noise~\cite{therm_noise:2007}.
-They can affect the state of processor latches and registers and may cause the application to crash or worse, to silently return incorrect results~\cite{mpi_ser:reed:2004}.
-As the feature sizes of electronic circuits shrink, technology scaling will make soft errors a larger problem~\cite{err_scaling:2012} due to the fact that each circuit element will hold less charge and can thus be disrupted more easily.
-These phenomena make it imperative to develop techniques to make HPC systems resilient to soft faults.
+They can affect processor latches and registers and may cause the application to crash or worse, to silently return incorrect results~\cite{mpi_ser:reed:2004}.
+As circuit feature sizes shrink, technology scaling will make soft faults a larger problem~\cite{err_scaling:2012}, %due to the fact that each circuit element will hold less charge and can thus be disrupted more easily.
+which makes it imperative to develop techniques to make HPC systems resilient to soft faults.
 
-Resilience is a problem that must be addressed at all levels.
-While materials science and circuit design techniques can be used to improve resilience, they come at a cost in reduced power efficiency and performance that can become prohibitive if processors need to be sufficiently reliable to build a large HPC system.
-Techniques such as error correcting codes have been very effective at making memories and caches resilient to soft faults~\cite{mem_errors:2010}.
-However, they are more expensive for protecting core-internal state such as latches and are significantly less effective for checking the correctness of computations.
-Traditional approaches like cross-core or node replication~\cite{rmpi:2011, dyn_cmp_repl:2007} are expensive in their use of computing resources and power and their long error detection latencies require processor state to be frequently checkpointed.
-Processor designs that incorporate instruction replication~\cite{repl_smt:2000} offer fine-grained error detection and rollback but require more power as well as novel hardware features that are not likely to be included in the commodity processors used in HPC systems for cost reduction reasons.
+Resilience must be addressed at all system levels.
+Materials science and circuit design techniques can improve resilience but at the cost of reduced power efficiency and performance that can become prohibitive at processor reliability levels needed to build a large HPC system.
+Error correcting codes are very effective at making memories and caches resilient to soft faults~\cite{mem_errors:2010} but are much more expensive for latches and significantly less effective for computational circuits.
+Replication across cores or nodes~\cite{rmpi:2011, dyn_cmp_repl:2007} more than double resource requirements and their long error detection latencies require frequent checkpointing.
+Processor designs that support instruction replication~\cite{repl_smt:2000} offer fine-grained error detection and rollback but require more power as well as novel hardware features unlikely to be included in commodity processors.
 
-\greg{The limitations of automated resilience techniques has motivated the need for algorithmic techniques that can detect and tolerate errors by taking advantage of application-specific properties and invariants~\cite{amg_abft:2012, robustification:2010, abft:1984}.
-Although these techniques can be very efficient, most research work on such techniques has primarily focused on how their algorithmic invariants can be leveraged to enable accurate error detection.
-While this work is crucial, it addresses only a portion of the full challenge of making scientific applications resilient to soft faults.
+The limitations of automated resilience techniques motivates work on algorithmic detection and tolerance techniques by leveraging application-specific properties and invariants~\cite{amg_abft:2012, robustification:2010, abft:1984}.
+Although very efficient, these techniques are not general and most research has focused on algorithmic resilience for numeric kernels.
+While this is crucial, it addresses only a portion of the full challenge of making scientific applications resilient to soft faults.
 A comprehensive resilience strategy must address three areas.
-First, it must deal all the possible application-level manifestations of soft faults including numeric errors (the focus of traditional algorithmic checkers), as well as corruptions in data structures and other application state and computations.
+First, it must deal all the possible application-level manifestations of soft faults including numeric errors (focus of traditional algorithmic checkers) as well as corruptions in data structures and other state and computations.
 Second, it must address tolerance to soft faults in addition to their detection.
-Finally, given the many ways in which resilience techniques can be configured (e.g.
-different detection tolerances for an algorithmic checker) it must provide ways to choose the best configuration for each system and for each user's performance and accuracy needs.}
+Finally, given the many ways to configure resilience techniques (e.g.
+detection tolerances) it must provide ways to choose the best configuration for each system and for each user's performance and accuracy needs.
 
-\greg{
-This paper presents a case study of how existing resilience techniques can be combined to make a usable resilience solution for scientific applications.
-Within the context of three real applications we consider various combinations of algorithmic error detection, light-weight checkpoint restart to tolerate detected errors, as well as replication of pointers within data structures.
-We focus on the Alternating Direction Method of Multipliers (ADDR) for Lasso problems~\cite{lasso:2011}, the Hattrick gravitational simulation~\cite{hattrick:2012} and the Digital Room Correction (DRC) acoustic correction application~\cite{drc:2012}.
-Each application requires different resilience techniques, each of which can be configured in different ways.
-We use extensive fault injection experiments to measure the performance/resilience properties of a large space of resilience configuration within each application.
-For each configuration we measure the slowdown induced by the execution of resilience logic and roll-backs due to error detections, as well as the distribution of magnitudes of error in the application's output.
-We then demonstrate how these measurements can be used to configure the resilience mechanisms used with each application to minimize overhead while meeting the user's accuracy requirements.
-Since in practice users do not have a clear idea of the level of accuracy needed from an algorithm's results, we demonstrate a novel alternative method to specify accuracy requirements as the level noise in the application's input required to deviate the output by a given magnitude (i.e.
+This paper presents a case study of how existing resilience techniques can be combined to make an efficient resilience solution for large-scale scientific applications.
+Our analysis focuses on linear solver, signal processing and n-body simulation applications, using the following applications as representatives: Alternating Direction Method of Multipliers (ADDR) for Lasso problems~\cite{lasso:2011}, the Hattrick gravitational simulation~\cite{hattrick:2012} and the Digital Room Correction (DRC) acoustic correction application~\cite{drc:2012}, respectively.
+We use small-scale runs of each application on multiple inputs to comprehensively measure their performance and resilience properties when protected by various various combinations of algorithmic error detection, light-weight checkpoint restart (tolerates detected errors), as well as pointer replication.
+We then extrapolate from these small-scale experiments to predict the performance and resilience properties of large-scale runs (1 million CPU hours) of applications within each domain.
+This is done by modeling large-scale runs as dependence graphs that connect small-scale runs into parallel reductions, sequential chains or a combination of both.
+Our analysis explicitly accounts for the accuracy that users may require from applications in terms of a bound on the Root Mean Squared (RMS) Error of the results as well as a bound on the probability that this target will not be met.
+To simplify the process of choosing the appropriate error bound we also provide a characterization of application output error in terms of the amount of measurement noise in application inputs that may produce the same deviation in application outputs (similar to backward error analysis).
+This makes it possible for application users to rationally trade off the accuracy of application results against the cost to achieve it (e.g. power, time or system purchase price).
+
+%Each application requires different resilience techniques, each of which can be configured in different ways.
+%We use extensive fault injection experiments to measure the performance/resilience properties of a large space of resilience configuration within each application.
+%For each application configuration we measure the slowdown induced by the execution of resilience logic and roll-backs due to error detections, as well as the distribution of magnitudes of error in the application's output.
+%We then demonstrate how these measurements can be used to configure the resilience mechanisms used with each application to minimize overhead while meeting the user's accuracy requirements.
+%Since in practice users do not have a clear idea of the level of accuracy needed from an algorithm's results, we demonstrate a novel alternative method to specify accuracy requirements as the level noise in the application's input required to deviate the output by a given magnitude (i.e.
 similar to backward error analysis).
-Finally, we extrapolate from the small-scale application runs considered in our fault injection campaign (the many runs required to cover the large space of resilience configurations limited the length of each run) to application runs with 1 million CPU hours via three different models of error propagation.
-Our results demonstrate that when applications use all the relevant resilience methods, configured for maximum efficiency, the overhead of resilient execution on expected Exascale designs is low.
+%Finally, we extrapolate from the small-scale application runs considered in our fault injection campaign (the many runs required to cover the large space of resilience configurations limited the length of each run) to application runs with 1 million CPU hours via three different models of error propagation.
+%Our results demonstrate that when applications use all the relevant resilience methods, configured for maximum efficiency, the overhead of resilient execution on expected Exascale designs is low.
 
@@
 Section~\ref{sec:soft_faults} discusses the problem of soft errors and presents the fault injection model we use to simulate them in our experiments.
-The applications to which we apply our resilience methodology are described in Section~\ref{sec:apps}.
-Section~\ref{sec:res_tech} summarizes the major routines of these applications and describes how these resilience techniques are applied to them.
-Further, it experimentally evaluates the resilience techniques, describing their overheads and the accuracy levels of the protected routines at a wide range of possible error rates.
-The effectiveness of our methodology for protecting the full applications is evaluated in Section~\ref{sec:eval}.
+The applications we focus on are described in Section~\ref{sec:apps}.
+Section~\ref{sec:res_tech} summarizes the major routines of these applications and describes the resilience techniques that are applied to them.
+It further details our methodology to evaluate application performance and resilience in the face of faults.
+Finally, Section~\ref{sec:eval} describes our procedure to use small-scale runs of our three target applications to predict the performance and resilience properties of large-scale runs of applications within their respective domains.
 
@@
 
-Cosmic rays or alpha particles hitting the computer circuits and accumulating a sufficient amount of energy may invert the device logic states from "0" to "1" or from "1" to "0"~\cite{Ziegler:1996:TCR:226354.226356,baumann:2005}.
-These inversions are termed as soft errors or soft faults.
-The shrinking processor and memory feature size and the lower threshold voltage coming with relentless manufacturing technology scaling make modern high performance computers are highly vulnerable to soft errors which propagate as the program is being executed.
-Prior studies have demonstrated that soft errors occurring in all the important bits that are required for Architecturally Correct Execution (ACE)~\cite{avf:2003} eventually change the program outcome, i.e., manifest errors to the application.
-The ACE bits include critical information to keep the program logic such as the program counter register for a committed instruction.
+The states of logic devices can be corrupted due to fluctuations in circuit voltages or strikes by charged particles from chip packaging or due to cosmic rays~\cite{Ziegler:1996:TCR:226354.226356,baumann:2005}.
+These ``soft faults'' are expected to become increasingly common as circuit scaling further shrinks circuit feature size and the threshold voltage.
+Prior studies have demonstrated that soft errors in processor state required for Architecturally Correct Execution (ACE)~\cite{avf:2003,mpi_ser:reed:2004} eventually change the program outcome.
+%The ACE bits include critical information to keep the program logic such as the program counter register for a committed instruction.
 In the past decades, soft errors have brought significant losses on commercial servers and high performance computers such as the recall of Sun servers due to memory errors~\cite{baumann:2005} or the cache corruptions in the ASCI Q system~\cite{asciQSER:2005}.
@@
 
-We use a fault injection infrastructure~\cite{relax:2010} that simulates the architecture-level effects of soft faults.
-It works by compiling application into the Low Level Virtual Machine (LLVM) typed bytecode~\cite{LLVM}.
-This bytecode is transformed to allow a bit flip in each instruction's output with some probability $P$.
-The time between injections is thus exponentially distributed with period $\frac{1}{P}$.
+We simulate soft faults using a fault injection infrastructure~\cite{relax:2010} that simulates the architecture-level effects of soft faults.
+It works by compiling application into the Low Level Virtual Machine (LLVM) typed bytecode~\cite{LLVM} and transforming it to allow bits in instruction outputs to be flipped with some probability.
+%The time between injections is thus exponentially distributed with period $\frac{1}{P}$.
 Our fault injector is a high-level approximation of the physical corruptions caused by soft errors that strikes a balance between model accuracy and cost, as do related approaches~\cite{fault_injection:iyer:1997, avf:2003}.
-Alternative models, such as injection into gate-level processor models~\cite{sesee:2004} or neutron beam irradiation~\cite{freq_dep:irom:2004} are possible but their significantly higher cost makes them impractical for conducting large-scale fault injection campaigns such as the ones reported in this paper.
-
-\sui{As a metric for quantifying errors in output caused by soft faults, we perturb inputs to numeric applications ...}
+Alternative models, such as injection into gate-level processor models~\cite{sesee:2004} or neutron beam irradiation~\cite{freq_dep:irom:2004} are possible but too expensive for large-scale fault injection campaigns such as the ones in this paper.
 
@@
 We evaluate our resilience methodology by applying it to the following three applications.
-Since they spend most of their execution time in library routines, we focus our resilience efforts on protecting these routines, which are described in Section~\ref{sec:res_tech:err_det:algo}.
-
-%\greg{This section needs to include an experimental breakdown of where the routines spent their time. We also need to describe the types of inputs we'll be giving them in our experiments.}
+Since they spend most of their execution time in library routines, we focus our resilience efforts on protecting these routines, as described in Section~\ref{sec:res_tech:err_det:algo}.
 
@@
 \label{sec:apps:lasso}
-The ADDR method solves under-unconstrained linear problems $Ax=b$ for $x$ ($A$ has fewer rows than columns) while minimizing the function $f(x)$.
-This is an iterative algorithm uses 64-bit precision and that spends most of its time in the following linear algebra operations, matrix-matrix multiplication, matrix-vector multiplication, rank-k update and Cholesky factorization, using the implementations of these routines from the GNU Scientific Library~\cite{gsl:2011}.
-We focus on a specific application that uses the ADDR method to solve Lasso problems, which are defined as minimizing the function $\frac{1}{2} \left|| Ax - b \right||_2^2 + \lambda*\left|| x \right||_1$.
-It is parallelized using MPI and our experiments focus on 4-processors runs of ADDR which computes linear systems of size $\{800, 1200, ..., 3600\} \times 500$ (800 to 3600 measurements with 500 variables each) and in the experiments below are denoted $S800$, $S1200$, ..., $S3600$.
-The values in $A$ and $b$ are generated by sampling a normal distribution with a mean of 0.
-According to our experiments, the running time of this solver is not dependent on the kind of distribution the inputs are sampled from.
+ADDR is an implementation of an iterative method for solving under-unconstrained linear problems $Ax=b$ for $x$ ($A$ has fewer rows than columns) while minimizing some function $f(x)$.
+It represents the linear solver application domain.
+It uses 64-bit precision and spends most of its time in the following linear algebra operations from the GNU Scientific Library~\cite{gsl:2011}: matrix-matrix multiplication (MMM), matrix-vector multiplication (MVM), rank-k update (RK) and Cholesky decomposition (CD).
+We use ADDR to solve Lasso problems, where the function $\frac{1}{2} \left|| Ax - b \right||_2^2 + \lambda*\left|| x \right||_1$ is minimized.
+It is parallelized using MPI and our experiments focus on 4-processors runs with linear systems of size $\{800, 1200, ..., 3600\} \times 500$ as input (denoted $S800$, ..., $S3600$ below).%(800 to 3600 measurements with 500 variables each) and in the experiments below are denoted $S800$, $S1200$, ..., $S3600$.
+The values in $A$ and $b$ are generated by sampling a normal distribution with a mean of 0. \greg{What's the $\sigma$?}
+According to our experiments, the running time of this solver is not dependent on the distribution the inputs are sampled from.
 
@@
 
-DRC is a sequential program generates filters for HiFi audio systems to compensate for the reflection of sounds in a room, using impulse response measurements of the audio equipment and considering the positions of the listeners.
-The inputs are stored in Pulse Code Modulation (PCM) format, which is an array of 32-bit floating point numbers representing the samples at each sampling time step and the computations are performed using 32-bit precision.
-Although the algorithm is divided into many phases the most computationally intensive step consists of the GSL's implementation of the Fast Fourier Transform and a DRC-internal implementation of Finite Impulse Response filter generation.
-This program accepts user-controllable configurations to decide the width and number of iterations of FFT transforms it uses to perform various passes on the input file, such as windowing, deconvolution, pre- and post-filtering.
+DRC is a sequential program that generates filters for HiFi audio systems to compensate for the reflection of sounds in a room, using impulse response measurements of the audio equipment and considering the positions of the listeners.
+It represents the signal processing application domain.
+DRC's inputs are stored in Pulse Code Modulation (PCM) format, which is an array of 32-bit floating point numbers representing the samples at each time step and the computations are performed using 32-bit precision.
+Most execution time is spent in the GSL implementation of Fast Fourier Transform (FFT) and a DRC-internal implementation of Finite Impulse Response (FIR) filter generation.
+%This program accepts user-controllable configurations to decide the width and number of iterations of FFT transforms it uses to perform various passes on the input file, such as windowing, deconvolution, pre- and post-filtering.
 The input used in this paper is an audio file of size 768 kilobytes, which is sampled at the rate of 30Khz, 40Khz, ..., 90Khz.
@@
 \label{sec:apps:hattrick}
-Hattrick is a sequential application simulates the motion of bodies under the effects of gravity using 64-bit precision.
-It is designed to help discover extra-solar planets by inferring their existence from Transit Timing Variations, where the difference between the times when a planet passes in front of its host star is used to infer the existence and properties of other planets in the system.
-This program spends most of its execution time in the GSL solver for Ordinary Differential Equations to solve the system's equations of motion.
+Hattrick is a sequential application that simulates the motion of bodies under the effects of gravity to help discover extra-solar planets by inferring their existence from Transit Timing Variations.
+%, where the difference between the times when a planet passes in front of its host star is used to infer the existence and properties of other planets in the system.
+It represents the n-body simulation application domain.
+Hattrick uses 64-bit precision and spends most of its execution time in the GSL solver for Ordinary Differential Equations to solve the system's equations of motion using the Runge-Kutta method (RK).
 A given input is described using three parameters: $P$=number of planets, $T$=amount of time to simulate, and $A$=the accuracy target.
 In our experiments we considered the following four inputs: $P2T2090A15$, $P2T3090A15$, $P2T4090A15$ and $P3T2090A11$, where an $A15$ and $A11$ denote accuracy targets of $1e-15$ and $1e-11$, respectively.
+\greg{Does it take transit timings as inputs? What are you perturbing?}
 
@@
 
-In this paper we consider a comprehensive resilience strategy that combines various types of error detection and recovery.
-Error detection techniques are discussed in Section~\ref{sec:res_tech:err_det}, with Section~\ref{sec:res_tech:err_det:algo} focusing on algorithmic techniques to detect errors in the numeric portions of application state, while Section~\ref{sec:res_tech:err_det:mem} focuses on protection of data structures.
-Section~\ref{sec:res_tech:cr} then considers recovery techniques based on both checkpoint-restart, which offers lower-overhead but higher-latency recovery, as well as replication of key pointers, which reduces performance but enables fast recovery.
-Finally, Section~\ref{sec:res_tech:eval} presents an experimental evaluation of the performance and accuracy properties of the resilient versions of the major routines used by our target applications, under a wide range of inputs and error injection rates.
+Our case-study focuses on a comprehensive resilience strategy that combines various types of error detection and recovery.
+Algorithmic error detection techniques of numeric state are discussed in Section~\ref{sec:res_tech:err_det}, while Section~\ref{sec:res_tech:pointers} focuses on protection of data structure pointers via replication.
+Section~\ref{sec:res_tech:cr} then considers recovery techniques based on both checkpoint-restart.%, which offers lower-overhead but higher-latency recovery, as well as replication of key pointers, which reduces performance but enables fast recovery.
+Finally, Section~\ref{sec:res_tech:eval} describes our methodology to analyze the resilience and performance properties of these mechanisms in the above applications across a broad range of configurations.
+This information is used in Section~\ref{sec:eval} to configure the resilience methods for modeled large-scale application runs to maximize the high performance across a broad range of fault rates and while satisfying user accuracy requirements.
 
@@
 
-%\sui{Performance experiments inserted here: running algorithmic checkers across different input sizes and fault rates: 62x62 to 500x500 matrices for MM, RK, CD, MV checkers; FFT sizes of $2^8$ to $2^{24}$ for the FFT checker; Error bars should apply here, especially for the Cholesky Decomposition checker}
-
-Each algorithm used by the above application was enhanced with an algorithm-specific checker that validated its results by exploiting some algorithmic identity.
-Each checker computes some aggregate property of the algorithm's results using two different algorithms to produce vectors $v$ and $v'$.
-If their relative difference by more than some threshold $\tau$, the checker declares that an error has occurred.
-The relative difference is computed as $\frac{\left|| v-v' \right||}{max(\left||v\right||, \left||v'\right||)}$.
+Algorithms used by the above application were enhanced with checkers of their algorithmic invariants.
+Each checker computes some aggregate property of the algorithm's output in two different ways to produce values $v$ and $v'$.
+An error is detected if their relative difference $\frac{\left|| v-v' \right||}{max(\left||v\right||, \left||v'\right||)}$ is larger than some threshold $\tau$.
 Each error detection causes the routine's execution to be restarted using the rollback mechanism described in Section~\ref{sec:res_tech:cr}.
 
-\vspace{-5pt}
+%\vspace{-5pt}
 \paragraph{Matrix-matrix multiplication (MMM)}
-The operation $A \cdot B$ is checked using a matrix vector multiplication (MVM), via the identity: $(A \cdot B) \cdot x = A \cdot (B \cdot x)$, where $x$ is an error-checking vector.
-Since we set the check vector $x$ to be all 1s this is equivalent to both sides computing the sums of the columns of matrix $A \cdot B$ using a single MVM on the left and two MVMs on the right.
-Most errors in the computation of $A \cdot B$ will affect the sums of its columns and will thus be detected.
-This check is efficient because the checker is asymptotically faster than MMM, with MVM taking $O(n^2)$ time and MMM $O(n^3)$ time.
+$A \cdot B$ is checked using a matrix vector multiplication, via the identity: $(A \cdot B) \cdot x = A \cdot (B \cdot x)$, where $x$ is an error-checking vector (we use the vector of all 1s).
+%Since we set the check vector $x$ to be all 1s this is equivalent to both sides computing the sums of the columns of matrix $A \cdot B$ using a single MVM on the left and two MVMs on the right.
+%Most errors in the computation of $A \cdot B$ will affect the sums of its columns and will thus be detected.
+The checker is asymptotically faster than MMM, with MVM taking $O(n^2)$ time and MMM $O(n^3)$ time.
 
-\vspace{-5pt}
+%\vspace{-5pt}
+\paragraph{Symmetric Rank-K update (SYRK)}
+Since this is a special case of MMM where $B = \alpha A \cdot A^T + \beta B$, %where an $n \times k$ matrix $A$ is multiplied by itself to produce an $n \times n$ matrix with rank $k$, which is incremented into the $n \times n$ matrix $B$.
+%Its results are checked using the same algorithm as MMM by checking the identity $B x = \alpha A \cdot (A^Tx) + \beta B x$, where $x$ is the vector of all 1's.
+%This check runs in time $O(n^2)$, as compared to the $O(n^3)$ cost of SYRK.
+we use the same check.
+
+%\vspace{-5pt}
 \paragraph{Matrix-vector multiplication (MVM)}
-The MVM algorithm $Ax=b$ is checked using a similar identity $(x^TA)x = x^T(Ax) = x^Tb$.
-We again set $x$ to be the vector of 1's, making the product $x^TA$ equivalent to sum of the columns of matrix $A$.
-Although the computation of $x^TA$ involves additions rather than multiplications, the complexity of computing $x^TA$ is $O(n^2)$, the same as the original MVM operation.
+$Ax=b$ is checked using a similar identity $(x^TA)x = x^T(Ax) = x^Tb$.
+%We again set $x$ to be the vector of 1's, making the product $x^TA$ equivalent to sum of the columns of matrix $A$.
+The complexity of computing $x^TA$ is $O(n^2)$, the same as the original MVM but using additions rather than multiplications.
 %While this cost can be reduced significantly by computing $x^TA$ once and reusing it in subsequent multiplications by the same matrix, we did not consider it in this paper because MVM took up little application execution time.
-However, for applications such as ADDR where the MVM is applied to same matrix many times with different vectors, the column sums can be reused, amortizing the cost of this computation.
-Our experimental evaluation below thus focuses on the overhead of the $x^T(Ax)$ computation, which has $O(n)$ complexity.
+Since in ADDR MVM is applied many times to the same matrix with different vector, the vector $x^TA$ can be reused, amortizing the cost of computing it.
+Our experiments below thus focus on the cost of computing $x^T(Ax)$, which has $O(n)$ complexity.
 
-\vspace{-5pt}
-\paragraph{Symmetric Rank-K update (SYRK)}
-The SYRK update is a special case of matrix-matrix multiplication $B = \alpha A \cdot A^T + \beta B$, where an $n \times k$ matrix $A$ is multiplied by itself to produce an $n \times n$ matrix with rank $k$, which is incremented into the $n \times n$ matrix $B$.
-Its results are checked using the same algorithm as MMM by checking the identity $B x = \alpha A \cdot (A^Tx) + \beta B x$, where $x$ is the vector of all 1's.
-This check runs in time $O(n^2)$, as compared to the $O(n^3)$ cost of SYRK.
+%\vspace{-5pt}
+\paragraph{Cholesky Decomposition (CD)}
+CD decomposes matrix $A$ into $L \cdot L^T$ where $L$ is lower-triangular with a positive diagonal.
+It is checked via the identity $Ax = L \cdot (L^T x)$, which at $O(n^2)$ complexity is cheaper than the $O(n^3)$ complexity of the deterministic CD algorithm.
+GSL implements an iterative algorithm that runs faster than $O(n^3)$ time but as our experiments show, our checker is still significantly faster.
 
-\vspace{-5pt}
-\paragraph{Cholesky Decomposition (CD)}
-In this decomposition matrix $A$ is decomposed into $L \cdot L^T$ where $L$ is a lower-triangular matrix with a positive diagonal.
-This routine is checked just like MMM by checking the identity $Ax = L \cdot (L^T x)$, which is cheaper than the $O(n^3)$ complexity of the deterministic CD algorithm.
-The GSL implements an iterative algorithm that runs faster than $O(n^3)$ time but as shown in the experiments below, it is significantly slower than our checker.
-
-\vspace{-5pt}
+%\vspace{-5pt}
 \paragraph{Fast Fourier Transform (FFT)}
-The FFT algorithm decomposes a function into a sum of sine waves of different frequencies: $f(x) = \sum_{n=0}^{N-1} x_n e^{-i2\pi k n / N}$ for some constant $k$.
-We check the results of FFT by using Parseval's theorem, which states that $\sum_{n=0}^{N-1} \left| x[n] \right|^2 = \frac{1}{N} \sum_{k=0}^{N-1} \left| X[k] \right|^2$, where $x$ is the original function and $X$ is its transform.
+Decomposes a function into a sum of sine waves of different frequencies: $f(x) = \sum_{n=0}^{N-1} x_n e^{-i2\pi k n / N}$ for some constant $k$.
+It is checked using Parseval's theorem: $\sum_{n=0}^{N-1} \left| x[n] \right|^2 = \frac{1}{N} \sum_{k=0}^{N-1} \left| X[k] \right|^2$, where $x$ is the original function and $X$ is its transform.
 Intuitively it means that the energy of the original function is preserved by the transform.
-GSL implements optimized radix 2, 3, 4, 5, 6 and 7 subtransform routines as well as a general radix-N FFT routine.
+%GSL implements optimized radix 2, 3, 4, 5, 6 and 7 subtransform routines as well as a general radix-N FFT routine.
 The subtransforms are optimized and consume $O(n log(n))$ time and the general one consumes $O(n^2)$ time.
 Therefore, GSL has a scheduling algorithm that factorizes the length of the input FFT, running the subtransforms whenever possible, using the fallback general radix-N routine only for the remaining factors.
-The theorem applies to FFT of any radix and provides an $O(n)$ time check for the $O(n log(n))$ or $O(n^2)$ FFT algorithms.
+%The theorem applies to FFT of any radix and provides an $O(n)$ time check for the $O(n log(n))$ or $O(n^2)$ FFT algorithms.
+This check runs in $O(n)$ time, as compared to $O(n log(n))$ or $O(n^2)$ for the FFT algorithm itself (depends on the type of FFT problem).
 
-\vspace{-5pt}
+%\vspace{-5pt}
 \paragraph{Finite Impulse Response Filter Generation (FIR)}
-The FIR filter generation algorithm used in DRC generates a series of samples over the function $sinc(x)=\frac{sin(x)}{x}$
+The DRC FIR filter algorithm generates a series of samples over the function $sinc(x)=\frac{sin(x)}{x}$
 %(\text{if}~x \ne 0)$ and $sinc(x)=1 (\text{if}~x=0)$
 and modulates it with a Blackman window.
-This function has the property that $\int_{-\infty}^{\infty} sinc(x)dx = 1$.
-The Blackman window - modulated output of this routine preserves this same property for most of the time, making it possible to check the output by computing the sum and comparing it to 1.
+It is checked using the invariant $\int_{-\infty}^{\infty} sinc(x)dx = 1$, which is preserved by the Blackman window in most cases.
 Computing the sum requires $O(n)$ additions, as compared to the $O(n)$ trigonometric function evaluations of the non-iterative FIR generation algorithm.
@@
 
-\vspace{-5pt}
+%\vspace{-5pt}
 \paragraph{Runge-Kutta PDE Solver (RK)}
-RK is a 4-th order Runge-Kutta method for solving Ordinary Differential Equations of the form $\frac{dy}{dx} = f(y, x)$.
-It advances the variable $x$ by steps of size $h$ and uses estimates of the derivative $\frac{dy}{dx}$ at each point $x$ to find the value of $y$ at the next point $x+h$.
-The implementation of RK in GSL uses adaptive step-size control where it simultaneously performs the RK method with two step sizes $h$ and $\frac{h}{2}$ (more precise).
-If the difference between the results obtained with the two step sizes are larger than a threshold $\tau$, the algorithm switches to the smaller step size to maintain accuracy and resumes.
+A 4-th order Runge-Kutta method for solving Ordinary Differential Equations of the form $\frac{dy}{dx} = f(y, x)$.
+It advances the variable $x$ by steps of size $h$ and computes the value $y$ at the next point $x+h$ using the derivative $\frac{dy}{dx}$ at $x$.
+The GSL implementation of RK uses adaptive step-size control where it simultaneously uses two step sizes $h$ and $\frac{h}{2}$ (more precise).
+If the difference between the two computations exceeds a threshold $\tau$, it switches to the smaller step size to maintain accuracy.% and resumes.
 If it is smaller than $\frac{\tau}{2}$, the algorithm switches to a larger step size.
-Because soft errors will usually cause significant differences between the two computations, the adaptive step size control algorithm is naturally fault tolerant.
-It will react to such errors by rolling back, temporarily increasing its step size and then reverting to the original step size when it observes that this is safe.
-As such, we utilize this algorithm to protect the computations within individual RK steps from soft errors.
-However, computations in user-provided function derivative functions are not checked since their invariants are notknown.
+Because soft faults usually cause significant differences between the two computations, the adaptive step size control algorithm naturally tolerates them by temporarily increasing its step size and then reverting to the original step size when it observes that this is safe.
+We thus protect the results of user-provided derivative functions within RK's steps using this algorithm.
+%However, computations in user-provided function derivative functions are not checked since their invariants are not known.
 
-%Experiment 2: Measure the error in the outputs of the algorithms with and without the algorithmic detectors.
-%First, show a probability distribution histogram for all the codes at one error rate, input size and algorithmic detection threshold.
-%Then show a summary of the entire space using the same heatmaps that we're using for final application results.
-%Put the input size and error rate on the y-axis and the detection threshold on the x-axis.
-%The color in each box should be the fraction of entries in the output that have an error of a given magnitude.
-%Show a separate plot for a few different magnitudes.
-
-% 4.1.2 is way too verbose, vague and not understandable
-%\subsubsection{Memory Fault Detection}
-%\label{sec:res_tech:err_det:mem}
-%
-%Detecting errors in application data structures requires checkers dedicated for this purpose.
-%Depending on the algorithm's properties different error detectors are appropriate.
-%In our work we considered the use of access protection hardware and replication of key data.
-%
-%Since modern systems use 64-bit addressing the range of values that can be assumed by a given pointer is much larger than the amount of memory that can ever be assigned to %an application.
-%This means that there is a very high probability that the corruption of a given pointer will cause it to point to an unallocated address.
-%Once this address is accessed the resulting violation will be detected by the memory protection hardware and communicated to the application via a Segmentation Fault or %Bus Error signal.
-%This mechanism is cheap but suffers from a long delay between an error and its detection.
-%
-%An alternative mechanism is to replicate key application variables and to compare the values of the replicas on each access to the variable.
-%While more expensive, this approach is effective if the number of key variables is small and the cost of rolling back computation is high.
-%
-%The low cost of hardware memory protection mechanisms means that they are deployed in protecting all our routines.
-%In contrast, replication is only used to protect the RK routine because it is more sensitive to soft errors than the others and the low detection latency provided by %replication ensures that any state corruptions are quickly corrected and have little impact on execution time.
-%This is important for resilience as well as performance since slowdowns due to frequent rollbacks increase the total number of errors the application is exposed to.
+\subsection{Detection and Tolerance of Pointer Corruptions}
+\label{sec:res_tech:pointers}
+In 64-bit address spaces corruptions of pointers almost always cause them to refer to un-allocated addresses.
+Any accesses via such pointers will be detected by the hardware and result in a Segmentation Fault or Bus Error.
+In applications that uses relatively few pointers we detect pointer errors by using these hardware checks and tolerate them via checkpoint-restart.
+This is done for ADDR and DRC.
+Applications that use pointer-based data structures may experience frequent rollbacks due to pointer errors.
+The cost of these rollbacks can be reduced by triplicating the pointers, which is done for Hattrick.
+The different pointer replicas can be checked on either some or all accesses.
+The former takes less time while the latter has a shorter detection latency.
+Every time an error is detected it is immediately corrected, with a rollback required only in the very rare case where multiple replicas are corrupted.
 
@@
 \label{sec:res_tech:cr}
-Since bit-flip errors may corrupt the address of a load or store instruction and a bit-flip in a 64-bit address can easily become an invalid value pointing to
-unallocated memory, a Segmentation Fault or Bus Error can be easily detected by the operating system.\\
-
-When errors are detected we employ a recovery method based on checkpoint-restart, where routine state is recorded at key locations and whenever an algorithmic or memory error is detected we roll back to that point in its execution.
-Rollbacks are implemented via the \texttt{sigsetjmp}/\texttt{siglongjmp} \cite{sigsetjmp:1997} routines.
+We tolerate detected but uncorrectable errors via checkpoint-restart.
+Each routine's state is recorded at key locations and we roll back to that execution point whenever an error is detected.
+Rollbacks are implemented via \texttt{sigsetjmp} and \texttt{siglongjmp}. %\cite{sigsetjmp:1997} routines.
 \texttt{sigsetjmp} records the application's execution state (registers and stack pointer) at the time of the checkpoint.
-On the detection of an error \texttt{siglongjmp} returns the application to the same execution point, unrolling any function calls between the checkpoint and the error detection.
-This method is useful for rolling back to checkpoints recorded in the same function or in a caller function.
+\texttt{siglongjmp} returns the application to the same execution point, unrolling any function calls between the checkpoint and the error detection.
+This light-weight method can roll back to checkpoints recorded in the same function or in a caller function.
 When execution returns to the \texttt{sigsetjmp} call the checkpointed portions of the routine's state are recovered and execution repeats.
-% The overhead of this API call is having to copy the processor state into memory, therefore installing them between computation-intensive method calls adds to negligible performance degradation.
-
-Checkpoint data is protected using two different mechanisms:
-Pointers and scalars are replicated such that the checkpoint contains two copies of each.
-Arrays of floating point numbers are protected via a block-checksum-based error correcting approach.
-By treating the input array as a matrix and comparing row sums and column sums of the submatrices of that matrix, it detects the existence of errors and fixes them.
-This algorithm takes an array of numbers as input, divides it into blocks of $N \times N$ numbers and for each block it computes and stores the row and column sums, which are used for recovery.
+Pointers and scalars within checkpoints are protected by duplicating them.
+Arrays of floating point numbers are protected via a block checksum-based error correction.
+Each array is treated as a matrix and divided into blocks of $N \times N$ numbers.
+A checksum is computed for each block's rows and columns and used to detect and correct any corrupted entries.
+(\greg{Is this a configurable parameter in our experiments?})
 
 The structure of each routine determines how checkpointing is deployed.
-All routines except RK are monolithic code blocks that take a relatively large amount of data as input, operate on it and then return.
-In contrast, RK takes in very little input and operates as a long sequence of steps, each of which involves a call to a user-provided derivative routine.
-As such, we checkpoint the state of non-RK routines immediately after the call to the routine and include in this checkpoint the routine's arguments and contents of the floating point input buffers.
-The one exception is the FIR routine, which takes no input and thus requires only the processor's state to be checkpointed.
-Because RK's checkpoints are very small and it is more sensitive to errors, it is checkpointed periodically during its execution with a fixed number of steps between checkpoints.
-As our experiments in \ref{sec:res_tech:eval:rk} show, the choice of the RK checkpoint period significantly affects its efficiency with longer periods suffering from more overhead due to rollbacks and short period spending more time checkpointing.
+For routines that take a relatively large amount of data as input and return the result we record a checkpoint at the routine's start, including its arguments and inputs, and roll back to this starting point on any error during the routine.
+%The one exception is the FIR routine, which takes no input and thus requires only the processor's state to be checkpointed.
+In contrast, RK in Hattrick takes in very little input and operates as a long sequence of steps, each of which involves a call to a user-provided derivative routine.
+Because RK's checkpoints are very small and it is more sensitive to errors, it is checkpointed periodically during its execution with a configurable number of steps between checkpoints.
+%Our experiments in \ref{sec:res_tech:eval:rk} show that the choice of the RK checkpoint period significantly affects its efficiency, with longer periods suffering from more overhead due to rollbacks and short period spending more time checkpointing.
+As RK checkpoint periods grow, the time spent in checkpointing shrinks while the cost of each rollback grows.
 
@@
 
-\subsubsection{Linear Algebra and Integer Routines}
-\label{sec:res_tech:eval:la_int}
+To analyze the performance and resilience properties the key routines in ADDR, DRC and Hattrick we conducted a comprehensive fault injection campaign where we employed the above resilience techniques with each routine, as applicable,
+We used fault rates between 1e-11 and 1e-8 errors per clock cycle \greg{(is this correct)}, probing each application's behavior as it was impacted by either one or multiple errors in a single run.
+Our experiments span a wide range of input sizes and a broad range of resilience mechanism configurations, with 100 (\greg{???}) runs for each combination of application, input, fault rate and resilience configuration.
+Table~\ref{tbl:configs} specifies the techniques applied to each routine and its configuration space.
+\begin{table}
+  \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    \multicolumn{2}{c}{Routine}          & Algorithmic Detector             & Checkpointing         & Pointer Replication  \\
+    \hline
+    \multirow{4}{*}{ADDR}      & MM   & Linear encoding                     & Inputs                & No \\
+                               & SYRK & Detection Thresholds:               & Inputs                & No \\
+                               & MVM  & 1e-4 to 1e-7                        & Inputs                & No \\
+                               & CD   &                                     & Inputs                & No \\
+    \hline
+    \multirow{2}{*}{FRC}       & FFT  & Parseval's theorem.                 & Inputs                & No \\
+                               & FIR  & Sum conservation.                   & Inputs                & No \\
+                               &      & Detection Thresholds: 1e-6 to 3e-8. &                            \\
+    \hline
+    \multirow{3}{*}{Hattrick}  & \multirow{3}{*}{RK} & \multirow{3}{*}{Variable step-size} & Periodic Timesteps    & None  \\
+                               &      &                                   & Period:  1, 1e3, 1e5 & All pointers checked at one code location \\
+                               &      &                                   &                       & All pointers checked on each use \\
+    \hline
+  \end{tabular}
+  \caption{The resilience techniques applied to each major routine of each application.}
+  \label{tbl:configs}
+\end{table}
 
-This section evaluates the routines MMM, MVM, SYRK, CD, FFT and FIR.
+Section~\ref{sec:res_tech:eval:perf} summarizes the impact of faults on routine execution time, while Section~\ref{sec:res_tech:eval:accuracy} focuses on the accuracy of routine outputs.
+Our experiments were performed on compute nodes with two 2.33 GHz dual-core Xeon E5345 processors with 4GB Ram and RHEL 4 Linux.
 
-\paragraph{Overheads}
-%\label{sec:res_tech:eval:ovhd}
-
-%\begin{figure}[ht!]
-%\centering
-%\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp1_linalg}
-%\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp1_fft}
-%\caption{Overhead of resilience techniques when errors are not injected}
-%\label{fig:routine_detect_ovhd}
-%\end{figure}
-
-%Experiment 1: Measure performance degradation of using error detectors as input sizes increase.
-In case of RK if we can run a version that doesn't do the multiple step sizes, do so but if this is difficult then don't worry about it.
-We evaluate the costs of the above algorithmic error checkers by measuring their costs and effectiveness when running at different error rates on different types of inputs.
-%Figure \ref{fig:routine_detect_ovhd} measures the overheads of employing the algorithmic error detectors when no errors are injected,
-We report the averages of 100 runs.
-The linear algebra routines were executed on $n \times n$ square matrices where $n$ varies from 100 to 1000.
-FFT ran on inputs vectors with 128K to 16M entries and FIR on window widths of 1M to 8M.
-RK ran on an integral with intervals $[0, t_1]$ where $t_1 \in \{10, 15, 20, 25, 30, 50, 75, 100\}$.
-Its ODE system is a second-order nonlinear Van der Pol oscillator equation.
-Our experiments show that the cost of algorithmic error detection is low (< 10\%) for small and drops with increasing input sizes.
+\subsubsection{Performance Impact}
+\label{sec:res_tech:eval:perf}
 
@@
 
-Figure~\ref{fig:routine_all_ovhd} shows in the bars the overhead of all resilience techniques with no fault injection, separating into overheads due to algorithmic error detection, the checkpointing of routine inputs as well as checkpointing processor state.
-The data shows that checkpointing the input of routines consumes more time than algorithmic error detection and that the cost of checkpointing processor state is negligible.
-In all cases, these overheads decrease as the input set size increases.
+Figure~\ref{fig:routine_all_ovhd} reports the relative increase in running time of MM, SYRK, MVM, CD, FFT and FIR as their input sizes are varied and no faults are injected.
+MM, SYRK, MVM and CD were executed on square matrixes.
+%The linear algebra routines were executed on $n \times n$ square matrices where $n$ varies from 100 to 1000.
+%FFT ran on inputs vectors with 128K to 16M entries and FIR on window widths of 1M to 8M.
+The experiments show that algorithmic error detection has a low overhead ($<10\%$) that shrinks with increasing input sizes.
+The cost of checkpointing inputs is typically somewhat higher but also shrinks with increasing input size.
+Checkpointing processor state (\texttt{sigsetjmp}) has a negligible cost.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph{Output Accuracy}
-%\label{sec:res_tech:eval:accu}
+\greg{Where is data for RK}
+%RK ran on an integral with intervals $[0, t_1]$ where $t_1 \in \{10, 15, 20, 25, 30, 50, 75, 100\}$.
+%Its ODE system is a second-order nonlinear Van der Pol oscillator equation.
 
@@
 \centering
-\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp2_1_Example.png}
-\caption{Error magnitude histogram showing the effectiveness of the error detector}
-\label{fig:algo_err_dist}
+\includegraphics[width=1.00\columnwidth]{figs/4_1_2_Exp2_Expected_Running_Time_Needed.png}
+\caption{Overhead of resilience techniques for RK when errors are injected}
+\label{fig:rk_routine_exp_exec}
 \end{figure}
+
+When errors are injected there is an additional cost of rolling the computation back due to error detections.
+Figure~\ref{fig:rk_routine_exp_exec} shows the effect on RK (similar for other routines (\greg{is this true})), when between 0 and 10 errors per run are injected and RK is protected using Checkpointing, Replication or both.
+(\greg{What type of replication was used and what checkpointing period?})
+The data shows that the best configuration varies with the number of errors.
+\texttt{NoRe\_NoCkpt} (no resilience) is the most efficient option when only a few errors are injected but above 2.5 errors per run this configuration is aborted and restarted so often that resilience techniques are needed to achieve the best performance.
+Replication without checkpointing is more effective than checkpointing without replication and for large error counts (>8 per run), both must be used.
+
+\subsubsection{Result Accuracy}
+\label{sec:res_tech:eval:accuracy}
+
+Even though an application may complete its execution, errors injected into its numeric state may produce incorrect output.
+In each experimental configuration we measured the distribution of errors in application output, measured as the Root Mean Square (RMS) Error: $\sqrt{\sigma_i (x_i-x'_i)^2}/n$, where $x_i$ is the value of output element $i$ in a fault-free run, $x'_i$ its value in a fault-injected run and $n$ is the total number of output elements (e.g.
+size of the output matrix).
+
+Figure~\ref{fig:algo_err_dist} shows a sample of our results, focusing on the linear algebra routines (restarted upto 5 times after an error detection).
+It presents the probability distribution of the routine's results having errors of various magnitudes.
+Each tile corresponds to a different routine (vertical order) and different algorithmic error detection threshold used (horizontal order).
+Table~\ref{fig:algo_err_heatmap} specifies the exact threshold values used.
+Within each tile the vertical axis corresponds different error rates.
+The axis numbers are normalized to the execution time of the configuration with no resilience techniques (\greg{errors injected or not?}) and specify the average number of errors injected during this time period at this rate.
+The horizontal axis corresponds to the error magnitude ranging from 1e-14 to 1e-1.
+For each error injection count and output error magnitude we report the probability of the output having error of \emph{larger than} this magnitude as a different shade of gray, with darker shades indicating higher probabilities.
+Thus, darker plots indicate that error injections make larger output errors more likely and errors $\ge$ a smaller magnitude are always less likely than errors $\ge$ a larger one (shades grow lighter from left to right).
+
+As error count increases the probability of larger output errors rises.
+Further, tighter (smaller) error detection thresholds reduce this probability.
+Error distributions vary widely, with MM being most resilient to errors (the probability of having large errors is low) and CD least.
+These experiments guide the choice of how each routine's resilience technique is configured to meet user-specified accuracy bounds.
+
+Figures~\ref{fig:algo_err_dist} show the fraction of entries in the outputs of each routine (y-axis) that have an error of a given magnitude (x-axis) as error magnitudes span from 1e-14 to 1e-4.
+This figure is a focused view on few representative cases (each routine run on a single input size and error injection rate) across different resilience configurations: no error detection as well as three error detection thresholds.
+As the error magnitude increases the fraction of output entries with errors of that magnitude drops.
+Importantly, the fraction of erroneous entries is reduced at all magnitudes as a result of error detection and restart.
+Further, this reduction is more significant as the detection threshold shrinks since smaller magnitude errors are detected.
+
+Figure~\ref{fig:algo_err_heatmap} shows the same information but for all experimental configurations: (i) input sizes as above (ii) error injection rates of 1e-7 to 1e-10.
+The graph is a grid, with results for different routines listed vertically and different resilience configurations (no detection and the different detection thresholds shown in the table at the bottom) listed horizontally.
+The graph for each routine and resilience configuration presents the curves from Figure~\ref{fig:algo_err_dist} as a heatmap.
+The y-axis shows different routine input sizes and error injection rates, sorted according to the average number of error injections per non-fault injected run, in decreasing order from top to bottom.
+The x-axis shows the different error magnitudes and the shade of each point is the fraction of the routine's entries that have a given error magnitude, with darker shades indicating more such entries (each shade corresponds to a fraction that is some power of 10).
+Thus, each row of Figure~\ref{fig:algo_err_heatmap} represents the same information as each curve in Figure~\ref{fig:algo_err_dist}, with the shades of the squares representing the same information as the height of the line's points.
+
+%\begin{figure}[ht!]
+%\centering
+%\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp2_1_Example.png}
+%\caption{Error magnitude histogram showing the effectiveness of the error detector}
+%\label{fig:algo_err_dist}
+%\end{figure}
 
@@
 
-We now evaluate the effectiveness of our algorithmic detectors when errors are injected.
-In each case when an error is detected by our algorithmic detectors the routine's execution would restart from the beginning upto five restarts.
-Figures~\ref{fig:algo_err_dist} show the fraction of entries in the outputs of each routine (y-axis) that have an error of a given magnitude (x-axis) as error magnitudes span from 1e-14 to 1e-4.
-This figure is a focused view on few representative cases (each routine run on a single input size and error injection rate) across different resilience configurations: no error detection as well as three error detection thresholds.
-As the error magnitude increases the fraction of output entries with errors of that magnitude drops.
-Importantly, the fraction of erroneous entries is reduced at all magnitudes as a result of error detection and restart.
-Further, this reduction is more significant as the detection threshold shrinks since smaller magnitude errors are detected.
-
-Figure~\ref{fig:algo_err_heatmap} shows the same information but for all experimental configurations: (i) input sizes as above (ii) error injection rates of 1e-7 to 1e-10.
-The graph is a grid, with results for different routines listed vertically and different resilience configurations (no detection and the different detection thresholds shown in the table at the bottom) listed horizontally.
-The graph for each routine and resilience configuration presents the curves from Figure~\ref{fig:algo_err_dist} as a heatmap.
-The y-axis shows different routine input sizes and error injection rates, sorted according to the average number of error injections per non-fault injected run, in decreasing order from top to bottom.
-The x-axis shows the different error magnitudes and the shade of each point is the fraction of the routine's entries that have a given error magnitude, with darker shades indicating more such entries (each shade corresponds to a fraction that is some power of 10).
-Thus, each row of Figure~\ref{fig:algo_err_heatmap} represents the same information as each curve in Figure~\ref{fig:algo_err_dist}, with the shades of the squares representing the same information as the height of the line's points.
-
-The data shows that in general as the error detection threshold drops there are fewer entries that have an error of a given magnitude.
-However, there are exceptions where the fraction of erroneous entries rises as the detection threshold shrinks (e.g. MMM and FT between NoFT and Threshold 1).
-This is caused by the fact that we restart each routine on an error detection, which can causes the routine's execution to be exposed to additional error injections.
-As the threshold is further tightened the amount of error in the output shrinks, indicating that error detection is ultimately effective if it is configured to be sufficiently accurate.
-Another phenomenon that is observed is that as the number of errors injected in a given run decreases the degree of error in routine outputs also generally decreases.
-However, this is not a fully consistent phenomenon, with some configurations that have more error injections being less vulnerable than others with a higher number of injections because the former had a higher injection rate but the latter a larger input size.
-This indicates that while the number of injected errors is strongly correlated with overall application vulnerability to errors, other properties such as input size also play a role.
-
-%We dont have the following figure and there is no time to generate it.
-%Figure~\ref{fig:algo_err_heatmap} shows the same information across the entire experimental space.
-%The y-axis includes all algorithms, input sizes and error rates.
-%The x-axis includes several segments for different degrees of error checking: none, as well as checking with thresholds \greg{???} \sui{1e-06, 1e-07 and 1e-08}.
-In each segment the x-axis spans over error magnitudes from \greg{???} \sui{1e-14} to \greg{???} \sui{1e-01}.
-%The box associated with each combination of the four parameters the figure indicates the fraction of outputs that have an error of this magnitude by the shade of gray it is colored with, as indicated in the legend.
-%The data shows that the accurate of application results is directly related to the number of errors that impact a given application run, rather than their rate.
-%Further, it shows that a reduction of detection threshold is consistently effective at reducing the fraction of errors in each algorithm's output.
-%However, the effectiveness of algorithmic checks has a limit, with CD and FFT showing moderate error vulnerability even at threshold 1e-8.
-%\sui{(Rephrase) However, the algorithmic checkers cannot cover 100\% of the errors for CD, FFT and FIR, mainly fixing errors of large magnitudes which turned out to be effective in real applications (refer to the imperfect run rate plot for DRC.)}
-%MM, MV and SYRK are significantly less vulnerable across the full range of error magnitudes.
-%\sui{(Rephrase) The errors shown in MM, MV and SYRK are much more noticably fixed.}
-%\greg{should reference vulnerability numbers more specifically}
-
-%Experiment 3: Measure the fraction fraction of routine runs that are interrupted by the detection of an error.
-%Display this information via the same heatmaps as above.
-
-%\greg{For each experiment we need to include a discussion of what it means for how these checkers behave within the full-application runs.}
-
-%\sui{The checkers have performed well in reducing erroneous results from the application outputs.
-Checkers for linear algebra are useful in removing errors from Lasso, as one could see from the results.}
-
-%Those checkers can detect errors greater than a certain numeric threshold in those computations (determined by the user-defined error checker threshold).
-Once an error beyond the threshold has occurred, a re-calculation would be initiated.
-%\greg{You need to be more specific.
-Each routine produces a result, your checker produces a result.
-How are these results compared and how is the threshold used to determine if there is an error or not.}
-
-%In the following sections we would show that the error tolerance could not be arbitrarily small in order to be helpful in providing fault-tolerance when we take into account the checker itself may be unreliable since it's also vulnerable to soft errors.
-%\greg{How is this shown? We certainly do show tighter detection tolerances cause longer executions, which cause more vulnerability. However, this is not the same as what you've just described}
-
-%Although checkers are faster than the routines they protect, the unreliability of those checkers and the ensuing ``false alarms'' might trigger redundant re-calculations that are unnecessary, causing the application to run for much longer.
-This is specially true with an overly small error detection threshold.
-On the other hand, an overly loose detection threshold may fail to detect data corruption in output files.
-The results would be discussed in the results section.
-
-\subsubsection{Runge-Kutta}
-\label{sec:res_tech:eval:rk}
-
-The difference between RK routine and others motivates special discussion of its resilience properties.
-Since RK's adaptive step size mechanism is already effective at tolerating many numerical errors we used it as the error detector for this routine without adjusting its tolerance parameters.
-The configurable portions of the RK resilience method are thus the choice of checkpointing and/or replication strategy.
-Configurations \texttt{Ckpt} and \texttt{NoChkpt} denote whether checkpointing was used and configurations \texttt{Re} and \texttt{NoRe} denote whether replication was used.
-We consider checkpointing periods equal to 1, $10^2$, $10^3$, $10^4$ and $10^5$ calls to the RK stepping function.
-Further, we consider the following replication configurations: (i) ``No replication'', (ii) `Medium replication'', where routine pointers are replicated but the replicas of each pointer are compared at a single code location where it is used, and (iii) ``Maximum replication'', where the comparison is performed on every pointer use.
-
-\paragraph{Overheads}
-%\label{sec:res_tech:eval:ovhd}
-
-%\begin{figure}[ht!]
-%%\vspace{-20pt}
-%\centering
-%\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Overall_Breakdown_RK4}
-%%\vspace{-5pt}
-%\caption{Overhead of resilience techniques for RK when errors are not injected}
-%\label{fig:rk_routine_detect_ovhd}
-%\end{figure}
-
-%Figure~\ref{fig:rk_routine_detect_ovhd} shows the overhead of employing each of the four RK resilience techniques when no errors are injected.
-We use here the same format as in figure\ref{fig:routine_all_ovhd}.
-%%MARC: I decided to remove the above figure because it doesnt make sense to me and the same information is displayed in the following figure with more readibility.
-
-\begin{figure}[ht!]
-\centering
-\includegraphics[width=1.00\columnwidth]{figs/4_1_2_Exp2_Expected_Running_Time_Needed.png}
-\caption{Overhead of resilience techniques for RK when errors are injected}
-\label{fig:rk_routine_exp_exec}
-\end{figure}
-
-Figure~\ref{fig:rk_routine_exp_exec} shows the RK's overhead when between 0 and 10 errors per run are injected.%, using the same format as Figure~\ref{fig:routine_exp_exec}.
-Although the \texttt{NoRe\_NoCkpt} configuration has the lowest overhead when only a few errors are injected (low error rate or small execution times), as the number of errors per run increases this configuration quickly becomes extremely slow because an error is detected in almost every run.
-Since each such detection requires a restart of the routine, at higher error rates this causes RK to restart many times before it completes.
-In contrast, while the other configurations have a higher overhead for a few errors per run, their execution time scales much better with increasing error rates, making them significantly more practical.
-In particular, the overheads of the configurations with more replication and checkpointing scale better as the number of errors per run increase.
-By varying the checkpointing interval and the degree to which various RK pointers are replicated it is possible to achieve a wide range of design points that provide the best performance at various error rates.
-
-\paragraph{Output Accuracy}
-%\label{sec:res_tech:eval:accu}
-
-\begin{figure}[ht!]
-\centering
-\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp2_2_Heatmap_Error_ConfSpace_RK4.png}
-\caption{Error magnitude histogram of the entire configuration space of RK}
-\label{fig:rt_algo_err_heatmap}
-\end{figure}
-
-\begin{figure}[ht!]
-\centering
-\includegraphics[width=1.00\columnwidth]{figs/4_1_2_Exp2_Effectiveness.png}
-\caption{Fraction of RK runs that achieve low error (smaller than 1e-7).}
-\label{fig:rk_effectiveness}
-\end{figure}
-
-Figure~\ref{fig:rt_algo_err_heatmap} shows the fraction of errors in RK's outputs for various checkpointing and replication configurations, in the same format and using the same experimental setup as in Figure~\ref{fig:algo_err_heatmap}.
-It shows that while the use replication reduces the amount of error in RK's outputs, increased checkpointing doesn't have a significant effect on this accuracy metric.
-Figure~\ref{fig:rk_effectiveness} evaluates the effect of checkpointing on accuracy from another perspective: the fraction of RK runs where the root-mean-squared error of the output vector is below 1e-7 (y-axis), relative to the number of error injections per run (x-axis).
-It shows that RK configurations with higher level of resilience and in particular, more frequent checkpoints, are able to achieve this accuracy target more frequently.
-This is primarily because by avoiding frequent rollbacks due to crashes these techniques reduce the total number of errors that are injected into the application run, which also reduces the probability that one such error will be missed by RK's algorithmic error detector.
-As with overhead, the choice of checkpointing interval and replication degree make it possible to choose a wide range accuracy levels and probabilities that these targets will be satisfied.
-
-%
-%???\sui{the effectiveness of the error handlers and the error checkers. Note that the graph shows fault injection experiments at various rates (defined by errors per run).}
-%
-%\sui{There are roughly three types of fault rates: 1) very high fault rates (more than tens of errors per run), under which the application would barely complete running; 2) medium fault rates (1 to 10 errors per run), where the rate that the application completes running is very sensitive to the fault rate (because for a non-resilient application, just one error is enough to cause a segmentation fault and abort the application) and 3) low error rates, smaller than 1 errors per run.
-The effectiveness of the fault detector is most visible at fault rates from under 1 faults per run to several errors per run, which is visible from the histograms that the vertical bars indicating a run with algorithmic detectors is generally brighter, therefore less erroneous; when the error rates goes beyond 10 errors per run, the effectiveness of the signal handlers become more noticeable, rising the chance the application would complete.}
-%
-%\sui{At higher error rates, the signal handlers enable more runs to complete as well as allowing erroneous results to go into the output.
-More erroneous results are allowed into the final output by the signal handler than the erroneous intermediate corrected by the algorithmic checker.
-That is why from the histograms, some vertical bars indicating fault-tolerant runs at a very high error rate would look darker.}
-%
-%
-%
-%
-%
-%
-%Errors that do not cause the application to access an invalid memory location are detected by replicating key pointers and loop iteration variables.
-%The Hattrick application uses a high-accuracy ODE solver that is capable of estimating errors and detecting and reducing errors to a certain level adaptively.
-%Therefore we would only focus on the application's vulnerability to pointer and index corruptions.
-%Replicating critical variables (pointers and indices) and leveraging Byzantine fault tolerance and correcting those pointers and indices during run-time would raise the applications' resistance to soft errors \sui{and increase the chance this application would output correct results.}
-%
-%\sui{Unlike the numerical and FFT routines used in the other two applications, the ODE solver has an stepping function called on the evaluation(s) of each integration step (GSL performs two evaluations per timestep in order to estimate integration error and perform adaptive step size control), which is called much more frequently, on the magnitude of millions of times during the lifetime of the application.
-Contrarily, the linear algebra routines are only called less than a hundred times during the execution of Lasso with a 10x input, and the FFT routines are called only several hundred times during the execution of DRC with a 30x input.
-Both are moderate-sized inputs in our experiments.}
-%
-%\greg{Need to explain why we only do this for Hattrick.} \sui{The checkpointing scheme used here has overhead of having to set signal handlers and perform up-to-date checksum computation of critical data structures.
-The more often the routines are called, the more frequent the critical data structures need be kept up-to-date, therefore this overhead is larger (non-trivial compated to that for FFT and linear algebra) for a frequently called routine like the integrator.
-This overhead makes the fault tolerance technique less efficient as it was in linear algebra and FFT routines, bringing it to the same ballpark as the expensive pointer replciation technique.
-Although it is possible to reduce the frequency of updating the checkpoints, it would also reduce the effectiveness of this method sharply.}
-%
-%\sui{
-%The other reason is, pointer replication performs a better job at maintaining the chance the routine generates correct results (See figure \ref{fig:rk4_eff_ovhd}).
-As mentioned before, the RK4 integrator does not have an effective checker that takes its output and tells the user whether the output is correct.
-As the user can't know whether a re-run is necessary and the user should be concerned by the expected time by which s/he can expect a correct output.
-Therefore, a higher chance of producing correct results would lower the expected cost the user would have to pay to get the correct result despite the overhead.
-The method of computing this expected running time is described in greater detail in section \ref{sec:eval:perf}'s Markov Model.
-From \ref{fig:rk4_eff_ovhd} one could see the ratio of perfect runs out of all runs (that are completed or incomplete) and the number of faults per run forms an almost linear relationship.
-The linear regression model is used to evaluate the position of the guide points needed for the following experiments.
-%}
-%
-%\sui{Refer to Figure \ref{fig:rk4_eff_ovhd} one could see the number of expected running time of different configurations if one wants to get a correct output from the application in a faulty environments.
-As the number of faults injected per run increases, the expected running time of the original, non-fault-tolerant application increases at a much faster rate than the other fault-tolerant counterparts.
-In real life scientific applications that run much longer, it's likely that a fault could hit the application several times before it completes.
-In such situations, the replication-only resilience technology produces the least expected number of runs needed to get a correct answer.}
-%
-%To determine a replication strategy we have to determine which variables to replicate and how many replicas to use for replicating them.
-The following two subsections would be focused on these problems.
-%\greg{The report spends a lot of space talking about this but in the end the conclusion is that we just a pick a few options and evaluate them experimentally.
-How much space should we devote to this discussion.}
-%
-%Experiment 2: Measure the effectiveness and performance costs of replication. Need to explain why we don't use it for Lasso and DRC.
 \section{Evaluation of Resilience Methodology in Applications}
@@
 
-Having evaluated the resilience properties of the individual application routines and resilience techniques we now consider our three target applications ADDR, Hattrick and DRC.
-Section~\ref{sec:eval:perf} reports on the impact of errors and resilience techniques on application performance.
-Section~\ref{sec:eval:acc} describes the distribution of errors in the results of these applications.
-These results are based on a fault injection campaign that includes a range of fault injection rates, input sizes and configurations of resilience mechanisms, with 100 application runs for each combination of parameters.
-This corresponds to a total of 66,100 fault injection experiments for ADDR (4-process parallel), 43,200 for DRC (sequential), 102,000 for Hattrick (sequential).
-Our experiments were performed on compute nodes with two 2.33 GHz dual-core Xeon E5345 processors with 4GB Ram and RHEL 4 OS.
+To effectively and efficiently protect applications from soft errors we must consider each application's and routine's unique resilience needs as well as the user's accuracy requirements, the error rate of the hardware and the scale at which the application run (total execution time and number of compute cores).
+We now build upon the analysis of individual routines summarized in Section~\ref{sec:res_tech:eval} to show how our methodology can be applied in the context of large-scale HPC applications.
+The most accurate way to configure a large-scale application run on a given platform with a given input is to conduct a thorough fault injection campaign at this scale, injecting large numbers of errors into various code regions and processors.
+Since this would require thousands of times more compute hours than a single large-scale application run, the only practical option is to extrapolate from small-scale application executions to large-scale ones.
+This case study addresses this challenge by modeling large-scale instances of linear solver, signal processing and n-body simulation applications as combinations of small instances of these applications, that can indeed be comprehensively studied via a fault injection campaign.
 
-A given application run may terminate successfully or crash due to an unrecoverable error.
-Unrecoverable errors occur when there is a segmentation fault in code outside our protected routines or if five rollbacks of a routine have been observed in a given run since such repeated rollbacks almost always indicate that the corruption in application state cannot be repaired.
-Since we assume that the user needs application results regardless of the error rate, if the application crashes it is restarted from the beginning as many times as needed for it to complete.
+This section reports the anticipated behavior of large-scale (1 million CPU-hours) runs of applications in these three domains.
+We begin by describing our methodology to extrapolate the fault injection results for individual small-scale runs of ADDR, DRC and Hattrick to large-scale runs that connect many small-scale instances in various parallel and sequential dependence patterns.
+We then apply analysis methodology described in Section~\ref{sec:res_tech:eval} to each of the three applications on various small-scale inputs to relate the various hardware error rate and the configurations of their resilience mechanisms to the resulting overhead and distribution of output errors.
 
-To evaluate the behavior of these applications in a broad range of error environments we consider several error injection rates.
-For reference, assuming a typical FIT of 100-100,000 for individual processing cores (FIT = failure in one billion hours of operation and this range has been experimentally observed for commodity electronics~\cite{mem_errors:2010, dram_error:2009}), an HPC system with one million cores will experience .1-100 errors per hour, which is equivalent to 3e-11 - 3e-14 errors per processor cycle on a 1Ghz processor.
-\sui{Hence, the experiments are referred to as ``accelerated experiments.''}
-The range of error injection rates used used in our experiments were chosen such that (i) on average one or more errors were injected in the executions of our target applications and (ii) at least some application runs completed without crashing.
-Results for error rates that induce fewer than one error per run were extrapolated from our experiments with higher error rates \sui{as described in \ref{sec:eval:extrapolation}}.
-This was done by considering only the runs where exactly one error was injected and trivially scaling the results to correspond to 1 in every $n$ runs being injected with an error while remaining $n-1$ runs complete with no injection.\\
+The results of these experiments are used to select the most efficient configuration of the resilience mechanisms for each given hardware error rate and user accuracy requirement, focusing on the specific use-case of application runs that use 1 million CPU hours.
+User accuracy requirements are expressed as a cap on the RMS error an application output may have and a bound on the probability with which this bound may be violated.
+Since users rarely have a clear idea of how much RMS error is acceptable, we provide an alternative way to specify such bounds in terms of the amount of uncertainty about the application's inputs that would have generated a given amount of RMS difference between outputs.
+By expressing the effects of soft faults using the same concepts as measurement error, this representation makes it easier to select an appropriate output accuracy and predict the effects of a given selection on other software that reads the application's output.
 
@@
 \label{sec:eval:extrapolation}
-\sui{
-This section explains our modelling of massively parallel and long-running applications with the experiments mentioned in page \pageref{sec:eval}, which in nature
-are accelerated experiments. We model larger scale experiments by extrapolating our experiments in the following ways:\\
-}
+
+Figure~\ref{fig:extrapolations} illustrates how individual runs of ADDR, DRC and Hattrick are combined to model large-scale instances of linear solver, signal processing and n-body simulation applications.
+We observe that most large-scale applications can be described simple dependence graphs of identical tasks, such as the execution of a single time-step calculation in a single point in space.
+The shape of the graph controls its total execution time and the way errors propagate.
+Our model focuses on two major types of task graphs: parallel reduction and sequential chaining.
+
+In the parallel reduction model all tasks run in parallel and their results are reduced into a common aggregate.
+As such, the total execution time of the graph is the maximum execution time of all the tasks (approximated as the top 10 percentile as discussed below).
+The RMS error of the aggregate is the combination of the RMS errors of the individual task outputs, which depends on the exact aggregation operation used.
+Since averaging is a very common but not universal aggregation operation, in reductions we model the RMS error of the aggregate as the average of the RMS errors of individual tasks, which is an over-estimate of the RMS errors of their average.
+In the sequential chaining model each task's output is used by the next task, forcing them to run in sequence.
+The total execution time of task chains is thus the sum of the execution times of the individual tasks.
+Tasks may either amplify or dampen errors in their inputs (output errors may be larger or smaller, respectively).
+Given the complexity of dampening and amplification behavior that may occur in practice our model assumes that no dampening or amplification occurs and the error at the end of a task chain is equal to the maximum error of any task within the chain (also approximated as the top 10 percentile).
+
+\greg{TODO: In addition to pure parallel and pure sequential dependence graphs we also considered sequences of parallel reductions, with the number of tasks in a given reduction equal to the number of reduction steps.}
 
@@
 
-\begin{figure}[ht!]
-\centering
-\begin{tabular}{|p{1.0in}|p{2.0in}|}
-\hline
- Model Type & Description \\
-\hline
-Accelerated Experiment & FIT > 1e+7, containing > 1e10 dynamic instructions, running on 1 node, wall time < 10 min \\
-\hline
-Serial Extrapolation by factor N & FIT > $\frac{1e+7}{N}$, containing > $N \cdot 1e10$ dynamic instructions, running on < 1 node \\
-\hline
-Parallel Extrapolation by factor N & FIT > $\frac{1e+7}{N}$, containing > $N \cdot 1e10$ dynamic instructions, running on N nodes \\
-\hline
-\end{tabular}
-\end{figure}
+We model large-scale linear solvers as dependence graphs of individual runs of ADDR, signal processing applications in terms of runs of DRC and n-body simulations in terms of runs of Hattrick.
+Our experiments were performed on compute nodes with two 2.33 GHz dual-core Xeon E5345 processors with 4GB Ram and RHEL 4 Linux.
+We run ADDR, DRC and Hattrick no fault injection on the inputs described in Section~\ref{sec:apps} and configuration options listed in Table~\ref{tbl:configs} to measure their fault-free execution time.
+We also executed each configuration while injecting faults at rates to produce one or more injections per run.% (since different configurations have different execution times, the number of injections per run varies).
+\greg{Since some runs
+In each accelerated fault injection experiment, we ran each application configuration to completion, recording whether it completed and if so, the magnitude of the
+	error.
+	We calculate the execution time of each configuration, accounting for restarting the application in cases of unrecoverable failures via the formula $\frac{t_c}{1-p_f}$, where $t_c$ is the average completion time of such runs and $p_f$ is the probability of an unrecoverable failure.}
+For each combination of application, input and configuration we performed 100 fault injections, for a total of 66,100 fault injection experiments for ADDR (4-process parallel runs), 43,200 for DRC (sequential runs), 102,000 for Hattrick (sequential runs).
+For reference, assuming a typical FIT of 100-100,000 for individual processing cores (FIT = failure in one billion hours of operation and this range has been experimentally observed for commodity electronics~\cite{mem_errors:2010, dram_error:2009}), an HPC system with one million cores will experience .1-100 errors per hour, which is equivalent to 3e-11 - 3e-14 errors per processor cycle on a 1Ghz processor.
+To make our analysis comprehensive we consider error rates between 1e-10 and 1e-16 (\greg{is this right?}).
 
-\sui{
-There is 1 instance drawn from the accelerated tests in both serial and parallel extrapolation models. Each time one instance is sampled
-from the accelerated tests, it is combined with N-1 fault-free instances to become a large serial or parallel execution. The value of N is determined
-by each fault rate that needs be extrapolated to. For example, a value N=10000 is equivalent to extrapolating from 1e+7 FIT to 1e+3 FIT.
-In order to determine the magnitude of output errors from the combined executions, we make the following assumptions to the extrapolation models
-in order to simplify our analysis:
-}
+We model the RMS error and expected execution time of a task graph that consists of executions of a given application, input and configuration combination (denoted a ``task''), on a system with a given per-processor error rate as follows.
+Let $N$ be the total number of task runs needed to reach 1 million CPU hours and let $NF$ be the number of runs that are expected to have faults injected into them at the given error rate.
+We sample $NF$ runs from the 100 fault injection runs conducted for the given configuration with replacement.
+Replacement is used since faults are memory-less, they do not depend on prior faults.
+As such, $NF$ may be larger than 100 without biasing the statistical results since the 100 runs are representative of the overall population.
+We then add $N-NF$ copies of the non-fault injected variant of the given task to produce a complete sample of a large-scale application execution.
+Finally, we compute the model application's metrics as:
+\begin{eqnarray}
+ExecTime_{Parallel} &= \text{Top 10 percentile of} ExecTime \text{of tasks in sample}\\
+RMSE_{Parallel} &= \text{Average} ExecTime \text{of tasks in sample}\\
+ExecTime_{Sequence} & = \text{Average} ExecTime \text{of tasks in sample}\\
+RMSE_{Sequence} & = \text{Top 10 percentile of} ExecTime \text{of tasks in sample}
+\end{eqnarray}
+This procedure is repeated \greg{???} times to compute the expected value of the quantities.
+We also use these samples to compute the 95\% confidence interval of the expected value (range where 95\% of the samples are located).
 
-\sui{
-\begin{itemize}
-\item{
-The magnitude of output errors in a combined serial execution is that of the accelerated instance with the largest output error magnitude.
-Analysis of the output error magnitude is in \ref{sec:eval:errorsensitivity}.\\
-$RMSD_{serial} = max_{i=1}^{N} (RMSD_n)$}
-\item{The magnitude of output errors in a combined parallel execution is the error magnitude of the average of all outputs.
-This reflects a classic reducing access pattern frequently seen in parallel programs. Since all instances except the first one
-are fault-less, the error magnitude, and the RMSD, of the combined instance would effectively be shrunk by a factor of N.\\
-$RMSD_{parallel} = \sum_{i=1}^{N}{RMSD_n}/N$
-}
-\item{The running time of a combined serial execution is the sum of every individual instance.}
-\item{The running time of a combined parallel execution is the running of the longest individual instance. }
-\item{At least 1 of the individual instances is injected with one error.}
-\end{itemize}
-}
+The reason why we use the top 10 percentile rather than the maximum is because the maximum metric is extremely sensitive to sampling.
+Since there is no bound in the difference between the maximum of a given sample and the overall maximum, the only way to estimate it accurately is to sample the entire space of possibilities, which is prohibitively expensive.
+In contrast, metrics such as average and percentiles are robust against such effects and can be accurately estimated from small samples.
 
-\subsection{Error sensitivity}
+\subsection{Intuitive Accuracy Metric}
 \label{sec:eval:errorsensitivity}
-\sui{
-	I intend to show with the figure that 1) error propagation at an early point of time in a long serial run will result in an error that keeps being oscillated and magnified, resulting in numerical instability or simulation explosion,
-	and be terminated or 2) the speed the magnitude of errors grows keeps dropping until the magnitude of errors stabilize and converge. \\
-	In the first case, the magnitude of errors in the output should be small enough not to trigger numerical unstability or simulation explosion. In the second case,
-	the magnitude of error would peak the instant after the errors occur. Either way, we can assume the magnitude of errors in a serial run to be the largest among
-	individual instances.
-}
+
+Even in the absence of hardware faults scientific applications face many sources of error, including errors in input data, discretization error and modeling error.
+All these sources of error are actively managed by application developers to ensure that the application's output remains useful.
+Importantly, it means that hardware faults are only important if they cause errors larger than the ambient error that already exists in a given application, with a sufficiently high probability to make it a threat to decision-making.
+While the probability of an inaccurate result makes intuitive sense to users, bounds on error are less intuitive and thus difficult to set.
+As such, in this case study we relate output errors due to soft faults to measurement errors in application inputs, something that developers currently need to consider.
+To this end we took the inputs to ADDR, DRC and Hattrick (all are matrixes or vectors of floating point numbers) and perturbed them by multiplying the values a random amount of noise sampled from a Gaussian distribution with mean 1 and a standard deviation that varies from 1e-9 to 1e0.
+Figure~\ref{fig:inputvarianceoutputrmsd} shows the RMS difference (vertical axis) between the outputs of each application on every input, when noise with each standard deviation (horizontal axis) is and is not injected.
+The data shows that DRC is largely insensitive to noise in its inputs, with RME differences of 1e-3 to 1e-2 typical across the noise spectrum.
+This means that fault-induced RMS errors smaller than 1e-3 are completely irrelevant to this application and errors between 1e-3 and 1e-2 are typical for it.
+In contrast, the output difference in ADDR and Hattrick are very sensitive to noise in the input, making it critical to set their target RMS error to be consistent with the uncertainty in the inputs.
 
@@
 
-\subsection{Experiment setup}
-\label{sec:eval:confs}
-
-\sui{
-As we have seen in section \ref{sec:res_tech:eval}, varying the error checker thresholds in linear algebra routines and the checkpointing intervals and number of
-replicas in the RK4 integrator results in different levels of accuracy. As they are deployed in applications, choice of these parameters may influence how often
-the applications need to rollback in order to obtain correct results. An intuition is we should strike a balance between the tradeoff between correctness
-and performance.
+%\subsection{Experiment setup}
+%\label{sec:eval:confs}
+%
+%\sui{
+%As we have seen in section \ref{sec:res_tech:eval}, varying the error checker thresholds in linear algebra routines and the checkpointing intervals and number of
+%replicas in the RK4 integrator results in different levels of accuracy. As they are deployed in applications, choice of these parameters may influence how often
+%the applications need to rollback in order to obtain correct results. An intuition is we should strike a balance between the tradeoff between correctness
+%and performance.
 To explore the possible tradeoff in applications, we consider a range of parameters for our resilience techniques as described in section \ref{sec:res_tech:eval} when deploying the algorithmic checkers in these applications.
 The variable parameters and fixed parameters are listed as below.
-}
-
-\begin{figure}[ht!]
-\centering
-\begin{tabular}{|p{0.7in}|p{2.0in}|}
-\hline
-{\bf Application} & {\bf Resilience Techniques} \\
-\hline
-\hline
-Lasso & Checkers for MM, MV, CD, SYRK; input backup; protection for critical data structures \\
-\hline
-DRC & Checkers for FFT, FIR; input backup; protection for critical data structures \\
-\hline
-Hattrick & Checkpointing; protection for critical data structures; replication of pointers \\
-\hline
-\hline
-{\bf Application} & {\bf Variable Parameters} \\
-\hline
-\hline
-Lasso & Checker thresholds 1e-4 to 1e-7 \\
-\hline
-DRC   & Checker thresholds 1e-6 to 3e-8 \\
-\hline
-Hattrick & Checkpointing intervals 0, 1e3, 1e5; 0, 1, 2 replicas for each pointer \\
-\hline
-\end{tabular}
-\caption{Experiment Configurations}
-\label{fig:exp_confs}
-\end{figure}
-
-\sui{
-	In each accelerated fault injection experiment, we ran each application configuration to completion, recording whether it completed and if so, the magnitude of the
-	error.
-	We calculate the execution time of each configuration, accounting for restarting the application in cases of unrecoverable failures via the formula $\frac{t_c}{1-p_f}$, where $t_c$ is the average completion time of such runs and $p_f$ is the probability of an unrecoverable failure.
-}
+%}
+%
+%\begin{figure}[ht!]
+%\centering
+%\begin{tabular}{|p{0.7in}|p{2.0in}|}
+%\hline
+%{\bf Application} & {\bf Resilience Techniques} \\
+%\hline
+%\hline
+%Lasso & Checkers for MM, MV, CD, SYRK; input backup; protection for critical data structures \\
+%\hline
+%DRC & Checkers for FFT, FIR; input backup; protection for critical data structures \\
+%\hline
+%Hattrick & Checkpointing; protection for critical data structures; replication of pointers \\
+%\hline
+%\hline
+%{\bf Application} & {\bf Variable Parameters} \\
+%\hline
+%\hline
+%Lasso & Checker thresholds 1e-4 to 1e-7 \\
+%\hline
+%DRC   & Checker thresholds 1e-6 to 3e-8 \\
+%\hline
+%Hattrick & Checkpointing intervals 0, 1e3, 1e5; 0, 1, 2 replicas for each pointer \\
+%\hline
+%\end{tabular}
+%\caption{Experiment Configurations}
+%\label{fig:exp_confs}
+%\end{figure}
+%
+%\sui{
+%	In each accelerated fault injection experiment, we ran each application configuration to completion, recording whether it completed and if so, the magnitude of the
+%	error.
+%	We calculate the execution time of each configuration, accounting for restarting the application in cases of unrecoverable failures via the formula $\frac{t_c}{1-p_f}$, where $t_c$ is the average completion time of such runs and $p_f$ is the probability of an unrecoverable failure.
+%}
 

### 2013-10-17 paper.tex
Another pass by Greg

@@
 
-To analyze the performance and resilience properties the key routines in ADDR, DRC and Hattrick we conducted a comprehensive fault injection campaign where we employed the above resilience techniques with each routine, as applicable,
-We used fault rates between 1e-11 and 1e-8 errors per clock cycle \greg{(is this correct)}, probing each application's behavior as it was impacted by either one or multiple errors in a single run.
-Our experiments span a wide range of input sizes and a broad range of resilience mechanism configurations, with 100 (\greg{???}) runs for each combination of application, input, fault rate and resilience configuration.
+We now summarize our methodology for analyzing the performance and resilience properties of applications when injected with errors.
+For clarity the analysis in this section focuses on the individual key routines in ADDR, DRC and Hattrick, since analyzing the entire applications obscures the properties of the routines.
+The same analyses are were conducted on the full applications and used to predict the properties of large-scale runs, as Section~\ref{sec:eval} reports.
+Due to space constraints this section reports representative experimental results, rather than all of them.
+
+We conducted a comprehensive fault injection campaign where we employed the above resilience techniques with each routine, as applicable.
+Fault injection rates between 1e-11 and 1e-8 errors per clock cycle were used\greg{(is this correct)} to probe the effect of one or multiple errors per routine execution.
+Our experiments span a wide range of input sizes and a broad range of resilience mechanism configurations, with 100 (\greg{???}) runs for each combination of routine, input, fault rate and resilience configuration.
 Table~\ref{tbl:configs} specifies the techniques applied to each routine and its configuration space.
@@
     \hline
-    \multicolumn{2}{c}{Routine}          & Algorithmic Detector             & Checkpointing         & Pointer Replication  \\
+    \multicolumn{2}{|c|}{Routine}          & Algorithmic Detector             & Checkpointing         & Pointer Replication  \\
     \hline
     \multirow{4}{*}{ADDR}      & MM   & Linear encoding                     & Inputs                & No \\
-                               & SYRK & Detection Thresholds:               & Inputs                & No \\
+                               & SYRK & Thresholds:                         & Inputs                & No \\
                                & MVM  & 1e-4 to 1e-7                        & Inputs                & No \\
@@
                                & FIR  & Sum conservation.                   & Inputs                & No \\
-                               &      & Detection Thresholds: 1e-6 to 3e-8. &                            \\
+                               &      & Thresholds: 1e-6 to 3e-8.           &                            \\
     \hline
-    \multirow{3}{*}{Hattrick}  & \multirow{3}{*}{RK} & \multirow{3}{*}{Variable step-size} & Periodic Timesteps    & None  \\
-                               &      &                                   & Period:  1, 1e3, 1e5 & All pointers checked at one code location \\
-                               &      &                                   &                       & All pointers checked on each use \\
+    \multirow{5}{*}{Hattrick}  & \multirow{5}{*}{RK} & \multirow{5}{*}{Variable step-size} & Periodic Timesteps    & None  \\
+                               &      &                                   & Period: 1, 1e3, 1e5   & All pointers, \\
+                               &      &                                   &                       & checked at one code location \\
+                               &      &                                   &                       & All pointers, \\
+                               &      &                                   &                       & checked on each use \\
     \hline
@@
 
-Even though an application may complete its execution, errors injected into its numeric state may produce incorrect output.
-In each experimental configuration we measured the distribution of errors in application output, measured as the Root Mean Square (RMS) Error: $\sqrt{\sigma_i (x_i-x'_i)^2}/n$, where $x_i$ is the value of output element $i$ in a fault-free run, $x'_i$ its value in a fault-injected run and $n$ is the total number of output elements (e.g.
+Even though a routine may complete its execution, errors injected into its numeric state may produce incorrect output.
+In each experimental configuration we measured the distribution of errors in routine output, quantified as the Root Mean Square (RMS) Error: $\sqrt{\sigma_i (x_i-x'_i)^2}/n$, where $x_i$ is the value of output element $i$ in a fault-free run, $x'_i$ its value in a fault-injected run and $n$ is the total number of output elements (e.g.
 size of the output matrix).
 
-Figure~\ref{fig:algo_err_dist} shows a sample of our results, focusing on the linear algebra routines (restarted upto 5 times after an error detection).
-It presents the probability distribution of the routine's results having errors of various magnitudes.
-Each tile corresponds to a different routine (vertical order) and different algorithmic error detection threshold used (horizontal order).
-Table~\ref{fig:algo_err_heatmap} specifies the exact threshold values used.
+Figure~\ref{fig:algo_err_dist} shows a sample of our results, focusing on the linear algebra routines (runs that restarted more than 5 times were excluded).
+It presents the probability distribution of errors with various magnitudes in each routine's output.
+Each tile corresponds to a different routine (vertical order) and different algorithmic error detection threshold used (horizontal order). (\greg{What input sizes})
+Table~\ref{fig:algo_err_heatmap} specifies the exact threshold values used for each routine.
 Within each tile the vertical axis corresponds different error rates.
-The axis numbers are normalized to the execution time of the configuration with no resilience techniques (\greg{errors injected or not?}) and specify the average number of errors injected during this time period at this rate.
-The horizontal axis corresponds to the error magnitude ranging from 1e-14 to 1e-1.
-For each error injection count and output error magnitude we report the probability of the output having error of \emph{larger than} this magnitude as a different shade of gray, with darker shades indicating higher probabilities.
-Thus, darker plots indicate that error injections make larger output errors more likely and errors $\ge$ a smaller magnitude are always less likely than errors $\ge$ a larger one (shades grow lighter from left to right).
+The axis values are the number of errors that would be injected at each injection rate into a single run of the routine with no resilience techniques (\greg{errors injected or not?}).
+The horizontal axis corresponds to the RMS error magnitude, ranging from 1e-14 to 1e-1.
+For each error injection count and output error magnitude we report the probability of the output having error $\ge$ this magnitude as a different shade of gray, with darker shades indicating higher probabilities on a log scale (base 10).
+Thus, darker plots indicate that error injections make larger output errors more likely and shades grow lighter from left to right since errors $\ge$ a smaller magnitude are always less likely than errors $\ge$ a larger one.
 
 As error count increases the probability of larger output errors rises.
-Further, tighter (smaller) error detection thresholds reduce this probability.
+Tighter (smaller) error detection thresholds reduce this probability.
 Error distributions vary widely, with MM being most resilient to errors (the probability of having large errors is low) and CD least.
 These experiments guide the choice of how each routine's resilience technique is configured to meet user-specified accuracy bounds.
-
-Figures~\ref{fig:algo_err_dist} show the fraction of entries in the outputs of each routine (y-axis) that have an error of a given magnitude (x-axis) as error magnitudes span from 1e-14 to 1e-4.
-This figure is a focused view on few representative cases (each routine run on a single input size and error injection rate) across different resilience configurations: no error detection as well as three error detection thresholds.
-As the error magnitude increases the fraction of output entries with errors of that magnitude drops.
-Importantly, the fraction of erroneous entries is reduced at all magnitudes as a result of error detection and restart.
-Further, this reduction is more significant as the detection threshold shrinks since smaller magnitude errors are detected.
-
-Figure~\ref{fig:algo_err_heatmap} shows the same information but for all experimental configurations: (i) input sizes as above (ii) error injection rates of 1e-7 to 1e-10.
-The graph is a grid, with results for different routines listed vertically and different resilience configurations (no detection and the different detection thresholds shown in the table at the bottom) listed horizontally.
-The graph for each routine and resilience configuration presents the curves from Figure~\ref{fig:algo_err_dist} as a heatmap.
-The y-axis shows different routine input sizes and error injection rates, sorted according to the average number of error injections per non-fault injected run, in decreasing order from top to bottom.
-The x-axis shows the different error magnitudes and the shade of each point is the fraction of the routine's entries that have a given error magnitude, with darker shades indicating more such entries (each shade corresponds to a fraction that is some power of 10).
-Thus, each row of Figure~\ref{fig:algo_err_heatmap} represents the same information as each curve in Figure~\ref{fig:algo_err_dist}, with the shades of the squares representing the same information as the height of the line's points.
-
-%\begin{figure}[ht!]
-%\centering
-%\includegraphics[width=1.00\columnwidth]{figs/4_1_1_Exp2_1_Example.png}
-%\caption{Error magnitude histogram showing the effectiveness of the error detector}
-%\label{fig:algo_err_dist}
-%\end{figure}
 
@@
 
-To effectively and efficiently protect applications from soft errors we must consider each application's and routine's unique resilience needs as well as the user's accuracy requirements, the error rate of the hardware and the scale at which the application run (total execution time and number of compute cores).
-We now build upon the analysis of individual routines summarized in Section~\ref{sec:res_tech:eval} to show how our methodology can be applied in the context of large-scale HPC applications.
-The most accurate way to configure a large-scale application run on a given platform with a given input is to conduct a thorough fault injection campaign at this scale, injecting large numbers of errors into various code regions and processors.
-Since this would require thousands of times more compute hours than a single large-scale application run, the only practical option is to extrapolate from small-scale application executions to large-scale ones.
-This case study addresses this challenge by modeling large-scale instances of linear solver, signal processing and n-body simulation applications as combinations of small instances of these applications, that can indeed be comprehensively studied via a fault injection campaign.
-
-This section reports the anticipated behavior of large-scale (1 million CPU-hours) runs of applications in these three domains.
-We begin by describing our methodology to extrapolate the fault injection results for individual small-scale runs of ADDR, DRC and Hattrick to large-scale runs that connect many small-scale instances in various parallel and sequential dependence patterns.
-We then apply analysis methodology described in Section~\ref{sec:res_tech:eval} to each of the three applications on various small-scale inputs to relate the various hardware error rate and the configurations of their resilience mechanisms to the resulting overhead and distribution of output errors.
-
-The results of these experiments are used to select the most efficient configuration of the resilience mechanisms for each given hardware error rate and user accuracy requirement, focusing on the specific use-case of application runs that use 1 million CPU hours.
+To effectively and efficiently protect applications from soft errors we must consider each application's unique resilience needs as well as the user's accuracy requirements, the hardware fault rate and the scale at which the application runs (total execution time and number of compute cores).
+%We now build upon the analysis of individual routines summarized in Section~\ref{sec:res_tech:eval} to show how our methodology can be applied in the context of large-scale HPC applications.
+We now predict how large-scale runs (1 million CPU-hours) of applications in the linear solver, signal processing and n-body simulation application domains will behave on faulty hardware.
+The most accurate way to do this is to perform many large-scale runs with many inputs and configurations of resilience techniques, while injecting errors at different rates.
+Since this requires hundreds of thousands of application runs, this is infeasible in practice and thus the only practical option is to extrapolate from small-scale application runs to large-scale ones.
+This case study addresses this challenge by modeling large-scale runs of applications in each target domain as combinations of small instances of representative applications, that can indeed be comprehensively studied via a fault injection campaign.
+First, we performed the fault injection-based analysis detailed in Section~\ref{sec:res_tech:eval} on ADDR, DRC and Hattrick, identifying for each combination of application, input, configuration of resilience techniques and error injection rate the application's the execution time and distribution of RMS errors in its outputs.
+We then use the methodology described in Section~\ref{sec:eval:extrapolation} to extrapolate the fault injection results for individual small-scale runs of ADDR, DRC and Hattrick to large-scale runs that connect many small-scale instances in various parallel and sequential dependence patterns.
+Our comprehensive fault injection experiments make it possible to select the most efficient configuration of resilience techniques for each input, error rate and user accuracy target, making it possible to reason about realistic deployments of these techniques.
+%The results of these experiments are used to select the most efficient configuration of the resilience mechanisms for each given hardware error rate and user accuracy requirement, focusing on the specific use-case of application runs that use 1 million CPU hours.
 User accuracy requirements are expressed as a cap on the RMS error an application output may have and a bound on the probability with which this bound may be violated.
-Since users rarely have a clear idea of how much RMS error is acceptable, we provide an alternative way to specify such bounds in terms of the amount of uncertainty about the application's inputs that would have generated a given amount of RMS difference between outputs.
+Since users rarely have a clear idea of how much RMS error is acceptable, Section~\ref{sec:eval:errorsensitivity} describes an alternative way to specify such bounds in terms of the amount of uncertainty about the application's inputs that would have generated a given amount of RMS difference between outputs.
 By expressing the effects of soft faults using the same concepts as measurement error, this representation makes it easier to select an appropriate output accuracy and predict the effects of a given selection on other software that reads the application's output.
+Finally, Sections~\ref{sec:eval:perf} and \ref{sec:eval:acc} discuss the performance and accuracy properties, respectively, of large scale runs of applications in each target application domain, as well as the best configurations of resilience techniques.
 
@@
 Figure~\ref{fig:extrapolations} illustrates how individual runs of ADDR, DRC and Hattrick are combined to model large-scale instances of linear solver, signal processing and n-body simulation applications.
-We observe that most large-scale applications can be described simple dependence graphs of identical tasks, such as the execution of a single time-step calculation in a single point in space.
-The shape of the graph controls its total execution time and the way errors propagate.
-Our model focuses on two major types of task graphs: parallel reduction and sequential chaining.
+We observe that most large-scale applications can be described simple dependence graphs of identical tasks (e.g. single time-step calculation on a point in space).
+The graph's shape controls its execution time and the way errors propagate.
+Our model focuses on three major types of task graphs (i) parallel reduction, (ii) sequential chaining and (iii) sequence of parallel reductions.
 
-In the parallel reduction model all tasks run in parallel and their results are reduced into a common aggregate.
-As such, the total execution time of the graph is the maximum execution time of all the tasks (approximated as the top 10 percentile as discussed below).
+In the parallel reduction model the results of concurrent tasks are reduced into a common aggregate.
+The graph's execution time is the maximum across all the tasks (approximated as the top 10 percentile, as discussed below).
 The RMS error of the aggregate is the combination of the RMS errors of the individual task outputs, which depends on the exact aggregation operation used.
-Since averaging is a very common but not universal aggregation operation, in reductions we model the RMS error of the aggregate as the average of the RMS errors of individual tasks, which is an over-estimate of the RMS errors of their average.
+Since averaging is a very common but not universal aggregation operation, we model the RMS error of the reduced aggregate as the average of the RMS errors of individual tasks, which over-approximates the RMS error of their average.
 In the sequential chaining model each task's output is used by the next task, forcing them to run in sequence.
-The total execution time of task chains is thus the sum of the execution times of the individual tasks.
+The chain's execution time is the sum of its member tasks.
 Tasks may either amplify or dampen errors in their inputs (output errors may be larger or smaller, respectively).
-Given the complexity of dampening and amplification behavior that may occur in practice our model assumes that no dampening or amplification occurs and the error at the end of a task chain is equal to the maximum error of any task within the chain (also approximated as the top 10 percentile).
+Given the complexity of dampening and amplification behavior that may occur in practice our model assumes that no dampening or amplification occurs and the RMS error at the end of a task chain is equal to the maximum RMS error of any task within the chain (also approximated as the top 10 percentile).
 
@@
 Our experiments were performed on compute nodes with two 2.33 GHz dual-core Xeon E5345 processors with 4GB Ram and RHEL 4 Linux.
-We run ADDR, DRC and Hattrick no fault injection on the inputs described in Section~\ref{sec:apps} and configuration options listed in Table~\ref{tbl:configs} to measure their fault-free execution time.
-We also executed each configuration while injecting faults at rates to produce one or more injections per run.% (since different configurations have different execution times, the number of injections per run varies).
-\greg{Since some runs
-In each accelerated fault injection experiment, we ran each application configuration to completion, recording whether it completed and if so, the magnitude of the
-	error.
-	We calculate the execution time of each configuration, accounting for restarting the application in cases of unrecoverable failures via the formula $\frac{t_c}{1-p_f}$, where $t_c$ is the average completion time of such runs and $p_f$ is the probability of an unrecoverable failure.}
-For each combination of application, input and configuration we performed 100 fault injections, for a total of 66,100 fault injection experiments for ADDR (4-process parallel runs), 43,200 for DRC (sequential runs), 102,000 for Hattrick (sequential runs).
-For reference, assuming a typical FIT of 100-100,000 for individual processing cores (FIT = failure in one billion hours of operation and this range has been experimentally observed for commodity electronics~\cite{mem_errors:2010, dram_error:2009}), an HPC system with one million cores will experience .1-100 errors per hour, which is equivalent to 3e-11 - 3e-14 errors per processor cycle on a 1Ghz processor.
+We run ADDR, DRC and Hattrick with no fault injection on the inputs described in Section~\ref{sec:apps} and configuration options listed in Table~\ref{tbl:configs} to measure their fault-free execution time.
+The reliability of today's processing cores is approximately 100-100,000 FIT~\cite{mem_errors:2010, dram_error:2009} (failure in one billion hours of operation).
+An HPC system with one million cores will thus experience .1-100 errors per hour, equivalent to 3e-11 - 3e-14 errors per processor cycle on a 1Ghz processor.
 To make our analysis comprehensive we consider error rates between 1e-10 and 1e-16 (\greg{is this right?}).
+The modeled this range of rates by performing fault injection at various rates to produce one or more error injections per run, upto 1e-10 errors per cycle.
+% (since different configurations have different execution times, the number of injections per run varies).
+%\greg{Since some runs
+%In each accelerated fault injection experiment, we ran each application configuration to completion, recording whether it completed and if so, the magnitude of the	error.
+%We calculate the execution time of each configuration, accounting for restarting the application in cases of unrecoverable failures via the formula $\frac{t_c}{1-p_f}$, where $t_c$ is the average completion time of such runs and $p_f$ is the probability of an unrecoverable failure.}
+Error rates that produce fewer than 1 injection per run were simulated by combining the behavior of runs with 1 error injection and runs with no errors injected, as described below.
+For each combination of application, input, resilience configuration and error rate we performed 100 fault injections, for a total of 66,100 fault injection experiments for ADDR (4-process parallel runs), 43,200 for DRC (sequential runs), 102,000 for Hattrick (sequential runs).
 
@@
 Let $N$ be the total number of task runs needed to reach 1 million CPU hours and let $NF$ be the number of runs that are expected to have faults injected into them at the given error rate.
+$NF=N$ is the expected number of error injections per run is >1 and $NF<N$ otherwise.
 We sample $NF$ runs from the 100 fault injection runs conducted for the given configuration with replacement.
-Replacement is used since faults are memory-less, they do not depend on prior faults.
-As such, $NF$ may be larger than 100 without biasing the statistical results since the 100 runs are representative of the overall population.
-We then add $N-NF$ copies of the non-fault injected variant of the given task to produce a complete sample of a large-scale application execution.
-Finally, we compute the model application's metrics as:
-\begin{eqnarray}
-ExecTime_{Parallel} &= \text{Top 10 percentile of} ExecTime \text{of tasks in sample}\\
-RMSE_{Parallel} &= \text{Average} ExecTime \text{of tasks in sample}\\
-ExecTime_{Sequence} & = \text{Average} ExecTime \text{of tasks in sample}\\
-RMSE_{Sequence} & = \text{Top 10 percentile of} ExecTime \text{of tasks in sample}
-\end{eqnarray}
+Because the 100 runs are representative of the overall population of runs and sampling is done with replacement (faults do not depend on prior history) $NF$ may be larger than 100.
+We then add $N-NF$ copies of the task's non-fault injected variant to produce a complete sample from the population of large-scale application executions.
+
+Finally, we compute the model application's metrics from the sample as follows:
+\begin{tabular}{|c|c|c|}
+\hline
+               & Parallel          & Chain             \\
+\hline
+Execution Time & Top 10 percentile & Average           \\
+\hline
+RMS error      & Average           & Top 10 percentile \\
+\hline
+\end{tabular}
+%\begin{eqnarray}
+%ExecTime_{Parallel} &= \text{Top 10 percentile of} ExecTime \text{of tasks in sample}\\
+%RMSE_{Parallel} &= \text{Average} ExecTime \text{of tasks in sample}\\
+%ExecTime_{Sequence} & = \text{Average} ExecTime \text{of tasks in sample}\\
+%RMSE_{Sequence} & = \text{Top 10 percentile of} ExecTime \text{of tasks in sample}
+%\end{eqnarray}
 This procedure is repeated \greg{???} times to compute the expected value of the quantities.
@@
 
-The reason why we use the top 10 percentile rather than the maximum is because the maximum metric is extremely sensitive to sampling.
-Since there is no bound in the difference between the maximum of a given sample and the overall maximum, the only way to estimate it accurately is to sample the entire space of possibilities, which is prohibitively expensive.
+The reason why we use the top 10 percentile rather than the maximum is because the maximum metric is extremely sensitive to sampling error.
+Since there is no bound in the difference between the maximum of a given sample and the overall maximum, the only way to estimate this metric accurately is to sample the entire space of possibilities, which is prohibitively expensive.
 In contrast, metrics such as average and percentiles are robust against such effects and can be accurately estimated from small samples.
@@
 Even in the absence of hardware faults scientific applications face many sources of error, including errors in input data, discretization error and modeling error.
-All these sources of error are actively managed by application developers to ensure that the application's output remains useful.
-Importantly, it means that hardware faults are only important if they cause errors larger than the ambient error that already exists in a given application, with a sufficiently high probability to make it a threat to decision-making.
-While the probability of an inaccurate result makes intuitive sense to users, bounds on error are less intuitive and thus difficult to set.
-As such, in this case study we relate output errors due to soft faults to measurement errors in application inputs, something that developers currently need to consider.
-To this end we took the inputs to ADDR, DRC and Hattrick (all are matrixes or vectors of floating point numbers) and perturbed them by multiplying the values a random amount of noise sampled from a Gaussian distribution with mean 1 and a standard deviation that varies from 1e-9 to 1e0.
-Figure~\ref{fig:inputvarianceoutputrmsd} shows the RMS difference (vertical axis) between the outputs of each application on every input, when noise with each standard deviation (horizontal axis) is and is not injected.
-The data shows that DRC is largely insensitive to noise in its inputs, with RME differences of 1e-3 to 1e-2 typical across the noise spectrum.
+These errors are actively managed by application developers to ensure that the application's output remains useful.
+Errors smaller than these or larger errors that are extremely unlikely do not affect the output's utility.
+The probability of an inaccurate result can be readily incorporated in to user decision making to trade off a result's value against the cost of its computation.
+However, explicit bounds on RMS error are less intuitive and thus difficult to specify.
+As such, in this case study we relate output errors due to soft faults to measurement errors in application inputs, which users currently account for.
+To this end we took the inputs to ADDR, DRC and Hattrick (all are matrixes or vectors of floating point numbers) and perturbed them by multiplying the values by a random value sampled from a Gaussian distribution with mean 1 and a standard deviation that varies from 1e-9 to 1e0.
+Figure~\ref{fig:inputvarianceoutputrmsd} shows for each application and input the RMS difference (vertical axis) between the application outputs, when perturbation with each standard deviation (horizontal axis) is and is not applied.
+The data shows that DRC is largely insensitive input noise, with RME differences of 1e-3 to 1e-2 typical across the spectrum.
 This means that fault-induced RMS errors smaller than 1e-3 are completely irrelevant to this application and errors between 1e-3 and 1e-2 are typical for it.

### 2013-10-19 paper.tex
Formatting

@@
 
+\vspace{-10pt}
 \section{Introduction}
+\vspace{-10pt}
 \label{sec:intro}
@@
 
+\vspace{-10pt}
 \section{Soft Faults}
+\vspace{-10pt}
 \label{sec:soft_faults}
@@
 
+\vspace{-10pt}
 \section{Target Applications}
+\vspace{-10pt}
 \label{sec:apps}
@@
 
+\vspace{-10pt}
 \subsection{Alternating Direction Method of Multipliers (ADDR)}
+\vspace{-10pt}
 \label{sec:apps:lasso}
@@
 
+\vspace{-10pt}
 \subsection{DRC: Digital Room Correction}
+\vspace{-10pt}
 \label{sec:apps:drc}
@@
 
+\vspace{-10pt}
 \subsection{Hattrick}
+\vspace{-10pt}
 \label{sec:apps:hattrick}
@@
 
+\vspace{-10pt}
 \section{Resilience Techniques}
+\vspace{-10pt}
 \label{sec:res_tech}
@@
 
+\vspace{-10pt}
 \subsection{Error Detection}
+\vspace{-10pt}
 \label{sec:res_tech:err_det}
 
-\vspace{-5pt}
 \subsubsection{Algorithmic Detection}
+\vspace{-10pt}
 \label{sec:res_tech:err_det:algo}
@@
 
+\vspace{-10pt}
 \subsection{Detection and Tolerance of Pointer Corruptions}
+\vspace{-10pt}
 \label{sec:res_tech:pointers}
@@
 
+\vspace{-10pt}
 \subsection{Recovery via Checkpoint-Restart and \\Replication}
+\vspace{-10pt}
 \label{sec:res_tech:cr}
@@
 
+\vspace{-10pt}
 \subsection{Evaluation of Subroutine Resilience}
+\vspace{-10pt}
 \label{sec:res_tech:eval}
@@
     \hline
-    \multicolumn{2}{|c|}{Routine}          & Algorithmic Detector             & Checkpointing         & Pointer Replication  \\
+    \multicolumn{2}{|c|}{Routine}          & Algorithmic                    & Checkpointing           & Pointer \\
+    \multicolumn{2}{|c|}{}                 & Detector                       & Checkpointing           & Replication  \\
     \hline
-    \multirow{4}{*}{ADDR}      & MM   & Linear encoding                     & Inputs                & No \\
-                               & SYRK & Thresholds:                         & Inputs                & No \\
-                               & MVM  & 1e-4 to 1e-7                        & Inputs                & No \\
-                               & CD   &                                     & Inputs                & No \\
+    \multirow{4}{*}{ADDR}      & MM   & Linear encoding                     & \multirow{4}{*}{Inputs} & \multirow{4}{*}{No} \\
+                               & SYRK & Thresholds:                         &                         &  \\
+                               & MVM  & 1e-4 to 1e-7                        &                         &  \\
+                               & CD   &                                     &                         &  \\
     \hline
-    \multirow{2}{*}{FRC}       & FFT  & Parseval's theorem.                 & Inputs                & No \\
-                               & FIR  & Sum conservation.                   & Inputs                & No \\
-                               &      & Thresholds: 1e-6 to 3e-8.           &                            \\
+    \multirow{2}{*}{FRC}       & FFT  & Parseval's theorem.                 & \multirow{4}{*}{Inputs} & \multirow{4}{*}{No} \\
+                               & FIR  & Sum conservation.                   &                         &   \\
+                               &      & Thresholds:                         &                         &   \\
+                               &      & 1e-6 to 3e-8.                       &                         &   \\
     \hline
-    \multirow{5}{*}{Hattrick}  & \multirow{5}{*}{RK} & \multirow{5}{*}{Variable step-size} & Periodic Timesteps    & None  \\
-                               &      &                                   & Period: 1, 1e3, 1e5   & All pointers, \\
-                               &      &                                   &                       & checked at one code location \\
-                               &      &                                   &                       & All pointers, \\
-                               &      &                                   &                       & checked on each use \\
+    \multirow{5}{*}{Hattrick}  & \multirow{5}{*}{RK} & \multirow{5}{*}{Variable step-size} & Periodic & \multicolumn{1}{|l|}{1. None}  \\
+                               &      &                                   & Timesteps      & \multicolumn{1}{|l|}{2. All pointers,} \\
+                               &      &                                   & Period:        & \multicolumn{1}{|l|}{checked at}  \\
+                               &      &                                   & 1, 1e3, 1e5    & \multicolumn{1}{|l|}{one code location}\\
+                               &      &                                   &                & \multicolumn{1}{|l|}{3. All pointers,} \\
+                               &      &                                   &                & \multicolumn{1}{|l|}{checked on each use} \\
     \hline
   \end{tabular}
+  \vspace{-10pt}
   \caption{The resilience techniques applied to each major routine of each application.}
+  \vspace{-15pt}
   \label{tbl:configs}
@@
 
+\vspace{-10pt}
 \subsubsection{Performance Impact}
+\vspace{-10pt}
 \label{sec:res_tech:eval:perf}
@@
 
+\vspace{-10pt}
 \subsubsection{Result Accuracy}
+\vspace{-10pt}
 \label{sec:res_tech:eval:accuracy}
@@
 
+\vspace{-10pt}
 \section{Evaluation of Resilience Methodology in Applications}
+\vspace{-10pt}
 \label{sec:eval}
@@
 
+\vspace{-10pt}
 \subsection{Extrapolation Model}
+\vspace{-10pt}
 \label{sec:eval:extrapolation}
@@
 \centering
-\includegraphics[width=1.00\columnwidth]{figs/extrapolates.png}
+\includegraphics[width=1.00\columnwidth]{figs/extrapolates.pdf}
 \caption{Extrapolation models}
@@
 
+\vspace{-10pt}
 \subsection{Intuitive Accuracy Metric}
+\vspace{-10pt}
 \label{sec:eval:errorsensitivity}
@@
 % Configurations are described twice.
+\vspace{-10pt}
 \subsection{Performance Impact}
+\vspace{-10pt}
 \label{sec:eval:perf}
@@
 
+\vspace{-10pt}
 \subsection{Result Accuracy}
+\vspace{-10pt}
 \label{sec:eval:acc}
@@
 
+\vspace{-10pt}
 \subsection{Choice of Optimal Fault Tolerance Techniques}
+\vspace{-10pt}
 \label{sec:eval:optchoice}
@@
 
+\vspace{-10pt}
 \section{Summary}
+\vspace{-10pt}
 \label{sec:summary}
@@
 \bibliography{bibs}
-
 \end{document}
